---
title: "Read Me"
date: "2026-03-04"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(igraph)
library(tibble)
library(dplyr)
library(ggraph)
library(tidygraph)
library(ggplot2)
```

# Project Overview
This document serves as a continuous research log for the Project. It tracks any important change, updates or key findings.

---

# Model

## Summary of the Current Base Code
The current implementation simulates a multi-community Liquid Democracy 
system in which lay agents probabilistically decide whether to vote directly 
or delegate their vote to neighbours or experts. Delegation decisions depend 
on ideological similarity and perceived competence, modelled through a 
logistic attractiveness function that incorporates both preference distance 
and accumulated voting power. The simulation runs over T rounds, dynamically 
updating delegation chains, computing transitive voting power, and propagating 
votes through the resulting delegation network.

---

### Core Technical Components

#### 1. Agent Architecture
* **Types**: The population consists of **Laypersons** (regular voters) and 
  **Experts** (specialised nodes). Only laypersons actively decide whether to 
  vote directly or delegate. Experts always vote directly with their own 
  preference and can receive delegations but never issue them.
* **Preferences**: Every agent is assigned a fixed random preference value 
  $pref_i \in [0,1]$, representing their ideological position. Preferences 
  do not change over the course of the simulation.
* **Communities**: Agents are organised into distinct communities to simulate 
  local social structures. Each community contains a mixture of laypersons 
  and a small number of experts.

---

#### 2. Network Topology

##### Friendship Network — Watts-Strogatz Small-World
The social network is constructed once at the start of the simulation and 
remains static across all T rounds. Construction proceeds in two stages.

**Stage 1 — Regular ring lattice:** Within each community, every lay agent 
is connected to their $k = \lfloor node\_degree / 2 \rfloor$ nearest 
neighbours on each side, forming a ring. All lay-to-lay edges are 
bidirectional, so any lay agent can delegate to any connected neighbour.

**Stage 2 — Watts-Strogatz rewiring:** Each directed edge is independently 
rewired to a uniformly chosen lay agent with probability $p_{rewire}$. This 
introduces long-range shortcuts while preserving local clustering.

| $p_{rewire}$ | Network structure |
|---|---|
| 0 | Pure ring — high clustering, long average paths |
| ~0.05 | Small-world — high clustering, short average paths |
| 1 | Random graph |

##### Expert Connectivity
Each expert is connected to $\lceil expert\_connectedness \times 
n\_per\_community \rceil$ randomly selected lay agents within their community. 
These edges run in one direction only (lay → expert), reflecting the 
assumption that experts always vote directly and are never in need of a 
delegate.

---

#### 3. Delegation Logic

At the start of every round $t$, each lay agent independently selects a 
delegation target from their candidate set: themselves (direct vote) or any 
out-neighbour in the friendship network.

##### Self-retention weight
The weight for voting directly is a logistic function of the agent's own 
current power:

$$
w_{self} = \frac{1}{1 + e^{-resp \cdot power_i}}
$$

An agent who already attracts many delegations (high power) is therefore more 
likely to continue voting directly. In round 1, all agents have $power = 1$, 
so $w_{self} = 0.5$ for every agent.

##### Neighbour attractiveness weight
The weight for delegating to neighbour $j$ combines ideological proximity 
with a competence signal:

$$
A_{ij} = \frac{1 - |pref_i - pref_j|}{1 + e^{-resp \cdot (power_j - power_i)}}
$$

* The first term equals 1 when $i$ and $j$ share the same preference and 
  falls to 0 for maximally different agents.
* The second term exceeds 0.5 whenever $j$ holds more power than $i$, 
  and drops below 0.5 otherwise.
* The parameter $resp$ controls the steepness of the competence sigmoid — 
  higher values make agents more sensitive to power differences.

A target is drawn randomly from the candidate set with probabilities 
proportional to these weights. This produces **endogenous persistence**: 
agents who accumulate power become more attractive as delegates in subsequent 
rounds, creating self-reinforcing delegation chains without any explicit 
memory parameter.

---

#### 4. Delegation Chains and Power Accumulation

Once delegation targets are chosen, a directed **delegation graph** $G_D$ is 
constructed. An edge $i \rightarrow j$ means agent $i$ delegates to agent $j$. 
Agents who vote directly have no outgoing edge.

This structure allows the formation of delegation chains, where votes pass 
through multiple intermediaries before reaching a final decision-maker:

$$A \rightarrow B \rightarrow C$$

Here $A$ delegates to $B$, who further delegates to $C$. Agent $C$ is the 
**chain root** — the agent who actually casts the vote. Every agent starts 
with $power = 1$ (their own vote); the root of a chain of length $k$ 
accumulates $power = k + 1$.

Roots are identified using a vectorized pointer-jumping algorithm: in each 
iteration, every agent simultaneously advances one step along their chain 
using integer indexing, requiring at most $L$ iterations where $L$ is the 
length of the longest chain. This avoids per-agent R loops and scales 
efficiently to large networks.

##### Cycles
If a group of agents form a mutual delegation cycle with no external root:

$$A \rightarrow B \rightarrow C \rightarrow A$$

no agent in the cycle casts a final vote. Agents in cycles retain $power = 1$ 
and their votes are recorded as lost ($my\_vote = \text{NA}$).

---

#### 5. Vote Propagation

After roots are identified, votes propagate as follows:

* **Direct voters and experts**: cast their own preference $pref_i$ as their 
  vote.
* **Delegating agents**: inherit the preference of their chain's root.
* **Cycle agents**: $my\_vote = \text{NA}$ (vote is lost).

When evaluating binary outcomes, preferences are interpreted using a threshold 
of $0.5$ (yes if $pref \geq 0.5$, no otherwise).

# Metrics (Summary Function)
To evaluate the state of the simulation, a comprehensive `summary_metrics` function is implemented. It summarizes key structural and behavioral properties of the final delegation network produced by the simulation.

## 1. network_description
This section characterizes the structural properties of the delegation network.

* **lost_vote_rate**: Measures the percentage of agents whose votes do not reach a final representative. Votes are lost when delegation chains form cycles.
* **vote_attribution_rate**: The proportion of agents whose votes successfully reach a final representative. Defined as $1 - \text{lost\_vote\_rate}$.
* **mean_power**: The average voting power across all agents in the network. Each agent starts with one vote; agents who receive delegations accumulate additional voting power.
* **max_power**: The highest voting power held by any agent in the network, indicating the maximum degree of vote concentration.
* **Gini Coefficient (Power)**: Measures inequality in the distribution of voting power across agents. A value of **0** represents perfect equality, while **1** indicates complete concentration of voting power in a single agent.
* **avg_chain_length**: The average number of delegation hops from an agent to its chain root, computed over all agents (including direct voters, for whom this value is 0). Higher values indicate a more indirect democracy.
* **total_components**: The number of weakly connected components in the delegation graph among agents who delegated or received a delegation. Isolated direct voters are excluded, as each would otherwise trivially inflate this count.
* **largest_voting_bloc_size**: The size of the largest weakly connected component in the delegation network, representing the largest single power bloc.

---

## 2. dynamic_evaluation
This section evaluates how delegation affects ideological representation and collective decisions.

* **avg_ideological_drift**: The average absolute difference between an agent's own preference and the final vote cast on their behalf, computed over represented agents only. Values near 0 indicate high-quality representation.
* **direct_yes / direct_no**: The number of yes (≥ 0.5) and no (< 0.5) votes under **direct democracy**, where every agent votes their own preference.
* **liquid_yes / liquid_no**: The number of yes and no votes under **liquid democracy**, based on the votes of final representatives after delegation chains are resolved. Note that `liquid_yes + liquid_no` will be smaller than `direct_yes + direct_no` whenever votes are lost to cycles, so the two vote counts are not directly commensurable in absolute terms.
* **liquid_margin**: The majority margin under liquid democracy, defined as `liquid_yes - liquid_no`. Positive values indicate a Yes majority; negative values indicate a No majority.
* **direct_margin**: The majority margin under direct democracy, defined as `direct_yes - direct_no`. Positive values indicate a Yes majority; negative values indicate a No majority.

Comparing `liquid_margin` to `direct_margin` shows whether delegation changes the direction or strength of the collective majority. Because cycle losses reduce the liquid vote count, a shift in margin does not necessarily reflect a genuine opinion change — it may partly reflect which side loses more votes to cycles.

---

## 3. power_by_type
This section compares voting power between different agent types.

* **mean_power**: The average voting power for each agent type (Laypersons vs. Experts).
* **max_power**: The maximum voting power held by any agent within each type.

This analysis helps identify whether experts systematically accumulate more voting power than regular voters within the delegation system.


# Visualisation

To complement the numerical metrics, a visualisation toolkit has been implemented
to evaluate the structural and qualitative performance of the model across different
complementary views.

```{r Functions}
library(here)
source("../scripts/Network.R")
source("../scripts/Summary.R")
source("../scripts/Visualisations.R")
```

## Example Simulation 
```{r run-simulation}
res <- simulate_liquid_democracy(
  seed                    = 123,
  n_per_community         = 20,
  n_communities           = 5,
  node_degree             = 6,
  n_experts_per_community = 2,
  expert_connectedness    = 0.2,
  p_rewire                = 0.05,
  responsiveness          = 1,
  T                       = 200
)
```
---

## 1. Multi-Layer Network Mapping

The structural visualisation distinguishes between the static social fabric and
the dynamic flow of delegated votes in the final round.

* **Community & Role**: Agents are coloured by their social community, while
  shapes distinguish between **Experts (Triangles)** and **Laypersons (Circles)**.
* **Dual-Edge Logic**: Faint grey lines represent static friendship ties, while
  dark arrows track the actual delegation edges resolved in round T.
* **Power Labels**: Numeric values on final voters (chain roots) indicate the
  total voting weight they carry into the collective decision.
```{r plot-network, fig.width=8, fig.height=6}
plot_social_and_delegation(res)
```

## 2. Summary Table

The summary table aggregates all key performance indicators across four
dimensions: representation quality, power inequality, opinion dynamics, and
network macro-structure.
```{r summary-table}
summary_table <- get_summary_table(res)

knitr::kable(
  summary_table,
  col.names = c("Dimension", "Metric", "Value"),
  caption   = "Liquid Democracy: System Performance Overview"
)
```

## 3. Distributional Diagnostics

Two charts provide a structural health check of the simulation in its final
state.

* **Ideological Drift (Representation Loss)**: Restricted to laypersons only,
  since experts always cast their own preference and have drift = 0 by
  definition. Measures the absolute distance between each layperson's true
  preference and the vote eventually cast on their behalf via delegation. A
  distribution concentrated near 0 indicates high-quality representation.
* **Power Concentration**: Ranks all agents by their accumulated voting power.
  Visually identifies the emergence of super-voters and shows whether experts
  are systematically accumulating more influence than laypersons as the model
  intends.
```{r plot-diagnostics, fig.width=8, fig.height=8}
plot_summary_stats(res)
```