---
title: "Visualtisation Report"
output: html_document
date: "2026-03-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(igraph)
library(tibble)
library(dplyr)
library(ggraph)
library(tidygraph)
library(ggplot2)
```

## Simulation run for Visualisation
```{r Functions}
library(here)
source("../scripts/Network.R")
source("../scripts/Summary.R")
source("../scripts/Visualisations.R")
```

## Simulation Execution
```{r run-simulation}
res <- simulate_liquid_democracy(
  seed                    = 123,
  n_per_community         = 20,
  n_communities           = 5,
  node_degree             = 6,
  n_experts_per_community = 2,
  expert_connectedness    = 0.2,
  p_rewire                = 0.05,
  responsiveness          = 1,
  T                       = 200
)
```
---

## [Update] - 2026-03-05: Visualisation

To complement the numerical metrics, a visualisation toolkit has been implemented
to evaluate the structural and qualitative performance of the model across three
complementary views.

### 1. Multi-Layer Network Mapping

The structural visualisation distinguishes between the static social fabric and
the dynamic flow of delegated votes in the final round.

* **Community & Role**: Agents are coloured by their social community, while
  shapes distinguish between **Experts (Triangles)** and **Laypersons (Circles)**.
* **Dual-Edge Logic**: Faint grey lines represent static friendship ties, while
  dark arrows track the actual delegation edges resolved in round T.
* **Power Labels**: Numeric values on final voters (chain roots) indicate the
  total voting weight they carry into the collective decision.
```{r plot-network, fig.width=8, fig.height=6}
plot_social_and_delegation(res)
```

### 2. Summary Table

The summary table aggregates all key performance indicators across four
dimensions: representation quality, power inequality, opinion dynamics, and
network macro-structure.
```{r summary-table}
summary_table <- get_summary_table(res)

knitr::kable(
  summary_table,
  col.names = c("Dimension", "Metric", "Value"),
  caption   = "Liquid Democracy: System Performance Overview"
)
```

### 3. Distributional Diagnostics

Two charts provide a structural health check of the simulation in its final
state.

* **Ideological Drift (Representation Loss)**: Restricted to laypersons only,
  since experts always cast their own preference and have drift = 0 by
  definition. Measures the absolute distance between each layperson's true
  preference and the vote eventually cast on their behalf via delegation. A
  distribution concentrated near 0 indicates high-quality representation.
* **Power Concentration**: Ranks all agents by their accumulated voting power.
  Visually identifies the emergence of super-voters and shows whether experts
  are systematically accumulating more influence than laypersons as the model
  intends.
```{r plot-diagnostics, fig.width=8, fig.height=8}
plot_summary_stats(res)
```