---
title: "Read Me"
date: "2026-03-28"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(igraph)
library(tibble)
library(dplyr)
library(ggraph)
library(tidygraph)
library(ggplot2)
```

# Model

## Summary of the Current Base Code
The current implementation simulates a multi-community Liquid Democracy
system in which lay agents probabilistically decide whether to vote directly
or delegate their vote to neighbours or experts. Delegation decisions depend
on ideological similarity and perceived competence, modelled through a
logistic attractiveness function that incorporates both preference distance
and accumulated voting power. The simulation runs over T rounds, dynamically
updating delegation chains, computing transitive voting power, and propagating
votes through the resulting delegation network.

---

### Core Technical Components

#### 1. Agent Architecture
* **Types**: The population consists of **Laypersons** (regular voters) and
  **Experts** (specialised nodes). Only laypersons actively decide whether to
  vote directly or delegate. Experts always vote directly with their own
  preference and can receive delegations but never issue them.
* **Preferences**: Every agent is assigned a fixed random preference value
  $pref_i \in [0,1]$, representing their ideological position. Preferences
  do not change over the course of the simulation.
* **Communities**: Agents are organised into distinct communities to simulate
  local social structures. Each community contains a mixture of laypersons
  and a small number of experts.

---

#### 2. Network Topology

##### Friendship Network — Watts-Strogatz Small-World
The social network is constructed once at the start of the simulation and
remains static across all T rounds. Construction proceeds in two stages.

**Stage 1 — Regular ring lattice:** Within each community, every lay agent
is connected to their $k = \lfloor node\_degree / 2 \rfloor$ nearest
neighbours on each side, forming a ring. All lay-to-lay edges are
bidirectional, so any lay agent can delegate to any connected neighbour.

**Stage 2 — Watts-Strogatz rewiring:** Each directed edge is independently
rewired to a uniformly chosen lay agent with probability $p_{rewire}$. This
introduces long-range shortcuts while preserving local clustering.

| $p_{rewire}$ | Network structure |
|---|---|
| 0 | Pure ring — high clustering, long average paths |
| ~0.05 | Small-world — high clustering, short average paths |
| 1 | Random graph |

##### Expert Connectivity
Each expert is connected to $\lceil expert\_connectedness \times
n\_per\_community \rceil$ randomly selected lay agents within their community.
These edges run in one direction only (lay → expert), reflecting the
assumption that experts always vote directly and are never in need of a
delegate.

---

#### 3. Delegation Logic

At the start of every round $t$, each lay agent independently selects a
delegation target from their candidate set: themselves (direct vote) or any
out-neighbour in the friendship network. The previous round's power values
are used when computing weights, so agents react to the state of the system
as it was, not as it is being updated.

##### Neighbour attractiveness weight

The weight for delegating to neighbour $j$ combines ideological proximity
with a competence signal:

$$
A_{ij} = \frac{1 - |pref_i - pref_j|}{1 + e^{-resp \cdot \log(power_j / power_i)}}
$$

* The numerator equals 1 when $i$ and $j$ share the same preference and
  falls to 0 for maximally different agents.
* The sigmoid term exceeds 0.5 whenever $j$ holds more power than $i$,
  and drops below 0.5 otherwise.
* The log-ratio input ensures that $resp$ retains a consistent interpretation
  across all rounds: a fixed power ratio (e.g., 5:1) produces the same sigmoid
  input regardless of the absolute scale of power, so $resp$ does not lose
  meaning after an early hierarchy forms.
* The parameter $resp$ controls the steepness of the sigmoid — higher
  values make agents more sensitive to relative power differences.

A target is drawn randomly from the candidate set with probabilities
proportional to these weights. This produces **endogenous persistence**:
agents who accumulate power become more attractive as delegates in subsequent
rounds, creating self-reinforcing delegation chains without any explicit
memory parameter.

##### Self-weight

The self-weight is computed in two steps. First, the agent uses $A_{ij}$
to identify $j^*$, the single most attractive available delegate — the
neighbour who best combines ideological proximity and accumulated power.
Then the agent compares its own power against $j^*$:

$$
j^* = \arg\max_{j \in N_i} A_{ij}
$$

$$
w_{self,i} = \frac{1}{1 + e^{-resp \cdot \log(power_i / power_{j^*})}}
$$

The behavioural assumption is: *"is there anyone who can represent me
better than I represent myself?"* Only the best available representative
is relevant to this comparison.

* If $power_i > power_{j^*}$: self-weight $> 0.5$, agent tends to vote directly.
* If $power_i < power_{j^*}$: self-weight $< 0.5$, agent tends to delegate.
* If no ideologically close neighbour exists, $A_{ij} \approx 0$ for all $j$,
  so all neighbour weights collapse to zero — the agent votes directly with
  near-certainty regardless of $w_{self}$.

When all agents have equal power (as in round 1), $\log(power_i / power_j) = 0$
and $sigmoid(0) = 0.5$ for every agent regardless of $resp$. For isolated agents
with no neighbours the fallback is $w_{self} = 0.5$.

Both $w_{self}$ and $A_{ij}$ use the same log-ratio sigmoid structure and the
same parameter $resp$, which therefore has a single consistent interpretation
throughout: **sensitivity to relative power differences**. Power must be $\geq 1$
at all times (every agent always holds at least their own vote), which guarantees
the log-ratio is well-defined.

##### Perception Noise (optional)

By default agents observe the true opinions and power of their neighbours.
Two optional parameters introduce noise into these perceptions:

| Parameter | Default | Description |
|---|---|---|
| `sigma_opinion` | 0 | SD of Gaussian noise added in logit space to perceived opinions; output is projected back to $[0,1]$ via $\text{plogis}$ |
| `sigma_pow` | 0 | SD of log-space multiplicative noise on perceived power; equivalent to scaling true power by $e^{\mathcal{N}(0, \sigma^2)}$ |

Setting both to 0 (the default) reproduces the original model where
perceived values equal true values. Each agent always perceives its own
opinion and power correctly; noise applies only to neighbours.

##### Formula Versioning

The attractiveness and self-weight formulas are passed as first-class
function arguments (`attractiveness_fn`, `selfweight_fn`) to
`simulate_liquid_democracy`. All formula versions — including archived
variants — are stored in `scripts/FunctionVersions.R`:

| Function | Type | Status |
|---|---|---|
| `attractiveness_absolute` | Neighbour attractiveness | Archived — uses raw power difference; saturates at high absolute power |
| `attractiveness_log` | Neighbour attractiveness | **Current default** — uses log-ratio; $resp$ stays interpretable across rounds |
| `selfweight_absolute_power` | Self-weight | Archived — proportional to own absolute power, no social comparison |
| `selfweight_mean_contest` | Self-weight | Archived — averages win-probabilities against all neighbours, ideology-blind |
| `selfweight_argmax` | Self-weight | Archived — argmax with absolute power differences |
| `selfweight_argmax_log` | Self-weight | **Current default** — argmax with log-ratio power; same scale invariance as `attractiveness_log` |

To switch formulas, pass a different function to the corresponding argument.
Old versions are preserved and must never be deleted.

##### Inertia (optional)
An optional `inertia` parameter (default 0) gives each lay agent a
probability of retaining their current delegate without re-evaluating
alternatives. At `inertia = 0` the system is fully memoryless — agents
reconsider completely every round.

---

#### 4. Delegation Chains and Power Accumulation

Once delegation targets are chosen, a directed **delegation graph** $G_D$ is
constructed. An edge $i \rightarrow j$ means agent $i$ delegates to agent $j$.
Agents who vote directly have no outgoing edge.

This structure allows the formation of delegation chains, where votes pass
through multiple intermediaries before reaching a final decision-maker:

$$A \rightarrow B \rightarrow C$$

Here $A$ delegates to $B$, who further delegates to $C$. Agent $C$ is the
**chain root** — the agent who actually casts the vote. Every agent starts
with $power = 1$ (their own vote); the root of a chain accumulates power
equal to 1 plus the number of agents in their chain.

Roots are identified using a vectorized pointer-jumping algorithm: in each
iteration, every agent simultaneously advances one step along their chain
using integer indexing, requiring at most $L$ iterations where $L$ is the
length of the longest chain.

##### Cycles
If a group of agents form a mutual delegation cycle with no external root:

$$A \rightarrow B \rightarrow C \rightarrow A$$

no agent in the cycle casts a final vote. Agents in cycles retain $power = 1$
and their votes are recorded as lost ($my\_vote = \text{NA}$).

---

#### 5. Vote Propagation

After roots are identified, votes propagate as follows:

* **Direct voters and experts**: cast their own preference $pref_i$ as their vote.
* **Delegating agents**: inherit the preference of their chain's root.
* **Cycle agents**: $my\_vote = \text{NA}$ (vote is lost).

When evaluating binary outcomes, preferences are interpreted using a threshold
of $0.5$ (yes if $pref \geq 0.5$, no otherwise).

---

# Simulation Output

`simulate_liquid_democracy()` returns a named list with the following elements:

| Element | Type | Description |
|---|---|---|
| `agents` | tibble | Final agent state (round T): `id`, `type`, `opinion`, `community`, `power`, `my_vote`, `delegated` |
| `history_lost` | numeric(T) | Fraction of agents with a lost vote (NA) at each round |
| `history_drift` | numeric(T) | Mean ideological drift among represented agents at each round |
| `history_delegation` | numeric(T) | Fraction of lay agents who delegated at each round |
| `history_stability` | numeric(T) | Fraction of lay agents keeping the same delegation target vs. the previous round |
| `snapshots` | tibble | Structured metrics recorded at 25%, 50%, 75%, and 100% of T (see below) |
| `delegation_graphs` | list(T) | Full igraph delegation graph for every round |
| `final_graph` | igraph | Delegation graph at round T (shorthand for `delegation_graphs[[T]]`) |
| `friendship_graph` | igraph | Static social network (Watts-Strogatz + expert edges) |

## Snapshots

The `snapshots` tibble provides a compact longitudinal view with one row per
snapshot round (25%, 50%, 75%, 100% of T):

| Column | What it records |
|---|---|
| `round`, `pct_T` | Round number and fraction of total T |
| `lost_vote_rate` | Fraction of agents with a lost vote |
| `avg_drift` | Mean ideological drift (represented agents) |
| `delegation_rate` | Fraction of lay agents who delegated |
| `gini_power` | Gini coefficient of power (represented agents) |
| `top5_power_share` | Top-5% power share (represented agents) |
| `avg_chain_length_delegators` | Mean chain length among delegating agents |
| `largest_voting_bloc_share` | Fraction of all agents in the largest delegation tree |
| `total_components` | Total independent voting groups |
| `delegation_stability` | Fraction of lay agents retaining the same target vs. prior round |
| `direct_yes / direct_no` | Yes/No counts under direct democracy |
| `direct_margin` | Raw majority margin under direct democracy |
| `liquid_yes / liquid_no` | Yes/No counts under liquid democracy |
| `liquid_margin` | Raw majority margin under liquid democracy |

---

# Metrics (Summary Function)

`summary_metrics(res)` evaluates the state of the simulation at round T across
four sections. Throughout, **represented agents** are agents with a valid
$my\_vote$ (non-NA) — i.e., all agents except those in delegation cycles.

---

## 1. Network Description

Structural properties of the final delegation graph.

| Metric | Agents included | What it measures |
|---|---|---|
| `lost_vote_rate` | Lay agents only | Fraction of lay agents in a delegation cycle (vote lost) |
| `delegation_rate` | Lay agents only | Fraction of lay agents who delegated in the final round |
| `max_power` | All agents | Absolute vote count of the most powerful agent |
| `max_power_share` | All agents | `max_power / total attributed votes` — normalised, comparable across simulation sizes |
| `max_indegree` | All agents | Maximum number of agents delegating *directly* (non-transitively) to one agent |
| `gini_power_nominal` | Represented agents only | Power inequality within the delegation hierarchy that operated; cycle agents excluded |
| `top5_power_share_nominal` | Represented agents only | Fraction of attributed votes held by the top 5%; cycle agents excluded |
| `gini_power_effective` | All lay agents | Power inequality across the full lay population; cycle agents enter as power = 0 |
| `top5_power_share_effective` | All lay agents | Top-5% power share over all lay agents; cycle agents enter as power = 0 |
| `avg_chain_length_all` | All agents (NA excluded) | Mean delegation depth; direct voters = 0, cycle agents excluded (NA) |
| `avg_chain_length_delegators` | Delegating agents only | Mean delegation depth among agents who delegated; more comparable across conditions with varying delegation rates |
| `max_chain_length` | Delegating agents only | Longest delegation chain in the final graph |
| `largest_voting_bloc_share` | All agents | Fraction of all agents in the largest connected delegation tree |
| `total_components` | All agents | Independent voting groups = delegation tree components + isolated direct voters |

**Nominal vs. effective power metrics:** Nominal metrics condition on
successfully represented agents and describe inequality within the operating
hierarchy. Effective metrics include cycle agents at power = 0 and reflect
system-wide inequality. Both should be reported together.

---

## 2. Dynamic Evaluation

How delegation affects ideological representation and collective decisions.

| Metric | Agents included | What it measures |
|---|---|---|
| `avg_ideological_drift` | Represented agents only | Mean \|preference − attributed vote\|; near 0 = high-quality representation |
| `med_ideological_drift` | Represented agents only | Median drift (robust to outliers) |
| `max_ideological_drift` | Represented agents only | Worst-case individual misrepresentation |
| `pct_high_drift` | Represented agents only | Fraction with drift > 0.3 |
| `misclassification_rate` | Represented agents only | Fraction whose attributed vote is on the *wrong side* of 0.5 relative to their own preference — outcome-relevant measure |
| `efficiency` | Lay agents | `(liquid_yes + liquid_no) / n_lay` — fraction of lay votes actually cast |
| `direct_yes / direct_no` | All agents | Yes/No vote counts under direct democracy (all agents vote own preference) |
| `liquid_yes / liquid_no` | Represented agents only | Yes/No vote counts under liquid democracy (cycle votes lost) |
| `direct_margin` | All agents | Raw Yes − No majority margin under direct democracy |
| `liquid_margin` | Represented agents only | Raw Yes − No margin under liquid democracy — **not directly comparable** to `direct_margin` due to different electorate sizes |
| `direct_margin_pct` | All agents | Proportional margin = `direct_margin / n_agents` |
| `liquid_margin_pct` | Represented agents only | Proportional margin = `liquid_margin / (liquid_yes + liquid_no)` — comparable to `direct_margin_pct` |
| `outcome_reversal` | — | Whether liquid and direct democracy disagree on Yes/No outcome; based on proportional margins |

**Drift caveat:** Drift is computed only for represented agents. If cycle
agents are non-randomly distributed across the preference space (which
depends on network topology), the estimate is biased. Always interpret
drift alongside `lost_vote_rate`.

---

## 3. Convergence

| Metric | What it measures |
|---|---|
| `convergence_sd` | SD of `lost_vote_rate` over the last 20 rounds — low = outcome has stabilised |
| `delegation_stability` | Mean fraction of lay agents keeping the same delegation target between consecutive rounds, averaged over the last 20 round-pairs |
| `edge_similarity_final` | Mean Jaccard similarity of the delegation graph between consecutive rounds, averaged over the last 20 round-pairs |
| `convergence_round` | First round from which `delegation_stability ≥ 0.99` holds for 50 consecutive rounds; NA if never reached within T |
| `cycle_indicator` | TRUE if stability < 0.99 but edge similarity > 0.95 — suggests oscillation of a subset of agents between fixed targets |
| `converged_to_dd` | TRUE if convergence was detected but `delegation_rate < 0.05` — flags degenerate convergence to near-direct-democracy |

---

## 4. Type-Specific Analysis

| Metric | What it measures |
|---|---|
| `mean_power` | Average voting power per agent type (lay vs. expert) |
| `max_power` | Maximum voting power per agent type |

Identifies whether experts systematically accumulate more voting power than
lay agents.

---

# Visualisation

To complement the numerical metrics, a visualisation toolkit has been
implemented to evaluate the structural and qualitative performance of the
model.

```{r Functions}
library(here)
source(here::here("scripts/FunctionVersions.R"))   # formula version library
source(here::here("scripts/Network.R"))
source(here::here("scripts/Summary.R"))
source(here::here("scripts/Visualisations.R"))
```

## Example Simulation
```{r run-simulation}
res <- simulate_liquid_democracy(
  seed                    = 123,
  n_per_community         = 20,
  n_communities           = 5,
  node_degree             = 6,
  n_experts_per_community = 2,
  expert_connectedness    = 0.2,
  p_rewire                = 0.05,
  responsiveness          = 1,
  inertia                 = 0,
  T                       = 200,
  selfweight_fn           = selfweight_argmax_log,  # default: log-ratio argmax
  attractiveness_fn       = attractiveness_log,     # default: log-ratio attractiveness
  sigma_opinion           = 0,   # perception noise on opinions (0 = no noise)
  sigma_pow               = 0    # perception noise on power    (0 = no noise)
)
```
---

## 1. Multi-Layer Network Mapping

The structural visualisation distinguishes between the static social fabric and
the dynamic flow of delegated votes in the final round.

* **Community & Role**: Agents are coloured by their social community, while
  shapes distinguish between **Experts (Triangles)** and **Laypersons (Circles)**.
* **Dual-Edge Logic**: Faint grey lines represent static friendship ties, while
  dark arrows track the actual delegation edges resolved in round T.
* **Power Labels**: Numeric values on final voters (chain roots) indicate the
  total voting weight they carry into the collective decision.
```{r plot-network, fig.width=8, fig.height=6}
plot_social_and_delegation(res)
```

## 2. Summary Table

The summary table aggregates key performance indicators across representation
quality, power inequality, opinion dynamics, and network macro-structure.
```{r summary-table}
summary_table <- get_summary_table(res)

knitr::kable(
  summary_table,
  col.names = c("Dimension", "Metric", "Value"),
  caption   = "Liquid Democracy: System Performance Overview"
)
```

## 3. Distributional Diagnostics

Two charts provide a structural health check of the simulation in its final
state.

* **Ideological Drift (Representation Loss)**: Restricted to represented agents
  only (cycle agents excluded). Measures the absolute distance between each
  agent's true preference and the vote eventually cast on their behalf. A
  distribution concentrated near 0 indicates high-quality representation.
* **Power Concentration**: Ranks all agents by their accumulated voting power.
  Visually identifies the emergence of super-voters and shows whether experts
  are systematically accumulating more influence than laypersons.
```{r plot-diagnostics, fig.width=8, fig.height=8}
plot_summary_stats(res)
```
