---
title: "Project Log: Liquid Democracy Simulation"
date: "2026-02-20"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(igraph)
library(tibble)
library(dplyr)
library(ggraph)
library(tidygraph)
library(ggplot2)
```

# Project Overview
This document serves as a continuous research log for the whole Project. It tracks any important change, parameter experiments, and key findings.

---

# [Baseline] - Initial Model Implementation

## Summary of the Current Base Code
The baseline version implements a multi-community Liquid Democracy environment. The simulation explores how voting power concentrates as agents delegate their votes to peers or experts based on ideological proximity and perceived influence.

### Core Technical Components:

#### 1. Agent Architecture
* **Types**: The population consists of **Laypersons** (regular voters) and **Experts** (specialized nodes).
* **Preferences**: Every agent is assigned a random preference value between $0$ and $1$, representing their ideological position.
* **Communities**: Agents are organized into distinct communities to simulate local social structures.

#### 2. Network Topologies
* **Friendship Rings**: Each community starts with a ring-lattice structure, representing a local social circle where everyone is connected to their immediate neighbors.
* **Expert Connectivity**: A defined percentage of laypersons in each community is connected to local experts.
* **Maslov-Sneppen Rewiring**: The model includes an optional rewiring mechanism to transform the local rings into Small-World networks, allowing for cross-community interactions.

#### 3. Delegation Logic
The decision to delegate is stochastic and based on an **Attractiveness Score**. An agent $i$ considers delegating to a friend $j$ based on:
* **Preference Similarity**: Agents prefer delegates with similar ideological positions.
* **Power Responsiveness**: Agents are attracted to delegates who already hold significant voting power.

The probability of delegation is calculated using a logistic function:
$$Attractiveness = \frac{1 - |pref_i - pref_j|}{1 + e^{-resp \cdot (power_j - power_i)}}$$

---

## Baseline Metrics (Summary Function)
To evaluate the state of the simulation, a comprehensive `summary_metrics` function has been implemented. It tracks the following key indicators:

### 1. network_description
* **lost_vote_rate**: Measures the percentage of agents whose votes do not reach a final representative (due to cycles or isolation).
* **vote_attribution_rate**: Measures the percentages of agents whose vote do reach a final representative (1 - lost_vote_rate).
* **mean_power**: Measures the average power of all agents within the network. 
* **max_power**: Identifies the agent with max power in the network.
* **Gini Coefficient (Power)**: Quantifies the inequality of voting power across the network ($0$ = perfect equality, $1$ = total concentration).
* **avg_chain_length**: The average length of the delegation chains, indicating how "liquid" or direct the democracy is.
* **total_components**: Counts the number of isolated delegation clusters.
* **largest_voting_bloc_size**: Measures the size of the largest voting bloc.

### 2. dynamic_evaluation
* **avg_ideological_drift**: Measures the average absolute deviation between an agent’s original preference and the final vote cast on their behalf (individual-level).
* **systemic_bias**: Measures difference between collective mean final vote and collective mean original preference. (collective-level).
* **direct_yes / direct_no**: Counts of yes (≥ 0.5) and no (< 0.5) votes under direct democracy, where each agent votes according to their own preference.
* **liquid_yes / liquid_no**: Counts of yes and no votes under liquid democracy, based on the final representative votes.
**direct_liquid_shift**: Measures the difference between direct and liquid majority outcomes.
It captures whether delegation changes the collective decision compared to direct voting.

### 3. power_by_type
* **mean_power & max_power**: Type-specific analysis function (lay/expert) in terms of power.

---


## Simulation run for Visualisation
```{r Functions}
source("Network.R") 
source("Summary.R")
source("Visualisations.R")
```

## Simulation Execution

```{r run-simulation}
set.seed(123)

res <- simulate_liquid_democracy(
  seed = 123,
  number_of_nodes_per_community = 20,
  number_of_communities = 5,
  node_degree = 6,
  number_of_experts_per_community = 2,
  expert_connectedness = 0.2,
  MS_rewiring_iterations = 50,
  responsiveness_to_power = 1,
  T = 200 # number of delegation updates
)
```
---

## [Update] - 2026-02-22: Visualisation

To complement the numerical metrics, a visualization toolkit has been implemented to evaluate the structural and qualitative performance of the models.

### 1. Multi-Layer Network Mapping
The structural visualization distinguishes between the static social fabric and the dynamic flow of power.
* **Community & Role**: Agents are colored by their social community, while shapes distinguish between **Experts (Triangles)** and **Laypersons (Circles)**.
* **Dual-Edge Logic**: Faint grey lines represent potential social ties (Friendship), while dark arrows track the actual path of delegated votes.
* **Power Labels**: Numeric values indicate the final voting weight concentrated in the hands of the representatives.
```{r plot-social-and_delegation, fig.width=8, fig.height=6}
# Visualizing the Network Structure
plot_social_and_delegation(res)
```

### 2. Summary Table
Contains all relevant metrics.
```{r summary-table}
summary_table <- get_summary_table(res)

knitr::kable(
  summary_table,
  col.names = c("Dimension", "Metric", "Value"),
  caption = "Liquid Democracy: System Performance Overview"
)
```

### 3. Distributional Diagnostics
Two key charts provide a "health check" of the current simulation state:
* **Ideological Drift (Representation Loss)**: This focuses exclusively on laypersons. It measures the distance between an agent's true preference and the vote eventually cast by their delegate. A concentration near 0 indicates high-quality representation.
* **Power Concentration**: This ranks all agents by influence. It visually identifies the emergence of "Super-Voters" and confirms whether experts are accumulating the majority of the community's influence as intended.

```{r plot-summary-stats, fig.width=8, fig.height=8}
plot_summary_stats(res)
```

