---
title: "Project Log: Liquid Democracy Simulation"
date: "2026-02-22"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(igraph)
library(tibble)
library(dplyr)
library(ggraph)
library(tidygraph)
library(ggplot2)
```

# Project Overview
This document serves as a continuous research log for the whole Project. It tracks any important change, parameter experiments, and key findings.

---

# [Baseline] - Initial Model Implementation

## Summary of the Current Base Code
The baseline version implements a multi-community Liquid Democracy environment. The simulation explores how voting power concentrates as agents delegate their votes to peers or experts based on ideological proximity and perceived influence.

### Core Technical Components:

#### 1. Agent Architecture
* **Types**: The population consists of **Laypersons** (regular voters) and **Experts** (specialized nodes).
* **Preferences**: Every agent is assigned a random preference value between $0$ and $1$, representing their ideological position.
* **Communities**: Agents are organized into distinct communities to simulate local social structures.

#### 2. Network Topologies
* **Friendship Rings**: Each community starts with a ring-lattice structure, representing a local social circle where everyone is connected to their immediate neighbors.
* **Expert Connectivity**: A defined percentage of laypersons in each community is connected to local experts.
* **Maslov-Sneppen Rewiring**: The model includes an optional rewiring mechanism to transform the local rings into Small-World networks, allowing for cross-community interactions.

#### 3. Delegation Logic
The decision to delegate is stochastic and based on an **Attractiveness Score**. An agent $i$ considers delegating to a friend $j$ based on:
* **Preference Similarity**: Agents prefer delegates with similar ideological positions.
* **Power Responsiveness**: Agents are attracted to delegates who already hold significant voting power.

The probability of delegation is calculated using a logistic function:
$$Attractiveness = \frac{1 - |pref_i - pref_j|}{1 + e^{-resp \cdot (power_j - power_i)}}$$

#### 4. Vote Propagation & Metrics
* **Power Counting**: The model recursively counts how many people have delegated (directly or indirectly) to a specific node.
* **Vote Attribution**: Preferences are propagated from the final representatives back through the delegation chain.
* **Lost Votes**: The system tracks "lost votes," which occur when delegation cycles form or when agents have no path to a final representative.

---

## Baseline Metrics (Summary Function)
To evaluate the state of the simulation, a comprehensive `summary_metrics` function has been implemented. It tracks the following key indicators:

### 1. Representation & Efficiency
* **Lost Vote Rate**: Measures the percentage of agents whose votes do not reach a final representative (due to cycles or isolation).
* **Average Delegation Distance**: The average length of the delegation chains, indicating how "liquid" or direct the democracy is.

### 2. Power Distribution & Inequality
* **Gini Coefficient (Power)**: Quantifies the inequality of voting power across the network ($0$ = perfect equality, $1$ = total concentration).
* **Max/Mean Power**: Identifies the influence of the strongest "super-voters" compared to the average agent.

### 3. Network Macro-Structure
* **Components**: Counts the number of isolated delegation clusters.
* **Largest Component Size**: Monitors the size of the dominant voting bloc.

### 4. Expert vs. Layperson Dynamics
* **Comparative Power**: Groups results by agent type to analyze if experts actually accumulate the majority of the community's influence as intended by the model design.

---

## [Update] - 2026-02-22: Visualisation

To complement the numerical metrics, a visualization toolkit has been implemented to evaluate the structural and qualitative performance of the models.

### 1. Multi-Layer Network Mapping
The structural visualization distinguishes between the static social fabric and the dynamic flow of power.
* **Community & Role**: Agents are colored by their social community, while shapes distinguish between **Experts (Triangles)** and **Laypersons (Circles)**.
* **Dual-Edge Logic**: Faint grey lines represent potential social ties (Friendship), while dark arrows track the actual path of delegated votes.
* **Power Labels**: Numeric values indicate the final voting weight concentrated in the hands of the representatives.
```{r update_2026_02_22, echo=TRUE}
# 1. Visualizing the Network Structure
# Shows who knows whom vs. who actually votes for whom
plot_social_and_delegation(res)
```
### 2. Summary Table
Contains all relevant metrics.
```{r update_2026_02_22, echo=TRUE}
# 3. Quantitative Performance Summary
# Generates a clean table for the report
summary_table <- get_summary_table(res)
knitr::kable(summary_table, 
             col.names = c("Dimension", "Metric", "Value"),
             caption = "Liquid Democracy: System Performance Overview")
```

### 3. Distributional Diagnostics
Two key charts provide a "health check" of the current simulation state:
* **Ideological Drift (Representation Loss)**: This focuses exclusively on laypersons. It measures the distance between an agent's true preference and the vote eventually cast by their delegate. A concentration near 0 indicates high-quality representation.
* **Power Concentration**: This ranks all agents by influence. It visually identifies the emergence of "Super-Voters" and confirms whether experts are accumulating the majority of the community's influence as intended.

```{r update_2026_02_22, echo=TRUE}
# 2. Statistical Analysis of Drift and Power
# Top: How much is the opinion distorted for laypersons?
# Bottom: How concentrated is the power among experts?
plot_summary_stats(res)
```

