---
title: "Delegation Mechanics: Legacy Dual-Sigmoid vs. Gaussian/Mobius Decision Model"
subtitle: "A Technical Reference for the Liquid Democracy Simulation (accompanies Report 20)"
date: "2026-07-30"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: show
    theme: flatly
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 9, fig.height = 5.5, fig.align = "center")
library(tidyverse)
library(here)
library(patchwork)
library(knitr)

theme_set(theme_minimal(base_size = 12))

# This document sources the ACTUAL production code -- it is not a
# reimplementation. Every formula plotted or tabulated below is the exact
# function used inside simulate_liquid_democracy() in Network.R. If the
# formulas in FunctionVersions.R / Network.R ever change, re-knitting this
# document will automatically reflect the update (or fail loudly if an
# assumption below no longer holds).
source(here::here("scripts/Network.R"))
```

# Purpose and Scope

This note is a self-contained technical reference explaining the
delegation-decision mechanics of the liquid-democracy simulation, comparing
the **legacy** dual-sigmoid model (used in Reports 15-19) against the **new**
Gaussian-opinion / Mobius-self-vote model introduced for Report 20
(`decision_model = "gaussian_mobius"` in `simulate_liquid_democracy()`).

Both formula sets are switched with a single argument; everything else in
the simulation -- network construction, the trust update, ingroup
preference, cycle handling, homophily shuffling, and all recorded metrics --
is **completely unchanged** between the two models. This note therefore
focuses exclusively on the one part of the model that *did* change: how
attractive a neighbour looks to an agent, and how that agent decides whether
to vote directly or delegate.

Throughout, $i$ is the focal agent, $j$ ranges over $i$'s network
neighbours, $o$ denotes opinion, $p$ denotes current voting power, and
$j^\*= \arg\max_j A_{ij}$ is $i$'s single most attractive neighbour.

---

# Part I -- The New Decision Model (`gaussian_mobius`)

## Overview: the three-step pipeline

Every round, for every lay agent $i$ with at least one neighbour, the
simulation resolves a delegation decision in three steps:

```{r pipeline-diagram-new, fig.height=3.2, echo=FALSE}
draw_pipeline <- function(labels, fills, title) {
  n <- length(labels)
  boxes <- tibble(
    x    = seq_len(n),
    label = labels,
    fill  = fills
  )
  arrows <- tibble(x = boxes$x[-n] + 0.38, xend = boxes$x[-1] - 0.38, y = 1, yend = 1)

  ggplot() +
    geom_rect(data = boxes, aes(xmin = x - 0.38, xmax = x + 0.38,
                                 ymin = 0.6, ymax = 1.4, fill = fill),
              colour = "grey30", linewidth = 0.6) +
    geom_text(data = boxes, aes(x = x, y = 1, label = label),
              size = 3.4, lineheight = 0.95) +
    geom_segment(data = arrows, aes(x = x, xend = xend, y = y, yend = yend),
                 arrow = arrow(length = unit(0.2, "cm")), linewidth = 0.6, colour = "grey30") +
    scale_fill_identity() +
    coord_cartesian(xlim = c(0.4, n + 0.6), ylim = c(0.3, 1.7)) +
    theme_void() +
    labs(title = title) +
    theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 12))
}

draw_pipeline(
  labels = c("Step 1\nAttractiveness\nA_ij for every\nneighbour j",
             "Step 2\nSelf-vote prob.\nP_self(m),\nm = max_j A_ij",
             "Step 3\nResolve:\nself-vote w.p. P_self,\nelse delegate ~ A_ij"),
  fills = c("#cfe8f3", "#f6d6ac", "#d9ead3"),
  title = "gaussian_mobius pipeline"
)
```

Step 1 computes a full vector of attractivenesses (one per neighbour). Step
2 collapses that vector to a single scalar $m$ (the best neighbour's
attractiveness) and maps it to a self-vote probability. Step 3 draws a
Bernoulli($P_{\text{self}}$) coin; on failure, agent $i$ delegates to a
neighbour sampled proportional to attractiveness. This is a genuine
information bottleneck: **Step 2 never looks at *who* $j^\*$ is (their
opinion, their power) -- only at *how attractive* they are.** Keep this in
mind; it is the single biggest structural difference from the legacy model,
discussed at length in Part III.

## Step 1: Attractiveness

$$A_{ij} = 2\,\tau_{ij}\,a^{\text{op}}_{ij}\,a^{\text{pw}}_{ij}$$

implemented as `attractiveness_gaussian_log()` in `FunctionVersions.R`. It
returns $2\,a^{\text{op}}_{ij}\,a^{\text{pw}}_{ij}$; the trust factor
$\tau_{ij}$ is applied afterwards, exactly as for every other attractiveness
formula in this codebase (see Part I.5).

### Opinion factor: a Gaussian kernel

$$a^{\text{op}}_{ij} = \exp\!\left(-\tfrac{1}{2}\,r_{op}^2\,(o_i - o_j)^2\right)$$

This is a Gaussian bump centred on $o_i$ with standard deviation
$\sigma_{\text{op}} = 1/r_{op}$: it equals exactly $1$ when $j$'s opinion is
identical to $i$'s, and decays smoothly to $0$ as the opinion gap grows.
$r_{op}$ is an inverse *width*: larger $r_{op}$ means a narrower, more
selective bump (only near-identical neighbours count as attractive);
smaller $r_{op}$ means a wide, forgiving bump (almost any neighbour scores
close to 1 on the opinion dimension).

```{r new-op-kernel-plot, fig.height=4.5}
op_dist <- seq(0, 1, length.out = 300)
r_op_vals <- c(0.5, 1, 2, 4, 8)

df_op_new <- expand_grid(delta = op_dist, r_op = r_op_vals) |>
  # attractiveness_gaussian_log(op_i, op_j, pow_i, pow_j, r_op, r_pw) returns
  # 2*a_op*a_pw. With pow_i = pow_j = 1, a_pw = sig(0) = 0.5 exactly, so
  # 2*a_pw = 1 -- the real production function's output already equals
  # a_op(delta) in isolation, with NO further division needed. (Do not divide
  # by 0.5 here: that would double-count the leading "2*" already baked into
  # the function, unlike attractiveness_dual_sigmoid below which has no such
  # factor -- verified against the reference ldmodel.py numerically.)
  mutate(a_op = attractiveness_gaussian_log(0, delta, 1, 1, r_op, 1),
         r_op_label = factor(paste0("r_op = ", r_op), levels = paste0("r_op = ", r_op_vals)))

ggplot(df_op_new, aes(delta, a_op, colour = r_op_label)) +
  geom_line(linewidth = 1) +
  scale_colour_viridis_d(end = 0.85) +
  labs(title = "New opinion factor a_op(Delta) -- Gaussian kernel",
       subtitle = "a_op = exp(-0.5 * r_op^2 * Delta^2); Delta = |o_i - o_j|",
       x = expression(Delta == "|"*o[i]-o[j]*"|"), y = expression(a[op]),
       colour = NULL) +
  theme(legend.position = "bottom")
```

**Numeric example.** At $r_{op}=4$:

```{r new-op-kernel-table}
tibble(Delta = c(0, 0.1, 0.25, 0.5, 0.75, 1)) |>
  mutate(a_op = attractiveness_gaussian_log(0, Delta, 1, 1, 4, 1)) |>
  kable(digits = 4, col.names = c("Opinion distance $\\Delta$", "$a^{\\text{op}}$ (r_op = 4)"))
```

A neighbour half a unit away on the $[0,1]$ opinion scale is already down to
about 14% attractiveness at this responsiveness level (`r round(attractiveness_gaussian_log(0, 0.5, 1, 1, 4, 1), 3)`
exactly), and a neighbour three-quarters of a unit away is down to about 1%
-- opinion gating is aggressive once $r_{op}$ is moderately large.

### Power factor: unchanged from the legacy model

$$a^{\text{pw}}_{ij} = \sigma\!\left(r_{pw}\,\log\frac{p_j}{p_i}\right), \qquad \sigma(x) = \frac{1}{1+e^{-x}}$$

This half of the formula is **identical in form** to the legacy model (Part
II) -- a sigmoid of the log power ratio. It equals $0.5$ when $i$ and $j$
hold equal power, saturates toward $1$ as $j$ becomes arbitrarily more
powerful, and toward $0$ as $j$ becomes negligible. $r_{pw}$ is the
steepness of that saturation.

```{r new-pw-kernel-plot}
pow_ratio <- 10^seq(-2, 2, length.out = 300)
r_pw_vals <- c(0.5, 1, 2, 4)

df_pw <- expand_grid(ratio = pow_ratio, r_pw = r_pw_vals) |>
  # Delta_op = 0 forces a_op = 1 exactly, so the real function's output /
  # 2 isolates a_pw(ratio). This is the SAME function used for the legacy
  # power factor -- see Part II.2.
  mutate(a_pw = attractiveness_gaussian_log(0, 0, 1, ratio, 1, r_pw) / 2,
         r_pw_label = factor(paste0("r_pw = ", r_pw), levels = paste0("r_pw = ", r_pw_vals)))

ggplot(df_pw, aes(ratio, a_pw, colour = r_pw_label)) +
  geom_line(linewidth = 1) +
  geom_hline(yintercept = 0.5, linetype = "dotted", colour = "grey50") +
  geom_vline(xintercept = 1, linetype = "dotted", colour = "grey50") +
  scale_x_log10() +
  scale_colour_viridis_d(end = 0.85, option = "magma") +
  labs(title = "Power factor a_pw(power ratio) -- identical in old and new models",
       subtitle = "a_pw = sig(r_pw * log(p_j / p_i)); = 0.5 at equal power (dotted lines)",
       x = expression(p[j]/p[i]~"(log scale)"), y = expression(a[pw]), colour = NULL) +
  theme(legend.position = "bottom")
```

```{r new-pw-kernel-table}
tibble(ratio = c(0.1, 0.5, 1, 2, 10)) |>
  mutate(a_pw = attractiveness_gaussian_log(0, 0, 1, ratio, 1, 2) / 2) |>
  kable(digits = 4, col.names = c("Power ratio $p_j/p_i$", "$a^{\\text{pw}}$ (r_pw = 2)"))
```

## Step 2: Self-vote probability -- the Mobius curve

Let $m_i = \max_j A_{ij}$ (the attractiveness of $i$'s single best
neighbour, on the $[0,2]$ scale established above). The self-vote
probability is:

$$P_{\text{self}}(m) = \frac{2 - m}{\,2 + \dfrac{1-2c}{c}\,m\,}$$

implemented as `p_self_mobius()`. It is a Mobius (rational-linear)
transform pinned at three anchors:

| best neighbour   | $m$ | $P_{\text{self}}$ | meaning              |
|---               |---  |---                 |---                   |
| nightmare        | $0$ | $1$                | always self-vote     |
| identical to me  | $1$ | $c$                | self-reliance knob    |
| perfect          | $2$ | $0$                | always delegate       |

$c \in (0,1)$ is the **entire** free parameter of this step: it is set
directly (`self_reliance_c` in `simulate_liquid_democracy()`) and does not
depend on $r_{op}$, $r_{pw}$, or anything about $j^\*$ specifically -- see
Part III.2 for why this matters.

```{r new-selfvote-plot, fig.height=4.5}
m_grid <- seq(0, 2, length.out = 300)
c_vals <- c(0.1, 0.2, 0.5, 0.8)

df_pself <- expand_grid(m = m_grid, c = c_vals) |>
  mutate(P_self = p_self_mobius(m, c),
         c_label = factor(paste0("c = ", c), levels = paste0("c = ", c_vals)))

ggplot(df_pself, aes(m, P_self, colour = c_label)) +
  geom_line(linewidth = 1) +
  geom_point(data = tibble(m = c(0, 1, 2), P_self = c(1, NA, 0)),
             aes(m, P_self), inherit.aes = FALSE, colour = "grey30", size = 0, alpha = 0) +
  annotate("point", x = 0, y = 1, colour = "grey20", size = 2) +
  annotate("point", x = 2, y = 0, colour = "grey20", size = 2) +
  annotate("text", x = 0.05, y = 1.04, label = "m=0 -> always self-vote", hjust = 0, size = 3) +
  annotate("text", x = 1.95, y = 0.06, label = "m=2 -> always delegate", hjust = 1, size = 3) +
  scale_colour_viridis_d(end = 0.85, option = "plasma") +
  labs(title = "Self-vote probability P_self(m) -- Mobius curve",
       subtitle = "m = attractiveness of the single best neighbour, on [0,2]",
       x = expression(m == max[j]~A[ij]), y = expression(P[self]), colour = "self-reliance c") +
  theme(legend.position = "bottom")
```

**Numeric example.** At $m=1$ (an exactly identical best neighbour),
$P_{\text{self}} = c$ *by construction, for every value of $r_{op}$ and
$r_{pw}$*:

```{r new-selfvote-table}
tibble(c = c(0.1, 0.2, 0.5, 0.8)) |>
  mutate(`P_self(m=1)` = p_self_mobius(1, c)) |>
  kable(digits = 3)
```

## Step 3: Delegate choice

Conditional on not self-voting, $i$ delegates to a neighbour sampled with
probability proportional to $A_{ij}$ among its real (non-padding)
neighbours -- a straightforward weighted draw, **structurally identical**
to the legacy model's delegate-choice step (`nb[sample.int(length(w), 1L,
prob = w)]` in `Network.R`, unchanged by `decision_model`). What differs
between the two models is only the *shape* of the weights being sampled
from, not the sampling rule itself.

```{r new-delegate-choice-example, fig.height=4}
A_vals <- c(j1 = 0.2, j2 = 0.8, j3 = 1.5, j4 = 0.1)
df_choice <- tibble(neighbour = names(A_vals), A = A_vals) |>
  mutate(P_delegate_to = A / sum(A))

ggplot(df_choice, aes(neighbour, P_delegate_to, fill = neighbour)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = sprintf("A=%.1f\nP=%.2f", A, P_delegate_to)), vjust = -0.15, size = 3.3) +
  scale_fill_viridis_d(end = 0.85) +
  coord_cartesian(ylim = c(0, max(df_choice$P_delegate_to) * 1.35)) +
  labs(title = "Delegate choice: weighted sampling proportional to A_ij (illustrative example)",
       subtitle = "4 neighbours with attractivenesses 0.2, 0.8, 1.5, 0.1 -> delegation probabilities, given not self-voting",
       x = NULL, y = "P(delegate to j | delegating)")
```

## Trust interaction (unchanged mechanism, recap only)

Trust is **not** part of the new formula set -- it is the existing Report
18 mechanism, applied identically regardless of `decision_model`:

$$\texttt{trust\_mod}_{ij} = 2\,\sigma(\gamma\,\tau_{ij}), \qquad w_{ij} = \texttt{trust\_mod}_{ij} \times A_{ij}$$

with $\tau_{ij}$ updated each round via the `punish_only` / `reward_punish`
rules documented in `Network.R`. `trust_mod = 1` (neutral) when $\gamma=0$
or $\tau_{ij}=0$; it can exceed $1$ (boosting a well-trusted neighbour) or
fall toward $0$ (suppressing a mistrusted one).

```{r trust-mod-plot, fig.height=4}
tau_grid <- seq(-3, 3, length.out = 200)
gamma_vals <- c(0.5, 1, 2, 4)

df_trust <- expand_grid(tau = tau_grid, gamma = gamma_vals) |>
  mutate(trust_mod = 2 * .sig(gamma * tau),
         gamma_label = factor(paste0("gamma = ", gamma), levels = paste0("gamma = ", gamma_vals)))

ggplot(df_trust, aes(tau, trust_mod, colour = gamma_label)) +
  geom_line(linewidth = 1) +
  geom_hline(yintercept = 1, linetype = "dotted", colour = "grey50") +
  geom_hline(yintercept = 2, linetype = "dashed", colour = "firebrick3") +
  annotate("text", x = -2.8, y = 2.05, label = "trust_mod ceiling = 2", colour = "firebrick3", size = 3, hjust = 0) +
  scale_colour_viridis_d(end = 0.85, option = "cividis") +
  labs(title = "trust_mod(tau) = 2*sig(gamma*tau) -- unchanged in both decision models",
       x = expression(tau[ij]), y = "trust_mod", colour = NULL) +
  theme(legend.position = "bottom")
```

Note the dashed ceiling at $\texttt{trust\_mod}=2$: this will matter in Part
III.6, because the new model's $m$ is computed from the *trust-adjusted*
weight $w_{j^\*}$, not the raw attractiveness.

## Parameter glossary (new model)

```{r param-glossary-new}
tibble(
  Parameter = c("r_op", "r_pw", "self_reliance_c", "gamma", "lambda", "trust_mode"),
  Controls = c(
    "Inverse width of the Gaussian opinion kernel a_op. Larger = narrower = more opinion-selective.",
    "Steepness of the power-ratio sigmoid a_pw. Larger = more power-deferential.",
    "Self-reliance knob c in the Mobius P_self curve. Larger = more self-reliant (delegates only to clearly-better neighbours). Independent of r_op/r_pw.",
    "Trust sensitivity: how strongly accumulated trust tau rescales attractiveness (unchanged mechanism).",
    "Trust decay/momentum: how much of last round's tau carries over (unchanged mechanism).",
    "'punish_only' or 'reward_punish' trust update rule (unchanged mechanism)."
  ),
  `Typical range` = c("0.5 - 8", "0.5 - 8", "0.1 - 0.9 (open interval)", "0 (off) - 8", "0 - 1", "categorical")
) |>
  kable()
```

---

# Part II -- Recap: The Legacy Dual-Sigmoid Model

Reports 15-19 use `decision_model = "legacy"` (the default), summarised
here for direct comparison.

## Attractiveness (dual sigmoid)

$$A_{ij} = \sigma\big(r_{op}(1-2|o_i-o_j|)\big)\,\times\,\sigma\!\left(r_{pw}\log\frac{p_j}{p_i}\right)$$

implemented as `attractiveness_dual_sigmoid()`. Both factors are sigmoids:
opinion similarity $1-2|o_i-o_j|$ is remapped from $[-1,1]$ (identical to
maximally distant) through a sigmoid, and the power factor is the same
log-ratio sigmoid used in the new model.

```{r old-op-kernel-plot, fig.height=4.5}
df_op_old <- expand_grid(delta = op_dist, r_op = r_op_vals) |>
  mutate(a_op = attractiveness_dual_sigmoid(0, delta, 1, 1, r_op, 1) / 0.5,
         r_op_label = factor(paste0("r_op = ", r_op), levels = paste0("r_op = ", r_op_vals)))

ggplot(df_op_old, aes(delta, a_op, colour = r_op_label)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = 0.5, linetype = "dotted", colour = "grey50") +
  annotate("text", x = 0.51, y = 0.95, label = "fixed midpoint\nDelta = 0.5", hjust = 0, size = 3) +
  scale_colour_viridis_d(end = 0.85) +
  labs(title = "Legacy opinion factor a_op(Delta) -- sigmoid of similarity",
       subtitle = "a_op = sig(r_op * (1 - 2*Delta)); always crosses exactly 0.5 at Delta = 0.5",
       x = expression(Delta == "|"*o[i]-o[j]*"|"), y = expression(a[op]), colour = NULL) +
  theme(legend.position = "bottom")
```

The power factor $a^{\text{pw}}$ is the **identical function** already
plotted in Part I.2 -- the legacy and new models do not differ on the power
dimension at all.

## Self-weight (dual-sigmoid argmax) and the "confidence" patch

Given $j^\*=\arg\max_j A_{ij}$ (found using the same attractiveness weights
as Step 3), the legacy self-weight is a **second, independent** formula
that re-derives a weight from $j^\*$'s specific opinion and power gap --
it does *not* reuse $A_{ij^\*}$ directly:

$$w^{\text{raw}}_{\text{self}} = \sigma\big(r_{op}(2|o_i-o_{j^\*}|-1)\big) \times \sigma\!\left(r_{pw}\log\frac{p_i}{p_{j^\*}}\right)$$

This is a mirror image of the attractiveness formula (opinion sign
flipped: far away $\Rightarrow$ high self-confidence). `self_weight_mode`
controls what happens next:

* `"raw"`: use $w^{\text{raw}}_{\text{self}}$ as-is.
* `"confidence"` (used in Reports 17-19): blend it toward "always
  self-vote" using a saturating function of total responsiveness
  $r_{\text{tot}} = r_{op}+r_{pw}+r_{ingroup}$ (or their mean, per
  `confidence_agg`):

$$\text{conf} = \frac{r_{\text{tot}}}{1+r_{\text{tot}}}, \qquad
  w_{\text{self}} = (1-\text{conf}) + \text{conf}\cdot w^{\text{raw}}_{\text{self}}$$

This blending step exists to fix a specific, documented artefact: at
$r_{op}=r_{pw}=0$ (no responsiveness at all), $w^{\text{raw}}_{\text{self}}
= \sigma(0)\cdot\sigma(0) = 0.25$ exactly -- a **floor** that keeps roughly a
quarter of agents self-voting even when nobody cares about opinion or power
at all, which is not a meaningful behavioural claim, just an artefact of
multiplying two sigmoids evaluated at zero. `"confidence"` mode papers over
this by interpolating toward $1$ (self-vote) as responsiveness $\to 0$.

```{r old-selfweight-floor-table}
tibble(r = c(0, 0.5, 1, 2, 4, 8)) |>
  mutate(
    `w_self_raw (identical j*)` = .sig(r * (2 * 0 - 1)) * .sig(r * log(1)),
    `confidence blend (r_ingroup=0)` = {
      conf <- (2 * r) / (1 + 2 * r)  # r_tot = r_op + r_pw = 2r here
      (1 - conf) + conf * `w_self_raw (identical j*)`
    }
  ) |>
  kable(digits = 4)
```

```{r old-confidence-blend-plot, fig.height=4.5}
r_seq <- seq(0, 8, length.out = 200)
df_conf <- tibble(r = r_seq) |>
  mutate(
    w_raw = .sig(r * (2 * 0 - 1)) * .sig(r * log(1)),
    conf  = (2 * r) / (1 + 2 * r),
    w_confidence = (1 - conf) + conf * w_raw
  ) |>
  pivot_longer(c(w_raw, w_confidence), names_to = "mode", values_to = "w_self") |>
  mutate(mode = recode(mode, w_raw = "raw", w_confidence = "confidence"))

ggplot(df_conf, aes(r, w_self, colour = mode)) +
  geom_line(linewidth = 1) +
  geom_hline(yintercept = 0.25, linetype = "dotted", colour = "grey50") +
  annotate("text", x = 0.1, y = 0.27, label = "0.25 floor (raw mode)", size = 3, hjust = 0) +
  scale_colour_manual(values = c(raw = "#d95f02", confidence = "#1b9e77")) +
  labs(title = "Legacy self-weight at an identical best neighbour, vs. responsiveness r = r_op = r_pw",
       subtitle = "'raw' mode floors at 0.25 as r -> 0; 'confidence' mode fixes this by interpolating toward self-vote",
       x = "r (= r_op = r_pw)", y = expression(w[self]), colour = "self_weight_mode") +
  theme(legend.position = "bottom")
```

## Delegate choice

Identical mechanism to Part I.4 -- weighted sampling proportional to
$A_{ij}$, unaffected by `decision_model`.

## Parameter glossary (legacy model)

```{r param-glossary-old}
tibble(
  Parameter = c("r_op", "r_pw", "self_weight_mode", "confidence_agg", "r_ingroup", "gamma / lambda / trust_mode"),
  Controls = c(
    "Steepness of the opinion-similarity sigmoid AND (via the mirrored formula) part of self-confidence.",
    "Steepness of the power-ratio sigmoid AND (via the mirrored formula) part of self-confidence.",
    "'raw' (use w_self_raw as-is) or 'confidence' (blend toward always-self-vote at low responsiveness).",
    "'sum' or 'mean' -- how r_op/r_pw/r_ingroup combine into r_tot for the confidence blend.",
    "Ingroup responsiveness -- multiplies attractiveness by sig(+-r_ingroup), also feeds r_tot.",
    "Same trust mechanism as the new model, unaffected by decision_model."
  )
) |>
  kable()
```

---

# Part III -- Detailed Comparison

## III.1 What changed, in one table

```{r formula-comparison-table}
tibble(
  Component = c("Opinion factor", "Power factor", "Self-vote step",
                "Self-vote inputs", "Delegate choice", "Trust interaction"),
  Legacy = c(
    "$\\sigma(r_{op}(1-2\\Delta))$ -- sigmoid, fixed midpoint at $\\Delta=0.5$",
    "$\\sigma(r_{pw}\\log(p_j/p_i))$",
    "Re-derived formula, mirror of attractiveness, evaluated at $j^*$'s raw $(o_{j^*}, p_{j^*})$",
    "2 numbers: $\\Delta(i,j^*)$ and power ratio $(i,j^*)$, both fed through $r_{op}, r_{pw}$ again",
    "Sample $\\propto A_{ij}$ among neighbours",
    "trust_mod multiplies $A_{ij}$ (affects $j^*$ selection and sampling only)"
  ),
  `Gaussian/Mobius` = c(
    "$\\exp(-\\tfrac12 r_{op}^2\\Delta^2)$ -- Gaussian, $r_{op}$-dependent half-width",
    "$\\sigma(r_{pw}\\log(p_j/p_i))$ -- unchanged",
    "$P_{\\text{self}}(m)$, Mobius curve, single free parameter $c$",
    "1 number: $m = A_{ij^*}$ (already includes trust/ingroup)",
    "Sample $\\propto A_{ij}$ among neighbours -- unchanged",
    "trust_mod multiplies $A_{ij}$, feeding directly into $m$ (see III.6)"
  )
) |>
  kable(escape = FALSE)
```

The power factor and the delegate-choice mechanism are **literally the same
code** in both models. Everything else changed. The next five subsections
unpack why each change matters.

## III.2 Ceiling non-equivalence: what does "identical neighbour" mean?

An *identical* neighbour ($\Delta=0$, equal power) is the natural reference
point for "as attractive as it gets, ideologically." The two models treat
it very differently:

```{r ceiling-comparison-plot, fig.height=4.5}
r_seq2 <- seq(0.25, 16, length.out = 200)
df_ceiling <- tibble(r = r_seq2) |>
  mutate(
    old = attractiveness_dual_sigmoid(0, 0, 1, 1, r, r),
    new = attractiveness_gaussian_log(0, 0, 1, 1, r, r)
  ) |>
  pivot_longer(c(old, new), names_to = "model", values_to = "A_identical")

ggplot(df_ceiling, aes(r, A_identical, colour = model)) +
  geom_line(linewidth = 1.1) +
  geom_hline(yintercept = c(0.5, 1), linetype = "dotted", colour = "grey50") +
  scale_colour_manual(values = c(old = "#d95f02", new = "#1b9e77"),
                       labels = c(old = "legacy (dual sigmoid)", new = "gaussian_mobius")) +
  labs(title = "Attractiveness of an IDENTICAL neighbour (Delta=0, equal power), vs. r = r_op = r_pw",
       subtitle = "New model: always exactly 1. Legacy model: rises toward 0.5 only as r -> infinity",
       x = "r (= r_op = r_pw)", y = expression(A[ij]~"(identical neighbour)"), colour = NULL) +
  theme(legend.position = "bottom")
```

At $r=4$ (a typical "responsive" setting used throughout Reports 17-19),
an identical neighbour scores $A \approx$ `r round(attractiveness_dual_sigmoid(0,0,1,1,4,4), 3)`
under the legacy model but exactly $1$ under the new one. **The new model's
"1" is an invariant reference point, independent of $r_{op}, r_{pw}$; the
legacy model's ceiling is itself a function of responsiveness, and never
actually reaches even $0.5$ at finite $r$.** This is a direct, mechanical
consequence of multiplying two sigmoids that individually saturate below 1
rather than a Gaussian that reaches exactly 1 at zero distance.

## III.3 Decoupling: is self-reliance its own parameter, or entangled with $r$?

This is the most consequential conceptual change. Evaluate each model's
self-vote quantity specifically at the "identical best neighbour" case
($m=1$ for the new model; $\Delta=0$, ratio$=1$ for the legacy formula):

```{r selfvote-decoupling-plot, fig.height=4.5}
df_decouple_old <- tibble(r = r_seq2) |>
  mutate(w_self = .sig(r * (2 * 0 - 1)) * .sig(r * log(1)), model = "legacy")

df_decouple_new <- expand_grid(r = r_seq2, c = c(0.2, 0.5, 0.8)) |>
  mutate(w_self = p_self_mobius(1, c), model = paste0("gaussian_mobius, c=", c))

ggplot() +
  geom_line(data = df_decouple_old, aes(r, w_self, colour = model), linewidth = 1.1) +
  geom_line(data = df_decouple_new, aes(r, w_self, colour = model), linewidth = 1.1, linetype = "dashed") +
  scale_colour_manual(values = c("legacy" = "#d95f02",
                                 "gaussian_mobius, c=0.2" = "#1b9e77",
                                 "gaussian_mobius, c=0.5" = "#377eb8",
                                 "gaussian_mobius, c=0.8" = "#984ea3")) +
  labs(title = "Self-vote probability at an IDENTICAL best neighbour, vs. background responsiveness r",
       subtitle = "Legacy: entangled with r, collapses toward 0 as r grows. New: flat -- exactly c, for any r",
       x = "r (= r_op = r_pw, legacy only)", y = expression(P[self]~"or"~w[self]~"(identical best neighbour)"),
       colour = NULL) +
  theme(legend.position = "bottom")
```

Under the legacy model, self-reliance when facing an identical neighbour is
not a free choice -- it is *whatever falls out* of $r_{op}, r_{pw}$, which
also simultaneously control how sharply opinion/power gate attractiveness.
Turning up "how much do I care about opinion similarity" (raising $r_{op}$)
**automatically and inescapably** turns down "how likely am I to still vote
for myself when facing someone just like me" (from $0.25$ at $r=0$ down to
essentially $0$ by $r\approx8$, confirmed numerically above). There is no
way, in the legacy model, to have a highly opinion-selective agent who is
*also* strongly self-reliant, short of the `"confidence"` patch, which
addresses only the $r\to0$ floor, not this coupling.

The new model removes this coupling entirely: $c$ is set independently and
$P_{\text{self}}(m{=}1)=c$ for *every* $r_{op}$. An agent can be extremely
opinion-selective ($r_{op}=8$) and still self-reliant ($c=0.8$)
simultaneously -- a combination that is difficult to reach in the legacy
model and impossible to reach *exactly* by construction.

## III.4 Where does $r_{op}$ actually act? A hidden non-equivalence

Both models expose an argument literally called `r_op`, swept over the same
values in Report 20 as in Report 19 ($1, 4, 8$). **These are not
comparable in the same units.** Consider the opinion distance at which
$a^{\text{op}}=0.5$ (the "half-attractiveness point"):

* **Legacy:** $\sigma(r_{op}(1-2\Delta))=0.5 \iff \Delta = 0.5$, **for every
  value of $r_{op}$.** $r_{op}$ only controls how *sharply* the curve
  transitions around that fixed midpoint -- it never moves.
* **New:** $\exp(-\tfrac12 r_{op}^2\Delta^2)=0.5 \iff \Delta =
  \sqrt{2\ln2}/r_{op} \approx 1.1774/r_{op}$ -- the half-attractiveness
  point *moves inward* as $r_{op}$ grows.

```{r half-width-plot, fig.height=4.5}
r_seq3 <- seq(0.25, 16, length.out = 200)
df_half <- tibble(r = r_seq3) |>
  mutate(legacy = 0.5, new = sqrt(2 * log(2)) / r) |>
  pivot_longer(c(legacy, new), names_to = "model", values_to = "half_distance")

ggplot(df_half, aes(r, half_distance, colour = model)) +
  geom_line(linewidth = 1.1) +
  scale_colour_manual(values = c(legacy = "#d95f02", new = "#1b9e77")) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(title = "Half-attractiveness distance (where a_op = 0.5), vs. r_op",
       subtitle = "Legacy: r_op controls STEEPNESS at a FIXED reference point (Delta=0.5).\nNew: r_op controls the LOCATION of the reference point (an inverse bandwidth).",
       x = "r_op", y = expression("opinion distance"~Delta~"where"~a[op]==0.5), colour = NULL) +
  theme(legend.position = "bottom")
```

**Practical consequence:** "$r_{op}=8$" means "a sharp step exactly at
$\Delta=0.5$" in the legacy model, but "a narrow bump that has already
collapsed by $\Delta\approx0.15$" in the new model. Report 20 sweeps the
same numeric values as Report 19 for continuity, but any claim of the form
"the effect of $r_{op}$ is X in both models" should be read as
label-comparable, not magnitude-comparable, unless the sweep is
re-calibrated via the mapping above (see Part VII.3 for a concrete fix).

## III.5 Sharper echo chambers?

Because the new model's opinion gate is genuinely bounded away from the
legacy sigmoid's shallow tails (a Gaussian's tail decays like
$e^{-\Delta^2}$, faster than a sigmoid's like $e^{-\Delta}$ for large
arguments), a highly polarised, highly opinion-responsive population is
expected to show attractiveness collapsing *faster* across an opinion gap
under `gaussian_mobius` than under `legacy`, at matched nominal $r_{op}$.
Combined with III.4 (the new model's bandwidth also *shrinks* with
$r_{op}$, compounding the effect), this predicts: **more baseline
opinion-driven self-segregation under the new model, before ingroup
preference ($r_{ingroup}$) is even switched on.** This is directly relevant
to Report 20's hypotheses -- see Part V.

## III.6 A genuine edge case: trust can push $m$ past its designed ceiling

Recall from Part I.5 that $\texttt{trust\_mod} = 2\sigma(\gamma\tau)$ can
itself range up to (not including) $2$. In the **legacy** model, trust only
ever affects (a) *which* neighbour becomes $j^\*$ (the argmax is computed
over trust-adjusted weights) and (b) the delegate-choice sampling weights.
The self-weight formula in Step 2 is *recomputed from scratch* from
$j^\*$'s raw opinion/power gap -- it never sees the trust-adjusted number
itself, so it is mathematically guaranteed to stay in $[0,1]$ (a product of
two sigmoids) regardless of how extreme trust becomes.

The **new** model's Step 2 uses $m = w_{j^\*}$ directly -- the
*trust-adjusted* attractiveness. If trust pushes a neighbour's weight above
the formula's designed ceiling of $2$ (entirely possible: $A_{ij}\le2$
times $\texttt{trust\_mod}<2$ can exceed $2$), the Mobius formula
extrapolates outside its intended domain:

```{r edge-case-mobius-plot, fig.height=4.5}
m_ext <- seq(0, 3, length.out = 300)
df_edge <- tibble(m = m_ext) |> mutate(P_self = p_self_mobius(m, 0.5))

ggplot(df_edge, aes(m, P_self)) +
  geom_rect(aes(xmin = 2, xmax = 3, ymin = -Inf, ymax = Inf), fill = "#fdecea", alpha = 0.6, inherit.aes = FALSE) +
  geom_line(linewidth = 1.1, colour = "#377eb8") +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey40") +
  geom_vline(xintercept = 2, linetype = "dashed", colour = "firebrick3") +
  annotate("text", x = 2.05, y = 0.35, label = "beyond designed ceiling\n(trust-boosted m > 2)\nP_self < 0 -> treated\nas 'never self-vote'",
           hjust = 0, size = 3, colour = "firebrick4") +
  labs(title = "p_self_mobius(m, c=0.5) extrapolated beyond the nominal [0,2] domain",
       x = expression(m == w[j^"*"]~"(trust-adjusted)"), y = expression(P[self]))
```

```{r edge-case-table}
tibble(m = c(0, 1, 1.5, 2, 2.5, 3)) |>
  mutate(P_self = p_self_mobius(m, 0.5)) |>
  kable(digits = 4, caption = "P_self(m, c=0.5): note the sign flip past m=2")
```

Negative $P_{\text{self}}$ is not a crash -- `runif(1) < P_self` is simply
always `FALSE` for a negative threshold, so the practical effect is "always
delegate," which is arguably the *intended direction* (a
strongly-trust-boosted neighbour should be very delegate-worthy) but is
reached through an out-of-domain formula evaluation rather than by design.
This is flagged explicitly as a concrete, fixable weakness in Part VII.

---

# Part IV -- Advantages and Disadvantages

```{r adv-disadv-table}
tibble(
  Criterion = c("Realism", "Interpretability", "Flexibility", "Robustness",
                "Computational complexity", "Ease of parameter tuning"),
  `Legacy (dual sigmoid)` = c(
    "Sigmoid saturation is a common choice-model primitive (logistic response); opinion cutoff shape is arguably less 'natural' than a bell curve of similarity.",
    "Single family of parameters (r_op, r_pw, r_ingroup) drives everything, but the self-weight formula's meaning is only clear once you see it's a 'mirror' of attractiveness -- non-obvious on first read.",
    "self_weight_mode/confidence_agg add two more categorical switches on top of r_op/r_pw to reach a full delegation-rate range -- flexible, but via a layered patch rather than a single mechanism.",
    "Self-weight is always a valid probability in [0,1] by construction, regardless of how extreme trust or attractiveness become -- no known domain violations.",
    "One extra sigmoid evaluation per agent per round vs. the new model -- negligible difference either way.",
    "r_op/r_pw jointly govern opinion gating AND self-reliance -- tuning one for a target delegation rate can silently distort the other."
  ),
  `Gaussian/Mobius` = c(
    "Gaussian opinion kernel matches a natural 'similarity as closeness in a metric space' intuition; Mobius curve is a clean, purpose-built interpolant with named anchors rather than an emergent by-product of two sigmoids.",
    "Anchors (m=0->1, m=1->c, m=2->0) are directly readable off the formula; c has a literal meaning ('self-vote probability when facing an identical neighbour') independent of anything else.",
    "c is a single, dedicated self-reliance knob -- can be swept independently of r_op/r_pw without any patch or blending logic.",
    "m can exceed the formula's designed [0,2] ceiling under strong reward+punish trust, producing negative (out-of-domain) probabilities -- see III.6. Rare, but real and currently unguarded.",
    "Same order of cost (one exp(), one division) -- negligible difference either way.",
    "c alone controls self-reliance; r_op/r_pw only shape the opinion/power kernels -- but r_op's meaning is not comparable across models (III.4), complicating any joint calibration against legacy results."
  )
) |>
  kable(escape = FALSE)
```

**When legacy may still be preferable.** If the research question is
specifically about the interaction between *responsiveness* and
*self-reliance being the same underlying quantity* (e.g. modelling agents
whose confidence and pickiness are genuinely the same trait), the legacy
formula's entanglement is a feature, not a bug. It is also the
already-published, already-validated basis for Reports 15-19; switching
wholesale would break continuity with that body of results unless
`decision_model = "legacy"` is retained for replication (which it is).

---

# Part V -- Impact on the Simulation

## What we expect to change

* **Cross-group delegation baseline (H1c/H2c, Report 20).** Per III.5, the
  Gaussian kernel plus its shrinking bandwidth should produce *lower
  baseline cross-group attractiveness* under polarisation, independent of
  $r_{ingroup}$. If Report 20 finds a *smaller marginal effect* of
  $r_{ingroup}$ on cross-group delegation than Report 19 did, that is not
  necessarily evidence against H1c/H2c -- it may simply mean the opinion
  kernel is already doing part of that work, leaving less room for
  $r_{ingroup}$ to add on top. This is a genuine confound to check before
  interpreting effect sizes across the two reports.
* **Delegation rate range.** The `"confidence"` patch was built specifically
  to let legacy delegation rates span the full $[0,1]$ range; the Mobius
  curve does this natively via $c$. We do not expect a qualitative
  difference in *achievable* delegation rates, but the *shape* of how
  delegation rate responds to $r_{op}$ should differ, because $r_{op}$ now
  also reshapes the opinion kernel's bandwidth (III.4), not just its
  steepness at a fixed point.
* **Voting power concentration (Gini, top-5% share).** Indirectly affected
  through the delegation-rate and chain-length channels above -- no direct
  mechanical change, since `compute_power()` itself is untouched.
* **Trust dynamics.** The trust *update* rule is identical, but its
  *consequence* differs: under legacy, trust cannot destabilise the
  self-weight formula (III.6); under the new model, strong reward+punish
  trust interacts with the Mobius ceiling. Expect this to matter mainly in
  Report 18/20-style trust-heavy diagnostic sweeps (high $\gamma$, high
  $\lambda$), not in the baseline Report 20 grid where $\gamma=r_{background}
  \le 8$ and $\lambda=0.5$.
* **Stability / convergence.** No structural reason to expect a change:
  both models resolve the delegation decision independently each round from
  the current opinion/power/trust state; neither introduces new
  path-dependence.

## Small-scale sensitivity check (this document's own simulation)

To make the discussion above concrete rather than purely formulaic, this
section runs a small, self-contained sweep with the *actual* simulation
code (not a toy re-derivation) -- deliberately with **no group structure**
(uniform-random opinions, no minority split), so that only the
decision-model effect is visible, uncontaminated by Report 20's separate
ingroup-preference analysis.

```{r sim-sensitivity, cache=TRUE}
sens_seeds <- 1:3
r_vals     <- c(0.5, 1, 2, 4, 8)

run_sens <- function(seed, r, model) {
  res <- simulate_liquid_democracy(
    seed = seed, n_per_community = 120, n_communities = 1, node_degree = 6,
    T = 12, r_op = r, r_pw = r, p_rewire = 0.05,
    decision_model = model,
    self_reliance_c = 0.5,       # ignored when model == "legacy"
    self_weight_mode = "confidence",  # ignored when model == "gaussian_mobius"
    gamma = r, lambda = 0.5, trust_mode = "reward_punish",
    sigma_opinion = "auto"
  )
  snaps <- res$snapshots |> dplyr::filter(round >= 12 - 4)
  tibble(seed = seed, r = r, model = model,
         delegation_rate  = mean(snaps$delegation_rate, na.rm = TRUE),
         avg_drift        = mean(snaps$avg_drift, na.rm = TRUE),
         gini_power       = mean(snaps$gini_power, na.rm = TRUE),
         top5_power_share = mean(snaps$top5_power_share, na.rm = TRUE))
}

grid_sens <- expand_grid(seed = sens_seeds, r = r_vals, model = c("legacy", "gaussian_mobius"))
sens_raw  <- pmap_dfr(grid_sens, run_sens)

sens_agg <- sens_raw |>
  group_by(r, model) |>
  summarise(across(c(delegation_rate, avg_drift, gini_power, top5_power_share), mean), .groups = "drop")
```

```{r sim-sensitivity-plot, fig.height=6}
sens_long <- sens_agg |>
  pivot_longer(c(delegation_rate, avg_drift, gini_power, top5_power_share),
               names_to = "metric", values_to = "value") |>
  mutate(metric = recode(metric,
                         delegation_rate = "Delegation rate", avg_drift = "Avg. ideological drift",
                         gini_power = "Gini (power)", top5_power_share = "Top-5% power share"))

ggplot(sens_long, aes(r, value, colour = model)) +
  geom_line(linewidth = 1) + geom_point(size = 2) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_colour_manual(values = c(legacy = "#d95f02", gaussian_mobius = "#1b9e77")) +
  scale_x_log10(breaks = r_vals) +
  labs(title = "Decision-model sensitivity check (no group structure; N=120, T=12, 3 seeds)",
       subtitle = "Same r_op = r_pw = r swept for both models -- see III.4 caveat on comparability",
       x = "r (= r_op = r_pw = gamma)", y = NULL, colour = "decision_model") +
  theme(legend.position = "bottom")
```

## The new model's independent self-reliance axis

This is the sensitivity surface that **cannot be produced by the legacy
model at all** without conflating it with $r_{op}/r_{pw}$ (III.3) -- a
direct visualisation of what the new parameter buys:

```{r new-model-heatmap, cache=TRUE, fig.height=5}
c_vals_heat <- c(0.1, 0.3, 0.5, 0.7, 0.9)

run_heat <- function(seed, r, c_self) {
  res <- simulate_liquid_democracy(
    seed = seed, n_per_community = 120, n_communities = 1, node_degree = 6,
    T = 12, r_op = r, r_pw = r, p_rewire = 0.05,
    decision_model = "gaussian_mobius", self_reliance_c = c_self,
    gamma = r, lambda = 0.5, trust_mode = "reward_punish",
    sigma_opinion = "auto"
  )
  snaps <- res$snapshots |> dplyr::filter(round >= 12 - 4)
  tibble(seed = seed, r = r, c_self = c_self,
         delegation_rate = mean(snaps$delegation_rate, na.rm = TRUE))
}

grid_heat <- expand_grid(seed = sens_seeds, r = r_vals, c_self = c_vals_heat)
heat_raw  <- pmap_dfr(grid_heat, run_heat)

heat_agg <- heat_raw |> group_by(r, c_self) |> summarise(delegation_rate = mean(delegation_rate), .groups = "drop")

ggplot(heat_agg, aes(factor(r), factor(c_self), fill = delegation_rate)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = sprintf("%.2f", delegation_rate)), size = 3.4,
            colour = ifelse(heat_agg$delegation_rate > 0.5, "white", "black")) +
  scale_fill_viridis_c(limits = c(0, 1), option = "viridis") +
  labs(title = "gaussian_mobius: delegation rate over (r_op = r_pw, self_reliance_c)",
       subtitle = "r_op/r_pw shape the opinion/power kernels; c independently sets self-reliance -- two orthogonal axes",
       x = "r (= r_op = r_pw = gamma)", y = "self_reliance_c", fill = "delegation\nrate")
```

Reading across a row (fixed $r$, varying $c$) isolates the self-reliance
axis; reading down a column (fixed $c$, varying $r$) isolates the
opinion/power responsiveness axis. There is no equivalent clean
decomposition available for the legacy model.

---

# Part VI -- Consolidated Visual Comparison

## Side-by-side decision flow

```{r flow-diagram-both, fig.height=6.5, echo=FALSE}
p_legacy <- draw_pipeline(
  labels = c("Attractiveness\nA_ij = sig(op)*sig(pw)",
             "Find j* = argmax A_ij\n(trust/ingroup applied)",
             "Re-derive w_self\nfrom (o_j*, p_j*)\n['raw' or 'confidence']",
             "Bernoulli(w_self):\nself-vote, else\nsample ~ A_ij"),
  fills = c("#f6d6ac", "#f6d6ac", "#fbb4ae", "#d9ead3"),
  title = "legacy pipeline"
)

p_new <- draw_pipeline(
  labels = c("Attractiveness\nA_ij = 2*gauss(op)*sig(pw)",
             "Find j* = argmax A_ij\n(trust/ingroup applied)",
             "m = A_ij* directly\nP_self = mobius(m, c)",
             "Bernoulli(P_self):\nself-vote, else\nsample ~ A_ij"),
  fills = c("#cfe8f3", "#cfe8f3", "#b3d9c9", "#d9ead3"),
  title = "gaussian_mobius pipeline"
)

p_legacy / p_new
```

The structural difference is entirely in the third box: **legacy
re-derives** a fresh quantity from $j^\*$'s identity; **the new model
reuses** the number already computed in box 1/2. This single design choice
is the source of both the model's main advantage (decoupled, clean
self-reliance parameter, III.3) and its main new risk (out-of-domain $m$,
III.6).

## Combined kernel comparison (opinion + power, both models)

```{r capstone-patchwork, fig.height=8}
p_op_overlay <- ggplot() +
  geom_line(data = df_op_old |> filter(r_op %in% c(1, 4)),
            aes(delta, a_op, colour = r_op_label, linetype = "legacy"), linewidth = 1) +
  geom_line(data = df_op_new |> filter(r_op %in% c(1, 4)),
            aes(delta, a_op, colour = r_op_label, linetype = "gaussian_mobius"), linewidth = 1) +
  scale_colour_viridis_d(end = 0.7) +
  scale_linetype_manual(values = c(legacy = "dashed", gaussian_mobius = "solid")) +
  labs(title = "Opinion kernel: legacy (dashed) vs. new (solid), r_op in {1, 4}",
       x = expression(Delta), y = expression(a[op]), colour = "r_op", linetype = "model") +
  theme(legend.position = "bottom")

p_pself_vs_old <- ggplot() +
  geom_line(data = df_conf, aes(r, w_self, linetype = mode), colour = "#d95f02", linewidth = 1) +
  geom_hline(data = tibble(c = c(0.2, 0.5, 0.8)), aes(yintercept = c),
             colour = "#1b9e77", linetype = "dotted") +
  annotate("text", x = 6, y = c(0.2, 0.5, 0.8) + 0.03,
           label = paste0("new model, c=", c(0.2, 0.5, 0.8)), colour = "#1b9e77", size = 3) +
  labs(title = "Self-vote at identical best neighbour: legacy (orange, varies with r) vs. new (green, flat at c)",
       x = "r (= r_op = r_pw)", y = expression(w[self]~"or"~P[self]), linetype = "legacy mode") +
  theme(legend.position = "bottom")

p_op_overlay / p_pself_vs_old
```

---

# Part VII -- Overall Assessment

## Summary comparison table

```{r final-summary-table}
tibble(
  Dimension = c("Opinion gating", "Power gating", "Self-vote mechanism",
                "Self-reliance tunability", "Ceiling behaviour", "Known edge cases",
                "Continuity with Reports 15-19", "Recommended for"),
  `Legacy (dual sigmoid)` = c(
    "Sigmoid, fixed midpoint at Delta=0.5, steepness varies with r_op",
    "Sigmoid of log power ratio (shared with new model)",
    "Two-sigmoid mirror of attractiveness, evaluated at j*'s raw values",
    "Entangled with r_op/r_pw; patched via self_weight_mode='confidence'",
    "Approaches but never reaches 0.5 (identical neighbour) at finite r",
    "None known -- self-weight always in [0,1] regardless of trust",
    "Exact replication of all published results",
    "Continuity work; studies where responsiveness and self-reliance are meant to be the same trait"
  ),
  `Gaussian/Mobius` = c(
    "Gaussian, r_op-dependent bandwidth (half-width shrinks as 1/r_op)",
    "Sigmoid of log power ratio (shared with legacy model)",
    "Single Mobius curve of m = A_ij*, with named anchors",
    "Fully decoupled via dedicated parameter c",
    "Exactly 1 (identical neighbour), invariant to r_op/r_pw",
    "m can exceed 2 under strong reward+punish trust -> P_self < 0 (III.6)",
    "New for Report 20; r_op is not magnitude-comparable to legacy (III.4)",
    "New work specifically about self-reliance as an independent axis; cleaner default going forward"
  )
) |>
  kable(escape = FALSE)
```

## Recommendation

For a thesis chapter whose explicit interest is *how self-reliance shapes
delegation independently of opinion/power responsiveness* -- which is what
Report 20's design suggests -- **`gaussian_mobius` is the better-justified
default going forward**: it removes an ad hoc patch (`"confidence"` mode),
gives the self-reliance parameter a literal, checkable meaning ($c =
P_{\text{self}}$ at an identical neighbour, for *any* $r$), and its ceiling
behaviour is exact rather than asymptotic. `legacy` remains valuable and
should stay available (as it already is, via one argument) for
(a) exact replication of Reports 15-19, and (b) any analysis that
specifically wants responsiveness and self-confidence to be the same
underlying quantity.

## Weaknesses of the new implementation, and concrete fixes

1. **Out-of-domain $m$ under strong trust (III.6).** *Fix:* clamp before
   evaluating the curve, preserving the formula's intent while removing
   the sign flip:

   ```r
   p_self_mobius(pmin(w[j_idx], 2), self_reliance_c)
   ```

   A one-line change in `Network.R`'s `gaussian_mobius` branch. As a
   diagnostic, it would be worth logging how often the clamp actually
   binds across Report 20's grid (likely rare at $\gamma\le8$,
   $\lambda=0.5$, but worth quantifying rather than assuming).

2. **$r_{op}$ not comparable across models (III.4).** *Fix:* either (a)
   explicitly caveat any cross-model $r_{op}$ comparison in the thesis
   text (the honest, zero-code option), or (b) recalibrate the
   `gaussian_mobius` sweep so its half-attractiveness distance matches the
   legacy sweep's fixed $\Delta=0.5$ reference by construction --
   i.e. choose $r_{op}$ values from $r_{op} = \sqrt{2\ln2}/\Delta^\star$
   for a chosen target $\Delta^\star$, rather than reusing Report 19's raw
   numbers verbatim.

3. **No sensitivity instrumentation yet for $c$ specifically.** Report 20's
   grid does not currently sweep `self_reliance_c` at all (it is fixed at
   $0.5$ throughout, per its setup chunk). Given III.3 identifies this as
   the headline new capability, a supplementary sweep over $c$ (holding
   $r_{ingroup}$ and $r_{background}$ fixed) would make direct use of the
   model's main advantage rather than leaving it implicit.
