---
title: "Weekly Report — Week 10 (24.04.2026 – 30.04.2026)"
output: github_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
```

<!--
  WORKFLOW NOTE
  ─────────────────────────────────────────────────────────────────
   Utility functions studied in isolation.
   No softmax, no self-weight, no simulation noise.
   Goal: understand the structural behaviour of the formulas
   before any decision mechanism is introduced.

  ─────────────────────────────────────────────────────────────────
-->

---

# Phase I — Utility Analysis (No Decision Rule)

---

## 1. Utility Functions: Model A and Model B


$$\text{opinion similarity:} \quad (1 - |o_i - o_j|) \in [0, 1]$$

$$\text{log power ratio:} \quad \log\!\left(\frac{p_j}{p_i}\right) \in \mathbb{R}$$

**Neighbour-Weight:**
$$A_{ij} = \sigma\!\left(r_{op} \cdot (1 - 2|o_i - o_j|)\right) \cdot \sigma\!\left(r_{pw} \cdot \log\frac{p_j}{p_i}\right)$$

**Self-Weight:**
$$w_{\text{self}}(i) = \sigma\!\left(r_{op} \cdot (2|op_i - op_{j^*}| - 1)\right) \cdot \sigma\!\left(r_{pw} \cdot \log\frac{pow_i}{pow_{j^*}}\right)$$


---

## 2. 1D Analysis — Effect of Opinion Similarity

We fix $\log(p_j/p_i)$ and study how utility varies with $(1-|o_i-o_j|)$.
Each panel holds $\alpha$, $\beta$, and $\log(p_j/p_i)$ constant; only
opinion similarity varies along the x-axis.

```{r plot-1d-similarity, fig.width=14, fig.height=9, dpi=300}
expand_grid(
  opinion_sim     = seq(0, 1, length.out = 200),
  log_power_ratio = c(-2, 0, 2),
  model           = c("A", "B"),
  alpha           = c(1, 2),
  beta            = c(1, 2)
) |>
  mutate(
    V = case_when(
      model == "A" ~ alpha * opinion_sim + beta * log_power_ratio,
      model == "B" ~ opinion_sim * (alpha + beta * log_power_ratio)
    ),
    lpr_label = paste0("log(p[j]/p[i]) == ", log_power_ratio),
    model     = paste("Model", model),
    param     = paste0("α = ", alpha, ", β = ", beta)
  ) |>
  ggplot(aes(opinion_sim, V, colour = lpr_label)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey50") +
  facet_grid(model ~ param) +
  scale_colour_brewer(
    palette = "Dark2",
    name    = NULL,
    labels  = scales::label_parse()
  ) +
  labs(
    x        = expression(1 - abs(o[i] - o[j])),
    y        = expression(V[ij]),
    title    = "Utility vs opinion similarity across parameter settings",
    subtitle = "Rows: Model A vs B | Columns: (α, β) combinations"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

**Model A** produces parallel lines: increasing opinion similarity raises
$V_{ij}$ at constant rate $\alpha$, while $\beta\log(p_j/p_i)$ adds a fixed
vertical offset independent of similarity. Even at zero similarity
($1-|o_i-o_j|=0$) the utility is nonzero whenever $\beta > 0$ — a
powerful but ideologically distant neighbour still attracts.
**Model B** forces all lines through the origin at zero similarity: the
entire expression is multiplied by $(1-|o_i-o_j|)$, so maximal dissimilarity
always yields $V_{ij}=0$ regardless of power or $\alpha$.
Larger $\beta$ (right columns) widens the gap between $\log(p_j/p_i)$ levels
in both models; larger $\alpha$ (bottom rows) shifts Model A upward uniformly
and steepens Model B's slope.

---

## 3. 1D Analysis — Effect of the Power Ratio

We fix $(1-|o_i-o_j|)$ and study how utility varies with $\log(p_j/p_i)$.
Each panel holds $\alpha$, $\beta$, and opinion similarity constant; only
the log power ratio varies along the x-axis.

```{r plot-1d-power, fig.width=14, fig.height=9, dpi=300}
expand_grid(
  log_power_ratio = seq(-3, 3, length.out = 200),
  opinion_sim     = c(0, 0.25, 0.5, 0.75, 1),
  model           = c("A", "B"),
  alpha           = c(1, 2),
  beta            = c(1, 2)
) |>
  mutate(
    V = case_when(
      model == "A" ~ alpha * opinion_sim + beta * log_power_ratio,
      model == "B" ~ opinion_sim * (alpha + beta * log_power_ratio)
    ),
    sim_label = paste0("1-abs(o[i]-o[j]) == ", opinion_sim),
    model     = paste("Model", model),
    param     = paste0("α = ", alpha, ", β = ", beta)
  ) |>
  ggplot(aes(log_power_ratio, V, colour = sim_label)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey50") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  facet_grid(model ~ param) +
  scale_colour_brewer(
    palette = "RdYlGn",
    direction = -1,
    name   = "Opinion similarity",
    labels = scales::label_parse()
  ) +
  labs(
    x        = expression(log(p[j] / p[i])),
    y        = expression(V[ij]),
    title    = "Utility vs log power ratio across parameter settings",
    subtitle = "Rows: Model A vs B | Columns: (α, β) combinations"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

**Model A** produces perfectly parallel lines with slope $\beta$: the log
power ratio has the same influence at every similarity level, and the lines
are shifted vertically by $\alpha \cdot (1-|o_i-o_j|)$. Agents with zero
similarity ($1-|o_i-o_j|=0$, green line) still respond to power.
**Model B** shows a fan converging to zero: at full similarity the slope
equals $\beta$ (matching Model A), but it contracts proportionally as
similarity falls, reaching a completely flat zero line for dissimilar agents.
The vertical dashed line marks equal power ($\log(p_j/p_i)=0$); the dotted
horizontal marks $V_{ij}=0$.

---

## 4. 2D Utility Surface

Both inputs vary simultaneously. Red indicates high utility (agent $i$ strongly
prefers to delegate to $j$), blue indicates low or negative utility, white marks
$V_{ij}=0$. Each panel is one $(\alpha,\beta)$ combination.

```{r plot-2d-heatmap, fig.width=9, fig.height=4, echo=FALSE}
expand_grid(
  opinion_sim     = seq(0, 1, length.out = 150),
  log_power_ratio = seq(-3, 3, length.out = 150),
  model           = c("A", "B"),
  alpha           = c(1, 2),
  beta            = c(1, 2)
) |>
  mutate(
    V = if_else(
      model == "A",
      alpha * opinion_sim + beta * log_power_ratio,
      opinion_sim * (alpha + beta * log_power_ratio)
    ),
    model = paste("Model", model),
    param = paste0("α = ", alpha, ", β = ", beta)
  ) |>
  ggplot(aes(opinion_sim, log_power_ratio, fill = V)) +
  geom_tile() +
  facet_grid(model ~ param) +
  scale_fill_gradient2(
    low = "steelblue",
    mid = "white",
    high = "firebrick",
    midpoint = 0,
    name = expression(V[ij])
  ) +
  labs(
    x        = expression(1 - abs(o[i] - o[j])),
    y        = expression(log(p[j] / p[i])),
    title    = "Utility surfaces across parameter settings",
    subtitle = "Rows: Model A vs B | Columns: (α, β) combinations"
  ) +
  theme_minimal(base_size = 12)
```

The left column ($1-|o_i-o_j|=0$) makes the key difference concrete:
Model A shows a continuous gradient driven by $\log(p_j/p_i)$;
Model B is uniformly white (zero), eliminating all dissimilar agents from
the choice set irrespective of their power.

---

## 5. Limit Behaviour

```{r limit-table, echo=FALSE}
tibble::tribble(
  ~Condition,
  ~V_A,
  ~V_B,
  ~Implication,

  "|o_i - o_j| = 1 (max. dissimilar)",
  "beta * log(p_j / p_i)",
  "0",
  "B ignores dissimilar agents entirely",

  "o_i = o_j (identical opinions)",
  "alpha + beta * log(p_j / p_i)",
  "alpha + beta * log(p_j / p_i)",
  "Models are identical",

  "beta = 0",
  "alpha * (1 - |o_i - o_j|)",
  "alpha * (1 - |o_i - o_j|)",
  "Models are identical; pure homophily",

  "alpha = 0",
  "beta * log(p_j / p_i)",
  "(1 - |o_i - o_j|) * beta * log(p_j / p_i)",
  "B still couples power to similarity",

  "p_j >> p_i, beta > 0, o_i ≈ o_j",
  "→ +∞",
  "→ +∞",
  "Both: powerful similar agents dominate",

  "p_j >> p_i, beta > 0, |o_i - o_j| = 1",
  "→ +∞",
  "0",
  "A: cross-ideological pull; B: no effect"
) |>
  knitr::kable(
    col.names = c("Condition", "V_A", "V_B", "Implication"),
    caption   = "Limit behaviour of both utility functions."
  )
```

Models diverge only when $\beta \neq 0$ **and** opinion heterogeneity is present.
The two bold rows define the critical qualitative difference.

