---
title: "Weekly Report -- Week 17: Testing the Confidence-Weighted Self-Weight"
output: github_document
date: "2026-07-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 6)
library(tidyverse)
library(here)
library(furrr)
library(patchwork)
library(digest)
library(knitr)

theme_set(theme_minimal(base_size = 12))
source(here::here("scripts/Network.R"))

# ── Simulation constants (same as Report 15) ────────────────────────────────
MINORITY_SHARE   <- 0.2
OP_SIGMA         <- 0.1
SIM_T            <- 20L
N_SEEDS          <- 50L
N_SIM            <- 500L
K_SIM            <- 14L
P_REWIRE         <- 0.05

DELTA_VALS        <- c(0.2, 0.4)
R_INGROUP_VALS    <- c(0, 1, 2, 4, 8)
SELF_WEIGHT_MODES <- c("raw", "confidence")

# ── Diagnostic-sweep constants ──────────────────────────────────────────────
R_DIAG_VALS   <- c(0, 0.5, 1, 2, 4, 8, 16, 32, 64, 128)
N_SEEDS_DIAG  <- 50L

# The full 0-128 sweep is dense near 0 (0, 0.5, 1) for resolution elsewhere,
# but that crowds the axis on delegation_rate/lost_vote_rate plots, which
# span the whole range -- those use this sparser subset of points instead.
R_SPARSE_VALS <- c(0, 2, 4, 8, 16, 32, 64, 128)

# ── Metrics actually implemented in Network.R's snapshot tibble ─────────────
# NOTE: minority_capture and minority_vote_share (MVS) are documented in
# Report 15 but not yet computed in Network.R, so they cannot be plotted here.
minority_metrics <- c("RR_m", "minority_self_rep_rate")

minority_lbl <- c(
  RR_m                   = "Relative Representation (RR_m)",
  minority_self_rep_rate = "Minority Self-Representation Rate (SR_m)"
)

# Metrics tracked in Section 2 (Testing the Self-Weight Fix)
rq1_metrics <- c("delegation_rate", "lost_vote_rate", "avg_chain_length_delegators",
                  "avg_drift", "top5_power_share", "share_50pct", "total_components")

rq1_lbl <- c(
  delegation_rate             = "Delegation Rate",
  lost_vote_rate              = "Lost Vote Rate",
  avg_chain_length_delegators = "Avg. Chain Length",
  avg_drift                   = "Avg. Ideological Drift",
  top5_power_share            = "Top-5% Power Share",
  share_50pct                 = "Agents for 50% Votes",
  total_components            = "Components"
)

# delegation_rate/lost_vote_rate always mean something; the rest are computed
# from represented (non-lost) agents only and become unreliable once most
# votes are lost -- masked to NA beyond LOST_THRESHOLD, mirroring Report 16.
ALWAYS_VALID   <- c("delegation_rate", "lost_vote_rate")
MASK_METRICS   <- setdiff(rq1_metrics, ALWAYS_VALID)
LOST_THRESHOLD <- 0.90

# delegation_rate/lost_vote_rate are shown over the full R_DIAG_VALS range
# (up to 128); every other metric is only shown up to R_SHORT_MAX, since
# beyond that they are already masked out by LOST_THRESHOLD for most of the
# range anyway and just crowd the axis.
R_SHORT_MAX <- 16
```

---

# 1  Purpose

The endogenous self-weight used throughout the model (`self_weight_mode = "raw"`)
computes

$$w_{\text{self}}(i) = \sigma\left(r_{op}(2|o_i-o_{j^\ast}|-1)\right)\cdot\sigma\left(r_{pw}\log\tfrac{p_i}{p_{j^\ast}}\right)$$

where $j^*$ is agent $i$'s most attractive neighbour. At
$r_{op}=r_{pw}=0$ both factors evaluate to $\sigma(0)=0.5$, so
$w_{self}^{raw}=0.25$ regardless of how unresponsive agents are — the
global delegation rate floors at roughly 75% and can never approach the
Direct Democracy limit (delegation rate $\to 0$).

New function:

$$r_{tot} = r_{op} + r_{pw} + r_{ingroup}, \qquad
  \text{conf} = \frac{r_{tot}}{1+r_{tot}}, \qquad
  w_{self} = (1-\text{conf}) + \text{conf} \cdot w_{self}^{raw}.$$

At $r_{tot}=0$: $\text{conf}=0 \Rightarrow w_{self}=1$ (Direct Democracy).
As $r_{tot}\to\infty$: $\text{conf}\to1 \Rightarrow w_{self}\to w_{self}^{raw}$,
recovering the original model exactly. `"raw"` remains the default
everywhere else in the codebase (Report 16 and earlier reports are
unaffected by this addition).

---

# 2  Testing the Self-Weight Fix

Following Report 16's RQ1 structure: Condition A isolates $r_{op}$;
Condition B isolates $r_{pw}$; Condition D varies trust sensitivity
$\gamma$ at fixed background levels $r_{op}=r_{pw}\in\{1,2,4,8\}$;
Condition C shows the full $r_{op}\times r_{pw}$ grid. $r_{ingroup}=0$
throughout (no minority extension). Conditions A, B and D each overlay
both `"raw"` and `"confidence"` in the same plot; Condition C instead
places the two models side by side per metric, since a 2-D grid
overlaid by colour would be unreadable.

```{r rq1-constants, include=FALSE}
GAMMA_VALS   <- c(0, 1, 2, 4, 8, 16)
R_TRUST_DIAG <- c(1, 2, 4, 8)
LAMBDA_DIAG  <- 0.5
```

```{r diag-run-one, include=FALSE}
run_diag_one <- function(seed, r_op, r_pw, gamma, self_weight_mode, lambda_val = 0) {
  res <- simulate_liquid_democracy(
    seed              = seed,
    n_per_community   = N_SIM,
    n_communities     = 1,
    node_degree       = K_SIM,
    T                 = SIM_T,
    r_op              = r_op,
    r_pw              = r_pw,
    p_rewire          = P_REWIRE,
    gamma             = gamma,
    lambda            = lambda_val,
    self_weight_mode  = self_weight_mode
  )
  snaps <- res$snapshots |> dplyr::filter(round >= SIM_T - 4L)
  tibble(
    seed                        = seed,
    r_op                        = r_op,
    r_pw                        = r_pw,
    gamma                       = gamma,
    self_weight_mode            = self_weight_mode,
    delegation_rate             = mean(res$history_delegation[(SIM_T - 4L):SIM_T], na.rm = TRUE),
    lost_vote_rate               = mean(res$history_lost[(SIM_T - 4L):SIM_T],       na.rm = TRUE),
    avg_chain_length_delegators = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
    avg_drift                   = mean(snaps$avg_drift,                  na.rm = TRUE),
    top5_power_share            = mean(snaps$top5_power_share,           na.rm = TRUE),
    share_50pct                 = mean(snaps$share_50pct,                na.rm = TRUE),
    total_components            = mean(snaps$total_components,           na.rm = TRUE)
  )
}
```

```{r diag-simulate, include=FALSE}
seeds_diag <- seq_len(N_SEEDS_DIAG)

grid_A <- expand_grid(r_op = R_DIAG_VALS, r_pw = 0, gamma = 0,
                       self_weight_mode = SELF_WEIGHT_MODES, seed = seeds_diag) |>
  mutate(condition = "A: r_op only")
grid_B <- expand_grid(r_op = 0, r_pw = R_DIAG_VALS, gamma = 0,
                       self_weight_mode = SELF_WEIGHT_MODES, seed = seeds_diag) |>
  mutate(condition = "B: r_pw only")
grid_C <- expand_grid(r_op = R_DIAG_VALS, r_pw = R_DIAG_VALS, gamma = 0,
                       self_weight_mode = SELF_WEIGHT_MODES, seed = seeds_diag) |>
  mutate(condition = "C: r_op x r_pw (grid)")
grid_D <- purrr::map_dfr(R_TRUST_DIAG, function(rv)
  expand_grid(r_op = rv, r_pw = rv, gamma = GAMMA_VALS,
              self_weight_mode = SELF_WEIGHT_MODES, seed = seeds_diag) |>
    mutate(condition = paste0("D: trust (r=", rv, ")"))
)

params_diag <- list(N = N_SIM, K = K_SIM, T = SIM_T, r_vals = R_DIAG_VALS,
                     gamma_vals = GAMMA_VALS, r_trust = R_TRUST_DIAG, lambda = LAMBDA_DIAG,
                     n_seeds = N_SEEDS_DIAG, modes = SELF_WEIGHT_MODES,
                     metrics = rq1_metrics, version = 3L)
hash_diag  <- digest::digest(params_diag)
cache_diag <- here::here("data/sim_report17_diag_cache.rds")

if (file.exists(cache_diag)) {
  cd <- readRDS(cache_diag)
  if (identical(cd$hash, hash_diag)) sim_diag <- cd$data else cd <- NULL
} else cd <- NULL

if (is.null(cd)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  plan(multisession, workers = parallel::detectCores() - 1)

  run_cond <- function(grid_df, lambda_val = 0) {
    future_pmap_dfr(
      list(seed = grid_df$seed, r_op = grid_df$r_op, r_pw = grid_df$r_pw,
           gamma = grid_df$gamma, self_weight_mode = grid_df$self_weight_mode,
           condition = grid_df$condition),
      function(seed, r_op, r_pw, gamma, self_weight_mode, condition)
        run_diag_one(seed, r_op, r_pw, gamma, self_weight_mode, lambda_val) |>
          mutate(condition = condition),
      .options = furrr_options(seed = NULL)
    )
  }

  sim_diag <- bind_rows(
    run_cond(grid_A),
    run_cond(grid_B),
    run_cond(grid_C),
    run_cond(grid_D, lambda_val = LAMBDA_DIAG)
  )
  plan(sequential)
  saveRDS(list(hash = hash_diag, data = sim_diag), cache_diag)
}

sim_diag_agg_raw <- sim_diag |>
  group_by(condition, r_op, r_pw, gamma, self_weight_mode) |>
  summarise(across(all_of(rq1_metrics), \(x) mean(x, na.rm = TRUE)), .groups = "drop")

# Mask concentration/chain-length/component metrics once too many votes are
# lost to be meaningful; delegation_rate and lost_vote_rate stay untouched.
sim_diag_agg <- sim_diag_agg_raw |>
  mutate(across(all_of(MASK_METRICS), \(x) ifelse(lost_vote_rate > LOST_THRESHOLD, NA_real_, x))) |>
  pivot_longer(all_of(rq1_metrics), names_to = "metric", values_to = "value") |>
  mutate(metric_label = factor(rq1_lbl[metric], levels = unname(rq1_lbl)))

# Unmasked counterpart -- identical to sim_diag_agg but without the
# LOST_THRESHOLD cutoff, used to draw a dotted continuation past the point
# where a metric gets masked out for being unreliable.
sim_diag_unmasked <- sim_diag_agg_raw |>
  pivot_longer(all_of(rq1_metrics), names_to = "metric", values_to = "value") |>
  mutate(metric_label = factor(rq1_lbl[metric], levels = unname(rq1_lbl)))
```

## 2.1  Condition A: $r_{op}$ only

Delegation rate and lost-vote rate, shown over the full $r_{op}$ range
(0-128, thinned to `r paste(R_SPARSE_VALS, collapse = ", ")`), with a
zoom to $r_{op} \in [0, 16]$ directly underneath. The remaining metrics
follow further below, capped at $r_{op} \le$ `r R_SHORT_MAX`.

```{r cond-a-plot, fig.width=10, fig.height=5}
df_a_primary <- sim_diag_agg |>
  filter(condition == "A: r_op only", metric %in% ALWAYS_VALID, r_op %in% R_SPARSE_VALS)

ggplot(df_a_primary, aes(r_op, value, colour = self_weight_mode, fill = self_weight_mode)) +
  geom_line(linewidth = 0.9, na.rm = TRUE) +
  geom_point(shape = 21, size = 2.3, stroke = 0.5, na.rm = TRUE) +
  geom_hline(yintercept = c(0, 1), linetype = "dotted", colour = "grey60") +
  facet_wrap(~ metric_label, scales = "free", ncol = 2) +
  scale_colour_manual(values = c(raw = "firebrick3", confidence = "steelblue4")) +
  scale_fill_manual(values   = c(raw = "firebrick1", confidence = "steelblue2")) +
  scale_x_continuous(breaks = R_SPARSE_VALS, labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "A: r_op only -- delegation_rate & lost_vote_rate (full range)",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG),
       x = "r_op", y = NULL, colour = "self_weight_mode", fill = "self_weight_mode") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

```{r cond-a-plot-zoom, fig.width=10, fig.height=5}
df_a_zoom <- sim_diag_agg |>
  filter(condition == "A: r_op only", metric %in% ALWAYS_VALID, r_op <= R_SHORT_MAX)

ggplot(df_a_zoom, aes(r_op, value, colour = self_weight_mode, fill = self_weight_mode)) +
  geom_line(linewidth = 0.9, na.rm = TRUE) +
  geom_point(shape = 21, size = 2.3, stroke = 0.5, na.rm = TRUE) +
  geom_hline(yintercept = c(0, 1), linetype = "dotted", colour = "grey60") +
  facet_wrap(~ metric_label, scales = "free", ncol = 2) +
  scale_colour_manual(values = c(raw = "firebrick3", confidence = "steelblue4")) +
  scale_fill_manual(values   = c(raw = "firebrick1", confidence = "steelblue2")) +
  scale_x_continuous(breaks = R_DIAG_VALS[R_DIAG_VALS <= R_SHORT_MAX], labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "A: r_op only -- zoom to r_op in [0, 16]",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG),
       x = "r_op", y = NULL, colour = "self_weight_mode", fill = "self_weight_mode") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

```{r cond-a-plot-other, fig.width=10, fig.height=12}
df_a_other_masked   <- sim_diag_agg      |> filter(condition == "A: r_op only", !metric %in% ALWAYS_VALID, r_op <= R_SHORT_MAX)
df_a_other_unmasked <- sim_diag_unmasked |> filter(condition == "A: r_op only", !metric %in% ALWAYS_VALID, r_op <= R_SHORT_MAX)

ggplot() +
  geom_line(data = df_a_other_unmasked, aes(r_op, value, colour = self_weight_mode, group = self_weight_mode),
            linetype = "dotted", linewidth = 1.1, alpha = 0.9, na.rm = TRUE) +
  geom_point(data = df_a_other_unmasked, aes(r_op, value, colour = self_weight_mode),
             size = 1.5, alpha = 0.9, na.rm = TRUE) +
  geom_line(data = df_a_other_masked, aes(r_op, value, colour = self_weight_mode, fill = self_weight_mode),
            linewidth = 0.9, na.rm = TRUE) +
  geom_point(data = df_a_other_masked, aes(r_op, value, colour = self_weight_mode, fill = self_weight_mode),
             shape = 21, size = 2.3, stroke = 0.5, na.rm = TRUE) +
  geom_hline(yintercept = c(0, 1), linetype = "dotted", colour = "grey60") +
  facet_wrap(~ metric_label, scales = "free", ncol = 2) +
  scale_colour_manual(values = c(raw = "firebrick3", confidence = "steelblue4")) +
  scale_fill_manual(values   = c(raw = "firebrick1", confidence = "steelblue2")) +
  scale_x_continuous(breaks = R_DIAG_VALS[R_DIAG_VALS <= R_SHORT_MAX], labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "A: r_op only -- other metrics",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds -- dotted = beyond %.0f%% lost-vote rate",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG, LOST_THRESHOLD * 100),
       x = "r_op", y = NULL, colour = "self_weight_mode", fill = "self_weight_mode") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

## 2.2  Condition B: $r_{pw}$ only

Delegation rate and lost-vote rate, shown over the full $r_{pw}$ range
(0-128, thinned to `r paste(R_SPARSE_VALS, collapse = ", ")`), with a
zoom to $r_{pw} \in [0, 16]$ directly underneath. The remaining metrics
follow further below, capped at $r_{pw} \le$ `r R_SHORT_MAX`.

```{r cond-b-plot, fig.width=10, fig.height=5}
df_b_primary <- sim_diag_agg |>
  filter(condition == "B: r_pw only", metric %in% ALWAYS_VALID, r_pw %in% R_SPARSE_VALS)

ggplot(df_b_primary, aes(r_pw, value, colour = self_weight_mode, fill = self_weight_mode)) +
  geom_line(linewidth = 0.9, na.rm = TRUE) +
  geom_point(shape = 21, size = 2.3, stroke = 0.5, na.rm = TRUE) +
  geom_hline(yintercept = c(0, 1), linetype = "dotted", colour = "grey60") +
  facet_wrap(~ metric_label, scales = "free", ncol = 2) +
  scale_colour_manual(values = c(raw = "firebrick3", confidence = "steelblue4")) +
  scale_fill_manual(values   = c(raw = "firebrick1", confidence = "steelblue2")) +
  scale_x_continuous(breaks = R_SPARSE_VALS, labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "B: r_pw only -- delegation_rate & lost_vote_rate (full range)",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG),
       x = "r_pw", y = NULL, colour = "self_weight_mode", fill = "self_weight_mode") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

```{r cond-b-plot-zoom, fig.width=10, fig.height=5}
df_b_zoom <- sim_diag_agg |>
  filter(condition == "B: r_pw only", metric %in% ALWAYS_VALID, r_pw <= R_SHORT_MAX)

ggplot(df_b_zoom, aes(r_pw, value, colour = self_weight_mode, fill = self_weight_mode)) +
  geom_line(linewidth = 0.9, na.rm = TRUE) +
  geom_point(shape = 21, size = 2.3, stroke = 0.5, na.rm = TRUE) +
  geom_hline(yintercept = c(0, 1), linetype = "dotted", colour = "grey60") +
  facet_wrap(~ metric_label, scales = "free", ncol = 2) +
  scale_colour_manual(values = c(raw = "firebrick3", confidence = "steelblue4")) +
  scale_fill_manual(values   = c(raw = "firebrick1", confidence = "steelblue2")) +
  scale_x_continuous(breaks = R_DIAG_VALS[R_DIAG_VALS <= R_SHORT_MAX], labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "B: r_pw only -- zoom to r_pw in [0, 16]",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG),
       x = "r_pw", y = NULL, colour = "self_weight_mode", fill = "self_weight_mode") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

```{r cond-b-plot-other, fig.width=10, fig.height=12}
df_b_other_masked   <- sim_diag_agg      |> filter(condition == "B: r_pw only", !metric %in% ALWAYS_VALID, r_pw <= R_SHORT_MAX)
df_b_other_unmasked <- sim_diag_unmasked |> filter(condition == "B: r_pw only", !metric %in% ALWAYS_VALID, r_pw <= R_SHORT_MAX)

ggplot() +
  geom_line(data = df_b_other_unmasked, aes(r_pw, value, colour = self_weight_mode, group = self_weight_mode),
            linetype = "dotted", linewidth = 1.1, alpha = 0.9, na.rm = TRUE) +
  geom_point(data = df_b_other_unmasked, aes(r_pw, value, colour = self_weight_mode),
             size = 1.5, alpha = 0.9, na.rm = TRUE) +
  geom_line(data = df_b_other_masked, aes(r_pw, value, colour = self_weight_mode, fill = self_weight_mode),
            linewidth = 0.9, na.rm = TRUE) +
  geom_point(data = df_b_other_masked, aes(r_pw, value, colour = self_weight_mode, fill = self_weight_mode),
             shape = 21, size = 2.3, stroke = 0.5, na.rm = TRUE) +
  geom_hline(yintercept = c(0, 1), linetype = "dotted", colour = "grey60") +
  facet_wrap(~ metric_label, scales = "free", ncol = 2) +
  scale_colour_manual(values = c(raw = "firebrick3", confidence = "steelblue4")) +
  scale_fill_manual(values   = c(raw = "firebrick1", confidence = "steelblue2")) +
  scale_x_continuous(breaks = R_DIAG_VALS[R_DIAG_VALS <= R_SHORT_MAX], labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "B: r_pw only -- other metrics",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds -- dotted = beyond %.0f%% lost-vote rate",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG, LOST_THRESHOLD * 100),
       x = "r_pw", y = NULL, colour = "self_weight_mode", fill = "self_weight_mode") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

## 2.3  Condition C: $r_{op} \times r_{pw}$ (grid) — models side by side

Unlike A, B and D, the two models are **not** overlaid here — each row is
one metric, each column is one `self_weight_mode`, so the same metric
appears twice next to each other under the two models. $r_{op}$ is on the
x-axis, $r_{pw}$ is shown as one coloured line per level (as in Report
16's grid line plot). Delegation_rate and lost_vote_rate are shown first,
over the full $r$ range (thinned to `r paste(R_SPARSE_VALS, collapse = ", ")`
on both axes), with a zoom to $r_{op}, r_{pw} \in [0, 16]$ directly
underneath; every other metric follows further below, capped at
$r_{op}, r_{pw} \le$ `r R_SHORT_MAX`.

```{r cond-c-plot, fig.width=11, fig.height=8}
# facet_grid only draws x-axis labels on its bottom row, even with free
# scales -- unusable here since rows have genuinely different ranges
# (0-128 for delegation/lost-vote vs. 0-16 for everything else). Using
# facet_wrap instead (which labels every panel), with metric and model
# interleaved so the two models still land next to each other per metric.
panel_levels <- as.vector(sapply(unname(rq1_lbl), function(m) paste(m, c("confidence", "raw"))))

mk_panel_primary <- \(df) df |>
  filter(condition == "C: r_op x r_pw (grid)", metric %in% ALWAYS_VALID,
         r_op %in% R_SPARSE_VALS, r_pw %in% R_SPARSE_VALS) |>
  mutate(r_pw_fac    = factor(r_pw),
         panel_label = factor(paste(metric_label, self_weight_mode), levels = panel_levels))

sim_diag_C_primary <- sim_diag_agg |> mk_panel_primary()

ggplot(sim_diag_C_primary, aes(r_op, value, colour = r_pw_fac, group = r_pw_fac)) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +
  geom_point(size = 1.8, na.rm = TRUE) +
  facet_wrap(~ panel_label, ncol = 2, scales = "free") +
  scale_colour_viridis_d(name = "r_pw", end = 0.88) +
  scale_x_continuous(breaks = R_SPARSE_VALS, labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "C: r_op x r_pw -- delegation_rate & lost_vote_rate (full range)",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG),
       x = "r_op", y = NULL) +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "right")
```

```{r cond-c-plot-zoom, fig.width=11, fig.height=9}
mk_panel_zoom <- \(df) df |>
  filter(condition == "C: r_op x r_pw (grid)", metric %in% ALWAYS_VALID,
         r_op <= R_SHORT_MAX, r_pw <= R_SHORT_MAX) |>
  mutate(r_pw_fac    = factor(r_pw),
         panel_label = factor(paste(metric_label, self_weight_mode), levels = panel_levels))

sim_diag_C_zoom <- sim_diag_agg |> mk_panel_zoom()

ggplot(sim_diag_C_zoom, aes(r_op, value, colour = r_pw_fac, group = r_pw_fac)) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +
  geom_point(size = 1.8, na.rm = TRUE) +
  facet_wrap(~ panel_label, ncol = 2, scales = "free") +
  scale_colour_viridis_d(name = "r_pw", end = 0.88) +
  scale_x_continuous(breaks = R_DIAG_VALS[R_DIAG_VALS <= R_SHORT_MAX], labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "C: r_op x r_pw -- zoom to [0, 16], raw vs. confidence, side by side",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG),
       x = "r_op", y = NULL) +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "right")
```

```{r cond-c-plot-other, fig.width=11, fig.height=19}
mk_panel_other <- \(df) df |>
  filter(condition == "C: r_op x r_pw (grid)", !metric %in% ALWAYS_VALID,
         r_op <= R_SHORT_MAX, r_pw <= R_SHORT_MAX) |>
  mutate(r_pw_fac    = factor(r_pw),
         panel_label = factor(paste(metric_label, self_weight_mode), levels = panel_levels))

sim_diag_C_other          <- sim_diag_agg      |> mk_panel_other()
sim_diag_C_other_unmasked <- sim_diag_unmasked |> mk_panel_other()

ggplot() +
  geom_line(data = sim_diag_C_other_unmasked, aes(r_op, value, colour = r_pw_fac, group = r_pw_fac),
            linetype = "dotted", linewidth = 1.0, alpha = 0.85, na.rm = TRUE) +
  geom_point(data = sim_diag_C_other_unmasked, aes(r_op, value, colour = r_pw_fac, group = r_pw_fac),
             size = 1.2, alpha = 0.85, na.rm = TRUE) +
  geom_line(data = sim_diag_C_other, aes(r_op, value, colour = r_pw_fac, group = r_pw_fac),
            linewidth = 0.8, na.rm = TRUE) +
  geom_point(data = sim_diag_C_other, aes(r_op, value, colour = r_pw_fac, group = r_pw_fac),
             size = 1.8, na.rm = TRUE) +
  facet_wrap(~ panel_label, ncol = 2, scales = "free") +
  scale_colour_viridis_d(name = "r_pw", end = 0.88) +
  scale_x_continuous(breaks = R_DIAG_VALS[R_DIAG_VALS <= R_SHORT_MAX], labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "C: r_op x r_pw -- other metrics",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds -- dotted = beyond %.0f%% lost-vote rate",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG, LOST_THRESHOLD * 100),
       x = "r_op", y = NULL) +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "right")
```

## 2.4  Condition D: Trust ($\gamma$)

```{r cond-d-plot, fig.width=13, fig.height=18}
mk_r_level <- \(df) df |> mutate(r_level = factor(paste0("r=", r_op), levels = paste0("r=", R_TRUST_DIAG)))

sim_diag_D          <- sim_diag_agg      |> filter(startsWith(condition, "D:")) |> mk_r_level()
sim_diag_D_unmasked <- sim_diag_unmasked |> filter(startsWith(condition, "D:")) |> mk_r_level()

ggplot() +
  geom_line(data = sim_diag_D_unmasked, aes(gamma, value, colour = self_weight_mode, group = self_weight_mode),
            linetype = "dotted", linewidth = 1.1, alpha = 0.9, na.rm = TRUE) +
  geom_point(data = sim_diag_D_unmasked, aes(gamma, value, colour = self_weight_mode),
             size = 1.5, alpha = 0.9, na.rm = TRUE) +
  geom_line(data = sim_diag_D, aes(gamma, value, colour = self_weight_mode, fill = self_weight_mode),
            linewidth = 0.9, na.rm = TRUE) +
  geom_point(data = sim_diag_D, aes(gamma, value, colour = self_weight_mode, fill = self_weight_mode),
             shape = 21, size = 2.3, stroke = 0.5, na.rm = TRUE) +
  facet_grid(metric_label ~ r_level, scales = "free_y") +
  scale_colour_manual(values = c(raw = "firebrick3", confidence = "steelblue4")) +
  scale_fill_manual(values   = c(raw = "firebrick1", confidence = "steelblue2")) +
  scale_x_continuous(breaks = GAMMA_VALS, labels = scales::label_number(drop0trailing = TRUE)) +
  labs(title = "D: trust  (gamma, r_op = r_pw = background level, lambda = 0.5)",
       subtitle = sprintf("N = %d | k = %d | T = %d | %d seeds -- dotted = beyond %.0f%% lost-vote rate",
                          N_SIM, K_SIM, SIM_T, N_SEEDS_DIAG, LOST_THRESHOLD * 100),
       x = "gamma", y = NULL, colour = "self_weight_mode", fill = "self_weight_mode") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

---
