---
title: "Weekly Report -- Week 13 (15.05.2026 -- 21.05.2026)"
output: github_document
date: "2026-05-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 9, fig.height = 5)
library(tidyverse)
library(here)
library(furrr)
library(patchwork)
library(lme4)
library(lmerTest)
library(mgcv)
library(broom)
library(broom.mixed)
library(MuMIn)
library(digest)

sigmoid <- function(x) 1 / (1 + exp(-x))
theme_set(theme_minimal(base_size = 12))

sig_stars <- function(p) dplyr::case_when(p < 0.01 ~ "**", p < 0.05 ~ "*", TRUE ~ "")
fmt_p     <- function(p) dplyr::case_when(is.na(p) ~ "", p < .001 ~ "< .001", TRUE ~ sprintf("%.3f", p))
fmt_coef  <- function(est, p) paste0(round(est, 4), sig_stars(p), " (", fmt_p(p), ")")

n_seeds_large <- 25L
n_seeds_rob   <- 100L

all_metrics <- c("delegation_rate", "lost_vote_rate", "avg_chain_len",
                 "avg_ideo_drift", "gini_coef", "top5_share", "enp",
                 "total_components")
all_metrics_lbl <- c(
  delegation_rate  = "Delegation Rate",
  lost_vote_rate   = "Lost Vote Rate",
  avg_chain_len    = "Avg. Chain Length",
  avg_ideo_drift   = "Avg. Ideological Drift",
  gini_coef        = "Gini Coefficient",
  top5_share       = "Top-5% Power Share",
  enp              = "ENP",
  total_components = "Total Components"
)
```

## Design

Two simulation sweeps underpin this report.

**Large sweep** ($r_{op}, r_{pw} \in \{0,\ldots,15\}$, N $\in$ \{300, 1000\}, node degree  = 6,
`r n_seeds_large` seeds per combination): used for GAM overview plots that capture the global,
potentially nonlinear response surface.

**Robustness sweep** ($r_{op}, r_{pw} \in \{0,1,2\}$,
N $\in$ \{100, 200, ..., 1000\}, node degree  = 6, `r n_seeds_rob` seeds per combination): used for
LMM inference (N = 300 subset) and N-scaling analysis across all ten network sizes.

All metrics are averaged over simulation rounds 15--20.

```{r large-sweep, include=FALSE}
source(here::here("scripts/Network.R"))

r_levels_large <- seq(0,15, by = 1) 
n_agents_large <- c(300L, 1000L)
sim_T          <- 20L
node_deg       <- 6L

sim_params_large <- list(r = r_levels_large, n = n_agents_large,
                         seeds = n_seeds_large, T = sim_T, deg = node_deg,
                         version = 4L)
hash_large       <- digest::digest(sim_params_large)
cache_large      <- here::here("data/sim_large_cache.rds")

if (file.exists(cache_large)) {
  cl <- readRDS(cache_large)
  if (identical(cl$hash, hash_large)) {
    sim_large_raw <- cl$data
    t_large       <- cl$elapsed
    message("Large-sweep cache hit.")
  } else { cl <- NULL }
} else { cl <- NULL }

if (is.null(cl)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  grid_large <- expand_grid(
    n_agents = n_agents_large,
    r_op     = r_levels_large,
    r_pw     = r_levels_large,
    seed     = seq_len(n_seeds_large)
  )
  plan(multisession, workers = parallel::detectCores() - 1)
  t0 <- proc.time()
  sim_large_raw <- future_pmap_dfr(
    grid_large,
    function(n_agents, r_op, r_pw, seed) {
      res   <- simulate_liquid_democracy(
        seed = seed, n_per_community = n_agents, n_communities = 1,
        node_degree = node_deg, T = sim_T, r_op = r_op, r_pw = r_pw)
      snaps <- tail(res$snapshots, 2)
      ag    <- res$agents
      tibble(
        n_agents        = n_agents, r_op = r_op, r_pw = r_pw, seed = seed,
        delegation_rate = mean(res$history_delegation[15:20], na.rm = TRUE),
        lost_vote_rate  = mean(res$history_lost[15:20],       na.rm = TRUE),
        avg_ideo_drift  = mean(res$history_drift[15:20],      na.rm = TRUE),
        gini_coef       = mean(snaps$gini_power,              na.rm = TRUE),
        avg_chain_len   = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
        top5_share      = mean(snaps$top5_power_share,        na.rm = TRUE),
        n_voters         = sum(!is.na(ag$my_vote)),
        enp              = { p <- ag$power[!is.na(ag$my_vote)]
                             if (!length(p) || sum(p) == 0) NA_real_
                             else 1 / sum((p / sum(p))^2) },
        total_components = mean(snaps$total_components, na.rm = TRUE)
      )
    },
    .options = furrr_options(seed = NULL)
  )
  plan(sequential)
  t_large <- proc.time() - t0
  saveRDS(list(hash = hash_large, data = sim_large_raw, elapsed = t_large), cache_large)
  message("Large sweep complete.")
}

df_large     <- sim_large_raw |> filter(!is.na(avg_chain_len))
df_large300  <- df_large      |> filter(n_agents == 300)
df_large1000 <- df_large      |> filter(n_agents == 1000)
```

```{r robust-sweep, include=FALSE}
r_levels_rob <- 0:2
n_agents_rob <- seq(100L, 1000L, by = 100L)

sim_params_rob <- list(r = r_levels_rob, n = n_agents_rob,
                       seeds = n_seeds_rob, T = sim_T, deg = node_deg,
                       version = 2L)
hash_rob       <- digest::digest(sim_params_rob)
cache_rob      <- here::here("data/sim_robust_cache.rds")

if (file.exists(cache_rob)) {
  cr <- readRDS(cache_rob)
  if (identical(cr$hash, hash_rob)) {
    sim_rob_raw <- cr$data
    t_rob       <- cr$elapsed
    message("Robust-sweep cache hit.")
  } else { cr <- NULL }
} else { cr <- NULL }

if (is.null(cr)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  grid_rob <- expand_grid(
    n_agents = n_agents_rob,
    r_op     = r_levels_rob,
    r_pw     = r_levels_rob,
    seed     = seq_len(n_seeds_rob)
  )
  plan(multisession, workers = parallel::detectCores() - 1)
  t0 <- proc.time()
  sim_rob_raw <- future_pmap_dfr(
    grid_rob,
    function(n_agents, r_op, r_pw, seed) {
      res   <- simulate_liquid_democracy(
        seed = seed, n_per_community = n_agents, n_communities = 1,
        node_degree = node_deg, T = sim_T, r_op = r_op, r_pw = r_pw)
      snaps <- tail(res$snapshots, 2)
      ag    <- res$agents
      tibble(
        n_agents        = n_agents, r_op = r_op, r_pw = r_pw, seed = seed,
        delegation_rate = mean(res$history_delegation[15:20], na.rm = TRUE),
        lost_vote_rate  = mean(res$history_lost[15:20],       na.rm = TRUE),
        avg_ideo_drift  = mean(res$history_drift[15:20],      na.rm = TRUE),
        gini_coef       = mean(snaps$gini_power,              na.rm = TRUE),
        avg_chain_len   = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
        top5_share      = mean(snaps$top5_power_share,        na.rm = TRUE),
        n_voters         = sum(!is.na(ag$my_vote)),
        enp              = { p <- ag$power[!is.na(ag$my_vote)]
                             if (!length(p) || sum(p) == 0) NA_real_
                             else 1 / sum((p / sum(p))^2) },
        total_components = mean(snaps$total_components, na.rm = TRUE)
      )
    },
    .options = furrr_options(seed = NULL)
  )
  plan(sequential)
  t_rob <- proc.time() - t0
  saveRDS(list(hash = hash_rob, data = sim_rob_raw, elapsed = t_rob), cache_rob)
  message("Robust sweep complete.")
}

df_rob    <- sim_rob_raw |> filter(!is.na(avg_chain_len))
df_rob300 <- df_rob      |> filter(n_agents == 300)
```

```{r sim-summary, include=TRUE}
secs_l <- as.numeric(t_large["elapsed"])
cat(sprintf(
  "Large sweep:    %d min %04.1f sec  |  %d parameter combinations  |  %d simulations\n",
  floor(secs_l / 60), secs_l %% 60,
  length(n_agents_large) * length(r_levels_large)^2,
  nrow(sim_large_raw)
))
secs_r <- as.numeric(t_rob["elapsed"])
cat(sprintf(
  "Robust sweep:   %d min %04.1f sec  |  %d parameter combinations  |  %d simulations\n",
  floor(secs_r / 60), secs_r %% 60,
  length(n_agents_rob) * length(r_levels_rob)^2,
  nrow(sim_rob_raw)
))
```

---

# Overview -- Delegation & Lost Vote Rate

```{r gam-ov-fit, cache=TRUE}
metrics_gam     <- c("delegation_rate", "lost_vote_rate")
metrics_gam_lbl <- c(delegation_rate = "Delegation Rate",
                     lost_vote_rate  = "Lost Vote Rate")
n_gam_vals <- c(300L, 1000L)

gam_ov <- purrr::map(n_gam_vals, function(n) {
  df_n <- df_large |> filter(n_agents == n) |> mutate(seed_f = factor(seed))
  purrr::map(metrics_gam, \(y)
    mgcv::gam(
      as.formula(paste0(y, " ~ s(r_op, r_pw, k = 25) + s(seed_f, bs = 're')")),
      data = df_n, method = "REML"
    )
  ) |> setNames(metrics_gam)
}) |> setNames(as.character(n_gam_vals))
```

```{r gam-ov-lineplot, fig.width=13, fig.height=10}
r_pw_show     <- c(0, 3, 6, 9, 12, 15)
pred_line_gam <- expand_grid(r_op = seq(0, 15, by = 0.4), r_pw = r_pw_show)

gam_line_plots <- purrr::map(metrics_gam, function(y) {
  lbl <- metrics_gam_lbl[y]
  purrr::map(n_gam_vals, function(n) {
    df_n  <- df_large |> filter(n_agents == n)
    mod   <- gam_ov[[as.character(n)]][[y]]
    sf_lv <- df_n |> mutate(seed_f = factor(seed)) |> pull(seed_f) |> levels()

    pred_df <- pred_line_gam |>
      mutate(seed_f = factor(sf_lv[1]),
             val    = predict(mod, newdata = pick(everything()),
                              exclude = "s(seed_f)", newdata.guaranteed = TRUE))

    obs_df <- df_n |>
      filter(r_pw %in% r_pw_show) |>
      group_by(r_op, r_pw) |>
      summarise(val = mean(.data[[y]], na.rm = TRUE), .groups = "drop")

    ggplot(pred_df, aes(r_op, val, colour = factor(r_pw), group = factor(r_pw))) +
      geom_point(data = obs_df, aes(x = r_op, y = val, colour = factor(r_pw)),
                 size = 1.8, alpha = 0.55, inherit.aes = FALSE) +
      geom_line(linewidth = 0.9) +
      scale_colour_viridis_d(name = expression(r[pw])) +
      scale_x_continuous(breaks = seq(0, 15, by = 3)) +
      labs(title = paste0(lbl, "  --  N = ", n),
           x = expression(r[op]), y = lbl) +
      theme(legend.position = "right")
  })
}) |> purrr::list_flatten()

wrap_plots(gam_line_plots, ncol = 2) +
  plot_annotation(
    title    = "GAM predicted lines -- delegation & lost vote rate  (r in [0, 15])",
    subtitle = "Points = observed cell means | Lines = GAM prediction | Colours = r_pw levels {0, 3, 6, 9, 12, 15}"
  )
```

Checked for reasonable r values. Plots show a very high delegation + lost vote rate for any value higher than r = 2. Therefore for the following simulations + models r values in {0, 1, 2}.

---

# Part 1 -- N = 300

## 0  Preliminary Check -- Seed-Level Bias

Each seed defines a unique network topology (Watts-Strogatz graph) and initial opinion
vector. If certain seeds consistently produce higher delegation rates or longer chains
regardless of parameter settings, ignoring this structure inflates model residuals.
The mixed model in Part 1 includes a seed-level random intercept `(1|seed)` to absorb
this variation.

```{r seed-bias-check, fig.width=12, fig.height=9}
seed_means <- df_rob300 |>
  group_by(seed) |>
  summarise(across(all_of(all_metrics), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  pivot_longer(-seed, names_to = "metric_key", values_to = "seed_mean") |>
  mutate(metric = factor(all_metrics_lbl[metric_key],
                         levels = all_metrics_lbl[all_metrics]))

seed_stats <- seed_means |>
  group_by(metric) |>
  summarise(grand_mean = mean(seed_mean),
            sd         = sd(seed_mean),
            cv         = sd(seed_mean) / mean(seed_mean),
            .groups    = "drop")

ggplot(seed_means, aes(x = seed_mean)) +
  geom_histogram(bins = 20, fill = "#3498DB", alpha = 0.65, colour = "white") +
  geom_vline(data = seed_stats, aes(xintercept = grand_mean),
             colour = "black", linewidth = 0.8) +
  geom_vline(data = seed_stats, aes(xintercept = grand_mean + sd),
             colour = "#E74C3C", linewidth = 0.6, linetype = "dashed") +
  geom_vline(data = seed_stats, aes(xintercept = grand_mean - sd),
             colour = "#E74C3C", linewidth = 0.6, linetype = "dashed") +
  facet_wrap(~metric, scales = "free", ncol = 4) +
  labs(
    title    = "Per-seed mean distributions -- N = 300, r in {0, 1, 2, 3}",
    subtitle = "Black = grand mean | Red dashed = +/- 1 SD",
    x = "Per-seed mean (averaged over all parameter cells)", y = "Count"
  ) +
  theme(strip.text = element_text(face = "bold"))
```

```{r seed-bias-table}
seed_stats |>
  mutate(across(c(grand_mean, sd, cv), \(x) round(x, 4))) |>
  rename(Metric = metric, `Grand mean` = grand_mean, SD = sd, `CV` = cv) |>
  knitr::kable(caption = "Seed-level variability -- CV > 0.05 indicates meaningful seed bias warranting a random intercept.")
```

CV values above 0.05 confirm that network topology and initial opinions introduce
non-negligible between-seed variation, justifying the random intercept `(1|seed)`.

---

## 1  LMM -- fine grid (r ∈ {0, 1, 2}, N = 300)

Within the approximately linear regime ($r \leq 2$) the interaction model

$$y_{ijk} = \beta_0 + \beta_1\,r_{op} + \beta_2\,r_{pw} + \beta_3\,(r_{op} \cdot r_{pw}) + u_k + \varepsilon_{ijk}$$

is fitted. $u_k \sim \mathcal{N}(0, \sigma_u^2)$ is the seed random intercept.


```{r lmm-marginal-plots, fig.width=13, fig.height=22}
obs_all <- df_rob300 |>
  group_by(r_op, r_pw) |>
  summarise(across(all_of(all_metrics), \(x) mean(x, na.rm = TRUE)), .groups = "drop")

lmm_plots <- purrr::map(all_metrics, function(y) {
  lbl    <- all_metrics_lbl[y]
  obs_df <- obs_all |> dplyr::select(r_op, r_pw, obs = all_of(y))

  ggplot(obs_df, aes(r_op, obs, colour = factor(r_pw), group = factor(r_pw))) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1, alpha = 0.9) +
    geom_point(size = 2.5, alpha = 0.85) +
    scale_colour_viridis_d(name = expression(r[pw])) +
    labs(title = lbl, x = expression(r[op]), y = lbl) +
    theme(legend.position = "right")
})

wrap_plots(lmm_plots, ncol = 2) +
  plot_annotation(
    title    = "LMM marginal effects  (N = 300, r in {0,1,2})",
    subtitle = "Points = seed-averaged cell means | Lines = OLS regression per r_pw level | Colours = r_pw levels"
  )
```


- **ENP** ($1/\sum s_i^2$): effective number of equally powerful agents; *higher = less concentrated*

```{r lmm-all-fit, cache=TRUE, dependson=c("setup","robust-sweep"), cache.extra=list(all_metrics)}
lmm_all <- purrr::map(all_metrics, \(y)
  lmer(reformulate(c("r_op * r_pw", "(1|seed)"), response = y),
       data = df_rob300, REML = TRUE)
) |> setNames(all_metrics)
```

```{r lmm-all-coef-table}
lmm_coef_tbl <- purrr::map_dfr(all_metrics, function(y) {
  m_ml  <- lmer(reformulate(c("r_op * r_pw", "(1|seed)"), response = y),
                data = df_rob300, REML = FALSE)
  m_add <- lmer(reformulate(c("r_op + r_pw",  "(1|seed)"), response = y),
                data = df_rob300, REML = FALSE)
  lrt   <- anova(m_add, m_ml)
  coefs <- broom.mixed::tidy(lmm_all[[y]], effects = "fixed", conf.int = TRUE)
  r2    <- MuMIn::r.squaredGLMM(lmm_all[[y]])

  b1 <- dplyr::filter(coefs, term == "r_op")
  b2 <- dplyr::filter(coefs, term == "r_pw")
  b3 <- dplyr::filter(coefs, term == "r_op:r_pw")

  lrt_chi2 <- lrt$Chisq[2]
  lrt_p    <- lrt[["Pr(>Chisq)"]][2]

  tibble::tibble(
    Metric           = all_metrics_lbl[y],
    `b1 (r_op)`      = fmt_coef(b1$estimate, b1$p.value),
    `b2 (r_pw)`      = fmt_coef(b2$estimate, b2$p.value),
    `b3 (r_op:r_pw)` = fmt_coef(b3$estimate, b3$p.value),
    `LRT χ²`         = paste0(round(lrt_chi2, 2), sig_stars(lrt_p),
                               " (", fmt_p(lrt_p), ")"),
    R2m              = round(r2[1, "R2m"], 3),
    R2c              = round(r2[1, "R2c"], 3)
  )
})

knitr::kable(lmm_coef_tbl,
  caption = "LMM fixed effects -- all metrics, r in {0,1,2}, N = 300. * p < 0.05, ** p < 0.01. LRT χ² tests interaction term (r_op:r_pw) against additive model. R2m = marginal, R2c = conditional.")
```

**Interpretation:**

**Note**  
- All results must be interpreted with considering a very high delegation rate and therefore a very high lost vote rate especially for higher r values. Meaning some results might be related to the enormous loss of votes instead of their "true" effects.  

**Delegation Dynamics: Delegation Rate and Lost Vote Rate**  
- Opinion responsiveness strongly increases both delegation activity and the likelihood of votes becoming lost within delegation chains. Agents are therefore substantially more likely to delegate when ideological similarity is prioritized, but this simultaneously creates more fragile delegation structures in which votes fail to reach an active representative. In contrast, power responsiveness slightly stabilizes the system by reducing lost votes, likely because if an agent is more powerful than his neighbours it is very unlikely that he/she will not cast his/her vote. Overall, the results suggest that ideological delegation promotes participation but also increases the lost vote rate due to cycles.

**Delegation Structure: Average Chain Length and Ideological Drift**  
- Higher opinion responsiveness produces longer delegation chains, indicating that agents delegate through sequences of ideologically similar intermediaries rather than directly to highly central actors. At the same time, both opinion responsiveness and power responsiveness increase (slightly) ideological drift. However, the negative interaction effect suggests that the simultaneous presence of both mechanisms partially constrains this increase in drift, whereas this effect is almost 0.  
Note: All effects are very very small! 

**Power Concentration: Gini Coefficient, Top-5% Share, and ENP**  
- Both responsiveness to opinion and responsiveness to power contribute to a small increase in overall inequality of political influence, although the effects are generally small. However, they act along different structural dimensions: opinion responsiveness strongly reduces the effective number of influential agents, which is most likely due to higher lost vote rate, while power responsiveness tends to distribute influence across multiple competing hubs rather than reinforcing a single dominant center. The absence of strong effects on the Top-5% power share suggests that concentration emerges more gradually across the system rather than through extreme domination by a very small elite. The absence of strong effects on the Top-5% power share suggests that concentration does not primarily take the form of extreme winner-takes-all (“rich-get-richer”) dynamics at the very top. Overall, the results are not what I expected especially from r_pw, I stil think the delegation rate has a strong effect on the power concentration.


```{r lmm-heatmaps, fig.width=11, fig.height=26}
heat_pairs <- purrr::map(all_metrics, function(y) {
  lbl     <- all_metrics_lbl[y]
  pred_df <- expand_grid(r_op = r_levels_rob, r_pw = r_levels_rob) |>
    mutate(val = predict(lmm_all[[y]], newdata = pick(everything()), re.form = NA))
  obs_df  <- obs_all |> dplyr::select(r_op, r_pw, val = all_of(y))
  zlim    <- range(c(pred_df$val, obs_df$val), na.rm = TRUE)

  p_pred <- ggplot(pred_df, aes(r_op, r_pw, fill = val)) +
    geom_tile() +
    scale_fill_viridis_c(name = lbl, limits = zlim) +
    labs(title = paste("Predicted --", lbl),
         x = expression(r[op]), y = expression(r[pw])) +
    theme(panel.grid = element_blank())

  p_obs <- ggplot(obs_df, aes(r_op, r_pw, fill = val)) +
    geom_tile() +
    scale_fill_viridis_c(name = lbl, limits = zlim) +
    labs(title = paste("Observed --", lbl),
         x = expression(r[op]), y = expression(r[pw])) +
    theme(panel.grid = element_blank())

  p_pred | p_obs
})

wrap_plots(heat_pairs, ncol = 1) +
  plot_annotation(
    title    = "Predicted vs observed -- all metrics  (fine grid, N = 300)",
    subtitle = "Each row uses its own colour scale | Left = LMM prediction | Right = observed cell means"
  )
```

Close agreement between predicted and observed heatmaps confirms that the linear
interaction model captures the dominant structure within $r \leq 2$. Systematic
deviations (e.g. curvature in corners) would indicate residual nonlinearity not
captured by $\beta_3$.

---

# Part 2 -- Robustness: Effect of Network Size N

The robustness sweep covers N $\in$ \{100, 200, \ldots, 1000\} at $r_{op}, r_{pw} \in \{0,1,2\}$.
Three complementary analyses address whether the parameter effects are N-dependent:
(1) raw metric trajectories at a fixed reference point, (2) N included directly as a
predictor to estimate the per-unit effect, and (3) separate LMMs at N = 200, 500, 1000
compared side by side.

ENP and Total Components are normalised for Part 2 comparisons across N: ENP is divided
by the number of direct voters ($n_\text{voters}$) and Total Components by $N$.

```{r part3-normalise, include=FALSE}
df_rob_n <- df_rob |>
  mutate(enp             = enp / n_voters,
         total_components = total_components / n_agents)
```

---

## 2.1  Metric trends across N (r_op = r_pw = 1)

```{r n-trends, fig.width=12, fig.height=10}
trend_df <- df_rob_n |>
  filter(r_op == 1, r_pw == 1) |>
  group_by(n_agents) |>
  summarise(across(all_of(all_metrics), \(x) mean(x, na.rm = TRUE)), .groups = "drop") |>
  pivot_longer(all_of(all_metrics), names_to = "metric_key", values_to = "value") |>
  mutate(
    metric = factor(all_metrics_lbl[metric_key],
                    levels = all_metrics_lbl[all_metrics])
  )

ggplot(trend_df, aes(n_agents, value)) +
  geom_line(colour = "#3498DB", linewidth = 1) +
  geom_point(colour = "#3498DB", size = 2.5) +
  scale_x_continuous(breaks = seq(100, 1000, by = 100)) +
  facet_wrap(~metric, scales = "free_y", ncol = 4) +
  labs(
    title    = "Metric values across N at r_op = r_pw = 1",
    subtitle = paste0("Each point = mean over ", n_seeds_rob, " seeds | N from 100 to 1000"),
    x = "N (agents per community)", y = NULL
  ) +
  theme(strip.text = element_text(face = "bold"))
```

Across the explored range (N = 200–1000), most delegation and concentration metrics remain relatively stable, suggesting that the system operates in a largely size-invariant regime under $r_{op} = r_{pw} = 1$.

The strongest size dependence appears in ENP, Top-5% Power Share and Avg. Chain Length (for smaller sizes).

---

## 2.2  Coefficient tables -- N = 200, 500, 1000

```{r n-coef-models, cache=TRUE, dependson="setup", cache.extra=list(all_metrics)}
n_compare <- c(200L, 500L, 1000L)

n_coef_raw <- purrr::map_dfr(n_compare, function(n) {
  df_n <- df_rob_n |> dplyr::filter(n_agents == n)
  purrr::map_dfr(all_metrics, function(y) {
    m_ml  <- lmer(reformulate(c("r_op * r_pw", "(1|seed)"), response = y),
                  data = df_n, REML = FALSE)
    m_add <- lmer(reformulate(c("r_op + r_pw",  "(1|seed)"), response = y),
                  data = df_n, REML = FALSE)
    m_r   <- lmer(reformulate(c("r_op * r_pw", "(1|seed)"), response = y),
                  data = df_n, REML = TRUE)
    lrt   <- anova(m_add, m_ml)
    coefs <- broom.mixed::tidy(m_ml, effects = "fixed", conf.int = TRUE)
    r2    <- MuMIn::r.squaredGLMM(m_r)
    b1 <- dplyr::filter(coefs, term == "r_op")
    b2 <- dplyr::filter(coefs, term == "r_pw")
    b3 <- dplyr::filter(coefs, term == "r_op:r_pw")
    tibble::tibble(
      N        = n, metric = y,
      b1_est   = b1$estimate, b1_p = b1$p.value,
      b1_lo    = b1$conf.low, b1_hi = b1$conf.high,
      b2_est   = b2$estimate, b2_p = b2$p.value,
      b2_lo    = b2$conf.low, b2_hi = b2$conf.high,
      b3_est   = b3$estimate, b3_p = b3$p.value,
      b3_lo    = b3$conf.low, b3_hi = b3$conf.high,
      lrt_chi2 = lrt$Chisq[2],
      lrt_p    = lrt[["Pr(>Chisq)"]][2],
      R2m      = round(r2[1, "R2m"], 3),
      R2c      = round(r2[1, "R2c"], 3)
    )
  })
})
```

```{r n-coef-table}
n_coef_raw |>
  mutate(
    metric_ord = factor(metric, levels = all_metrics),
    Metric     = all_metrics_lbl[metric],
    b1       = fmt_coef(b1_est, b1_p),
    b2       = fmt_coef(b2_est, b2_p),
    b3       = fmt_coef(b3_est, b3_p),
    `LRT χ²` = paste0(round(lrt_chi2, 2), sig_stars(lrt_p),
                       " (", fmt_p(lrt_p), ")")
  ) |>
  dplyr::arrange(metric_ord, N) |>
  dplyr::select(Metric, N, b1, b2, b3, `LRT χ²`, R2m, R2c) |>
  knitr::kable(
    col.names = c("Metric", "N", "b1 (r_op)", "b2 (r_pw)", "b3 (r_op:r_pw)",
                  "LRT χ²", "R2m", "R2c"),
    caption   = "LMM coefficients for N = 200, 500, 1000 (r in {0,1,2}). * p < 0.05, ** p < 0.01. LRT χ² tests interaction term against additive model.")
```

Across all system sizes, delegation rate and lost vote dynamics remain highly stable and strongly explained by the responsiveness parameters ($R^2_m \approx 0.95$), indicating robust and largely scale-independent delegation behavior. In contrast, chain length, ideological drift, and power concentration become increasingly sensitive to the parameters as $N$ grows, reflected in rising effect sizes and higher explanatory power at larger population sizes.

---

## 2.3  Coefficient plot -- N = 200, 500, 1000

```{r n-coef-plot, fig.width=12, fig.height=10}
n_coef_long <- n_coef_raw |>
  mutate(
    Metric = factor(all_metrics_lbl[metric],
                    levels = all_metrics_lbl[all_metrics])
  ) |>
  tidyr::pivot_longer(c(b1_est, b2_est, b3_est),
                      names_to = "coef", values_to = "estimate") |>
  mutate(
    ci_lo = dplyr::case_when(
      coef == "b1_est" ~ b1_lo,
      coef == "b2_est" ~ b2_lo,
      coef == "b3_est" ~ b3_lo
    ),
    ci_hi = dplyr::case_when(
      coef == "b1_est" ~ b1_hi,
      coef == "b2_est" ~ b2_hi,
      coef == "b3_est" ~ b3_hi
    ),
    coef_lbl = dplyr::recode(coef,
      b1_est = "b1 (r_op)", b2_est = "b2 (r_pw)", b3_est = "b3 (r_op:r_pw)")
  )

ggplot(n_coef_long,
       aes(x = coef_lbl, y = estimate, colour = factor(N), group = factor(N))) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_pointrange(
    aes(ymin = ci_lo, ymax = ci_hi),
    position = position_dodge(width = 0.4), linewidth = 0.8, size = 0.7
  ) +
  scale_colour_viridis_d(name = "N") +
  facet_wrap(~Metric, scales = "free_y", ncol = 4) +
  labs(
    title    = "LMM coefficients: N = 200, 500, 1000",
    subtitle = "All coefficients with 95% CI | Dashed = 0",
    x = NULL, y = "Estimate"
  ) +
  theme(strip.text       = element_text(face = "bold"),
        axis.text.x      = element_text(angle = 20, hjust = 1))
```

---

# Part 3 -- Trust-Weighted Delegation

## Formulas

**Trust update** (after each round, using the effective vote propagated through
$j$'s chain):

$$\tau_{ij}(t) = \lambda \cdot \tau_{ij}(t-1) - (1-\lambda)\cdot\bigl|o_i - v_j(t-1)\bigr|$$

**Modified attractiveness:**

$\tilde{A}_{ij}(t) = A_{ij} \cdot 2\sigma\bigl(\gamma\cdot\tau_{ij}(t)\bigr)$
where $A_{ij} = \sigma\left(r_{op}(1-2|o_i-o_j|)\right)\cdot\sigma\left(r_{pw}\log(p_j/p_i)\right)$
is the baseline attractiveness from Part 1. The factor $2\sigma(\cdot)$ ensures that
$\tau=0$ (no experience) gives a neutral modifier of $2\sigma(0)=1$, recovering the
unmodified $A_{ij}$ exactly. The self-weight $w_\text{self}$ is unchanged.

The simulations in Part 1 and Part 3 used $\lambda = 0$, $\gamma = 0$, which sets
the modifier to $2\sigma(0)=1$ for all agents and rounds, leaving the original
attractiveness intact.

Here, $\lambda = 0.5$ is held fixed (exponential moving average with half-life of one
round) and $\gamma$ is varied from 0 to 3 to examine how trust sensitivity modulates
delegation patterns.

```{r trust-sweep, include=FALSE, cache=TRUE}
gamma_levels  <- seq(0, 3, by = 0.5)
r_trust_vals  <- 0:2
lambda_fixed  <- 0.5
n_trust       <- 300L
n_seeds_trust <- 100L

sim_params_trust <- list(gamma = gamma_levels, r = r_trust_vals,
                         lambda = lambda_fixed, n = n_trust,
                         seeds = n_seeds_trust, T = sim_T, deg = node_deg,
                         version = 2L)
hash_trust  <- digest::digest(sim_params_trust)
cache_trust <- here::here("data/sim_trust_cache.rds")

if (file.exists(cache_trust)) {
  ct <- readRDS(cache_trust)
  if (identical(ct$hash, hash_trust)) {
    sim_trust_raw <- ct$data
    t_trust       <- ct$elapsed
    message("Trust-sweep cache hit.")
  } else { ct <- NULL }
} else { ct <- NULL }

if (is.null(ct)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  grid_trust <- expand_grid(
    gamma  = gamma_levels,
    r_op   = r_trust_vals,
    r_pw   = r_trust_vals,
    seed   = seq_len(n_seeds_trust)
  )
  plan(multisession, workers = parallel::detectCores() - 1)
  t0 <- proc.time()
  sim_trust_raw <- future_pmap_dfr(
    grid_trust,
    function(gamma, r_op, r_pw, seed) {
      res   <- simulate_liquid_democracy(
        seed = seed, n_per_community = n_trust, n_communities = 1,
        node_degree = node_deg, T = sim_T, r_op = r_op, r_pw = r_pw,
        lambda = lambda_fixed, gamma = gamma)
      snaps <- tail(res$snapshots, 2)
      ag    <- res$agents
      tibble(
        gamma           = gamma, r_op = r_op, r_pw = r_pw, seed = seed,
        delegation_rate = mean(res$history_delegation[15:20], na.rm = TRUE),
        lost_vote_rate  = mean(res$history_lost[15:20],       na.rm = TRUE),
        avg_ideo_drift  = mean(res$history_drift[15:20],      na.rm = TRUE),
        gini_coef       = mean(snaps$gini_power,              na.rm = TRUE),
        avg_chain_len   = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
        top5_share      = mean(snaps$top5_power_share,        na.rm = TRUE),
        enp              = { p <- ag$power[!is.na(ag$my_vote)]
                             if (!length(p) || sum(p) == 0) NA_real_
                             else 1 / sum((p / sum(p))^2) },
        total_components = mean(snaps$total_components, na.rm = TRUE)
      )
    },
    .options = furrr_options(seed = NULL)
  )
  plan(sequential)
  t_trust <- proc.time() - t0
  saveRDS(list(hash = hash_trust, data = sim_trust_raw, elapsed = t_trust),
          cache_trust)
  message("Trust sweep complete.")
}

df_trust <- sim_trust_raw |> filter(!is.na(avg_chain_len))
```

## Effect of Trust Sensitivity $\gamma$ on All Metrics

$\lambda = 0.5$ fixed; $\gamma \in [0, 3]$; $r_{op}, r_{pw} \in \{0, 1, 2\}$.
Points are seed-averaged cell means; lines connect them.
Colour encodes $r_{op}$, line type encodes $r_{pw}$.

```{r trust-gamma-plot, fig.width=14, fig.height=22}
trust_cell <- df_trust |>
  group_by(gamma, r_op, r_pw) |>
  summarise(across(all_of(all_metrics), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop")

lt_map <- c("0" = "solid", "1" = "dashed", "2" = "dotted")

trust_plots <- purrr::map(all_metrics, function(y) {
  lbl <- all_metrics_lbl[y]
  df_y <- trust_cell |>
    dplyr::select(gamma, r_op, r_pw, val = all_of(y))

  ggplot(df_y, aes(gamma, val,
                   colour   = factor(r_op),
                   linetype = factor(r_pw),
                   group    = interaction(r_op, r_pw))) +
    geom_point(size = 2.2, alpha = 0.85) +
    geom_line(linewidth = 0.85) +
    scale_colour_viridis_d(name = expression(r[op])) +
    scale_linetype_manual(name = expression(r[pw]),
                          values = lt_map) +
    scale_x_continuous(breaks = seq(0, 3, by = 0.5)) +
    labs(title = lbl, x = expression(gamma), y = lbl) +
    theme(legend.position = "right")
})

wrap_plots(trust_plots, ncol = 2) +
  plot_annotation(
    title    = paste0("Trust sensitivity γ vs all metrics  (λ = ", lambda_fixed,
                      ", N = ", n_trust, ", r_op/r_pw ∈ {0,1,2})"),
    subtitle = "Points = seed-averaged cell means | Lines connect gamma levels | Colour = r_op | Linetype = r_pw"
  )
```

## LMM with Trust -- Full Interaction Model

The same mixed-effects framework as in Part 1 is extended to include $\gamma$ as a
third factorial predictor:

$$y_{ijk} = \beta_0 + \beta_1 r_{op} + \beta_2 r_{pw} + \beta_\gamma \gamma + \beta_3 (r_{op} \cdot r_{pw}) + \beta_4 (r_{op} \cdot \gamma) + \beta_5 (r_{pw} \cdot \gamma) + \beta_6 (r_{op} \cdot r_{pw} \cdot \gamma) + u_k + \varepsilon_{ijk}$$

$\lambda = 0.5$ is held fixed throughout. The LRT tests whether the three-way
interaction $\beta_6$ significantly improves fit over the model containing all
two-way interactions but no three-way term.

```{r trust-coef-table, cache=TRUE, dependson=c("setup","trust-sweep"), cache.extra=list(all_metrics)}
trust_lmm <- purrr::map(all_metrics, \(y)
  lmer(reformulate(c("r_op * r_pw * gamma", "(1|seed)"), response = y),
       data = df_trust, REML = TRUE)
) |> setNames(all_metrics)

trust_coef_tbl <- purrr::map_dfr(all_metrics, function(y) {
  m_full <- lmer(reformulate(c("r_op * r_pw * gamma", "(1|seed)"), response = y),
                 data = df_trust, REML = FALSE)
  m_2way <- lmer(
    reformulate(c("r_op * gamma + r_pw * gamma + r_op * r_pw", "(1|seed)"),
                response = y),
    data = df_trust, REML = FALSE)
  lrt    <- anova(m_2way, m_full)
  coefs  <- broom.mixed::tidy(trust_lmm[[y]], effects = "fixed", conf.int = TRUE)
  r2     <- MuMIn::r.squaredGLMM(trust_lmm[[y]])

  gc <- function(trm) {
    row <- dplyr::filter(coefs, term == trm)
    if (nrow(row) == 0) return("—")
    fmt_coef(row$estimate, row$p.value)
  }

  lrt_chi2 <- lrt$Chisq[2]
  lrt_p    <- lrt[["Pr(>Chisq)"]][2]

  tibble::tibble(
    Metric             = all_metrics_lbl[y],
    `b1 (r_op)`        = gc("r_op"),
    `b2 (r_pw)`        = gc("r_pw"),
    `b_γ (γ)`          = gc("gamma"),
    `b3 (r_op:r_pw)`   = gc("r_op:r_pw"),
    `b4 (r_op:γ)`      = gc("r_op:gamma"),
    `b5 (r_pw:γ)`      = gc("r_pw:gamma"),
    `b6 (r_op:r_pw:γ)` = gc("r_op:r_pw:gamma"),
    `LRT χ²`           = paste0(round(lrt_chi2, 2), sig_stars(lrt_p),
                                 " (", fmt_p(lrt_p), ")"),
    R2m                = round(r2[1, "R2m"], 3),
    R2c                = round(r2[1, "R2c"], 3)
  )
})

knitr::kable(trust_coef_tbl,
  caption = paste0(
    "LMM with trust: y ~ r_op * r_pw * gamma + (1|seed), lambda = ", lambda_fixed,
    " fixed, N = ", n_trust, ". * p < 0.05, ** p < 0.01. ",
    "LRT χ² tests three-way term (b6) against full two-way model."))
```
**Interpretation of $\gamma$ (trust sensitivity):**
$\gamma$ acts consistently as a **decentralising signal**
across all metrics.
Higher trust sensitivity reduces delegation rate, chain length, lost votes, power
concentration (Gini, Top-5%), and ideological drift, whileraising ENP and total
components — indicating that agents become more selective:
they retain their voterather than delegate to an unproven agent, which distributes power more evenly.
The positive `r_op:γ` interaction (b4) shows that opinionsensitivity amplifies
this selectivity: agents who filter by both opinion similarity *and* trust history
delegate even less and to fewer hubs.
The three-way term b6 is significant for delegation, lostvotes, Gini, drift,
ENP, and components, confirming that the joint effect of all three parameters is not simply 
additive — $\gamma$ reshapes how $r_{op}$ and $r_{pw}$ interact ratherthan just shifting outcomes by a fixed offset.
---

# Part 4  Node Degree Sensitivity

`node_degree = 6` is the default WS ring-lattice parameter used throughout
Parts 1, 2, and 3. A higher degree means more potential delegation partners per
agent, which directly affects chain length and connectivity. This section
checks whether the key results generalise across a wide range of degree values.

The sweep varies `node_degree` from 6 to 30 across all nine combinations of
$r_{op}, r_{pw} \in \{0, 1, 2\}$ at $N = 300$.

```{r deg-sweep, include=FALSE}
deg_vals_s   <- c(6L, 10L, 14L, 18L, 22L, 26L, 30L)
r_vals_deg   <- 0:2
n_seeds_deg  <- 25L
n_agents_deg <- 300L

sim_params_deg <- list(deg = deg_vals_s, r = r_vals_deg,
                       n = n_agents_deg, seeds = n_seeds_deg,
                       T = sim_T, version = 1L)
hash_deg  <- digest::digest(sim_params_deg)
cache_deg <- here::here("data/sim_deg_cache.rds")

if (file.exists(cache_deg)) {
  cd <- readRDS(cache_deg)
  if (identical(cd$hash, hash_deg)) {
    sim_deg_raw <- cd$data; t_deg <- cd$elapsed
    message("Degree-sweep cache hit.")
  } else cd <- NULL
} else cd <- NULL

if (is.null(cd)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  grid_deg <- expand_grid(
    node_degree = deg_vals_s,
    r_op        = r_vals_deg,
    r_pw        = r_vals_deg,
    seed        = seq_len(n_seeds_deg)
  )
  plan(multisession, workers = parallel::detectCores() - 1)
  t0 <- proc.time()
  sim_deg_raw <- future_pmap_dfr(
    grid_deg,
    function(node_degree, r_op, r_pw, seed) {
      res   <- simulate_liquid_democracy(
        seed = seed, n_per_community = n_agents_deg, n_communities = 1,
        node_degree = node_degree, T = sim_T,
        r_op = r_op, r_pw = r_pw)
      snaps <- tail(res$snapshots, 2)
      ag    <- res$agents
      tibble(
        node_degree      = node_degree,
        r_op             = r_op,
        r_pw             = r_pw,
        seed             = seed,
        delegation_rate  = mean(res$history_delegation[15:20], na.rm = TRUE),
        lost_vote_rate   = mean(res$history_lost[15:20],       na.rm = TRUE),
        avg_chain_len    = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
        avg_ideo_drift   = mean(res$history_drift[15:20],      na.rm = TRUE),
        gini_coef        = mean(snaps$gini_power,              na.rm = TRUE),
        top5_share       = mean(snaps$top5_power_share,        na.rm = TRUE),
        enp              = { p <- ag$power[!is.na(ag$my_vote)]
                             if (!length(p) || sum(p) == 0) NA_real_
                             else 1 / sum((p / sum(p))^2) },
        total_components = mean(snaps$total_components,        na.rm = TRUE)
      )
    },
    .options = furrr_options(seed = NULL)
  )
  plan(sequential)
  t_deg <- proc.time() - t0
  saveRDS(list(hash = hash_deg, data = sim_deg_raw, elapsed = t_deg), cache_deg)
  message("Degree sweep complete.")
}
df_deg <- sim_deg_raw |> filter(!is.na(avg_chain_len))
```

## All Metrics vs Node Degree

```{r deg-all-metrics, fig.width=14, fig.height=20}
lt_map_deg <- c("0" = "solid", "1" = "dashed", "2" = "dotted")

cell_deg <- df_deg |>
  group_by(node_degree, r_op, r_pw) |>
  summarise(across(all_of(all_metrics), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop")

deg_plots <- purrr::map(all_metrics, function(y) {
  lbl <- all_metrics_lbl[y]
  ggplot(cell_deg, aes(node_degree, .data[[y]],
                       colour   = factor(r_op),
                       linetype = factor(r_pw),
                       group    = interaction(r_op, r_pw))) +
    geom_point(size = 1.8, alpha = 0.85) +
    geom_line(linewidth = 0.8) +
    scale_colour_viridis_d(name = expression(r[op])) +
    scale_linetype_manual(name = expression(r[pw]), values = lt_map_deg) +
    scale_x_continuous(breaks = deg_vals_s) +
    labs(title = lbl, x = "Node degree (k)", y = lbl)
})

wrap_plots(deg_plots, ncol = 2) +
  plot_annotation(
    title    = paste0("Node degree sensitivity  (N = ", n_agents_deg,
                      ", r_op/r_pw ∈ {0,1,2})"),
    subtitle = paste0("Points = mean over ", n_seeds_deg,
                      " seeds | Colour = r_op | Linetype = r_pw")
  )
```

## Delegation Rate and Chain Length: Core Degree Effects

```{r deg-focus, fig.width=12, fig.height=5}
p_dlg <- ggplot(
  cell_deg |> filter(r_pw == 1),
  aes(node_degree, delegation_rate,
      colour = factor(r_op), group = factor(r_op))
) +
  geom_point(size = 2.2) + geom_line(linewidth = 0.9) +
  scale_colour_viridis_d(name = expression(r[op])) +
  scale_x_continuous(breaks = deg_vals_s) +
  labs(title = "Delegation Rate vs degree  (r_pw = 1 fixed)",
       x = "Node degree (k)", y = "Delegation Rate")

p_chn <- ggplot(
  cell_deg |> filter(r_pw == 1),
  aes(node_degree, avg_chain_len,
      colour = factor(r_op), group = factor(r_op))
) +
  geom_point(size = 2.2) + geom_line(linewidth = 0.9) +
  scale_colour_viridis_d(name = expression(r[op])) +
  scale_x_continuous(breaks = deg_vals_s) +
  labs(title = "Avg. Chain Length vs degree  (r_pw = 1 fixed)",
       x = "Node degree (k)", y = "Avg. Chain Length")

(p_dlg | p_chn) + plot_annotation(
  subtitle = paste0("Mean over ", n_seeds_deg, " seeds | N = ", n_agents_deg)
)
```

---

# Part 5  Cycle Fallback -- Lost Vote Correction

Parts 1--4 use the default simulation where agents in delegation cycles
keep their vote lost (`my_vote = NA`).
This section tests three correction mechanisms based on between-round learning:
if an agent's vote was lost in round $t$, they adjust their delegation choice
in round $t+1$. The vote is genuinely lost in round $t$ and recorded as such;
the correction only takes effect from the following round onward.

**Option A -- Direct voting** (`cycle_fallback = "direct"`):
If an agent's vote was lost in the previous round, they cast their own opinion
directly in the current round, bypassing delegation entirely.
Requires minimal information — the agent only needs to know their own vote
was not represented. Delegation can resume in subsequent rounds if the agent
finds an attractive delegate again.

**Option B -- Re-delegation, excluding bad delegate** (`cycle_fallback = "redelegate"`):
If an agent's vote was lost in the previous round, their previous delegate is
excluded from the choice set. The agent then selects normally from all remaining
neighbours (including the self-weight), so they may still delegate or vote
directly depending on relative attractiveness.
Requires only local memory: the agent recalls who they delegated to last round.

**Option C -- Informed re-delegation** (`cycle_fallback = "informed"`):
If an agent's vote was lost in the previous round, only neighbours whose vote
*was represented* in the previous round are considered as delegation targets.
The agent selects among these verified-valid neighbours using the standard
attractiveness weights, and falls back to direct voting if no such neighbour
exists.
Requires local transparency: the agent can observe which neighbours had an
active, non-lost vote in the previous round — analogous to LiquidFeedback's
public delegation-status display.

All three options are compared to the **baseline** (no fallback) across the
same robustness sweep ($r_{op}, r_{pw} \in \{0,1,2\}$, N = 300, 25 seeds).

```{r fallback-sweep, include=FALSE}
n_seeds_fb  <- 25L
n_agents_fb <- 300L

run_fallback_sweep <- function(fb_type) {
  params <- list(r = r_levels_rob, n = n_agents_fb,
                 seeds = n_seeds_fb, T = sim_T, deg = node_deg,
                 fallback = fb_type, version = 2L)
  h    <- digest::digest(params)
  path <- here::here(paste0("data/sim_fallback_", fb_type, "_cache.rds"))

  if (file.exists(path)) {
    cc <- readRDS(path)
    if (identical(cc$hash, h)) {
      message(fb_type, " cache hit.")
      return(list(data = cc$data, elapsed = cc$elapsed))
    }
  }

  grid <- expand_grid(
    r_op = r_levels_rob, r_pw = r_levels_rob,
    seed = seq_len(n_seeds_fb)
  )
  plan(multisession, workers = parallel::detectCores() - 1)
  t0  <- proc.time()
  dat <- future_pmap_dfr(grid, function(r_op, r_pw, seed) {
    res   <- simulate_liquid_democracy(
      seed = seed, n_per_community = n_agents_fb, n_communities = 1,
      node_degree = node_deg, T = sim_T,
      r_op = r_op, r_pw = r_pw,
      cycle_fallback = fb_type)
    snaps <- tail(res$snapshots, 2)
    ag    <- res$agents
    tibble(
      r_op = r_op, r_pw = r_pw, seed = seed,
      delegation_rate  = mean(res$history_delegation[15:20], na.rm = TRUE),
      lost_vote_rate   = mean(res$history_lost[15:20],       na.rm = TRUE),
      avg_chain_len    = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
      avg_ideo_drift   = mean(res$history_drift[15:20],      na.rm = TRUE),
      gini_coef        = mean(snaps$gini_power,              na.rm = TRUE),
      top5_share       = mean(snaps$top5_power_share,        na.rm = TRUE),
      enp              = { p <- ag$power[!is.na(ag$my_vote)]
                           if (!length(p) || sum(p) == 0) NA_real_
                           else 1 / sum((p / sum(p))^2) },
      total_components = mean(snaps$total_components,        na.rm = TRUE)
    )
  }, .options = furrr_options(seed = NULL))
  plan(sequential)
  el <- proc.time() - t0
  saveRDS(list(hash = h, data = dat, elapsed = el), path)
  list(data = dat, elapsed = el)
}

res_direct   <- run_fallback_sweep("direct")
res_redeleg  <- run_fallback_sweep("redelegate")
res_informed <- run_fallback_sweep("informed")

df_direct   <- res_direct$data   |> filter(!is.na(avg_chain_len))
df_redeleg  <- res_redeleg$data  |> filter(!is.na(avg_chain_len))
df_informed <- res_informed$data |> filter(!is.na(avg_chain_len))
```

```{r fallback-cells, include=FALSE}
model_lvls <- c("Baseline", "Option A: Direct", "Option B: Re-delegate",
                "Option C: Informed")
model_cols <- c("Baseline"              = "#95a5a6",
                "Option A: Direct"      = "#3498DB",
                "Option B: Re-delegate" = "#E74C3C",
                "Option C: Informed"    = "#2ECC71")
lt_fb <- c("0" = "solid", "1" = "dashed", "2" = "dotted")

cell_fb <- bind_rows(
  df_rob300   |> mutate(model = "Baseline"),
  df_direct   |> mutate(model = "Option A: Direct"),
  df_redeleg  |> mutate(model = "Option B: Re-delegate"),
  df_informed |> mutate(model = "Option C: Informed")
) |>
  group_by(model, r_op, r_pw) |>
  summarise(across(all_of(all_metrics), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  mutate(model = factor(model, levels = model_lvls))
```

## All Metrics: Fallback Comparison ($r_{pw} = 1$ fixed)

$r_{pw} = 1$ is held fixed; $r_{op}$ varies on the x-axis.
Each panel shows one line per model; colour encodes the fallback strategy.

```{r fallback-rop, fig.width=14, fig.height=20}
cell_fb1 <- cell_fb |> filter(r_pw == 1)

fb_plots <- purrr::map(all_metrics, function(y) {
  lbl <- all_metrics_lbl[y]
  ggplot(cell_fb1,
         aes(r_op, .data[[y]], colour = model, group = model)) +
    geom_point(size = 2.2, alpha = 0.85) +
    geom_line(linewidth = 0.9) +
    scale_colour_manual(values = model_cols, name = "Model") +
    scale_x_continuous(breaks = r_levels_rob) +
    labs(title = lbl, x = expression(r[op]), y = lbl) +
    theme(legend.position = "right")
})

wrap_plots(fb_plots, ncol = 2) +
  plot_annotation(
    title    = paste0("Cycle fallback comparison  (r_pw = 1 fixed, N = ", n_agents_fb,
                      ", ", n_seeds_fb, " seeds)"),
    subtitle = "Colour = model | Points = seed-averaged cell means | r_pw = 1 fixed"
  )
```

## Delegation Rate & Lost Vote Rate Over Rounds

```{r fallback-history-sweep, include=FALSE}
n_seeds_fb_hist <- 25L

run_history_sweep <- function(fb_type) {
  params <- list(r = r_levels_rob, n = n_agents_fb,
                 seeds = n_seeds_fb_hist, T = sim_T, deg = node_deg,
                 fallback = fb_type, version = 2L, type = "history")
  h    <- digest::digest(params)
  path <- here::here(paste0("data/sim_fb_hist_", fb_type, "_cache.rds"))
  if (file.exists(path)) {
    cc <- readRDS(path)
    if (identical(cc$hash, h)) { message(fb_type, " hist cache hit."); return(cc$data) }
  }
  grid <- expand_grid(r_op = r_levels_rob, r_pw = 1L,
                      seed = seq_len(n_seeds_fb_hist))
  plan(multisession, workers = parallel::detectCores() - 1)
  dat <- future_pmap_dfr(grid, function(r_op, r_pw, seed) {
    res <- simulate_liquid_democracy(
      seed = seed, n_per_community = n_agents_fb, n_communities = 1,
      node_degree = node_deg, T = sim_T, r_op = r_op, r_pw = r_pw,
      cycle_fallback = fb_type)
    tibble(r_op = r_op, r_pw = r_pw, seed = seed,
           round           = seq_len(sim_T),
           delegation_rate = res$history_delegation,
           lost_vote_rate  = res$history_lost)
  }, .options = furrr_options(seed = NULL))
  plan(sequential)
  saveRDS(list(hash = h, data = dat), path)
  dat
}

hist_none     <- run_history_sweep("none")
hist_direct   <- run_history_sweep("direct")
hist_redel    <- run_history_sweep("redelegate")
hist_informed <- run_history_sweep("informed")
```

```{r fallback-history-plot, fig.width=13, fig.height=8}
hist_all <- bind_rows(
  hist_none     |> mutate(model = "Baseline"),
  hist_direct   |> mutate(model = "Option A: Direct"),
  hist_redel    |> mutate(model = "Option B: Re-delegate"),
  hist_informed |> mutate(model = "Option C: Informed")
) |>
  group_by(model, r_op, round) |>
  summarise(delegation_rate = mean(delegation_rate, na.rm = TRUE),
            lost_vote_rate  = mean(lost_vote_rate,  na.rm = TRUE),
            .groups = "drop") |>
  mutate(model = factor(model, levels = model_lvls))

hist_long <- hist_all |>
  pivot_longer(c(delegation_rate, lost_vote_rate),
               names_to = "metric_key", values_to = "value") |>
  mutate(metric = factor(all_metrics_lbl[metric_key],
                         levels = all_metrics_lbl[all_metrics]))

ggplot(hist_long, aes(round, value, colour = model, group = model)) +
  geom_line(linewidth = 0.85) +
  scale_colour_manual(values = model_cols, name = "Model") +
  scale_x_continuous(breaks = seq(0, sim_T, by = 5)) +
  facet_grid(metric ~ factor(r_op, labels = paste0("r_op = ", r_levels_rob)),
             scales = "free_y") +
  labs(
    title    = paste0("Delegation & Lost Vote Rate over rounds  (r_pw = 1 fixed, N = ",
                      n_agents_fb, ", ", n_seeds_fb_hist, " seeds)"),
    subtitle = "Lines = seed-averaged mean per round | Colour = model",
    x = "Round", y = NULL
  ) +
  theme(strip.text       = element_text(face = "bold"),
        legend.position  = "bottom")
```