---
title: "Weekly Report — Week 9 (17.04.2026 – 23.04.2026)"
output: github_document
date: "2026-04-17"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())

library(here)
library(tibble)
library(dplyr)
library(tidyverse)
library(ggplot2)

source(here::here("scripts/Network.R"))
source(here::here("scripts/Summary.R"))
```

## Summary

- Implemented structural homophily via a Metropolis opinion-shuffle step
- The shuffle runs once on the fixed network before the simulation loop;
  it stops as soon as a specified target assortativity is reached
- Added `compute_network_homophily()` to measure the degree of homophily
  of any friendship graph
- Ran the responsiveness experiment across three target assortativity
  levels (0.2, 0.4, 0.6) plus the random baseline

------------------------------------------------------------------------

## 1. Homophily Algorithm

The standard model assigns opinions randomly, so connected agents are
no more similar than any two strangers. The Metropolis opinion-shuffle
redistributes lay opinions on the **fixed** network until a target level
of structural homophily is reached.

**What stays fixed:** the entire network topology — edges, degree
sequence, clustering, path lengths. Experts are excluded throughout.

**What changes:** the assignment of opinion values to lay nodes. The
same set of values drawn from $\text{Uniform}[0,1]$ is preserved; only
their allocation is reshuffled.

**Algorithm.** Repeat up to `homophily_steps` times:

1. Draw two random lay agents $i$ and $j$
2. Compute total opinion disagreement on all lay-to-lay edges touching
   $i$ or $j$, excluding the $i$–$j$ edge (cancels):
   $$E = \sum_{v \in N_i \setminus \{j\}} |op_i - op_v| + \sum_{v \in N_j \setminus \{i\}} |op_j - op_v|$$
3. Propose swapping $op_i \leftrightarrow op_j$; compute $E'$
4. Accept if $E' < E$; otherwise accept with probability
   $e^{-(E'-E)\cdot \text{homophily\_t}}$
5. Every 100 steps: compute opinion assortativity on the lay subgraph;
   **stop** as soon as $r_{op} \geq \text{target\_homophily}$

The parameter `homophily_t` (Metropolis temperature) controls how
strictly improvements are enforced — higher values make the algorithm
more greedy and converge faster. The algorithm is disabled entirely when
`target_homophily = NULL`.

**Measuring homophily.** `compute_network_homophily(gF, agents)` returns
two metrics on the lay subgraph:

| Metric | Range | Random baseline | Perfect homophily |
|----|----|----|-----|
| `assortativity` | $[-1, 1]$ | $\approx 0$ | $1$ |
| `mean_edge_disagreement` | $[0, 1]$ | $\approx 0.33$ | $0$ |

The assortativity is the stopping criterion. The mean edge disagreement
is directly proportional to the energy $E$ the shuffle minimises,
making it the most natural companion metric.

------------------------------------------------------------------------

## 2. Results — Responsiveness × Target Homophily

Simulation parameters: n = 250 lay agents, 1 community, k = 6,
p_rewire = 0.10, T = 200, 100 seeds, `homophily_t` = 5.

```{r load-results}
rds_path <- here::here("simulations/output/results_sim_homophily.rds")
if (!file.exists(rds_path))
  stop("RDS not found — knit simulations/sim_03_homophily.Rmd first.")

saved <- readRDS(rds_path)

if (!"target_homophily" %in% names(saved$results_snapshots))
  stop("RDS uses old column name 'homophily_degree' — delete the file and re-knit sim_03_homophily.Rmd.")

results_snapshots <- saved$results_snapshots %>%
  mutate(homophily_label = ifelse(is.na(target_homophily), "baseline",
                                  as.character(target_homophily)))

x_scale    <- scale_x_log10(breaks = c(0.1, 0.5, 1, 2, 5, 10, 25),
                             labels = c("0.1","0.5","1","2","5","10","25"))
hom_colors <- c("baseline" = "grey60",
                "0.2"      = "#4DBBD5",
                "0.4"      = "#E64B35",
                "0.6"      = "#3C5488")
```

### Ideological Drift

```{r plot-drift}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = avg_drift,
             color = homophily_label)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  scale_color_manual(values = hom_colors) +
  labs(title  = "Average Ideological Drift at T = 50 / 100 / 150 / 200",
       x      = "Responsiveness to Power",
       y      = "Avg. Ideological Drift",
       color  = "Target Homophily") +
  theme_minimal()
```

### Lost Vote Rate

```{r plot-lost}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = lost_vote_rate,
             color = homophily_label)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  scale_color_manual(values = hom_colors) +
  labs(title  = "Lost Vote Rate at T = 50 / 100 / 150 / 200",
       x      = "Responsiveness to Power",
       y      = "Lost Vote Rate",
       color  = "Target Homophily") +
  theme_minimal()
```

### Delegation Rate

```{r plot-delegation}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = delegation_rate,
             color = homophily_label)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  scale_color_manual(values = hom_colors) +
  labs(title  = "Delegation Rate at T = 50 / 100 / 150 / 200",
       x      = "Responsiveness to Power",
       y      = "Share of Lay Agents Delegating",
       color  = "Target Homophily") +
  theme_minimal()
```
