---
title: "Weekly Report — Week 7 (03.04.2026 – 09.04.2026)"
output: github_document
date: "2026-04-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())

library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(purrr)

source(here::here("scripts/Network.R"))

f_abs <- readRDS(here::here("simulations/output/results_sim3_rewire_0.1_1000seeds_argmaxformula.rds"))
f_log <- readRDS(here::here("simulations/output/results_sim3_rewire_0.1_1000seeds_log_argmax.rds"))

combined <- bind_rows(
  f_abs$results_snapshots %>% mutate(formula = "Absolute (Formula 3)"),
  f_log$results_snapshots %>% mutate(formula = "Log-Ratio (Formula 4)")
)
combined_time <- bind_rows(
  f_abs$results_time %>% mutate(formula = "Absolute (Formula 3)"),
  f_log$results_time %>% mutate(formula = "Log-Ratio (Formula 4)")
)

x_scale <- scale_x_log10(
  breaks = c(0.1, 0.5, 1, 2, 5, 10, 25),
  labels = c("0.1", "0.5", "1", "2", "5", "10", "25")
)
col_scale <- scale_color_manual(values = c(
  "Absolute (Formula 3)" = "steelblue",
  "Log-Ratio (Formula 4)" = "tomato"
))
fill_scale <- scale_fill_manual(values = c(
  "Absolute (Formula 3)" = "steelblue",
  "Log-Ratio (Formula 4)" = "tomato"
))
```

## Summary

- Switched the power term in both the neighbour attractiveness and self-weight
  formulas from absolute differences to log-ratios
- Old formula: `sigmoid(r × (power_j − power_i))`
- New formula: `sigmoid(r × log(power_j / power_i))`
- Motivation: absolute differences cause sigmoid saturation once a power
  hierarchy forms, making `r` lose interpretive meaning after early rounds;
  log-ratios keep `r` consistent across all rounds
- Both formulas run on 1 000 seeds, p_rewire = 0.10, T = 200

---

## 1. Formula Comparison

**Formula 3 — Absolute Power Difference** (archived):
`A_ij = (1 - |pref_i - pref_j|) × sigmoid(r × (power_j - power_i))`

Once a hierarchy forms (e.g. round 20+), power differences easily reach 10–50.
For r = 1, sigmoid(10) ≈ 1 and sigmoid(50) ≈ 1 — indistinguishable.
The formula is effectively binary (always delegate / never delegate) long before T = 200.

**Formula 4 — Log-Ratio Power** (current default):
`A_ij = (1 - |pref_i - pref_j|) × sigmoid(r × log(power_j / power_i))`

A 5:1 power ratio produces log(5) ≈ 1.61 regardless of whether the absolute
values are 1 vs 5 or 100 vs 500. The sigmoid input stays on a bounded,
interpretable scale throughout the simulation.

---

## 2. Snapshot Comparisons (T = 50 / 100 / 150 / 200)

### Ideological Drift

```{r plot-drift, fig.width = 10, fig.height = 5}
combined %>%
  ggplot(aes(x = responsiveness, y = avg_drift, color = formula)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale + col_scale +
  labs(title    = "Average Ideological Drift at T = 50 / 100 / 150 / 200",
       subtitle = "Averaged over 1 000 seeds — absolute vs. log-ratio power",
       x = "Responsiveness to Power", y = "Avg. Ideological Drift",
       color = NULL) +
  theme_minimal() + theme(legend.position = "bottom")
```

### Lost Vote Rate

```{r plot-lost, fig.width = 10, fig.height = 5}
combined %>%
  ggplot(aes(x = responsiveness, y = lost_vote_rate, color = formula)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale + col_scale +
  labs(title    = "Lost Vote Rate at T = 50 / 100 / 150 / 200",
       subtitle = "Averaged over 1 000 seeds — absolute vs. log-ratio power",
       x = "Responsiveness to Power", y = "Lost Vote Rate",
       color = NULL) +
  theme_minimal() + theme(legend.position = "bottom")
```

### Delegation Rate

```{r plot-delegation, fig.width = 10, fig.height = 5}
combined %>%
  ggplot(aes(x = responsiveness, y = delegation_rate, color = formula)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale + col_scale +
  labs(title    = "Delegation Rate at T = 50 / 100 / 150 / 200",
       subtitle = "Averaged over 1 000 seeds — absolute vs. log-ratio power",
       x = "Responsiveness to Power", y = "Share of Lay Agents Delegating",
       color = NULL) +
  theme_minimal() + theme(legend.position = "bottom")
```

### Power Concentration — Gini

```{r plot-gini, fig.width = 10, fig.height = 5}
combined %>%
  ggplot(aes(x = responsiveness, y = gini_power, color = formula)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale + col_scale +
  labs(title    = "Gini Coefficient of Power at T = 50 / 100 / 150 / 200",
       subtitle = "Averaged over 1 000 seeds — absolute vs. log-ratio power",
       x = "Responsiveness to Power", y = "Gini (Power)",
       color = NULL) +
  theme_minimal() + theme(legend.position = "bottom")
```

### Power Concentration — Top-5% Share

```{r plot-top5, fig.width = 10, fig.height = 5}
combined %>%
  ggplot(aes(x = responsiveness, y = top5_power_share, color = formula)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale + col_scale +
  labs(title    = "Top-5% Power Share at T = 50 / 100 / 150 / 200",
       subtitle = "Averaged over 1 000 seeds — absolute vs. log-ratio power",
       x = "Responsiveness to Power", y = "Top-5% Power Share",
       color = NULL) +
  theme_minimal() + theme(legend.position = "bottom")
```


## 3. Time Series Comparisons (all 200 rounds)

### Ideological Drift over Time

```{r plot-drift-time, fig.width = 10, fig.height = 5}
combined_time %>%
  ggplot(aes(x = round, y = avg_drift, color = formula)) +
  geom_line(alpha = 0.8) +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 4) +
  col_scale +
  labs(title  = "Average Ideological Drift over Time",
       x = "Round", y = "Avg. Ideological Drift", color = NULL) +
  theme_minimal() + theme(legend.position = "bottom")
```

### Delegation Rate over Time

```{r plot-delegation-time, fig.width = 10, fig.height = 5}
combined_time %>%
  ggplot(aes(x = round, y = delegation_rate, color = formula)) +
  geom_line(alpha = 0.8) +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 4) +
  col_scale +
  labs(title  = "Delegation Rate over Time",
       x = "Round", y = "Share of Lay Agents Delegating", color = NULL) +
  theme_minimal() + theme(legend.position = "bottom")
```

### Lost Vote Rate over Time

```{r plot-lost-time, fig.width = 10, fig.height = 5}
combined_time %>%
  ggplot(aes(x = round, y = lost_vote_rate, color = formula)) +
  geom_line(alpha = 0.8) +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 4) +
  col_scale +
  labs(title  = "Lost Vote Rate over Time",
       x = "Round", y = "Lost Vote Rate", color = NULL) +
  theme_minimal() + theme(legend.position = "bottom")
```

