---
title: "Weekly Report — Week 11 (01.05.2026 – 07.05.2026)"
output: github_document
date: "2026-05-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(furrr)

sigmoid <- function(x) 1 / (1 + exp(-x))
```

## 1. Formulas

### Neighbour-Weight (Attractiveness)

$$A_{ij} = \sigma\!\left(r_{op} \cdot \bigl(1 -2\,|o_i - o_j|\bigr)\right) \cdot \sigma\!\left(r_{pw} \cdot \log\frac{p_j}{p_i}\right)$$

Defining the opinion similarity $s_{ij} = 1-2|o_i-o_j| \in [-1,1]$, this is
equivalent to $A_{ij} = \sigma(-r_{op}\,s_{ij})\cdot\sigma(r_{pw}\log\tfrac{p_j}{p_i})$.

### Self-Weight

$$j^* = \arg\max_{j \in N(i)} A_{ij}, \qquad w_{\text{self}}(i) = \sigma\!\left(r_{op} \cdot \bigl(1 - 2\,|o_i - o_{j^*}|\bigr)\right) \cdot \sigma\!\left(r_{pw} \cdot \log\frac{p_{j^*}}{p_i}\right)$$

Equivalently, $w_{\text{self}}(i) = \sigma(r_{op}\,s_{ij^*})\cdot\sigma(r_{pw}\log\tfrac{p_{j^*}}{p_i})$.
agent's own position.

---

## 2. Weight Surfaces (1-D Slices)

```{r plot-1d-power, fig.width=14, fig.height=9, dpi=300}
expand_grid(
  similarity      = seq(-1, 1, length.out = 200),
  log_power_ratio = c(-2, 0, 2),
  weight_type     = c("Neighbour-Weight", "Self-Weight"),
  r_op            = c(1, 2),
  r_pw            = c(1, 2)
) |>
  mutate(
    A_ij   = sigmoid( r_op * similarity) * sigmoid(r_pw * log_power_ratio),
    w_self = sigmoid( r_op * similarity) * sigmoid(r_pw * log_power_ratio),
    value  = if_else(weight_type == "Neighbour-Weight", A_ij, w_self),
    lpr_label = paste0("log(p[j]/p[i])==", log_power_ratio),
    param     = paste0("r[op]==", r_op, "~~r[pw]==", r_pw)
  ) |>
  ggplot(aes(similarity, value, colour = lpr_label)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0.5, linetype = "dotted", colour = "grey60") +
  facet_grid(weight_type ~ param,
             labeller = labeller(weight_type = label_value, param = label_parsed)) +
  scale_colour_brewer(
    palette = "Dark2",
    name    = expression(log(p[j] / p[i])),
    labels  = scales::label_parse()
  ) +
  labs(
    x     = expression(s[ij] == 1 - 2 * abs(o[i] - o[j])),
    y     = "Weight",
    title = "Weight vs. opinion similarity",
    subtitle = expression("Neighbour-Weight: " * A[ij] * "   |   Self-Weight: " * w[self](i))
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

## 3. Weight Surfaces (2-D Heatmaps)

```{r plot-2d-heatmap, fig.width=9, fig.height=4}
expand_grid(
  similarity      = seq(-1, 1,  length.out = 150),
  log_power_ratio = seq(-3, 3, length.out = 150),
  weight_type     = c("Neighbour-Weight", "Self-Weight"),
  r_op            = c(1, 2),
  r_pw            = c(1, 2)
) |>
  mutate(
    A_ij   = sigmoid( r_op * similarity) * sigmoid(r_pw * log_power_ratio),
    w_self = sigmoid( r_op * similarity) * sigmoid(r_pw * log_power_ratio),
    V      = if_else(weight_type == "Neighbour-Weight", A_ij, w_self),
    param  = paste0("r[op]==", r_op, "~~r[pw]==", r_pw)
  ) |>
  ggplot(aes(similarity, log_power_ratio, fill = V)) +
  geom_tile() +
  facet_grid(weight_type ~ param,
             labeller = labeller(weight_type = label_value, param = label_parsed)) +
  scale_fill_gradient(
  low  = "steelblue",
  high = "firebrick",
  name = "Weight"
  ) +
  labs(
    x     = expression(s[ij] == 1 - 2 * abs(o[i] - o[j])),
    y     = expression(log(p[j] / p[i])),
    title = "Weight surfaces",
    subtitle = expression(
      "Neighbour-Weight: " * A[ij] * "   |   Self-Weight: " * w[self](i)
    )
  ) +
  theme_minimal(base_size = 12)
```

## 4. Simulation — Delegation Rate & Lost Vote Rate

The simulation runs `simulate_liquid_democracy` from `Network.R` using the
Neighbour-Weight and Self-Weight formulas across a $3\times3$ grid of
$r_{op},r_{pw} \in \{0.5,\,1,\,2\}$ ($n=250$ agents, $T=100$ rounds,
averaged over 100 independent runs).

```{r sim-setup, include=FALSE}
source(here::here("scripts/Network.R"))

# Neighbour-Weight: A_ij = σ(r·(2|o_i−o_j|−1)) · σ(r·log(p_j/p_i))
#                        = σ(−r·s_ij)           · σ(r·log(p_j/p_i))
attractiveness <- function(op_i, op_j, pow_i, pow_j, r_op, r_pw) {
  sigmoid(r_op * (1 - 2 * abs(op_i - op_j))) *
  sigmoid(r_pw * log(pow_j / pow_i))
}

selfweight <- function(i, nb, op, pow, r_op, r_pw) {
  if (!length(nb)) return(0.5)
  
  a_nb <- attractiveness(op[i], op[nb], pow[i], pow[nb], r_op, r_pw)
  j_star <- nb[which.max(a_nb)]
  
  sigmoid(r_op * (1 - 2 * abs(op[i] - op[j_star]))) *
  sigmoid(r_pw * log(pow[j_star] / pow[i]))
}
```

```{r simulation, fig.width=10, fig.height=6, cache=TRUE}
plan(multisession, workers = parallel::detectCores() - 1)

full_grid <- expand_grid(
  r_op = c(0.5, 1, 2),
  r_pw = c(0.5, 1, 2),
  run  = seq_len(100)
)

sim_raw <- future_pmap_dfr(full_grid, function(r_op, r_pw, run) {

  res <- simulate_liquid_democracy(
    seed            = run,
    n_per_community = 250,
    n_communities   = 1,
    node_degree     = 6,
    T               = 100,
    responsiveness  = 1,

    attractiveness_fn = function(op_i, op_j, pow_i, pow_j, responsiveness) {
      sigmoid(r_op * (1 - 2 * abs(op_i - op_j))) *
      sigmoid(r_pw * log(pow_j / pow_i))
    },

    selfweight_fn = function(i, nb, op, pow, responsiveness) {
      if (!length(nb)) return(0.5)
      a_nb   <- sigmoid(r_op * (1 - 2 * abs(op[i] - op[nb]))) *
                sigmoid(r_pw * log(pow[nb] / pow[i]))
      j_star <- nb[which.max(a_nb)]
      sigmoid(r_op * (1 - 2 * abs(op[i] - op[j_star]))) *
      sigmoid(r_pw * log(pow[j_star] / pow[i]))
    }
  )

  tibble(
    r_op            = r_op,
    r_pw            = r_pw,
    run             = run,
    round           = seq_len(length(res$history_delegation)),
    delegation_rate = res$history_delegation,
    lost_vote_rate  = res$history_lost
  )
}, .options = furrr_options(seed = NULL))

plan(sequential)

# Average across runs
sim_results <- sim_raw |>
  group_by(r_op, r_pw, round) |>
  summarise(
    delegation_rate = mean(delegation_rate),
    lost_vote_rate  = mean(lost_vote_rate),
    .groups         = "drop"
  )

k <- 6  # node_degree used in simulate_liquid_democracy

# Theoretical bounds, scoped to the correct facet rows
r_op_labels <- paste0("r[op]==", c(0.5, 1, 2))

bounds <- bind_rows(
  # Delegation rate: lower = 1/2, upper = k/(k+1)
  tibble(metric = "Delegation rate", r_op_label = r_op_labels,
         yint = 1/2,       lty = "lower"),
  tibble(metric = "Delegation rate", r_op_label = r_op_labels,
         yint = k / (k+1), lty = "upper"),
  # Lost vote rate: upper = k/(k+1) only
  tibble(metric = "Lost vote rate",  r_op_label = r_op_labels,
         yint = k / (k+1), lty = "upper")
)

sim_results |>
  pivot_longer(c(delegation_rate, lost_vote_rate),
               names_to = "metric", values_to = "value") |>
  mutate(
    metric     = recode(metric,
      delegation_rate = "Delegation rate",
      lost_vote_rate  = "Lost vote rate"
    ),
    r_pw_label = paste0("r[pw]==", r_pw),
    r_op_label = paste0("r[op]==", r_op)
  ) |>
  ggplot(aes(round, value, colour = r_pw_label)) +
  geom_hline(data = filter(bounds, lty == "lower"),
             aes(yintercept = yint),
             linetype = "dotted", colour = "grey40", linewidth = 0.5) +
  geom_hline(data = filter(bounds, lty == "upper"),
             aes(yintercept = yint),
             linetype = "dashed", colour = "grey40", linewidth = 0.5) +
  geom_line(linewidth = 0.8, alpha = 0.9) +
  facet_grid(metric ~ r_op_label,
             scales   = "free_y",
             labeller = labeller(metric = label_value, r_op_label = label_parsed)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_colour_brewer(
    palette = "Set2",
    name    = expression(r[pw]),
    labels  = scales::label_parse()
  ) +
  labs(
    x       = "Round",
    y       = NULL,
    title   = "Delegation rate and lost vote rate over 100 rounds",
    caption = "Dashed: upper bound k/(k+1) ≈ 85.7 %;  dotted: lower bound 1/2 (delegation rate only)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        plot.caption    = element_text(size = 8, colour = "grey50"))
```

```{r simulation-summary}
sim_results |>
  group_by(r_op, r_pw) |>
  summarise(
    del_mean  = mean(delegation_rate),
    del_sd    = sd(delegation_rate),
    lost_mean = mean(lost_vote_rate),
    lost_sd   = sd(lost_vote_rate),
    .groups   = "drop"
  ) |>
  knitr::kable(
    digits    = 3,
    col.names = c(
      "r_op", "r_pw",
      "Del. mean", "Del. SD",
      "Lost mean", "Lost SD"
    )
  )
```