---
title: "Weekly Report — Week 5 (20.03.2026 – 26.03.2026)"
output: github_document
date: "2026-03-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())

library(here)
library(dplyr)
library(tidyr)
library(ggplot2)

saved             <- readRDS(here::here("simulations/output/results_sim3_rewire_0.1.rds"))
results_time      <- saved$results_time
results_snapshots <- saved$results_snapshots

x_scale <- scale_x_log10(
  breaks = c(0.1, 0.5, 1, 2, 5, 10, 25),
  labels = c("0.1", "0.5", "1", "2", "5", "10", "25")
)
```

## Summary

- 100-seed averaging framework, snapshot tracking at T = 50/100/150/200,
  per-round history for cheap metrics, parallel execution via `furrr`
- Added `inertia` parameter (default 0 = memoryless baseline)
- Experiment 3: effect of responsiveness to power, p\_rewire = 0.10,
  100 seeds, T = 200

---

## Experiment Design

| Parameter | Values |
|---|---|
| Responsiveness | 0.1, 0.5, 1, 2, 5, 10, 25 |
| Rewiring (p) | 0.10 |
| Seeds | 1 – 100 |
| Agents | 250 lay, 1 community |
| T | 200 (snapshots at 50, 100, 150, 200) |
| Inertia | 0 |

---

## Results

### Ideological Drift

Drift decreases monotonically with responsiveness and stabilises early — results are nearly identical across all four snapshots, indicating the system reaches its stochastic equilibrium well before T = 50.

```{r plot-drift, fig.width = 10, fig.height = 5}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = avg_drift)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  labs(title    = "Average Ideological Drift at T = 50 / 100 / 150 / 200",
       subtitle = "Averaged over 100 seeds",
       x = "Responsiveness to Power", y = "Avg. Ideological Drift") +
  theme_minimal()
```

### Lost Vote Rate

Vote loss drops sharply as responsiveness increases, driven by lower delegation rates reducing the number of mutual cycles that can form.

```{r plot-lost, fig.width = 10, fig.height = 5}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = lost_vote_rate)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  labs(title    = "Lost Vote Rate at T = 50 / 100 / 150 / 200",
       subtitle = "Averaged over 100 seeds",
       x = "Responsiveness to Power", y = "Lost Vote Rate") +
  theme_minimal()
```

### Delegation Rate

Delegation rate decreases with responsiveness as agents with above-average local power increasingly prefer to vote directly, consistent with the self-weight formula.

```{r plot-delegation, fig.width = 10, fig.height = 5}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = delegation_rate)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  labs(title    = "Delegation Rate at T = 50 / 100 / 150 / 200",
       subtitle = "Averaged over 100 seeds",
       x = "Responsiveness to Power", y = "Share of Lay Agents Delegating") +
  theme_minimal()
```

### Delegation Stability over Time

Stability never reaches the 99% threshold under the memoryless baseline, confirming the system settles into a stochastic equilibrium rather than a fixed delegation structure.

```{r plot-stability, fig.width = 10, fig.height = 5}
results_time %>%
  filter(!is.na(delegation_stability)) %>%
  ggplot(aes(x = round, y = delegation_stability)) +
  geom_line(alpha = 0.8) +
  geom_hline(yintercept = 0.99, linetype = "dashed", color = "red") +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 4) +
  labs(title    = "Delegation Stability over Time",
       subtitle = "Red line = 99% convergence threshold; averaged over 100 seeds",
       x = "Round", y = "Share of Agents Not Switching Target") +
  theme_minimal()
```

### Majority Margins

With 100 seeds the Direct Democracy margin shows noticeable sampling noise around zero (SE ≈ 1.58), while the Liquid Democracy margin tracks closely but varies with responsiveness.

```{r plot-margins, fig.width = 10, fig.height = 5}
results_snapshots %>%
  tidyr::pivot_longer(cols = c(liquid_margin, direct_margin),
                      names_to = "system", values_to = "margin") %>%
  mutate(system = recode(system,
    "liquid_margin" = "Liquid Democracy",
    "direct_margin" = "Direct Democracy")) %>%
  ggplot(aes(x = responsiveness, y = margin, color = system)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  labs(title    = "Majority Margins: Liquid vs. Direct Democracy",
       subtitle = "Averaged over 100 seeds (p_rewire = 0.10)",
       x = "Responsiveness to Power", y = "Majority Margin (Yes - No)",
       color = "System") +
  theme_minimal()
```

---

## Open Issues
