---
title: "Report 16 -- Systematic Hypothesis Plots"
output: github_document
date: "2026-06-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 14, fig.height = 10)
library(tidyverse)
library(here)
library(furrr)
library(patchwork)
library(digest)
library(knitr)
library(segmented)

theme_set(theme_minimal(base_size = 11))
source(here::here("scripts/Network.R"))

# ── Fixed simulation parameters ───────────────────────────────────────────────
N_SIM      <- 500L
K_SIM      <- 14L
P_REWIRE   <- 0.05
T_SIM      <- 20L
N_SEEDS    <- 50L
LAMBDA     <- 0.5
R_VALS     <- c(0, 1, 2, 4, 8, 16, 32)   # RQ1 only
R_TRUST    <- c(1, 2, 4, 8)             # r_op = r_pw levels shown in trust panels
GAMMA_VALS <- c(0, 1, 2, 4, 8, 16)
R_VALS_FINE <- c(0, 0.5, 1, 1.5, 2, 3, 4, 8)       # RQ2/3/4 fine grid
snap_idx   <- (T_SIM - 4L):T_SIM   # last 5 rounds: 16,17,18,19,20

# ── New metric: share of agents needed to jointly hold 50% of total power ─────
share_for_50pct <- function(power) {
  pv <- sort(power[power > 0], decreasing = TRUE)
  if (!length(pv)) return(NA_real_)
  k <- which(cumsum(pv) / sum(pv) >= 0.5)[1]
  k / length(pv)
}

# ── Outcome transform (logit for rates, log for positive-skewed) ──────────────
transform_y <- function(x, met) {
  if (met %in% c("delegation_rate", "lost_vote_rate", "top5_share", "share_50pct"))
    qlogis(pmin(pmax(x, 0.001), 0.999))
  else
    log(pmax(x, 1e-6))
}

# ── Metric catalogue ──────────────────────────────────────────────────────────
metric_labels <- c(
  delegation_rate           = "Delegation Rate",
  lost_vote_rate            = "Lost Vote Rate",
  effective_delegation_rate = "Effective Delegation Rate",
  share_50pct               = "Agents for 50% Votes",
  top5_share                = "Top-5% Power Share",
  avg_ideo_drift            = "Avg. Ideological Drift",
  avg_chain_len             = "Avg. Chain Length",
  total_components          = "Components"
)
ALL_METRICS <- names(metric_labels)

COND_LEVELS <- c(
  "A: r_op only",
  "B: r_pw only",
  "C: r_op x r_pw",
  "C: r_op x r_pw  (grid)",
  "D: trust (r=1)",
  "D: trust (r=2)",
  "D: trust (r=4)",
  "D: trust (r=8)"
)

# Metrics that remain valid regardless of lost vote rate
ALWAYS_VALID  <- c("delegation_rate", "lost_vote_rate", "effective_delegation_rate")
# Metrics set to NA when too many votes are lost (< 10% cast)
LOST_THRESHOLD <- 0.90
MASK_METRICS  <- setdiff(ALL_METRICS, ALWAYS_VALID)

RQ1_METRICS <- c("delegation_rate", "lost_vote_rate", "effective_delegation_rate", "avg_chain_len", "total_components")
RQ2_METRICS <- c("avg_chain_len", "avg_ideo_drift", "lost_vote_rate", "delegation_rate")
RQ3_METRICS <- c("top5_share", "total_components")
RQ4_METRICS <- c("share_50pct")
```

---
title: "Weekly Report -- Week 16: Research Questions + Hypotheses"
output: github_document
date: "2026-05-24"
---
# Model Parameters:

- Watts-Strogatz (Small-World) Network
- N = 500
- K = 14
- T = 20
- N_Seeds = 50

---

# 1  Research Questions

## Part 1: Core Network dynamics

**RQ1**: When are the effects of responsiveness parameters saturating?

- H1a:  The relationship between opinion responsiveness (r_op) and delegation network metrics is non-linear and saturates at a certain point.

- H2a: The relationship between power responsiveness (r_pw) and delegation network metrics is non-linear and saturates at a certain point.

- H3a: When both r_op and r_pw increase jointly (r_op = r_pw), the combined effect on delegation network metrics is non-linear and saturates at a certain point.

- H4a: The relationship between trust sensitivity (γ) and delegation network metrics is non-linear and saturates at a certain point.


**RQ2**: How far drifts the average opinion of an agent when he delegates?

- H1b: Increasing resp. to opinion increases the average ideological drift.

- H2b: Increasing resp. to power increases the average ideological drift.

- H3b Increasing resp- to power and resp to opinion increases the average ideological drift.

- H4b: Increasing resp. to trust reduces the influence of the effect of resp. to opinion and power on the average ideological drift.


**RQ3**: How much power do the most powerful 5% agent within a network have?

- H1c: Increasing resp. to opinion increases the top 5% share.

- H2c: Increasing resp. to power increases the top 5% share.

- H3c: Increasing resp. to opinion and power increases the the top 5% share.

- H4c: Increasing resp. to trust has a reducing influence of the effect of resp. to opinion and power.


**RQ4**: How many of the most influential agents do you need to get 50% of the votes? 

- H1d: Increasing resp. to opinion decreases the share of agents you need to get 50% of the vote.

- H2d: Increasing resp. to power decreases the share of agents you need to get 50% of the vote.

- H3d: Increasing resp. to opinion and power decreases the share of agents you need to get 50% votes.

- H4d: Increasing resp. to trust has a reducing influence of the effect of resp. to opinion and power.

---

```{r run-one, include=FALSE}
run_one <- function(seed, r_op, r_pw, gamma, lambda = 0) {
  res <- simulate_liquid_democracy(
    seed            = seed,
    n_per_community = N_SIM,
    n_communities   = 1,
    node_degree     = K_SIM,
    T               = T_SIM,
    r_op            = r_op,
    r_pw            = r_pw,
    p_self          = NULL,
    p_rewire        = P_REWIRE,
    gamma           = gamma,
    lambda          = lambda
  )
  # last-5-round snapshots for steady-state metric averages
  snaps <- res$snapshots |> dplyr::filter(round >= T_SIM - 4L)
  tibble(
    seed             = seed,
    r_op             = r_op,
    r_pw             = r_pw,
    gamma            = gamma,
    delegation_rate           = mean(res$history_delegation[snap_idx], na.rm = TRUE),
    lost_vote_rate            = mean(res$history_lost[snap_idx],       na.rm = TRUE),
    effective_delegation_rate = mean(res$history_delegation[snap_idx], na.rm = TRUE) -
                                mean(res$history_lost[snap_idx],       na.rm = TRUE),
    avg_ideo_drift   = {
      d <- res$history_drift[snap_idx]
      if (mean(is.na(d)) > 0.5) NA_real_ else mean(d, na.rm = TRUE)
    },
    avg_chain_len    = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
    top5_share       = mean(snaps$top5_power_share,            na.rm = TRUE),
    total_components = mean(snaps$total_components,            na.rm = TRUE),
    share_50pct      = mean(snaps$share_50pct,                 na.rm = TRUE)
  )
}
```

```{r simulations, include=FALSE}
seeds <- seq_len(N_SEEDS)

grid_rop <- expand_grid(
  r_op = R_VALS, r_pw = 0, gamma = 0, seed = seeds,
  condition = "A: r_op only"
)
grid_rpw <- expand_grid(
  r_op = 0, r_pw = R_VALS, gamma = 0, seed = seeds,
  condition = "B: r_pw only"
)
grid_diag <- expand_grid(
  r_op = R_VALS, r_pw = R_VALS, gamma = 0, seed = seeds,
  condition = "C: r_op x r_pw  (grid)"
)

grid_trust <- bind_rows(
  expand_grid(r_op = 1, r_pw = 1, gamma = GAMMA_VALS, seed = seeds,
              condition = "D: trust (r=1)"),
  expand_grid(r_op = 2, r_pw = 2, gamma = GAMMA_VALS, seed = seeds,
              condition = "D: trust (r=2)"),
  expand_grid(r_op = 4, r_pw = 4, gamma = GAMMA_VALS, seed = seeds,
              condition = "D: trust (r=4)"),
  expand_grid(r_op = 8, r_pw = 8, gamma = GAMMA_VALS, seed = seeds,
              condition = "D: trust (r=8)")
)

# ── Cache ─────────────────────────────────────────────────────────────────────
params_16 <- list(
  N = N_SIM, K = K_SIM, p_rewire = P_REWIRE,
  T = T_SIM, n_seeds = N_SEEDS, r_vals = R_VALS,
  gamma_vals = GAMMA_VALS, r_trust = R_TRUST, lambda = LAMBDA, version = 12L
)
hash_16  <- digest::digest(params_16)
cache_16 <- here::here("data/sim_report16_cache.rds")

if (file.exists(cache_16)) {
  c16 <- readRDS(cache_16)
  if (identical(c16$hash, hash_16)) {
    sim_raw <- c16$data
    message("Report 16: cache hit.")
  } else { c16 <- NULL }
} else { c16 <- NULL }

# Always defined so both simulation blocks can use it
run_cond <- function(grid_df, lambda_val = 0) {
  future_pmap_dfr(
    list(
      seed      = grid_df$seed,
      r_op      = grid_df$r_op,
      r_pw      = grid_df$r_pw,
      gamma     = grid_df$gamma,
      condition = grid_df$condition
    ),
    function(seed, r_op, r_pw, gamma, condition)
      run_one(seed, r_op, r_pw, gamma, lambda_val) |>
      dplyr::mutate(condition = condition),
    .options = furrr_options(seed = NULL)
  )
}

if (is.null(c16)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  message(sprintf("Report 16: starting simulations (%d rows total)...",
                  nrow(grid_rop) + nrow(grid_rpw) + nrow(grid_diag) + nrow(grid_trust)))
  t0 <- proc.time()
  plan(multisession, workers = parallel::detectCores() - 1)
  sim_raw <- bind_rows(
    run_cond(grid_rop),
    run_cond(grid_rpw),
    run_cond(grid_diag),
    run_cond(grid_trust, lambda_val = LAMBDA)
  )
  plan(sequential)
  elapsed <- proc.time() - t0
  saveRDS(list(hash = hash_16, data = sim_raw), cache_16)
  message(sprintf("Report 16: simulations complete — %.1f min (%.0f sec).",
                  elapsed["elapsed"] / 60, elapsed["elapsed"]))
}
```

```{r simulations-fine, include=FALSE}
# Fine-grid simulations for RQ2/3/4: r = 0, 0.5, 1, ..., 3
seeds_f <- seq_len(N_SEEDS)

grid_rop_f <- expand_grid(
  r_op = R_VALS_FINE, r_pw = 0, gamma = 0, seed = seeds_f,
  condition = "A: r_op only"
)
grid_rpw_f <- expand_grid(
  r_op = 0, r_pw = R_VALS_FINE, gamma = 0, seed = seeds_f,
  condition = "B: r_pw only"
)
grid_diag_f <- expand_grid(
  r_op = R_VALS_FINE, r_pw = R_VALS_FINE, gamma = 0, seed = seeds_f,
  condition = "C: r_op x r_pw  (grid)"
)
grid_trust_f <- bind_rows(
  expand_grid(r_op = 1, r_pw = 1, gamma = GAMMA_VALS, seed = seeds_f,
              condition = "D: trust (r=1)"),
  expand_grid(r_op = 2, r_pw = 2, gamma = GAMMA_VALS, seed = seeds_f,
              condition = "D: trust (r=2)"),
  expand_grid(r_op = 4, r_pw = 4, gamma = GAMMA_VALS, seed = seeds_f,
              condition = "D: trust (r=4)"),
  expand_grid(r_op = 8, r_pw = 8, gamma = GAMMA_VALS, seed = seeds_f,
              condition = "D: trust (r=8)")
)

params_fine <- list(
  N = N_SIM, K = K_SIM, p_rewire = P_REWIRE,
  T = T_SIM, n_seeds = N_SEEDS, r_vals = R_VALS_FINE,
  gamma_vals = GAMMA_VALS, lambda = LAMBDA, version = 8L
)
hash_fine  <- digest::digest(params_fine)
cache_fine <- here::here("data/sim_report16_fine_cache.rds")

if (file.exists(cache_fine)) {
  cf <- readRDS(cache_fine)
  if (identical(cf$hash, hash_fine)) {
    sim_raw_fine <- cf$data
    message("Report 16 fine: cache hit.")
  } else { cf <- NULL }
} else { cf <- NULL }

if (is.null(cf)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  message(sprintf("Report 16 fine: starting simulations (%d rows total)...",
                  nrow(grid_rop_f) + nrow(grid_rpw_f) + nrow(grid_diag_f) + nrow(grid_trust_f)))
  t0_f <- proc.time()
  plan(multisession, workers = parallel::detectCores() - 1)
  sim_raw_fine <- bind_rows(
    run_cond(grid_rop_f),
    run_cond(grid_rpw_f),
    run_cond(grid_diag_f),
    run_cond(grid_trust_f, lambda_val = LAMBDA)
  )
  plan(sequential)
  elapsed_f <- proc.time() - t0_f
  saveRDS(list(hash = hash_fine, data = sim_raw_fine), cache_fine)
  message(sprintf("Report 16 fine: simulations complete — %.1f min (%.0f sec).",
                  elapsed_f["elapsed"] / 60, elapsed_f["elapsed"]))
}
```

```{r aggregate, include=FALSE}
# Derive diagonal condition (r_op = r_pw) from the full grid without extra sims
sim_diag <- sim_raw |>
  filter(condition == "C: r_op x r_pw  (grid)", r_op == r_pw) |>
  mutate(condition = "C: r_op x r_pw")

sim_all <- bind_rows(sim_raw, sim_diag)

sim_agg <- sim_all |>
  group_by(condition, r_op, r_pw, gamma) |>
  summarise(across(all_of(ALL_METRICS), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  mutate(
    param_val = case_when(
      condition == "A: r_op only"           ~ r_op,
      condition == "B: r_pw only"           ~ r_pw,
      condition == "C: r_op x r_pw"         ~ r_op,
      condition == "C: r_op x r_pw  (grid)" ~ r_op,
      startsWith(condition, "D: trust")     ~ gamma
    ),
    x_label = case_when(
      condition == "A: r_op only"           ~ "r_op   (r_pw = 0, gamma = 0)",
      condition == "B: r_pw only"           ~ "r_pw   (r_op = 0, gamma = 0)",
      condition == "C: r_op x r_pw"         ~ "r_op = r_pw   (gamma = 0)",
      condition == "C: r_op x r_pw  (grid)" ~ "r_op  (grid: r_pw also varies)",
      startsWith(condition, "D: trust")     ~ "gamma   (lambda = 0.5)"
    ),
    condition = factor(condition, levels = COND_LEVELS)
  )

# sim_long = full unfiltered data (used as grey background in plots)
sim_long <- sim_agg |>
  pivot_longer(all_of(ALL_METRICS), names_to = "metric", values_to = "value") |>
  mutate(metric_label = factor(metric_labels[metric], levels = unname(metric_labels)))

# sim_long_masked = primary analysis data: MASK_METRICS set to NA when LVR > 0.90
sim_agg_masked <- sim_agg |>
  mutate(across(all_of(MASK_METRICS),
                \(x) ifelse(lost_vote_rate > LOST_THRESHOLD, NA_real_, x)))

sim_long_masked <- sim_agg_masked |>
  pivot_longer(all_of(ALL_METRICS), names_to = "metric", values_to = "value") |>
  mutate(metric_label = factor(metric_labels[metric], levels = unname(metric_labels)))

# ── Fine-grid aggregate for RQ2/3/4 ──────────────────────────────────────────
sim_diag_fine <- sim_raw_fine |>
  filter(condition == "C: r_op x r_pw  (grid)", r_op == r_pw) |>
  mutate(condition = "C: r_op x r_pw")

sim_all_fine <- bind_rows(sim_raw_fine, sim_diag_fine)

sim_agg_fine <- sim_all_fine |>
  group_by(condition, r_op, r_pw, gamma) |>
  summarise(across(all_of(ALL_METRICS), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  mutate(
    param_val = case_when(
      condition == "A: r_op only"           ~ r_op,
      condition == "B: r_pw only"           ~ r_pw,
      condition == "C: r_op x r_pw"         ~ r_op,
      condition == "C: r_op x r_pw  (grid)" ~ r_op,
      startsWith(condition, "D: trust")     ~ gamma
    ),
    x_label = case_when(
      condition == "A: r_op only"           ~ "r_op   (r_pw = 0, gamma = 0)",
      condition == "B: r_pw only"           ~ "r_pw   (r_op = 0, gamma = 0)",
      condition == "C: r_op x r_pw"         ~ "r_op = r_pw   (gamma = 0)",
      condition == "C: r_op x r_pw  (grid)" ~ "r_op  (grid: r_pw also varies)",
      startsWith(condition, "D: trust")     ~ "gamma   (lambda = 0.5)"
    ),
    condition = factor(condition, levels = COND_LEVELS)
  )

sim_long_fine <- sim_agg_fine |>
  pivot_longer(all_of(ALL_METRICS), names_to = "metric", values_to = "value") |>
  mutate(metric_label = factor(metric_labels[metric], levels = unname(metric_labels)))

sim_agg_fine_masked <- sim_agg_fine |>
  mutate(across(all_of(MASK_METRICS),
                \(x) ifelse(lost_vote_rate > LOST_THRESHOLD, NA_real_, x)))

sim_long_fine_masked <- sim_agg_fine_masked |>
  pivot_longer(all_of(ALL_METRICS), names_to = "metric", values_to = "value") |>
  mutate(metric_label = factor(metric_labels[metric], levels = unname(metric_labels)))
```

```{r plot-fn, include=FALSE}
SUBTITLE <- sprintf(
  "N = %d  |  k = %d  |  p_rewire = %.2f  |  %d seeds  |  endogenous p_self",
  N_SIM, K_SIM, P_REWIRE, N_SEEDS)

make_sweep_plot <- function(df_long, cond, focus = NULL, bg_long = NULL) {
  df <- df_long |>
    filter(condition == cond) |>
    mutate(is_focus = if (!is.null(focus)) metric %in% focus else FALSE)

  x_lab <- unique(df$x_label)

  p <- ggplot()

  # Grey background = unfiltered data (when provided)
  if (!is.null(bg_long)) {
    bg <- bg_long |> filter(condition == cond)
    p <- p +
      geom_line(data = bg, aes(x = param_val, y = value, group = metric_label),
                colour = "firebrick3", alpha = 0.75, linewidth = 0.9, linetype = "dotted",
                na.rm = TRUE)
  }

  p +
    geom_line(data = df, aes(param_val, value, colour = is_focus),
              linewidth = 0.9, show.legend = FALSE, na.rm = TRUE) +
    geom_point(data = df, aes(param_val, value, colour = is_focus, fill = is_focus),
               shape = 21, size = 2.5, stroke = 0.5, show.legend = FALSE,
               na.rm = TRUE) +
    scale_colour_manual(values = c(`FALSE` = "steelblue4", `TRUE` = "firebrick3")) +
    scale_fill_manual(  values = c(`FALSE` = "steelblue2", `TRUE` = "firebrick1")) +
    facet_wrap(~ metric_label, scales = "free_y", ncol = 2) +
    scale_x_continuous(breaks = sort(unique(df$param_val))) +
    labs(title = as.character(cond), subtitle = SUBTITLE, x = x_lab, y = NULL) +
    theme(
      strip.background = element_rect(fill = "grey93", colour = NA),
      strip.text       = element_text(face = "bold", size = 9),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold", size = 11),
      plot.subtitle    = element_text(size = 8.5, colour = "grey40")
    )
}

make_heatmap_plot <- function(df_long) {
  COND_GRID <- "C: r_op x r_pw  (grid)"
  df <- df_long |>
    filter(condition == COND_GRID) |>
    group_by(metric, metric_label) |>
    mutate(value_norm = {
      rng <- range(value, na.rm = TRUE)
      if (diff(rng) < 1e-9) 0.5 else (value - rng[1]) / diff(rng)
    }) |>
    ungroup()

  ggplot(df, aes(factor(r_op), factor(r_pw), fill = value_norm)) +
    geom_tile(colour = "white", linewidth = 0.5) +
    geom_text(aes(label = round(value, 3)), size = 2.4, colour = "grey10") +
    scale_fill_viridis_c(option = "plasma", begin = 0.1, end = 0.95,
                         name = "norm.\nvalue") +
    facet_wrap(~ metric_label, ncol = 2) +
    labs(
      title    = "C: r_op × r_pw  (full grid)",
      subtitle = SUBTITLE,
      x = "r_op", y = "r_pw"
    ) +
    theme(
      strip.background = element_rect(fill = "grey93", colour = NA),
      strip.text       = element_text(face = "bold", size = 9),
      panel.grid       = element_blank(),
      plot.title       = element_text(face = "bold", size = 11),
      plot.subtitle    = element_text(size = 8.5, colour = "grey40"),
      legend.position  = "right"
    )
}

# Trust panel: gamma on x-axis, one line per r_op = r_pw level
make_trust_plot <- function(df_long, focus = NULL, bg_long = NULL) {
  trust_r <- as.character(sort(R_TRUST))                 # "1" "2" "4"
  lvls    <- paste0("r=", trust_r)                       # "r=1" "r=2" "r=4"

  df <- df_long |>
    filter(startsWith(as.character(condition), "D: trust")) |>
    mutate(
      r_level = str_extract(as.character(condition), "r=(\\d+\\.?\\d*)"),
      r_level = factor(r_level, levels = lvls),
      is_focus = if (!is.null(focus)) metric %in% focus else FALSE
    ) |>
    filter(!is.na(r_level))

  p <- ggplot()

  if (!is.null(bg_long)) {
    bg <- bg_long |>
      filter(startsWith(as.character(condition), "D: trust")) |>
      mutate(r_level = str_extract(as.character(condition), "r=(\\d+\\.?\\d*)"),
             r_level = factor(r_level, levels = lvls)) |>
      filter(!is.na(r_level))
    p <- p +
      geom_line(data = bg, aes(x = param_val, y = value, group = interaction(r_level, metric_label)),
                colour = "firebrick3", alpha = 0.75, linewidth = 0.9, linetype = "dotted", na.rm = TRUE)
  }

  p +
    geom_line(data = df, aes(param_val, value, colour = r_level, group = r_level),
              linewidth = 0.9, na.rm = TRUE) +
    geom_point(data = df, aes(param_val, value, colour = r_level, fill = r_level, group = r_level),
               shape = 21, size = 2.5, stroke = 0.5, na.rm = TRUE) +
    scale_colour_viridis_d(name = "r_op = r_pw", end = 0.85) +
    scale_fill_viridis_d(name = "r_op = r_pw", end = 0.85) +
    facet_wrap(~ metric_label, scales = "free_y", ncol = 2) +
    scale_x_continuous(breaks = sort(unique(df$param_val))) +
    labs(title = paste0("D: trust  (gamma, r_op = r_pw = ",
                        paste(trust_r, collapse = " / "), ")"),
         subtitle = SUBTITLE,
         x = "gamma   (lambda = 0.5)", y = NULL) +
    theme(
      strip.background = element_rect(fill = "grey93", colour = NA),
      strip.text       = element_text(face = "bold", size = 9),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold", size = 11),
      plot.subtitle    = element_text(size = 8.5, colour = "grey40"),
      legend.position  = "right"
    )
}

# Condition C line plot: r_op on x-axis, r_pw as coloured lines
make_grid_lineplot <- function(df_long, focus = NULL, bg_long = NULL) {
  COND_GRID <- "C: r_op x r_pw  (grid)"
  df <- df_long |>
    filter(condition == COND_GRID) |>
    mutate(r_pw_fac = factor(r_pw),
           is_focus = if (!is.null(focus)) metric %in% focus else FALSE)

  p <- ggplot()

  if (!is.null(bg_long)) {
    bg <- bg_long |> filter(condition == COND_GRID) |> mutate(r_pw_fac = factor(r_pw))
    p <- p +
      geom_line(data = bg, aes(x = r_op, y = value, group = interaction(r_pw_fac, metric_label)),
                colour = "firebrick3", alpha = 0.75, linewidth = 0.9, linetype = "dotted", na.rm = TRUE)
  }

  p +
    geom_line(data = df, aes(x = r_op, y = value,
                 colour = r_pw_fac, group = r_pw_fac),
              linewidth = 0.9, na.rm = TRUE) +
    geom_point(data = df, aes(x = r_op, y = value,
                 colour = r_pw_fac, group = r_pw_fac),
               size = 2.2, na.rm = TRUE) +
    scale_colour_viridis_d(name = "r_pw", end = 0.88) +
    facet_wrap(~ metric_label, scales = "free_y", ncol = 2) +
    scale_x_continuous(breaks = sort(unique(df$r_op))) +
    labs(
      title    = "C: r_op × r_pw  (grid) — r_pw as lines",
      subtitle = SUBTITLE,
      x = "r_op", y = NULL
    ) +
    theme(
      strip.background = element_rect(fill = "grey93", colour = NA),
      strip.text       = element_text(face = "bold", size = 9),
      panel.grid.minor = element_blank(),
      plot.title       = element_text(face = "bold", size = 11),
      plot.subtitle    = element_text(size = 8.5, colour = "grey40"),
      legend.position  = "right"
    )
}
```

---

# RQ1 — Saturation of Responsiveness Effects

**When do the effects of responsiveness parameters on delegation network
metrics saturate?**

Each responsiveness parameter is varied across r ∈ {0, 1, 2, 4, 8, 16, 32}.
Condition A isolates r\_op (H1a); Condition B isolates r\_pw (H2a); Condition C shows the
full r\_op × r\_pw grid with r\_op on the x-axis and one coloured line per r\_pw level
(H3a — joint effect); Condition D varies gamma at four background r levels (H4a — trust saturation).
The **dashed red line** shows the full unfiltered curve; the **solid coloured line** shows
only the reliable range (lost vote rate ≤ `r LOST_THRESHOLD`).

```{r rq1-plot, fig.width=14, fig.height=70}
df_rq1    <- sim_long_masked |> filter(metric %in% ALL_METRICS)
df_rq1_bg <- sim_long        |> filter(metric %in% ALL_METRICS)

p_rq1_A    <- make_sweep_plot(df_rq1, "A: r_op only",  bg_long = df_rq1_bg)
p_rq1_B    <- make_sweep_plot(df_rq1, "B: r_pw only",  bg_long = df_rq1_bg)
p_rq1_grid <- make_grid_lineplot(df_rq1,               bg_long = df_rq1_bg)
p_rq1_D    <- make_trust_plot(df_rq1,                  bg_long = df_rq1_bg)

p_rq1_A / p_rq1_B / p_rq1_grid / p_rq1_D +
  plot_annotation(
    title = "RQ1 — Saturation of Responsiveness Effects",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  )
```

## RQ1 — GAM Saturation Analysis: Delegation Rate and Lost Vote Rate

### Statistical model: Generalised Additive Model (GAM)

A **Generalised Additive Model** replaces the straight line of OLS regression
with a flexible smooth curve $f(\cdot)$ estimated directly from the data,
without assuming a specific shape such as linear or exponential.
The smoothness of $f$ is controlled automatically by a penalty on curvature
(REML), preventing overfitting while allowing the curve to capture rises,
peaks, and plateaus.

For each metric $Y$ and each condition a separate GAM is fitted:

$$\boxed{Y_i = f\bigl(\log_2 r_i\bigr) + \varepsilon_i,
\qquad \varepsilon_i \sim \mathcal{N}(0, \sigma^2)}$$

- $Y_i$: metric value at parameter setting $r_i$, averaged over 50 seeds
- $f(\cdot)$: unknown smooth function — shape estimated from data, not assumed
- $\log_2 r$: the predictor is **log₂-transformed** because the simulation sweeps
  $r$ over doublings. On the original scale the steps are unequal; on the log₂
  scale they are equidistant, so each step carries equal information.
  One unit on the log₂ scale = one doubling of $r$. So the log transformation 
  makes equal multiplicative changes correspond to equal distances.
- $\varepsilon_i \sim \mathcal{N}(0,\sigma^2)$: residual error, assumed normally
  distributed with constant variance

The same GAM structure is used throughout RQ1–RQ4. Condition-specific model
specifications (1-D sweep, tensor-product interaction, trust moderation) are
described in the respective sections.

Both delegation rate and lost vote rate are monotonically increasing with r
and bounded in [0, 1]. The GAM on log(r) identifies where the slope
drops to near-zero — the practical saturation point r\* beyond which
further increases in r add negligible additional delegation or vote loss.

```{r rq1-sat-gam, include=FALSE}
library(mgcv)

RQ1_SAT_METRICS <- c("delegation_rate", "lost_vote_rate")
RQ1_SAT_CONDS   <- c("A: r_op only", "B: r_pw only")

rq1_sat_data <- sim_agg |>
  filter(condition %in% RQ1_SAT_CONDS, param_val > 0) |>
  mutate(log2_r = log2(param_val))

r_rq1_pos  <- R_VALS[R_VALS > 0]
r_rq1_seq  <- 2^seq(log2(min(r_rq1_pos)), log2(max(r_rq1_pos)), length.out = 300)
l2_rq1_seq <- log2(r_rq1_seq)

# ── Fit GAMs ──────────────────────────────────────────────────────────────────
rq1_sat_fits <- purrr::map(RQ1_SAT_METRICS, function(met) {
  purrr::map(RQ1_SAT_CONDS, function(cond) {
    d <- rq1_sat_data |> filter(condition == cond)
    if (sum(!is.na(d[[met]])) < 5) return(NULL)
    tryCatch(
      mgcv::gam(as.formula(paste0(met, " ~ s(log2_r, k = 6)")),
                data = d, method = "REML"),
      error = \(e) NULL
    )
  }) |> setNames(RQ1_SAT_CONDS)
}) |> setNames(RQ1_SAT_METRICS)

# ── Predictions ───────────────────────────────────────────────────────────────
rq1_sat_pred <- purrr::map_dfr(RQ1_SAT_METRICS, function(met) {
  purrr::map_dfr(RQ1_SAT_CONDS, function(cond) {
    fit <- rq1_sat_fits[[met]][[cond]]
    if (is.null(fit)) return(NULL)
    p <- predict(fit, newdata = data.frame(log2_r = l2_rq1_seq), se.fit = TRUE)
    tibble(r = r_rq1_seq, fit = p$fit,
           lwr = p$fit - 1.96 * p$se.fit,
           upr = p$fit + 1.96 * p$se.fit,
           metric = met, metric_label = metric_labels[met], condition = cond)
  })
})

# ── Derivatives + r* ─────────────────────────────────────────────────────────
get_deriv_rq1sat <- function(gfit) {
  h  <- 1e-5
  tryCatch({
    Xp  <- predict(gfit, newdata = data.frame(log2_r = l2_rq1_seq + h), type = "lpmatrix")
    Xm  <- predict(gfit, newdata = data.frame(log2_r = l2_rq1_seq - h), type = "lpmatrix")
    Xd  <- (Xp - Xm) / (2 * h)
    est <- as.vector(Xd %*% coef(gfit))
    se  <- sqrt(rowSums((Xd %*% vcov(gfit)) * Xd))
    tibble(r = r_rq1_seq, derivative = est,
           lower = est - 1.96 * se, upper = est + 1.96 * se)
  }, error = \(e) NULL)
}

rq1_sat_derivs <- purrr::map_dfr(RQ1_SAT_METRICS, function(met) {
  purrr::map_dfr(RQ1_SAT_CONDS, function(cond) {
    dv <- get_deriv_rq1sat(rq1_sat_fits[[met]][[cond]])
    if (is.null(dv)) return(NULL)
    dv |> mutate(metric = met, metric_label = metric_labels[met], condition = cond)
  })
})

rq1_sat_rstar <- rq1_sat_derivs |>
  group_by(metric, condition, metric_label) |>
  mutate(threshold = 0.05 * max(abs(derivative), na.rm = TRUE)) |>
  filter(abs(derivative) <= threshold) |>
  slice_min(r, n = 1, with_ties = FALSE) |>
  ungroup() |>
  rename(r_star = r)

# ── Summary table ─────────────────────────────────────────────────────────────
rq1_sat_summary <- purrr::map_dfr(RQ1_SAT_METRICS, function(met) {
  purrr::map_dfr(RQ1_SAT_CONDS, function(cond) {
    fit <- rq1_sat_fits[[met]][[cond]]
    if (is.null(fit)) return(NULL)
    s   <- summary(fit)
    r_s <- rq1_sat_rstar |> filter(metric == met, condition == cond) |> pull(r_star)
    tibble(
      Metric      = metric_labels[met],
      Condition   = cond,
      edf         = round(s$edf[1], 2),
      R2          = round(s$r.sq, 3),
      `p-value`   = signif(s$s.table[1, "p-value"], 2),
      `r* (sat.)` = if (length(r_s)) round(r_s, 1) else NA_real_
    )
  })
})
```

```{r rq1-sat-curves, fig.width=12, fig.height=8}
scatter_rq1sat <- rq1_sat_data |>
  pivot_longer(all_of(RQ1_SAT_METRICS), names_to = "metric", values_to = "value") |>
  mutate(metric_label = metric_labels[metric])

ggplot() +
  geom_ribbon(data = rq1_sat_pred,
              aes(x = r, ymin = lwr, ymax = upr),
              alpha = 0.15, fill = "steelblue") +
  geom_line(data = rq1_sat_pred,
            aes(x = r, y = fit), colour = "steelblue4", linewidth = 0.9) +
  geom_point(data = scatter_rq1sat,
             aes(x = param_val, y = value),
             size = 2.2, colour = "grey30", na.rm = TRUE) +
  scale_x_log10(breaks = r_rq1_pos, labels = r_rq1_pos) +
  facet_grid(metric_label ~ condition, scales = "free_y") +
  labs(title    = "RQ1 — GAM fitted curves: Delegation Rate and Lost Vote Rate",
       subtitle = "Blue band = 95% CI | Points = mean over seeds",
       x = "r (log scale)", y = NULL) +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text = element_text(face = "bold", size = 9),
        panel.grid.minor = element_blank())
```

```{r rq1-sat-deriv, fig.width=12, fig.height=8}
thresh_rq1sat <- rq1_sat_derivs |>
  group_by(metric, condition, metric_label) |>
  summarise(threshold = 0.05 * max(abs(derivative), na.rm = TRUE), .groups = "drop")

ggplot(rq1_sat_derivs, aes(x = r)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "steelblue", alpha = 0.15) +
  geom_line(aes(y = derivative), colour = "steelblue4", linewidth = 0.8) +
  geom_hline(yintercept = 0, colour = "grey50", linewidth = 0.4) +
  geom_hline(data = thresh_rq1sat, aes(yintercept =  threshold),
             linetype = "dashed", colour = "firebrick3", linewidth = 0.6) +
  geom_hline(data = thresh_rq1sat, aes(yintercept = -threshold),
             linetype = "dashed", colour = "firebrick3", linewidth = 0.6) +
  geom_vline(data = rq1_sat_rstar, aes(xintercept = r_star),
             linetype = "dotted", colour = "firebrick4", linewidth = 0.9) +
  scale_x_log10(breaks = r_rq1_pos, labels = r_rq1_pos) +
  facet_grid(metric_label ~ condition, scales = "free_y") +
  labs(title    = "RQ1 — GAM derivative: marginal effect per doubling of r",
       subtitle = "Dashed = 5% threshold | Dotted = saturation point r*",
       x = "r (log scale)", y = "dY / d(log₂r)") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text = element_text(face = "bold", size = 9),
        panel.grid.minor = element_blank())
```
- The derivative shows how much the fitted response changes for each doubling of r. The saturation point r*
 is the first value where this marginal effect falls below 5% of its maximum, indicating that further increases in r have little additional impact.
 
```{r rq1-sat-table}
kable(rq1_sat_summary,
      caption = paste0(
        "RQ1 GAM saturation summary — delegation rate and lost vote rate. ",
        "edf > 2 confirms nonlinearity. r* = practical saturation point ",
        "(first r where slope drops below 5% of its peak value)."),
      align = c("l", "l", "r", "r", "r", "r"))
```

---

# RQ2 — Ideological Drift

**What determines the average ideological drift between an agent and
the vote cast on their behalf?**

The same GAM framework introduced in the RQ1 saturation analysis
($Y_i = f(\log_2 r_i) + \varepsilon_i$) is applied here.

Condition A isolates r\_op (H1b); Condition B isolates r\_pw (H2b);
Condition C shows the full r\_op × r\_pw grid (H3b — joint effect);
Condition D varies gamma at r\_op = r\_pw ∈ {1, 2, 4, 8} (H4b — trust moderation).
The **dashed red line** shows the unfiltered curve; the **solid coloured line**
shows the reliable range (lost vote rate ≤ `r LOST_THRESHOLD`).

```{r rq2-plot, fig.width=14, fig.height=30}
df_rq2    <- sim_long_fine_masked |> filter(metric %in% RQ2_METRICS)
df_rq2_bg <- sim_long_fine        |> filter(metric %in% RQ2_METRICS)

p_rq2_A    <- make_sweep_plot(df_rq2, "A: r_op only", focus = "avg_ideo_drift", bg_long = df_rq2_bg)
p_rq2_B    <- make_sweep_plot(df_rq2, "B: r_pw only", focus = "avg_ideo_drift", bg_long = df_rq2_bg)
p_rq2_grid <- make_grid_lineplot(df_rq2, focus = "avg_ideo_drift",               bg_long = df_rq2_bg)
p_rq2_D    <- make_trust_plot(df_rq2, focus = "avg_ideo_drift",                  bg_long = df_rq2_bg)

p_rq2_A / p_rq2_B / p_rq2_grid / p_rq2_D +
  plot_annotation(
    title = "RQ2 — Avg. Chain Length & Ideological Drift",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  )
```

**Note on Condition B (r\_pw only):** Lost vote rate does not exceed ~0.45, even at the highest r\_pw values. So most votes still reach a direct voter.

## RQ2 — GAM Analysis

The analysis focuses on **avg. chain length** and **avg. ideological drift** —
the two metrics that directly reflect the quality of representation in the
delegation network. For each metric and each experimental condition a GAM of
the form

$$Y_i = f(x_i) + \varepsilon_i, \qquad \varepsilon_i \sim \mathcal{N}(0,\sigma^2)$$

is fitted, where $x$ is the log₂-transformed responsiveness parameter and
$f(\cdot)$ is a flexible penalised spline (REML). The table reports edf
(nonlinearity), R², p-value, and the predicted maximum within the simulated range.

```{r rq2-gam-setup, include=FALSE}
library(mgcv)

# Only chain length and ideological drift for the GAM analysis
GAM2_METRICS <- c("avg_chain_len", "avg_ideo_drift")

rq2_1d_data <- sim_agg_fine |>
  filter(condition %in% c("A: r_op only", "B: r_pw only", "C: r_op x r_pw"),
         param_val > 0) |>
  mutate(log2_r = log2(param_val))

# ── Prediction grid ───────────────────────────────────────────────────────────
fine_r_pos_rq2 <- R_VALS_FINE[R_VALS_FINE > 0]
r_pred_rq2     <- 2^seq(log2(min(fine_r_pos_rq2)), log2(max(fine_r_pos_rq2)),
                        length.out = 200)
log2_pred_rq2  <- log2(r_pred_rq2)

# ── Fit 1-D GAMs ──────────────────────────────────────────────────────────────
rq2_gam_1d <- purrr::map(GAM2_METRICS, function(met) {
  purrr::map(c("A: r_op only", "B: r_pw only", "C: r_op x r_pw"),
    function(cond) {
      d <- rq2_1d_data |> filter(condition == cond)
      if (sum(!is.na(d[[met]])) < 5) return(NULL)
      tryCatch(
        mgcv::gam(as.formula(paste0(met, " ~ s(log2_r, k = 6)")),
                  data = d, method = "REML"),
        error = \(e) NULL
      )
    }) |> setNames(c("A: r_op only", "B: r_pw only", "C: r_op x r_pw"))
}) |> setNames(GAM2_METRICS)

# ── Summary table ─────────────────────────────────────────────────────────────
gam_summary <- purrr::map_dfr(GAM2_METRICS, function(met) {
  purrr::map_dfr(c("A: r_op only", "B: r_pw only", "C: r_op x r_pw"),
    function(cond) {
      fit <- rq2_gam_1d[[met]][[cond]]
      if (is.null(fit)) return(tibble(Metric=metric_labels[met], Condition=cond,
                                      edf=NA_real_, R2=NA_real_,
                                      `p-value`=NA_real_, `Pred. max`=NA_real_))
      s        <- summary(fit)
      pred_all <- predict(fit, newdata=data.frame(log2_r=log2_pred_rq2))
      tibble(
        Metric      = metric_labels[met],
        Condition   = cond,
        edf         = round(s$edf[1], 2),
        R2          = round(s$r.sq,   3),
        `p-value`   = signif(s$s.table[1, "p-value"], 2),
        `Pred. max` = round(max(pred_all, na.rm = TRUE), 3)
      )
    })
})

# ── Predicted values at r = 0, 1, 2, 4, 8 with 95% CI ───────────────────────
ref_r_vals <- c(0, 1, 2, 4, 8)

gam_pred_ref <- purrr::map_dfr(GAM2_METRICS, function(met) {
  purrr::map_dfr(c("A: r_op only", "B: r_pw only", "C: r_op x r_pw"),
    function(cond) {
      fit <- rq2_gam_1d[[met]][[cond]]
      purrr::map_dfr(ref_r_vals, function(rv) {
        if (rv == 0 || is.null(fit)) {
          obs <- sim_agg_fine |> filter(condition == cond, param_val == 0)
          val <- if (nrow(obs) > 0) round(mean(obs[[met]], na.rm=TRUE), 3)
                 else NA_real_
          tibble(Metric=metric_labels[met], Condition=cond, r=rv,
                 Estimate=val, `CI low`=NA_real_, `CI high`=NA_real_)
        } else {
          nd <- data.frame(log2_r = log2(rv))
          p  <- predict(fit, newdata=nd, se.fit=TRUE)
          tibble(Metric=metric_labels[met], Condition=cond, r=rv,
                 Estimate  = round(p$fit, 3),
                 `CI low`  = round(p$fit - 1.96 * p$se.fit, 3),
                 `CI high` = round(p$fit + 1.96 * p$se.fit, 3))
        }
      })
    })
})

gam_pred_wide <- gam_pred_ref |>
  mutate(cell = ifelse(is.na(`CI low`), as.character(Estimate),
                       paste0(Estimate, " [", `CI low`, ", ", `CI high`, "]"))) |>
  dplyr::select(Metric, r, Condition, cell) |>
  tidyr::pivot_wider(names_from = Condition, values_from = cell)
```

```{r rq2-gam-table}
kable(gam_summary,
      caption = paste0(
        "RQ2 GAM summary — one row per metric × condition. ",
        "edf: effective degrees of freedom (≈1 = linear, >2 = clearly curved). ",
        "R²: share of variance explained. ",
        "p-value: test of smooth term against a flat line (note: near-zero values ",
        "are expected in a designed simulation and confirm non-flatness only). ",
        "Pred. max: highest predicted value of the fitted curve within the simulated range."),
      align = c("l", "l", "r", "r", "r", "r"))
```

```{r rq2-gam-pred-table}
kable(gam_pred_wide,
      caption = paste0(
        "RQ2 — Predicted metric values at representative r levels with 95% CI ",
        "[lower, upper]. r = 0 uses the observed mean (GAM fit starts at r = 0.5). ",
        "Reading across columns compares the effect of r_op (A), r_pw (B), and ",
        "both jointly (C) at each responsiveness level."),
      align = c("l", "r", rep("l", ncol(gam_pred_wide) - 2)))
```

```{r rq2-gam-curves, fig.width=14, fig.height=14}
# Fitted GAM curves overlaid on aggregated data
rq2_gam_pred <- purrr::map_dfr(GAM2_METRICS, function(met) {
  purrr::map_dfr(c("A: r_op only", "B: r_pw only", "C: r_op x r_pw"),
    function(cond) {
      fit <- rq2_gam_1d[[met]][[cond]]
      if (is.null(fit)) return(NULL)
      nd <- data.frame(log2_r = log2_pred_rq2)
      p  <- predict(fit, newdata = nd, se.fit = TRUE)
      tibble(r = r_pred_rq2, fit = p$fit,
             lwr = p$fit - 1.96 * p$se.fit,
             upr = p$fit + 1.96 * p$se.fit,
             metric = met, metric_label = metric_labels[met],
             condition = cond)
    })
})

# Unfiltered scatter (tomato = outside threshold); filtered (grey = within threshold)
scatter_rq2_all <- rq2_1d_data |>
  pivot_longer(all_of(GAM2_METRICS), names_to = "metric", values_to = "value") |>
  mutate(metric_label = metric_labels[metric])

scatter_rq2_ok <- sim_agg_fine_masked |>
  filter(condition %in% c("A: r_op only", "B: r_pw only", "C: r_op x r_pw"),
         param_val > 0) |>
  pivot_longer(all_of(GAM2_METRICS), names_to = "metric", values_to = "value") |>
  mutate(metric_label = metric_labels[metric])

ggplot() +
  geom_ribbon(data = rq2_gam_pred,
              aes(x = r, ymin = lwr, ymax = upr),
              alpha = 0.15, fill = "steelblue") +
  geom_line(data = rq2_gam_pred,
            aes(x = r, y = fit), colour = "steelblue4", linewidth = 0.9) +
  geom_point(data = scatter_rq2_all,
             aes(x = param_val, y = value),
             size = 2.5, colour = "tomato3", alpha = 0.45, na.rm = TRUE) +
  geom_point(data = scatter_rq2_ok,
             aes(x = param_val, y = value),
             size = 2, colour = "grey25", na.rm = TRUE) +
  scale_x_log10(breaks = fine_r_pos_rq2, labels = fine_r_pos_rq2) +
  facet_grid(metric_label ~ condition, scales = "free_y") +
  labs(title    = "RQ2 — GAM fitted curves (blue) with 95% CI",
       subtitle = paste0("Grey points = within LVR ≤ ", LOST_THRESHOLD,
                         " threshold  |  Red points = outside threshold (unreliable)"),
       x = "r (log scale)", y = NULL) +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text = element_text(face = "bold", size = 8.5),
        panel.grid.minor = element_blank())
```

---

# RQ3 — Top-5% Power Share & Network Components

**What share of total voting power do the most powerful 5% of agents hold,
and how does network fragmentation evolve?**

Condition A isolates r\_op (H1c — top-5% share increases slightly via lost votes,
components decrease); Condition B isolates r\_pw (H2c — strong hub concentration,
more components); Condition C\_diag shows the diagonal r\_op = r\_pw (H3c — joint);
the full r\_op × r\_pw grid is shown as a heatmap; Condition D adds trust at
r\_op = r\_pw ∈ {1, 4, 8} (H4c — trust reduces top-5% share).

```{r rq3-plot, fig.width=14, fig.height=30}
df_rq3    <- sim_long_fine_masked |> filter(metric %in% RQ3_METRICS)
df_rq3_bg <- sim_long_fine        |> filter(metric %in% RQ3_METRICS)

p_rq3_A    <- make_sweep_plot(df_rq3, "A: r_op only", focus = "top5_share", bg_long = df_rq3_bg)
p_rq3_B    <- make_sweep_plot(df_rq3, "B: r_pw only", focus = "top5_share", bg_long = df_rq3_bg)
p_rq3_grid <- make_grid_lineplot(df_rq3, focus = "top5_share",               bg_long = df_rq3_bg)
p_rq3_D    <- make_trust_plot(df_rq3, focus = "top5_share",                  bg_long = df_rq3_bg)

p_rq3_A / p_rq3_B / p_rq3_grid / p_rq3_D +
  plot_annotation(
    title = "RQ3 — Top-5% Power Share & Components",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  )
```
---

# RQ4 — Share of Agents Needed for 50% of Votes

**How many of the most influential agents are jointly needed to hold
50% of total voting power?**

`share_50pct` = k / N where k is the minimum number of agents (sorted
descending by power) whose cumulative power reaches 50% of the total.
A **lower** value means more concentrated power.

Condition A isolates r\_op (H1d — decreases: more concentrated);
Condition B isolates r\_pw (H2d — increases: power spread across hubs);
Condition C\_diag shows the diagonal r\_op = r\_pw (H3d — joint);
the full r\_op × r\_pw grid is shown as a heatmap;
Condition D adds trust at r\_op = r\_pw ∈ {1, 4, 8} (H4d — increases: more even distribution).

```{r rq4-plot, fig.width=14, fig.height=28}
df_rq4    <- sim_long_fine_masked |> filter(metric %in% RQ4_METRICS)
df_rq4_bg <- sim_long_fine        |> filter(metric %in% RQ4_METRICS)

p_rq4_A    <- make_sweep_plot(df_rq4, "A: r_op only", focus = "share_50pct", bg_long = df_rq4_bg)
p_rq4_B    <- make_sweep_plot(df_rq4, "B: r_pw only", focus = "share_50pct", bg_long = df_rq4_bg)
p_rq4_grid <- make_grid_lineplot(df_rq4, focus = "share_50pct",               bg_long = df_rq4_bg)
p_rq4_D    <- make_trust_plot(df_rq4, focus = "share_50pct",                  bg_long = df_rq4_bg)

p_rq4_A / p_rq4_B / p_rq4_grid / p_rq4_D +
  plot_annotation(
    title = "RQ4 — Share of Agents Needed for 50% of Votes",
    theme = theme(plot.title = element_text(face = "bold", size = 14))
  )
```
