---
title: "Weekly Report -- Week 19: Minority Representation in Liquid Democracy"
output: github_document
date: "2026-07-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 10, fig.height = 6)
library(tidyverse)
library(here)
library(furrr)
library(patchwork)
library(digest)
library(knitr)

theme_set(theme_minimal(base_size = 12))
source(here::here("scripts/Network.R"))

# ── Simulation constants (subset of Report 18 needed for Part II only;
# Part I -- confidence-aggregator and trust-update development -- is not
# repeated here, see Report 18) ──────────────────────────────────────────────
MINORITY_SHARE   <- 0.2
OP_SIGMA         <- 0.1
SIM_T            <- 20L
N_SIM            <- 500L
K_SIM            <- 14L
P_REWIRE         <- 0.05

DELTA_VALS       <- c(0.2, 0.4)
R_INGROUP_VALS   <- c(0, 1, 2, 4, 8)

LAMBDA_DIAG   <- 0.5   # trust decay, Report 18 Sec. 3
CYCLE_PENALTY <- 1     # kappa_cyc for reward+punish trust update, Report 18 Sec. 4

N_SEEDS_MINORITY <- 25L

# r_op = r_pw = r_trust ("background responsiveness") is swept as a
# robustness check: does the r_ingroup/condition story hold up at low (1),
# medium (4) and high (8) background responsiveness, or is it specific to
# one regime? Deliberately NOT averaged across these -- see Section 1.2.
R_BACKGROUND_VALS <- c(1, 4, 8)

# "homophily" reuses the existing opinion-based network shuffle
# (shuffle_opinions_homophily(), Report 9) instead of a group-based shuffle:
# it runs on top of "polarised" opinions, so making the network assortative
# on opinion indirectly clusters it by group too, since the two are
# correlated once minority/majority opinion means differ. Fixed at
# delta=0.4 only (strongest opinion-group correlation, clearest effect).
HOMOPHILY_TARGET_R <- 0.8
HOMOPHILY_DELTA    <- 0.4

# Trust: reward+punish update (Report 18 Sec. 4). Self-weight: confidence
# self-weight with the sum aggregator (Report 18 Sec. 1) -- see the recap
# in Section 1 below.
TRUST_MODE       <- "reward_punish"
SELF_WEIGHT_MODE <- "confidence"
CONFIDENCE_AGG   <- "sum"

# Colour convention used throughout: majority = blue, minority = red,
# population-wide/"all" = grey.
GROUP_COLOURS <- c(all = "grey40", minority = "firebrick3", majority = "steelblue4")
```

## Opinion Distributions by Condition

Opinions are drawn once at initialisation from group-specific normal
distributions,

$$o_i \sim \begin{cases}
  \mathcal{N}(0.5-\delta/2,\ \sigma^2) & g_i = \text{minority} \quad (20\%) \\
  \mathcal{N}(0.5+\delta/2,\ \sigma^2) & g_i = \text{majority} \quad (80\%)
\end{cases}, \qquad \sigma = OP\_SIGMA = 0.1,$$

with $\delta=0$ under `random` (both groups share $\mathcal{N}(0.5,\sigma^2)$),
$\delta\in\{0.2,0.4\}$ under `polarised`, and $\delta=0.4$ under `homophily`
(same draw as `polarised`, then reshuffled onto the network -- see the
*Implementation Note: Homophily* section below).

```{r opinion-dist-plot, fig.width=11, fig.height=5}
dist_conditions <- tribble(
  ~condition_label,                                   ~mu_min,                  ~mu_maj,
  "random",                                            0.5,                      0.5,
  "polarised (delta=0.2)",                             0.5 - 0.2 / 2,            0.5 + 0.2 / 2,
  "polarised (delta=0.4)",                             0.5 - 0.4 / 2,            0.5 + 0.4 / 2,
  sprintf("homophily (delta=%.1f)", HOMOPHILY_DELTA),  0.5 - HOMOPHILY_DELTA/2,  0.5 + HOMOPHILY_DELTA/2
) |>
  mutate(condition_label = factor(condition_label, levels = c(
    "random", "polarised (delta=0.2)", "polarised (delta=0.4)",
    sprintf("homophily (delta=%.1f)", HOMOPHILY_DELTA))))

x_grid <- seq(0, 1, length.out = 400)

df_dist <- dist_conditions |>
  reframe(x = x_grid,
          minority = dnorm(x_grid, mu_min, OP_SIGMA),
          majority = dnorm(x_grid, mu_maj, OP_SIGMA),
          .by = condition_label) |>
  pivot_longer(c(minority, majority), names_to = "group", values_to = "density") |>
  mutate(group = factor(group, levels = c("minority", "majority")))

ggplot(df_dist, aes(x, density, colour = group, fill = group)) +
  geom_area(position = "identity", alpha = 0.25, linewidth = 0) +
  geom_line(linewidth = 0.9) +
  facet_wrap(~ condition_label, nrow = 1) +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_fill_manual(values   = GROUP_COLOURS) +
  labs(title = "Initial opinion distributions by condition",
       subtitle = sprintf("o_i ~ N(0.5 -+ delta/2, %.2f^2); homophily reshuffles polarised (delta=%.1f) onto a network with opinion correlation r >= %.1f",
                          OP_SIGMA, HOMOPHILY_DELTA, HOMOPHILY_TARGET_R),
       x = "opinion o_i", y = "density", colour = "group", fill = "group") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

`random` and `homophily` share the same $\delta$-implied means as `random`
($\delta{=}0$) and `polarised (delta=0.4)` respectively -- homophily changes
*who sits next to whom* on the network, not the opinion draw itself, so its
marginal distribution is identical to `polarised (delta=0.4)`.

## Opinion Distributions by Condition (Uniform-Distribution Design)

A second, independent study design keeps the same model and metrics
throughout (Sections 3-5 below) but draws opinions from a **uniform**
rather than a normal distribution:

$$o_i \sim \begin{cases}
  \mathcal{U}(0,1)     & \text{random} \quad \text{(both groups, no group difference)} \\
  \mathcal{U}(0,\,0.5) & g_i=\text{minority} \quad \text{(polarised / homophily)} \\
  \mathcal{U}(0.5,\,1) & g_i=\text{majority} \quad \text{(polarised / homophily)}
\end{cases}$$

`random` is simply $\mathcal{U}(0,1)$ for every agent, regardless of group.
`polarised` puts minority and majority on opposite sides of a hard split at
$0.5$, with no overlap; since this split is already maximal within $[0,1]$,
there is no $\delta$ to sweep here (unlike the normal-distribution design
above). `homophily` uses the same `polarised` draw, then reshuffles it onto
the network exactly as described in the *Implementation Note: Homophily*
section below. Results for this design are reported in Section 5, using
the same plots as Sections 3-4.

```{r opinion-dist-plot-uniform, fig.width=9, fig.height=5}
dist_conditions_u <- tribble(
  ~condition_label, ~lo_min, ~hi_min, ~lo_maj, ~hi_maj,
  "random",          0,       1,       0,       1,
  "polarised",       0,       0.5,     0.5,     1,
  "homophily",       0,       0.5,     0.5,     1
) |>
  mutate(condition_label = factor(condition_label, levels = c("random", "polarised", "homophily")))

x_grid_u <- seq(0, 1, length.out = 400)

df_dist_u <- dist_conditions_u |>
  reframe(x = x_grid_u,
          minority = dunif(x_grid_u, lo_min, hi_min),
          majority = dunif(x_grid_u, lo_maj, hi_maj),
          .by = condition_label) |>
  pivot_longer(c(minority, majority), names_to = "group", values_to = "density") |>
  mutate(group = factor(group, levels = c("minority", "majority")))

ggplot(df_dist_u, aes(x, density, colour = group, fill = group)) +
  geom_area(position = "identity", alpha = 0.25, linewidth = 0) +
  geom_line(linewidth = 0.9) +
  facet_wrap(~ condition_label, nrow = 1) +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_fill_manual(values   = GROUP_COLOURS) +
  labs(title = "Initial opinion distributions by condition (uniform-distribution design)",
       subtitle = "random: U(0,1) for both groups; polarised/homophily: minority U(0,0.5), majority U(0.5,1)",
       x = "opinion o_i", y = "density", colour = "group", fill = "group") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

---

# Part II: Minority Representation in Liquid Democracy

## RQ1: At which point is the responsiveness to ingroup parameter saturating?

- **H1a.** The relationship between behavioural ingroup responsiveness
  ($r_{ingroup}$) and minority delegation behaviour is non-linear and
  exhibits a saturation point.

*Operationalized by:* $RR_m$, $PC_m$, $RC_m$, $CDR$, delegation rate,
representative share, average delegation chain length.

## RQ2: How does behavioural ingroup preference affect minority representation in liquid democracy?

- **H1b.** At $r_{ingroup}=0$ under random opinion distributions,
  minority representation is proportional to the minority's population
  share.
- **H2b.** Increasing behavioural ingroup responsiveness improves
  minority representation.
- **H3b.** Increasing behavioural ingroup responsiveness reduces
  minority power capture and minority representation capture.
- **H4b.** The effects of behavioural ingroup responsiveness on minority
  representation become stronger as opinion polarisation increases.

*Operationalized by:* $RR_m$, $PC_m$, $RC_m$.

## RQ3: Through which delegation mechanisms does behavioural ingroup preference influence minority representation?

- **H1c.** Increasing behavioural ingroup responsiveness reduces
  cross-group delegation.
- **H2c.** The reduction in cross-group delegation is stronger under
  greater opinion polarisation.
- **H3c.** Increasing behavioural ingroup responsiveness shortens
  delegation chains.
- **H4c.** Increasing behavioural ingroup responsiveness reduces
  ideological drift within both minority and majority groups.

*Operationalized by:* $CDR$, average delegation chain length, average
ideological drift.

---

# Implementation Note: Homophily

**Energy.** Total opinion disagreement across the friendship graph $G$:

$$E = \sum_{(i,j)\in E(G)} |o_i - o_j|$$

**Metropolis step** (`shuffle_opinions_homophily()`, repeated up to
`homophily_steps` $=10^5$ times):

1. Draw two lay agents $i,j$ uniformly at random.
2. Compute the *local* energy change of swapping $o_i \leftrightarrow o_j$
   (only terms touching $i$ or $j$ change):
$$E_{\text{before}} = \sum_{k\in \text{nbr}(i)}|o_i-o_k| + \sum_{k\in \text{nbr}(j)}|o_j-o_k|$$
$$E_{\text{after}} = \sum_{k\in \text{nbr}(i)}|o_j-o_k| + \sum_{k\in \text{nbr}(j)}|o_i-o_k|$$
$$\Delta E = E_{\text{after}} - E_{\text{before}}$$
3. Accept the swap if $\Delta E<0$; otherwise accept with probability
$$P(\text{accept}) = \exp(-\Delta E \cdot t), \qquad t = \texttt{homophily\_t} = 5.$$
4. Every 100 steps, check the **network opinion correlation** $r$ and stop
   once $r \ge r^\star$, $r^\star=$ `r HOMOPHILY_TARGET_R`. $r$ is the
   Pearson correlation between the opinions sitting at either end of an
   edge, i.e. the ordinary Pearson correlation coefficient applied to the
   list of edge endpoint pairs (each undirected edge counted once in each
   direction, so both sides of the pair share the same marginal
   distribution of opinions):
$$r = \frac{\sum_{(i,j)\in \vec E(G)} (o_i-\bar o)(o_j-\bar o)}
           {\sum_{(i,j)\in \vec E(G)} (o_i-\bar o)^2},
   \qquad \bar o = \frac{1}{N}\sum_{i} o_i,$$
   where $\vec E(G)$ lists every edge of $G$ in both directions. $r=1$
   means every edge connects two agents with identical opinions; $r=0$
   means opinions are uncorrelated with network position (the untouched,
   randomly wired network).

---

# 1  Model

## 1.1  Metrics

Averaged over the last 5 rounds (`round >= T-4`), then over
`r N_SEEDS_MINORITY` seeds.

**1. Voting power / share.**
$$P_M = \sum_{i \in M} p_i, \quad P_{Maj} = \sum_{i \in N\setminus M} p_i,
\quad \text{share}_M = \frac{P_M}{P_M+P_{Maj}}$$
Total power actually cast by each group's roots ($p_i$ = root power if
$i$ is a root, else 0), and each group's share of the total.

**2. Relative Representation.**
$$RR_m = \frac{\big(\sum_{i\in M} p_i \,/\, \sum_{i\in N} p_i\big)}{m},
\qquad m = |M|/|N|$$
Minority's power share divided by its population share; $1$ =
proportional representation.

**2b. Relative Representation, resolved-population denominator.**
$$RR_m^{\text{resolved}} = \frac{\big(\sum_{i\in M} p_i \,/\, \sum_{i\in N} p_i\big)}{m_{\text{resolved}}},
\qquad m_{\text{resolved}} = \frac{\#\{i\in M : \text{root}(i) \text{ defined}\}}{\#\{i\in N : \text{root}(i) \text{ defined}\}}$$
Same power-share numerator as $RR_m$, but divided by the minority's share
of the *resolved* population (agents whose delegation chain reaches a
root, i.e. not stuck in a cycle) instead of its total population share
$m$. $RR_m$ conflates two channels -- majority capturing minority power,
and minority votes lost to cycles (which shrink the numerator but not the
fixed denominator $m$) -- into one number; $RR_m^{\text{resolved}}$
isolates the first channel alone, directly comparable to $PC_m$ (which is
resolved-only by construction already). $RR_{Maj}^{\text{resolved}}$ is
the majority counterpart, analogous to $RR_{Maj}$.

**3. Power Capture.**
$$PC_m = \frac{\#\{i\in M : \text{root}(i)\in N\setminus M\}}{\#\{i\in M : \text{root}(i) \text{ defined}\}},
\qquad PC_{Maj} \text{ analogous with } M \leftrightarrow N\setminus M$$
Fraction of one group's agents whose delegation chain terminates at a representative from the other group.

**4. Cross-group delegation, both directions.**
$$CDR_{M\to Maj} = \frac{\#\{(i,j)\in E : i\in M,\, j\notin M\}}{\#\{(i,j)\in E: i \in M\}},
\qquad
CDR_{Maj\to M} = \frac{\#\{(i,j)\in E : i\notin M,\, j\in M\}}{\#\{(i,j)\in E: i \notin M\}}$$
Share of each group's own outgoing delegation edges that cross into the
other group.

**5. Group-specific rates.**
$$\text{DelRate}_G = \frac{\#\{i\in G : \text{delegated}(i)\}}{|G|}$$
Delegation rate within group $G$ ($G\in\{N, M, N\setminus M\}$);
lost-vote rate, chain length and ideological drift use the same
all/minority/majority split, restricted to each group's own agents.

**6. Representatives per group.**
$$n_G = \#\{i\in G : \text{root}(i)=i\}$$
Headcount of group $G$'s direct voters (roots), regardless of how many
others delegate to them.

## 1.2  Experimental Design

| Factor | Values |
|---|---|
| Condition | `random`, `polarised`, `homophily` |
| Polarisation $\delta$ | $\{0.2, 0.4\}$ under `polarised`; fixed at `r HOMOPHILY_DELTA` under `homophily` |
| Ingroup preference $r_{ingroup}$ | $\{0, 1, 2, 4, 8\}$ |
| Background responsiveness $r_{op}=r_{pw}=r_{trust}$ | $\{$ `r paste(R_BACKGROUND_VALS, collapse=", ")` $\}$ |

`random`: $\mathcal{N}(0.5,0.1^2)$ for both groups. `polarised`:
$\mathcal{N}(0.5\mp\delta/2,0.1^2)$. `homophily`: `polarised` at
$\delta=$ `r HOMOPHILY_DELTA` + network reshuffle to a network opinion
correlation $\ge$ `r HOMOPHILY_TARGET_R` (see the *Implementation Note:
Homophily* section above). Every plot below repeats across all 3
background-responsiveness levels (separate rows) rather than averaging
across them, since averaging would blend potentially different regimes
(e.g. saturation effects) into a number that matches none of them.

---

# 2  Simulation

```{r sim-run-one, include=FALSE}
run_cell <- function(seed, condition, delta, r_ingroup, r_background) {
  if (condition == "random") {
    min_mu <- 0.5
    maj_mu <- 0.5
    th     <- NULL
  } else if (condition == "polarised") {
    min_mu <- 0.5 - delta / 2
    maj_mu <- 0.5 + delta / 2
    th     <- NULL
  } else {  # homophily
    min_mu <- 0.5 - delta / 2
    maj_mu <- 0.5 + delta / 2
    th     <- HOMOPHILY_TARGET_R
  }

  res <- simulate_liquid_democracy(
    seed                    = seed,
    n_per_community         = N_SIM,
    n_communities           = 1,
    node_degree             = K_SIM,
    T                       = SIM_T,
    r_op                    = r_background,
    r_pw                    = r_background,
    p_rewire                = P_REWIRE,
    r_ingroup               = r_ingroup,
    minority_share          = MINORITY_SHARE,
    minority_opinion_mu     = min_mu,
    minority_opinion_sigma  = OP_SIGMA,
    majority_opinion_mu     = maj_mu,
    majority_opinion_sigma  = OP_SIGMA,
    self_weight_mode        = SELF_WEIGHT_MODE,
    confidence_agg          = CONFIDENCE_AGG,
    sigma_opinion           = "auto",
    target_homophily        = th,
    gamma                   = r_background,
    lambda                  = LAMBDA_DIAG,
    trust_mode              = TRUST_MODE,
    cycle_penalty           = CYCLE_PENALTY
  )

  snaps <- res$snapshots |> dplyr::filter(round >= SIM_T - 4L)

  tibble(
    seed                         = seed,
    condition                    = condition,
    delta                        = delta,
    r_ingroup                    = r_ingroup,
    r_background                 = r_background,
    delegation_rate              = mean(snaps$delegation_rate, na.rm = TRUE),
    lost_vote_rate                = mean(snaps$lost_vote_rate,  na.rm = TRUE),
    avg_chain_length_delegators  = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
    avg_drift                    = mean(snaps$avg_drift,        na.rm = TRUE),
    minority_delegation_rate     = mean(snaps$minority_delegation_rate, na.rm = TRUE),
    majority_delegation_rate     = mean(snaps$majority_delegation_rate, na.rm = TRUE),
    minority_lost_vote_rate      = mean(snaps$minority_lost_vote_rate,  na.rm = TRUE),
    majority_lost_vote_rate      = mean(snaps$majority_lost_vote_rate,  na.rm = TRUE),
    minority_chain_len           = mean(snaps$minority_chain_len, na.rm = TRUE),
    majority_chain_len           = mean(snaps$majority_chain_len, na.rm = TRUE),
    minority_drift                = mean(snaps$minority_drift,    na.rm = TRUE),
    majority_drift                = mean(snaps$majority_drift,    na.rm = TRUE),
    minority_power                = mean(snaps$minority_power,    na.rm = TRUE),
    majority_power                = mean(snaps$majority_power,    na.rm = TRUE),
    RR_m                          = mean(snaps$RR_m,              na.rm = TRUE),
    RR_m_resolved                 = mean(snaps$RR_m_resolved,     na.rm = TRUE),
    minority_n_resolved            = mean(snaps$minority_n_resolved, na.rm = TRUE),
    majority_n_resolved            = mean(snaps$majority_n_resolved, na.rm = TRUE),
    minority_capture               = mean(snaps$minority_capture, na.rm = TRUE),
    majority_capture               = mean(snaps$majority_capture, na.rm = TRUE),
    cross_group_dlg_rate          = mean(snaps$cross_group_dlg_rate, na.rm = TRUE),
    cdr_min_to_maj                = mean(snaps$cdr_min_to_maj,    na.rm = TRUE),
    cdr_maj_to_min                = mean(snaps$cdr_maj_to_min,    na.rm = TRUE),
    minority_n_reps                = mean(snaps$minority_n_reps,  na.rm = TRUE),
    majority_n_reps                = mean(snaps$majority_n_reps,  na.rm = TRUE)
  )
}
```

```{r sim-grid, include=FALSE}
seeds_sim <- seq_len(N_SEEDS_MINORITY)

grid_random    <- expand_grid(condition = "random", delta = NA_real_,
                               r_ingroup = R_INGROUP_VALS, r_background = R_BACKGROUND_VALS, seed = seeds_sim)
grid_polarised <- expand_grid(condition = "polarised", delta = DELTA_VALS,
                               r_ingroup = R_INGROUP_VALS, r_background = R_BACKGROUND_VALS, seed = seeds_sim)
grid_homophily <- expand_grid(condition = "homophily", delta = HOMOPHILY_DELTA,
                               r_ingroup = R_INGROUP_VALS, r_background = R_BACKGROUND_VALS, seed = seeds_sim)
grid_sim <- bind_rows(grid_random, grid_polarised, grid_homophily)

params_sim <- list(N = N_SIM, K = K_SIM, T = SIM_T, r_background_vals = R_BACKGROUND_VALS,
                    lambda = LAMBDA_DIAG, trust_mode = TRUST_MODE,
                    cycle_penalty = CYCLE_PENALTY,
                    delta_vals = DELTA_VALS, r_ingroup_vals = R_INGROUP_VALS,
                    homophily_target = HOMOPHILY_TARGET_R, homophily_delta = HOMOPHILY_DELTA,
                    minority_share = MINORITY_SHARE,
                    self_weight_mode = SELF_WEIGHT_MODE, confidence_agg = CONFIDENCE_AGG,
                    sigma_opinion = "auto", n_seeds = N_SEEDS_MINORITY, version = 9L)
hash_sim  <- digest::digest(params_sim)
cache_sim <- here::here("data/sim_report19_cache.rds")

if (file.exists(cache_sim)) {
  cs <- readRDS(cache_sim)
  if (identical(cs$hash, hash_sim)) sim_raw <- cs$data else cs <- NULL
} else cs <- NULL

if (is.null(cs)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  plan(multisession, workers = parallel::detectCores() - 1)

  sim_raw <- future_pmap_dfr(
    list(seed = grid_sim$seed, condition = grid_sim$condition,
         delta = grid_sim$delta, r_ingroup = grid_sim$r_ingroup,
         r_background = grid_sim$r_background),
    run_cell,
    .options = furrr_options(seed = NULL)
  )
  plan(sequential)
  saveRDS(list(hash = hash_sim, data = sim_raw), cache_sim)
}

# Average over seeds; build "condition_label"/"r_background_label" factors
# for plotting, plus derived group-share, deviation and
# total-representatives columns used below.
sim_agg <- sim_raw |>
  group_by(condition, delta, r_ingroup, r_background) |>
  summarise(across(where(is.numeric) & !any_of("seed"), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  mutate(
    condition_label = case_when(
      condition == "random"    ~ "random",
      condition == "polarised" ~ sprintf("polarised (delta=%.1f)", delta),
      condition == "homophily" ~ sprintf("homophily (delta=%.1f)", delta)
    ),
    condition_label = factor(condition_label, levels = c(
      "random", "polarised (delta=0.2)", "polarised (delta=0.4)",
      sprintf("homophily (delta=%.1f)", HOMOPHILY_DELTA))),
    r_background_label = factor(paste0("r=", r_background), levels = paste0("r=", R_BACKGROUND_VALS)),
    minority_power_share_pct = 100 * minority_power / (minority_power + majority_power),
    majority_power_share_pct = 100 - minority_power_share_pct,
    # Deviation from proportional share (in percentage points) -- lets
    # minority and majority sit in ONE panel on a shared, tight y-axis
    # instead of needing separate rows for a ~60pp baseline gap.
    minority_power_dev_pct   = minority_power_share_pct - 100 * MINORITY_SHARE,
    majority_power_dev_pct   = majority_power_share_pct - 100 * (1 - MINORITY_SHARE),
    all_n_reps                = minority_n_reps + majority_n_reps,
    # RR_maj: majority's power share over majority's population share --
    # the same formula as RR_m, just for the other group. Not independent
    # of RR_m with only two groups, but shown alongside it so both
    # group-specific lines are visible rather than only the minority one.
    RR_maj                    = (majority_power_share_pct / 100) / (1 - MINORITY_SHARE),
    # RR_maj_resolved: majority's counterpart to RR_m_resolved -- same
    # resolved-power numerator, divided by majority's share of the RESOLVED
    # population instead of its total population share.
    RR_maj_resolved            = (majority_power_share_pct / 100) /
      (majority_n_resolved / (minority_n_resolved + majority_n_resolved))
  )

SUBTITLE <- sprintf("N = %d | k = %d | T = %d | %d seeds | confidence (sum) | trust: reward+punish",
                    N_SIM, K_SIM, SIM_T, N_SEEDS_MINORITY)
```

---

# 3  Representation Overview

```{r plot-part1-rates, fig.width=13, fig.height=16}
rates_map <- tribble(
  ~metric,           ~scope,     ~column,
  "Delegation Rate", "all",      "delegation_rate",
  "Delegation Rate", "minority", "minority_delegation_rate",
  "Delegation Rate", "majority", "majority_delegation_rate",
  "Lost Vote Rate",  "all",      "lost_vote_rate",
  "Lost Vote Rate",  "minority", "minority_lost_vote_rate",
  "Lost Vote Rate",  "majority", "majority_lost_vote_rate"
)

df_part1_rates <- rates_map |>
  pmap_dfr(function(metric, scope, column)
    sim_agg |> transmute(condition_label, r_background_label, r_ingroup, value = .data[[column]], metric, scope)) |>
  mutate(metric = factor(metric, levels = unique(rates_map$metric)),
         scope  = factor(scope, levels = c("all", "minority", "majority")))

ggplot(df_part1_rates, aes(r_ingroup, value, colour = scope)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  facet_grid(r_background_label + metric ~ condition_label, scales = "free_y") +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_x_continuous(breaks = R_INGROUP_VALS) +
  labs(title = "Delegation rate & lost-vote rate -- all vs. minority vs. majority, by background r",
       subtitle = SUBTITLE, x = "r_ingroup", y = NULL, colour = "scope") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```
- all minority condition tend to have higher lost vote rates due to smaller population size. 

```{r plot-part1-rr-pc, fig.width=13, fig.height=16}
# RR_maj and PC_Maj are not independent of RR_m/PC_m (only two groups), but
# shown as their own red/blue lines so both groups are visible directly,
# matching the minority=red/majority=blue convention used everywhere else.
#
# RR_m/RR_maj (solid) use the total population share as their denominator,
# so they fold two channels together: majority capturing minority power,
# AND minority votes lost to cycles (which shrink the resolved-power
# numerator but not the fixed population-share denominator). RR_m_resolved/
# RR_maj_resolved (dashed) use the RESOLVED population share instead --
# same numerator, so the gap between solid and dashed lines is exactly the
# lost-vote/cycling contribution, and the dashed line isolates the capture
# channel alone, directly comparable to PC_m/PC_Maj (which are resolved-only
# by construction already).
df_rr_pc <- sim_agg |>
  select(condition_label, r_background_label, r_ingroup,
         RR_m, RR_maj, RR_m_resolved, RR_maj_resolved,
         PC_m = minority_capture, PC_Maj = majority_capture) |>
  pivot_longer(c(RR_m, RR_maj, RR_m_resolved, RR_maj_resolved, PC_m, PC_Maj),
               names_to = "series", values_to = "value") |>
  mutate(metric = factor(if_else(series %in% c("RR_m", "RR_maj", "RR_m_resolved", "RR_maj_resolved"), "RR", "PC"),
                          levels = c("RR", "PC")),
         scope  = factor(if_else(series %in% c("RR_m", "PC_m", "RR_m_resolved"), "minority", "majority"),
                          levels = c("minority", "majority")),
         denom  = factor(if_else(series %in% c("RR_m_resolved", "RR_maj_resolved"), "resolved", "population"),
                          levels = c("population", "resolved")))

ggplot(df_rr_pc, aes(r_ingroup, value, colour = scope, linetype = denom)) +
  geom_hline(data = tibble(metric = factor("RR", levels = c("RR", "PC")), value = 1),
             aes(yintercept = value), linetype = "dotted", colour = "grey50") +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  facet_grid(r_background_label + metric ~ condition_label, scales = "free_y") +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_linetype_manual(values = c(population = "solid", resolved = "dashed")) +
  scale_x_continuous(breaks = R_INGROUP_VALS) +
  labs(title = "RR_m / RR_maj (population vs. resolved denominator) and PC_m / PC_Maj, by background r",
       subtitle = SUBTITLE, x = "r_ingroup", y = NULL, colour = "scope", linetype = "RR denominator") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```
- Why is PCm and RRm relation ship positive? I would expect the behavior as it is for majority
---

# 4  Delegation Mechanisms

```{r plot-part2-cdr, fig.width=13, fig.height=9}
df_cdr <- sim_agg |>
  select(condition_label, r_background_label, r_ingroup, all = cross_group_dlg_rate,
         minority = cdr_min_to_maj, majority = cdr_maj_to_min) |>
  pivot_longer(c(all, minority, majority), names_to = "scope", values_to = "value") |>
  mutate(scope = factor(scope, levels = c("all", "minority", "majority")))

ggplot(df_cdr, aes(r_ingroup, value, colour = scope)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  facet_grid(r_background_label ~ condition_label, scales = "free_y") +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_x_continuous(breaks = R_INGROUP_VALS) +
  labs(title = "Cross-group delegation rate (all = population-wide; minority = M->Maj; majority = Maj->M), by background r",
       subtitle = SUBTITLE, x = "r_ingroup", y = "share of own edges crossing group", colour = "scope") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```
- Difference quiete small but visible between different conditions (direction of results as expected).
```{r plot-part2-chain-drift, fig.width=13, fig.height=16}
mech_map <- tribble(
  ~metric,             ~scope,     ~column,
  "Chain Length",      "all",      "avg_chain_length_delegators",
  "Chain Length",      "minority", "minority_chain_len",
  "Chain Length",      "majority", "majority_chain_len",
  "Ideological Drift", "all",      "avg_drift",
  "Ideological Drift", "minority", "minority_drift",
  "Ideological Drift", "majority", "majority_drift"
)

df_mech <- mech_map |>
  pmap_dfr(function(metric, scope, column)
    sim_agg |> transmute(condition_label, r_background_label, r_ingroup, value = .data[[column]], metric, scope)) |>
  mutate(metric = factor(metric, levels = unique(mech_map$metric)),
         scope  = factor(scope, levels = c("all", "minority", "majority")))

ggplot(df_mech, aes(r_ingroup, value, colour = scope)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  facet_grid(r_background_label + metric ~ condition_label, scales = "free_y") +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_x_continuous(breaks = R_INGROUP_VALS) +
  labs(title = "Chain length & ideological drift -- all vs. minority vs. majority, by background r",
       subtitle = SUBTITLE, x = "r_ingroup", y = NULL, colour = "scope") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```
- decrease of chain length of minority probably due to cycles.
- ideological drift not lowest for minority in the homophily conditions seems not right.

---

# 5  Robustness Check: Uniform Opinion Distribution

Same model, metrics and plots as Sections 3-4, using the
uniform-distribution study design introduced at the top of this report
(*Opinion Distributions by Condition (Uniform-Distribution Design)*).

```{r sim-run-one-uniform, include=FALSE}
run_cell_uniform <- function(seed, condition, r_ingroup, r_background) {
  if (condition == "random") {
    min_lo <- 0;   min_hi <- 1
    maj_lo <- 0;   maj_hi <- 1
    clustered <- FALSE  # no override -- relies on setup_agents()'s default U(0,1) draw
    th        <- NULL
  } else if (condition == "polarised") {
    min_lo <- 0;   min_hi <- 0.5
    maj_lo <- 0.5; maj_hi <- 1
    clustered <- TRUE
    th        <- NULL
  } else {  # homophily
    min_lo <- 0;   min_hi <- 0.5
    maj_lo <- 0.5; maj_hi <- 1
    clustered <- TRUE
    th        <- HOMOPHILY_TARGET_R
  }

  res <- simulate_liquid_democracy(
    seed                    = seed,
    n_per_community         = N_SIM,
    n_communities           = 1,
    node_degree             = K_SIM,
    T                       = SIM_T,
    r_op                    = r_background,
    r_pw                    = r_background,
    p_rewire                = P_REWIRE,
    r_ingroup               = r_ingroup,
    minority_share          = MINORITY_SHARE,
    opinion_dist            = "uniform",
    minority_opinion_mu     = if (clustered) (min_lo + min_hi) / 2 else NULL,
    minority_opinion_min    = min_lo,
    minority_opinion_max    = min_hi,
    majority_opinion_mu     = if (clustered) (maj_lo + maj_hi) / 2 else NULL,
    majority_opinion_min    = maj_lo,
    majority_opinion_max    = maj_hi,
    self_weight_mode        = SELF_WEIGHT_MODE,
    confidence_agg          = CONFIDENCE_AGG,
    sigma_opinion           = "auto",
    target_homophily        = th,
    gamma                   = r_background,
    lambda                  = LAMBDA_DIAG,
    trust_mode              = TRUST_MODE,
    cycle_penalty           = CYCLE_PENALTY
  )

  snaps <- res$snapshots |> dplyr::filter(round >= SIM_T - 4L)

  tibble(
    seed                         = seed,
    condition                    = condition,
    r_ingroup                    = r_ingroup,
    r_background                 = r_background,
    delegation_rate              = mean(snaps$delegation_rate, na.rm = TRUE),
    lost_vote_rate                = mean(snaps$lost_vote_rate,  na.rm = TRUE),
    avg_chain_length_delegators  = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
    avg_drift                    = mean(snaps$avg_drift,        na.rm = TRUE),
    minority_delegation_rate     = mean(snaps$minority_delegation_rate, na.rm = TRUE),
    majority_delegation_rate     = mean(snaps$majority_delegation_rate, na.rm = TRUE),
    minority_lost_vote_rate      = mean(snaps$minority_lost_vote_rate,  na.rm = TRUE),
    majority_lost_vote_rate      = mean(snaps$majority_lost_vote_rate,  na.rm = TRUE),
    minority_chain_len           = mean(snaps$minority_chain_len, na.rm = TRUE),
    majority_chain_len           = mean(snaps$majority_chain_len, na.rm = TRUE),
    minority_drift                = mean(snaps$minority_drift,    na.rm = TRUE),
    majority_drift                = mean(snaps$majority_drift,    na.rm = TRUE),
    minority_power                = mean(snaps$minority_power,    na.rm = TRUE),
    majority_power                = mean(snaps$majority_power,    na.rm = TRUE),
    RR_m                          = mean(snaps$RR_m,              na.rm = TRUE),
    RR_m_resolved                 = mean(snaps$RR_m_resolved,     na.rm = TRUE),
    minority_n_resolved            = mean(snaps$minority_n_resolved, na.rm = TRUE),
    majority_n_resolved            = mean(snaps$majority_n_resolved, na.rm = TRUE),
    minority_capture               = mean(snaps$minority_capture, na.rm = TRUE),
    majority_capture               = mean(snaps$majority_capture, na.rm = TRUE),
    cross_group_dlg_rate          = mean(snaps$cross_group_dlg_rate, na.rm = TRUE),
    cdr_min_to_maj                = mean(snaps$cdr_min_to_maj,    na.rm = TRUE),
    cdr_maj_to_min                = mean(snaps$cdr_maj_to_min,    na.rm = TRUE),
    minority_n_reps                = mean(snaps$minority_n_reps,  na.rm = TRUE),
    majority_n_reps                = mean(snaps$majority_n_reps,  na.rm = TRUE)
  )
}
```

```{r sim-grid-uniform, include=FALSE}
grid_random_u    <- expand_grid(condition = "random",
                                 r_ingroup = R_INGROUP_VALS, r_background = R_BACKGROUND_VALS, seed = seeds_sim)
grid_polarised_u <- expand_grid(condition = "polarised",
                                 r_ingroup = R_INGROUP_VALS, r_background = R_BACKGROUND_VALS, seed = seeds_sim)
grid_homophily_u <- expand_grid(condition = "homophily",
                                 r_ingroup = R_INGROUP_VALS, r_background = R_BACKGROUND_VALS, seed = seeds_sim)
grid_sim_u <- bind_rows(grid_random_u, grid_polarised_u, grid_homophily_u)

params_sim_u <- list(N = N_SIM, K = K_SIM, T = SIM_T, r_background_vals = R_BACKGROUND_VALS,
                      lambda = LAMBDA_DIAG, trust_mode = TRUST_MODE,
                      cycle_penalty = CYCLE_PENALTY,
                      r_ingroup_vals = R_INGROUP_VALS,
                      homophily_target = HOMOPHILY_TARGET_R,
                      minority_share = MINORITY_SHARE,
                      self_weight_mode = SELF_WEIGHT_MODE, confidence_agg = CONFIDENCE_AGG,
                      sigma_opinion = "auto", n_seeds = N_SEEDS_MINORITY,
                      opinion_dist = "uniform", version = 2L)
hash_sim_u  <- digest::digest(params_sim_u)
cache_sim_u <- here::here("data/sim_report19_uniform_cache.rds")

if (file.exists(cache_sim_u)) {
  cs_u <- readRDS(cache_sim_u)
  if (identical(cs_u$hash, hash_sim_u)) sim_raw_u <- cs_u$data else cs_u <- NULL
} else cs_u <- NULL

if (is.null(cs_u)) {
  dir.create(here::here("data"), showWarnings = FALSE)
  plan(multisession, workers = parallel::detectCores() - 1)

  sim_raw_u <- future_pmap_dfr(
    list(seed = grid_sim_u$seed, condition = grid_sim_u$condition,
         r_ingroup = grid_sim_u$r_ingroup, r_background = grid_sim_u$r_background),
    run_cell_uniform,
    .options = furrr_options(seed = NULL)
  )
  plan(sequential)
  saveRDS(list(hash = hash_sim_u, data = sim_raw_u), cache_sim_u)
}

sim_agg_u <- sim_raw_u |>
  group_by(condition, r_ingroup, r_background) |>
  summarise(across(where(is.numeric) & !any_of("seed"), \(x) mean(x, na.rm = TRUE)),
            .groups = "drop") |>
  mutate(
    condition_label = factor(condition, levels = c("random", "polarised", "homophily")),
    r_background_label = factor(paste0("r=", r_background), levels = paste0("r=", R_BACKGROUND_VALS)),
    minority_power_share_pct = 100 * minority_power / (minority_power + majority_power),
    majority_power_share_pct = 100 - minority_power_share_pct,
    minority_power_dev_pct   = minority_power_share_pct - 100 * MINORITY_SHARE,
    majority_power_dev_pct   = majority_power_share_pct - 100 * (1 - MINORITY_SHARE),
    all_n_reps                = minority_n_reps + majority_n_reps,
    RR_maj                    = (majority_power_share_pct / 100) / (1 - MINORITY_SHARE),
    RR_maj_resolved            = (majority_power_share_pct / 100) /
      (majority_n_resolved / (minority_n_resolved + majority_n_resolved))
  )

SUBTITLE_U <- sprintf("N = %d | k = %d | T = %d | %d seeds | uniform opinion distribution | confidence (sum) | trust: reward+punish",
                      N_SIM, K_SIM, SIM_T, N_SEEDS_MINORITY)
```

## 5.1  Representation Overview

```{r plot-part6-rates, fig.width=11, fig.height=16}
df_part1_rates_u <- rates_map |>
  pmap_dfr(function(metric, scope, column)
    sim_agg_u |> transmute(condition_label, r_background_label, r_ingroup, value = .data[[column]], metric, scope)) |>
  mutate(metric = factor(metric, levels = unique(rates_map$metric)),
         scope  = factor(scope, levels = c("all", "minority", "majority")))

ggplot(df_part1_rates_u, aes(r_ingroup, value, colour = scope)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  facet_grid(r_background_label + metric ~ condition_label, scales = "free_y") +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_x_continuous(breaks = R_INGROUP_VALS) +
  labs(title = "Delegation rate & lost-vote rate -- all vs. minority vs. majority, by background r (uniform distribution)",
       subtitle = SUBTITLE_U, x = "r_ingroup", y = NULL, colour = "scope") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

```{r plot-part6-rr-pc, fig.width=11, fig.height=16}
# See Section 3's plot-part1-rr-pc for the population-vs-resolved rationale.
df_rr_pc_u <- sim_agg_u |>
  select(condition_label, r_background_label, r_ingroup,
         RR_m, RR_maj, RR_m_resolved, RR_maj_resolved,
         PC_m = minority_capture, PC_Maj = majority_capture) |>
  pivot_longer(c(RR_m, RR_maj, RR_m_resolved, RR_maj_resolved, PC_m, PC_Maj),
               names_to = "series", values_to = "value") |>
  mutate(metric = factor(if_else(series %in% c("RR_m", "RR_maj", "RR_m_resolved", "RR_maj_resolved"), "RR", "PC"),
                          levels = c("RR", "PC")),
         scope  = factor(if_else(series %in% c("RR_m", "PC_m", "RR_m_resolved"), "minority", "majority"),
                          levels = c("minority", "majority")),
         denom  = factor(if_else(series %in% c("RR_m_resolved", "RR_maj_resolved"), "resolved", "population"),
                          levels = c("population", "resolved")))

ggplot(df_rr_pc_u, aes(r_ingroup, value, colour = scope, linetype = denom)) +
  geom_hline(data = tibble(metric = factor("RR", levels = c("RR", "PC")), value = 1),
             aes(yintercept = value), linetype = "dotted", colour = "grey50") +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  facet_grid(r_background_label + metric ~ condition_label, scales = "free_y") +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_linetype_manual(values = c(population = "solid", resolved = "dashed")) +
  scale_x_continuous(breaks = R_INGROUP_VALS) +
  labs(title = "RR_m / RR_maj (population vs. resolved denominator) and PC_m / PC_Maj, by background r (uniform distribution)",
       subtitle = SUBTITLE_U, x = "r_ingroup", y = NULL, colour = "scope", linetype = "RR denominator") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

## 5.2  Delegation Mechanisms

```{r plot-part6-cdr, fig.width=11, fig.height=9}
df_cdr_u <- sim_agg_u |>
  select(condition_label, r_background_label, r_ingroup, all = cross_group_dlg_rate,
         minority = cdr_min_to_maj, majority = cdr_maj_to_min) |>
  pivot_longer(c(all, minority, majority), names_to = "scope", values_to = "value") |>
  mutate(scope = factor(scope, levels = c("all", "minority", "majority")))

ggplot(df_cdr_u, aes(r_ingroup, value, colour = scope)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  facet_grid(r_background_label ~ condition_label, scales = "free_y") +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_x_continuous(breaks = R_INGROUP_VALS) +
  labs(title = "Cross-group delegation rate (all = population-wide; minority = M->Maj; majority = Maj->M), by background r (uniform distribution)",
       subtitle = SUBTITLE_U, x = "r_ingroup", y = "share of own edges crossing group", colour = "scope") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```

```{r plot-part6-chain-drift, fig.width=11, fig.height=16}
df_mech_u <- mech_map |>
  pmap_dfr(function(metric, scope, column)
    sim_agg_u |> transmute(condition_label, r_background_label, r_ingroup, value = .data[[column]], metric, scope)) |>
  mutate(metric = factor(metric, levels = unique(mech_map$metric)),
         scope  = factor(scope, levels = c("all", "minority", "majority")))

ggplot(df_mech_u, aes(r_ingroup, value, colour = scope)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 1.8) +
  facet_grid(r_background_label + metric ~ condition_label, scales = "free_y") +
  scale_colour_manual(values = GROUP_COLOURS) +
  scale_x_continuous(breaks = R_INGROUP_VALS) +
  labs(title = "Chain length & ideological drift -- all vs. minority vs. majority, by background r (uniform distribution)",
       subtitle = SUBTITLE_U, x = "r_ingroup", y = NULL, colour = "scope") +
  theme(strip.background = element_rect(fill = "grey93", colour = NA),
        strip.text       = element_text(face = "bold", size = 9),
        legend.position  = "bottom")
```
