---
title: "Weekly Report — Week 10 (24.04.2026 – 30.04.2026)"
output: github_document
date: "2026-04-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
```

<!--
  WORKFLOW NOTE
  ─────────────────────────────────────────────────────────────────
  Phase I  (Sections 1–5): Utility functions studied in isolation.
           No softmax, no self-weight, no simulation noise.
           Goal: understand the structural behaviour of the formulas
           before any decision mechanism is introduced.

  Phase II (Sections 6–9): The same functions are embedded in the
           full stochastic decision model from scripts/Network.R.
           Softmax and self-weight are reintroduced; outcomes are
           compared across a parameter grid.
  ─────────────────────────────────────────────────────────────────
-->

---

# Phase I — Utility Analysis (No Decision Rule)

---

## 1. Utility Functions: Model A and Model B

We compare two specifications for the utility agent $i$ assigns to delegating
to neighbour $j$. Both depend on the same two quantities:

$$\text{opinion similarity:} \quad (1 - |o_i - o_j|) \in [0, 1]$$

$$\text{log power ratio:} \quad \log\!\left(\frac{p_j}{p_i}\right) \in \mathbb{R}$$

**Model A — additive:**
$$V_{ij}^{(A)} \;=\; \alpha\,(1 - |o_i - o_j|) \;+\; \beta\,\log\!\left(\frac{p_j}{p_i}\right)$$

Opinion similarity and relative power contribute **independently**.
A powerful but ideologically distant neighbour can still accumulate a large
$\beta\log(p_j/p_i)$ term even when $(1-|o_i-o_j|)$ is small.

**Model B — interaction:**
$$V_{ij}^{(B)} \;=\; (1 - |o_i - o_j|)\,\left[\alpha \;+\; \beta\,\log\!\left(\frac{p_j}{p_i}\right)\right]$$

Opinion similarity acts as a **multiplicative gate** on the power term.
When agents are maximally dissimilar ($|o_i - o_j| = 1$), $V_{ij}^{(B)} = 0$
regardless of power. This encodes a strong behavioural assumption:
*power only matters if agents are similar*. Ideological distance is not
a penalty — it is a hard exclusion.

| | Model A | Model B |
|---|---|---|
| Structure | Additive, separable | $(1-\|o_i-o_j\|)$ gates the power term |
| Dissimilar, powerful neighbour | $V_{ij} = \beta\log(p_j/p_i) \neq 0$ | $V_{ij} = 0$ always |
| Models agree when | — | $\beta = 0$, $\log(p_j/p_i) = 0$, $o_i = o_j$  |

---

## 2. 1D Analysis — Effect of Opinion Similarity

We fix $\log(p_j/p_i)$ and study how utility varies with $(1-|o_i-o_j|)$.
Each panel holds $\alpha$, $\beta$, and $\log(p_j/p_i)$ constant; only
opinion similarity varies along the x-axis.

```{r plot-1d-similarity, fig.width=14, fig.height=9, dpi=300}
expand_grid(
  opinion_sim     = seq(0, 1, length.out = 200),
  log_power_ratio = c(-2, 0, 2),
  model           = c("A", "B"),
  alpha           = c(1, 2),
  beta            = c(1, 2)
) |>
  mutate(
    V = case_when(
      model == "A" ~ alpha * opinion_sim + beta * log_power_ratio,
      model == "B" ~ opinion_sim * (alpha + beta * log_power_ratio)
    ),
    lpr_label = paste0("log(p[j]/p[i]) == ", log_power_ratio),
    model     = paste("Model", model),
    param     = paste0("α = ", alpha, ", β = ", beta)
  ) |>
  ggplot(aes(opinion_sim, V, colour = lpr_label)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey50") +
  facet_grid(model ~ param) +
  scale_colour_brewer(
    palette = "Dark2",
    name    = NULL,
    labels  = scales::label_parse()
  ) +
  labs(
    x        = expression(1 - abs(o[i] - o[j])),
    y        = expression(V[ij]),
    title    = "Utility vs opinion similarity across parameter settings",
    subtitle = "Rows: Model A vs B | Columns: (α, β) combinations"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

**Model A** produces parallel lines: increasing opinion similarity raises
$V_{ij}$ at constant rate $\alpha$, while $\beta\log(p_j/p_i)$ adds a fixed
vertical offset independent of similarity. Even at zero similarity
($1-|o_i-o_j|=0$) the utility is nonzero whenever $\beta > 0$ — a
powerful but ideologically distant neighbour still attracts.
**Model B** forces all lines through the origin at zero similarity: the
entire expression is multiplied by $(1-|o_i-o_j|)$, so maximal dissimilarity
always yields $V_{ij}=0$ regardless of power or $\alpha$.
Larger $\beta$ (right columns) widens the gap between $\log(p_j/p_i)$ levels
in both models; larger $\alpha$ (bottom rows) shifts Model A upward uniformly
and steepens Model B's slope.

---

## 3. 1D Analysis — Effect of the Power Ratio

We fix $(1-|o_i-o_j|)$ and study how utility varies with $\log(p_j/p_i)$.
Each panel holds $\alpha$, $\beta$, and opinion similarity constant; only
the log power ratio varies along the x-axis.

```{r plot-1d-power, fig.width=14, fig.height=9, dpi=300}
expand_grid(
  log_power_ratio = seq(-3, 3, length.out = 200),
  opinion_sim     = c(0, 0.25, 0.5, 0.75, 1),
  model           = c("A", "B"),
  alpha           = c(1, 2),
  beta            = c(1, 2)
) |>
  mutate(
    V = case_when(
      model == "A" ~ alpha * opinion_sim + beta * log_power_ratio,
      model == "B" ~ opinion_sim * (alpha + beta * log_power_ratio)
    ),
    sim_label = paste0("1-abs(o[i]-o[j]) == ", opinion_sim),
    model     = paste("Model", model),
    param     = paste0("α = ", alpha, ", β = ", beta)
  ) |>
  ggplot(aes(log_power_ratio, V, colour = sim_label)) +
  geom_line(linewidth = 0.9) +
  geom_hline(yintercept = 0, linetype = "dotted", colour = "grey50") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey70") +
  facet_grid(model ~ param) +
  scale_colour_brewer(
    palette = "RdYlGn",
    direction = -1,
    name   = "Opinion similarity",
    labels = scales::label_parse()
  ) +
  labs(
    x        = expression(log(p[j] / p[i])),
    y        = expression(V[ij]),
    title    = "Utility vs log power ratio across parameter settings",
    subtitle = "Rows: Model A vs B | Columns: (α, β) combinations"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

**Model A** produces perfectly parallel lines with slope $\beta$: the log
power ratio has the same influence at every similarity level, and the lines
are shifted vertically by $\alpha \cdot (1-|o_i-o_j|)$. Agents with zero
similarity ($1-|o_i-o_j|=0$, green line) still respond to power.
**Model B** shows a fan converging to zero: at full similarity the slope
equals $\beta$ (matching Model A), but it contracts proportionally as
similarity falls, reaching a completely flat zero line for dissimilar agents.
The vertical dashed line marks equal power ($\log(p_j/p_i)=0$); the dotted
horizontal marks $V_{ij}=0$.

---

## 4. 2D Utility Surface

Both inputs vary simultaneously. Red indicates high utility (agent $i$ strongly
prefers to delegate to $j$), blue indicates low or negative utility, white marks
$V_{ij}=0$. Each panel is one $(\alpha,\beta)$ combination.

```{r plot-2d-heatmap, fig.width=9, fig.height=4, echo=FALSE}
expand_grid(
  opinion_sim     = seq(0, 1, length.out = 150),
  log_power_ratio = seq(-3, 3, length.out = 150),
  model           = c("A", "B"),
  alpha           = c(1, 2),
  beta            = c(1, 2)
) |>
  mutate(
    V = if_else(
      model == "A",
      alpha * opinion_sim + beta * log_power_ratio,
      opinion_sim * (alpha + beta * log_power_ratio)
    ),
    model = paste("Model", model),
    param = paste0("α = ", alpha, ", β = ", beta)
  ) |>
  ggplot(aes(opinion_sim, log_power_ratio, fill = V)) +
  geom_tile() +
  facet_grid(model ~ param) +
  scale_fill_gradient2(
    low = "steelblue",
    mid = "white",
    high = "firebrick",
    midpoint = 0,
    name = expression(V[ij])
  ) +
  labs(
    x        = expression(1 - abs(o[i] - o[j])),
    y        = expression(log(p[j] / p[i])),
    title    = "Utility surfaces across parameter settings",
    subtitle = "Rows: Model A vs B | Columns: (α, β) combinations"
  ) +
  theme_minimal(base_size = 12)
```

The left column ($1-|o_i-o_j|=0$) makes the key difference concrete:
Model A shows a continuous gradient driven by $\log(p_j/p_i)$;
Model B is uniformly white (zero), eliminating all dissimilar agents from
the choice set irrespective of their power.

---

## 5. Limit Behaviour

```{r limit-table, echo=FALSE}
tibble::tribble(
  ~Condition,
  ~V_A,
  ~V_B,
  ~Implication,

  "|o_i - o_j| = 1 (max. dissimilar)",
  "beta * log(p_j / p_i)",
  "0",
  "B ignores dissimilar agents entirely",

  "o_i = o_j (identical opinions)",
  "alpha + beta * log(p_j / p_i)",
  "alpha + beta * log(p_j / p_i)",
  "Models are identical",

  "beta = 0",
  "alpha * (1 - |o_i - o_j|)",
  "alpha * (1 - |o_i - o_j|)",
  "Models are identical; pure homophily",

  "alpha = 0",
  "beta * log(p_j / p_i)",
  "(1 - |o_i - o_j|) * beta * log(p_j / p_i)",
  "B still couples power to similarity",

  "p_j >> p_i, beta > 0, o_i ≈ o_j",
  "→ +∞",
  "→ +∞",
  "Both: powerful similar agents dominate",

  "p_j >> p_i, beta > 0, |o_i - o_j| = 1",
  "→ +∞",
  "0",
  "A: cross-ideological pull; B: no effect"
) |>
  knitr::kable(
    col.names = c("Condition", "V_A", "V_B", "Implication"),
    caption   = "Limit behaviour of both utility functions."
  )
```

Models diverge only when $\beta \neq 0$ **and** opinion heterogeneity is present.
The two bold rows define the critical qualitative difference.

---

# Phase II — Full Decision Model and Simulation

---

## 6. Softmax Decision Rule and Self-Weight

Having characterised the utility functions in isolation, we now embed them in
a stochastic choice model. Agent $i$ selects from
$\{\text{self}\} \cup N(i)$ by drawing proportional to unnormalised weights:

$$w_{\text{self},i} = k_i\,e^\delta, \qquad
  w_{ij} = \exp\!\left(V_{ij}\right)$$

where $k_i$ is agent $i$'s degree and $\delta$ is the **self-preference
parameter**. These weights are positive scores — not yet probabilities.
The softmax converts them into a proper distribution over the choice set:

$$P_i(k) = \frac{w_k}{w_{\text{self},i} + \displaystyle\sum_{j \in N(i)} w_{ij}}$$

so all options compete on a common scale and probabilities sum to 1.
In particular, $P_i(\text{self})$ and each $P_i(j)$ for $j \in N(i)$
are given by substituting the respective weight in the numerator.

At baseline ($\alpha = \beta = 0$, all $w_{ij}=1$):

$$P_i(\text{self}) \;=\; \frac{k_i e^\delta}{k_i e^\delta + k_i}
\;=\; \frac{e^\delta}{e^\delta + 1} \;=\; \sigma(\delta)$$

The degree term cancels, so $\sigma(\delta)$ is the baseline self-vote
probability regardless of network structure. This gives interpretable targets:

| $\delta$ | Baseline self-vote | Baseline delegation |
|---|---|---|
| $-2.2$ | 10 % | **~90 %** |
| $-1.4$ | 20 % | **~80 %** |
| $-0.85$ | 30 % | **~70 %** |
| $-0.4$ | 40 % | **~60 %** |

With $\alpha > 0$ or $\beta > 0$, neighbours become more attractive and actual
delegation rates rise above the baseline.

---


## 7. Simulation 

```{r source-network, include=FALSE}
source(here::here("scripts/Network.R"))
```


```{r wrapper-functions}
# Returns an attractiveness_fn for Model A (additive)
make_attract_A <- function(alpha, beta) {
  function(op_i, op_j, pow_i, pow_j, resp) {
    opinion_sim     <- 1 - abs(op_i - op_j)
    log_power_ratio <- log(pow_j / pow_i)
    exp(alpha * opinion_sim + beta * log_power_ratio)
  }
}

# Returns an attractiveness_fn for Model B (interaction)
make_attract_B <- function(alpha, beta) {
  function(op_i, op_j, pow_i, pow_j, resp) {
    opinion_sim     <- 1 - abs(op_i - op_j)
    log_power_ratio <- log(pow_j / pow_i)
    exp(opinion_sim * (alpha + beta * log_power_ratio))
  }
}

# Returns a selfweight_fn with degree-normalised self-preference delta
make_selfweight_delta <- function(delta) {
  function(i, nb, op, pow, resp) {
    length(nb) * exp(delta)   # = k_i * exp(delta)
  }
}
```

---

## 8. Running Both Models

```{r sim-grid, cache=TRUE}
# Parameter grid
# delta values calibrated to baseline delegation ~90 / 80 / 70 / 60 %
delta_map <- c("~90%" = -2.2, "~80%" = -1.4, "~70%" = -0.85, "~60%" = -0.4)

param_grid <- expand_grid(
  alpha    = c(0, 1, 2),
  beta     = c(0, 1, 2),
  baseline = names(delta_map)
) |>
  mutate(delta = delta_map[baseline])

# Run one simulation, return the final-round snapshot metrics
run_one <- function(alpha, beta, delta, model_label, seed = 42) {
  attr_fn <- if (model_label == "A") make_attract_A(alpha, beta)
             else                    make_attract_B(alpha, beta)

  res <- simulate_liquid_democracy(
    seed                    = seed,
    n_per_community         = 100,
    n_communities           = 1,
    node_degree             = 6,
    T                       = 100,
    attractiveness_fn       = attr_fn,
    selfweight_fn           = make_selfweight_delta(delta),
    responsiveness          = 1   # unused by our fns; kept for compatibility
  )

  snap <- res$snapshots |> slice_tail(n = 1)
  tibble(
    model           = model_label,
    delegation_rate = snap$delegation_rate,
    lost_vote_rate  = snap$lost_vote_rate
  )
}

# Run both models on every parameter combination (same seed = same network)
results <- param_grid |>
  rowwise() |>
  mutate(
    res_A = list(run_one(alpha, beta, delta, "A")),
    res_B = list(run_one(alpha, beta, delta, "B"))
  ) |>
  ungroup() |>
  pivot_longer(c(res_A, res_B), names_to = "tmp", values_to = "res") |>
  unnest(res) |>
  select(-tmp)
```

---

## 9. Results

### Comparison table

Each row is one $(\alpha, \beta, \delta)$ combination run on the same
network (seed = 42). **Del.** is the share of agents who delegated in
the final round; **Lost** is the share of votes that ended in a cycle
or reached a node with no outgoing delegation and were therefore lost.

```{r results-table}
results |>
  pivot_wider(
    names_from  = model,
    values_from = c(delegation_rate, lost_vote_rate)
  ) |>
  mutate(across(where(is.numeric) & !matches("delta"),
                \(x) round(x, 2))) |>
  select(alpha, beta, baseline, starts_with("delegation"), starts_with("lost")) |>
  arrange(baseline, alpha, beta) |>
  knitr::kable(
    col.names = c(
      "alpha", "beta", "Baseline del.",
      "Del. A", "Del. B", "Lost A", "Lost B"
    ),
    caption = "Final-round metrics for Model A and Model B (n = 100, T = 100 rounds)."
  )
```

---

## 10. Key Findings

**From the utility analysis (Phase I):**

- In **Model A**, opinion similarity and power contribute independently.
  A powerful but dissimilar agent retains a nonzero utility driven entirely
  by $\beta\log(p_j/p_i)$.
- In **Model B**, $(1-|o_i-o_j|)$ gates the power term. A dissimilar agent
  is assigned $V_{ij}=0$ regardless of power. This is not a technical
  variant but a distinct behavioural assumption: *trust must exist before
  power becomes relevant*.
- The two models coincide when $\beta = 0$ (power ignored) or when all
  agent pairs are opinion-identical.

**From the simulation (Phase II):**

- $\delta$ controls the baseline delegation volume; $\alpha$ and $\beta$
  shift rates upward by making neighbours more attractive.
- **Model B produces lower delegation rates and fewer lost votes** whenever
  $\beta > 0$. By excluding dissimilar agents, it keeps delegation chains
  shorter and within ideologically coherent sub-groups, reducing cycle risk.
- **Model A allows cross-ideological power-driven delegation**, which can
  form long chains spanning opinion clusters — the primary source of its
  elevated lost-vote rate at high $\beta$.
  
## 11. Model C — Soft Gate
    
$$
V_{ij}^{(C)} \;=\; \alpha\,(1 - |o_i - o_j|) \;+\; \beta\,(1 - |o_i - o_j|)^\gamma\,\log\!\left(\frac{p_j}{p_i}\right), \qquad \gamma > 0
$$

Model C nests A ($\gamma = 0$) and B ($\gamma = 1$) as special cases.
The exponent $\gamma$ continuously controls how aggressively ideological
distance suppresses power-driven delegation — interpolating between the
fully independent structure of A and the hard-exclusion logic of B.
