---
title: "Weekly Report — Week 12 (08.05.2026 – 14.05.2026)"
output: github_document
date: "2026-05-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(furrr)
library(patchwork)

sigmoid <- function(x) 1 / (1 + exp(-x))
```

## 1. Design

Full **6 × 6 parameter grid** ($r_{op}, r_{pw} \in \{0,\ldots,5\}$) ×
**4 agent counts** ($n \in \{100, 250, 1000, 2500\}$), 100 seeds each.
All metrics are averaged over rounds 15–20.

$$A_{ij} = \sigma\!\left(r_{op}(1-2|o_i-o_j|)\right)\cdot\sigma\!\left(r_{pw}\log\tfrac{p_j}{p_i}\right)$$

$$w_{\text{self}}(i) = \sigma\!\left(r_{op}(2|o_i-o_{j^\ast}|-1)\right)\cdot\sigma\!\left(r_{pw}\log\tfrac{p_i}{p_{j^\ast}}\right)$$

---

```{r grid-setup, include=FALSE}
source(here::here("scripts/Network.R"))

metric_info <- tribble(
  ~col,             ~label,                      ~pct,
  "delegation_rate","Delegation Rate",            TRUE,
  "lost_vote_rate", "Lost Vote Rate",             TRUE,
  "avg_chain_len",  "Avg. Chain Length",          FALSE,
  "avg_ideo_drift", "Avg. Ideological Drift",     FALSE,
  "gini_coef",      "Gini Coefficient (Power)",   FALSE
)
```

```{r grid-run, cache=TRUE}
plan(multisession, workers = parallel::detectCores() - 1)

full_grid <- expand_grid(
  n_agents = c(100L, 250L, 1000, 2500L),
  r_op     = 0:5,
  r_pw     = 0:5,
  seed     = seq_len(100)
)

.t0_grid <- Sys.time()

sim_grid_raw <- future_pmap_dfr(
  full_grid,
  function(n_agents, r_op, r_pw, seed) {

    res <- simulate_liquid_democracy(
      seed            = seed,
      n_per_community = n_agents,
      n_communities   = 1,
      node_degree     = 6,
      T               = 20,
      r_op            = r_op,
      r_pw            = r_pw
    )

    snaps <- tail(res$snapshots, 2)
    ag    <- res$agents

    tibble(
      n_agents        = n_agents,
      r_op            = r_op,
      r_pw            = r_pw,
      seed            = seed,
      delegation_rate = mean(res$history_delegation[15:20],  na.rm = TRUE),
      lost_vote_rate  = mean(res$history_lost[15:20],        na.rm = TRUE),
      avg_ideo_drift  = mean(res$history_drift[15:20],       na.rm = TRUE),
      gini_coef       = mean(snaps$gini_power,               na.rm = TRUE),
      avg_chain_len   = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
      pct_high_drift  = mean(abs(ag$opinion - ag$my_vote) > 0.3, na.rm = TRUE)
    )
  },
  .options = furrr_options(seed = NULL)
)

plan(sequential)
.t1_grid  <- Sys.time()
.el_grid  <- as.numeric(difftime(.t1_grid, .t0_grid, units = "mins"))
.n_grid   <- nrow(full_grid)
cat(sprintf(
  "\n── grid-run ──────────────────────────────────────────\n  Combinations : %d n_agents × 6 r_op × 6 r_pw = %d\n  Seeds        : 100\n  Total runs   : %s\n  Workers      : %d\n  Elapsed      : %.1f min  (%.2f s / run)\n──────────────────────────────────────────────────────\n",
  4L, 4L * 36L,
  formatC(.n_grid, format = "d", big.mark = ","),
  parallel::detectCores() - 1L,
  .el_grid, .el_grid * 60 / .n_grid
))

grid_summary <- sim_grid_raw |>
  group_by(n_agents, r_op, r_pw) |>
  summarise(
    across(
      delegation_rate:pct_high_drift,
      list(mean = \(x) mean(x, na.rm = TRUE),
           sd   = \(x) sd(x,   na.rm = TRUE)),
      .names = "{.col}__{.fn}"
    ),
    .groups = "drop"
  ) |>
  mutate(
    delegation_rate__cv = if_else(delegation_rate__mean > 1e-9, delegation_rate__sd / delegation_rate__mean, NA_real_),
    lost_vote_rate__cv  = if_else(lost_vote_rate__mean  > 1e-9, lost_vote_rate__sd  / lost_vote_rate__mean,  NA_real_),
    avg_ideo_drift__cv  = if_else(avg_ideo_drift__mean  > 1e-9, avg_ideo_drift__sd  / avg_ideo_drift__mean,  NA_real_),
    gini_coef__cv       = if_else(gini_coef__mean       > 1e-9, gini_coef__sd       / gini_coef__mean,       NA_real_),
    avg_chain_len__cv   = if_else(avg_chain_len__mean   > 1e-9, avg_chain_len__sd   / avg_chain_len__mean,   NA_real_)
  )

# Baseline for main analysis
g250 <- grid_summary |> filter(n_agents == 250)
```

---

## 2. Parameter Profiles ($n = 250$)

Solid lines: cross-seed mean. Ribbons: mean ± 1 SD.

```{r line-plots, fig.width=9, fig.height=5}
for (i in seq_len(nrow(metric_info))) {
  col      <- metric_info$col[i]
  label    <- metric_info$label[i]
  pct      <- metric_info$pct[i]
  mean_col <- paste0(col, "__mean")
  sd_col   <- paste0(col, "__sd")

  p <- g250 |>
    mutate(r_op_f = factor(paste0("r[op]==", r_op))) |>
    ggplot(aes(
      x = r_pw, y = .data[[mean_col]],
      ymin = .data[[mean_col]] - .data[[sd_col]],
      ymax = .data[[mean_col]] + .data[[sd_col]],
      colour = r_op_f, fill = r_op_f, group = r_op_f
    )) +
    geom_ribbon(alpha = 0.12, colour = NA) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 1.8) +
    scale_x_continuous(breaks = 0:5) +
    scale_colour_brewer(palette = "Dark2", name = expression(r[op]),
                        labels = scales::label_parse()) +
    scale_fill_brewer(palette = "Dark2", guide = "none") +
    labs(x = expression(r[pw]), y = label, title = label) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "right")

  if (pct) p <- p + scale_y_continuous(
    labels = scales::percent_format(accuracy = 1)
  )

  print(p)
}
```

---

## 3. Mean & Variability Heatmaps ($n = 250$)

Left: cross-seed mean. Right: coefficient of variation (CV = SD/mean).
CV peaks at phase-boundary cells — high CV indicates that small differences
in initialisation are sufficient to tip the system between regimes.

```{r heatmaps, fig.width=13, fig.height=5.5}
for (i in seq_len(nrow(metric_info))) {
  col      <- metric_info$col[i]
  label    <- metric_info$label[i]
  pct      <- metric_info$pct[i]
  mean_col <- paste0(col, "__mean")
  cv_col   <- paste0(col, "__cv")

  fmt_mean <- if (pct) function(x) sprintf("%.0f%%", x * 100) else
                        function(x) sprintf("%.2f", x)

  p_mean <- g250 |>
    mutate(lbl = fmt_mean(.data[[mean_col]])) |>
    ggplot(aes(factor(r_pw), factor(r_op), fill = .data[[mean_col]])) +
    geom_tile(colour = "white", linewidth = 0.4) +
    geom_text(aes(label = lbl), size = 2.8, colour = "white", fontface = "bold") +
    scale_fill_viridis_c(
      option = "C", name = "Mean",
      labels = if (pct) scales::percent_format(accuracy = 1) else waiver()
    ) +
    labs(x = expression(r[pw]), y = expression(r[op]),
         title = label, subtitle = "Mean") +
    theme_minimal(base_size = 11) +
    theme(panel.grid = element_blank())

  p_cv <- g250 |>
    mutate(lbl = sprintf("%.2f", pmin(.data[[cv_col]], 2, na.rm = TRUE))) |>
    ggplot(aes(factor(r_pw), factor(r_op), fill = .data[[cv_col]])) +
    geom_tile(colour = "white", linewidth = 0.4) +
    geom_text(aes(label = lbl), size = 2.8, colour = "white", fontface = "bold") +
    scale_fill_viridis_c(option = "A", name = "CV",
                         direction = -1, limits = c(0, NA)) +
    labs(x = expression(r[pw]), y = expression(r[op]),
         subtitle = "Coefficient of Variation (SD / mean)") +
    theme_minimal(base_size = 11) +
    theme(panel.grid = element_blank())

  print(p_mean + p_cv)
}
```

---

## 4. Varying N 


```{r robustness-heatmaps, fig.width=14, fig.height=8}
rob_metrics <- tribble(
  ~col,             ~label,
  "delegation_rate","Delegation Rate",
  "lost_vote_rate", "Lost Vote Rate",
  "avg_chain_len",  "Avg. Chain Len.",
  "avg_ideo_drift", "Avg. Ideo. Drift",
  "gini_coef",      "Gini Coefficient "
)

for (i in seq_len(nrow(rob_metrics))) {
  col      <- rob_metrics$col[i]
  label    <- rob_metrics$label[i]
  mean_col <- paste0(col, "__mean")
  pct      <- col == "delegation_rate"

  fmt <- if (pct) function(x) sprintf("%.0f%%", x * 100) else
                   function(x) sprintf("%.2f", x)

p <- grid_summary |>
  mutate(
    lbl   = fmt(.data[[mean_col]]),

    n_lab = factor(
      paste0("n = ", n_agents),
      levels = paste0("n = ", c(100, 250, 1000, 2500))
    )
  ) |>
  ggplot(aes(factor(r_pw), factor(r_op), fill = .data[[mean_col]])) +
    geom_tile(colour = "white", linewidth = 0.35) +
    geom_text(aes(label = lbl), size = 2.3, colour = "white", fontface = "bold") +
    facet_wrap(~ n_lab, ncol = 2) +
    scale_fill_viridis_c(
      option = "C", name = label,
      labels = if (pct) scales::percent_format(accuracy = 1) else waiver()
    ) +
    labs(x = expression(r[pw]), y = expression(r[op]), title = label) +
    theme_minimal(base_size = 11) +
    theme(panel.grid = element_blank(), legend.position = "right")

  print(p)
}
```

---

## 5. Extension: Outcome-Aware Delegation via Trust

### Formulas

**Trust update** (after each round, using the effective vote propagated through
$j$'s chain):

$$\tau_{ij}(t) = \lambda \cdot \tau_{ij}(t-1) \;-\; (1-\lambda)\cdot\bigl|o_i - v_j(t-1)\bigr|$$

**Modified attractiveness:**

$$\tilde{A}_{ij}(t) = A_{ij} \cdot 2\,\sigma\!\bigl(\gamma\cdot\tau_{ij}(t)\bigr)$$

where $A_{ij} = \sigma(r_{op}(1-2|o_i-o_j|))\cdot\sigma(r_{pw}\log(p_j/p_i))$
is the baseline attractiveness. The factor $2\sigma(\cdot)$ is
chosen so that $\tau=0$ (no experience) gives a neutral modifier of
$2\sigma(0)=1$, recovering the unmodified $A_{ij}$ exactly. The self-weight
$w_\text{self}$ is unchanged.

| Parameter | Interpretation | Values tested |
|---|---|---|
| $\lambda \in [0,1]$ | Memory length — how much past experience persists | 0.1, 0.5, 0.9 |
| $\gamma \geq 0$ | Trust sensitivity — how strongly trust history overrides static $A_{ij}$ | 0, 1, 2, 3 |
| $r_{op}$ | Opinion sensitivity (fixed) | 2 |
| $r_{pw}$ | Power sensitivity (fixed) | 2 |

---

```{r sim-function, include=FALSE}
# Custom simulation implementing the trust mechanism.
# Cannot use the attractiveness_fn interface from simulate_liquid_democracy()
# because trust requires agent *indices* (not just values) and per-round
# feedback from effective votes — neither is accessible through that API.

simulate_with_trust <- function(seed, n_agents, node_degree, T,
                                r_op, r_pw, lambda, gamma) {

  sig <- function(x) 1 / (1 + exp(-x))
  gini_fn <- function(x) {
    if (sum(x) == 0) return(0)
    x <- sort(x); n <- length(x)
    sum((2 * seq_len(n) - n - 1) * x) / (n * sum(x))
  }

  # ── Network setup ─────────────────────────────────────────────────────────
  st      <- setup_agents(n_agents, 1L, 0L, seed)
  agents  <- st$agents
  n       <- st$n_all

  gF      <- setup_friendship_network(agents, 1L, node_degree, 0.05, seed)
  adj     <- lapply(seq_len(n), \(v) as.integer(neighbors(gF, v, mode = "out")))
  lay_ids <- which(agents$type == "lay")
  op      <- agents$opinion

  # ── Trust initialisation ──────────────────────────────────────────────────
  # tau[[i]] is a numeric vector of length(adj[[i]]):
  # tau[[i]][k] = trust of agent i in its k-th neighbour adj[[i]][k]
  # Initialised at 0 → neutral modifier 2*sigma(0) = 1
  tau <- lapply(seq_len(n), \(i) rep(0.0, length(adj[[i]])))

  # ── History storage ───────────────────────────────────────────────────────
  hist_mat <- matrix(NA_real_, nrow = T, ncol = 7,
    dimnames = list(NULL, c("delegation_rate", "lost_vote_rate",
                            "avg_ideo_drift",  "gini_coef",
                            "avg_chain_len",   "stability",
                            "tau_mean")))
  prev_targets <- integer(n)

  # ── Main simulation loop ──────────────────────────────────────────────────
  for (t in seq_len(T)) {

    pow <- agents$power

    # Delegation decision
    targets <- vapply(lay_ids, function(i) {
      nb <- adj[[i]]
      if (!length(nb)) return(i)

      op_nb  <- op[nb]
      pow_nb <- pow[nb]

      # Baseline attractiveness (Report-12 formulas)
      a_nb <- sig(r_op * (1 - 2 * abs(op[i] - op_nb))) *
              sig(r_pw * log(pow_nb / pow[i]))

      # Self-weight: argmax formulation
      j_star <- which.max(a_nb)
      w_self <- sig(r_op * (2 * abs(op[i] - op_nb[j_star]) - 1)) *
                sig(r_pw * log(pow[i] / pow_nb[j_star]))

      # Trust-modified attractiveness: 2*sigma(gamma*tau) — neutral at tau=0
      w_nb <- a_nb * (2 * sig(gamma * tau[[i]]))

      w <- pmax(c(w_self, w_nb), 0)
      if (sum(w) == 0) w[] <- 1
      c(i, nb)[sample.int(length(w), 1L, prob = w)]
    }, integer(1L))

    # Delegation graph
    mask      <- targets != lay_ids
    edge_from <- lay_ids[mask]
    edge_to   <- targets[mask]

    if (length(edge_from)) {
      gD <- graph_from_edgelist(cbind(edge_from, edge_to), directed = TRUE)
      gD <- add_vertices(gD, max(0L, n - vcount(gD)))
    } else {
      gD <- make_empty_graph(n = n, directed = TRUE)
    }

    agents$power     <- compute_power(n, edge_from, edge_to)
    agents$my_vote   <- propagate_votes(op, n, edge_from, edge_to)
    agents$delegated <- seq_len(n) %in% edge_from

    # ── Trust update ─────────────────────────────────────────────────────────
    # v_j^eff = effective vote cast through j's delegation chain.
    # Disagreement with own opinion → accumulates negative trust.
    # Cycles (NA vote): treated as maximum disagreement (= 1).
    for (i in lay_ids) {
      nb <- adj[[i]]
      if (!length(nb)) next
      disagreement       <- abs(op[i] - agents$my_vote[nb])
      disagreement[is.na(disagreement)] <- 1
      tau[[i]] <- lambda * tau[[i]] - (1 - lambda) * disagreement
    }

    # ── Metrics ───────────────────────────────────────────────────────────────
    represented <- agents[!is.na(agents$my_vote), ]

    curr_targets <- integer(n)
    el_curr      <- as_edgelist(gD, names = FALSE)
    if (nrow(el_curr)) curr_targets[el_curr[, 1]] <- el_curr[, 2]

    hist_mat[t, "delegation_rate"] <- mean(agents$delegated[lay_ids])
    hist_mat[t, "lost_vote_rate"]  <- mean(is.na(agents$my_vote))
    hist_mat[t, "avg_ideo_drift"]  <- if (nrow(represented) > 0)
      mean(abs(represented$opinion - represented$my_vote)) else NA_real_
    hist_mat[t, "gini_coef"]       <- gini_fn(agents$power)
    hist_mat[t, "avg_chain_len"]   <- if (length(edge_from) > 0) {
      roots_d <- which(degree(gD, mode = "out") == 0)
      if (length(roots_d) > 0) {
        dm  <- distances(gD, v = edge_from, to = roots_d, mode = "out")
        cl  <- apply(dm, 1, \(d) { fd <- d[is.finite(d)]; if (length(fd)) min(fd) else NA_real_ })
        good <- cl[!is.na(cl)]
        if (length(good)) mean(good) else 0
      } else 0
    } else 0
    hist_mat[t, "stability"]       <- if (t > 1)
      mean(prev_targets[lay_ids] == curr_targets[lay_ids]) else NA_real_
    hist_mat[t, "tau_mean"]        <- mean(unlist(tau[lay_ids]))

    prev_targets <- curr_targets
  }

  as_tibble(hist_mat) |>
    mutate(round = seq_len(T), .before = 1)
}
```

```{r trust-setup, include=FALSE}
# Fixed baseline at phase boundary from Report 12
R_OP <- 2
R_PW <- 2
T_SIM  <- 20   # rounds per run
SS_WIN <- 15:20 # steady-state window

metric_info <- tribble(
  ~col,             ~label,                ~pct,
  "delegation_rate","Delegation Rate",      TRUE,
  "lost_vote_rate", "Lost Vote Rate",       TRUE,
  "avg_ideo_drift", "Avg. Ideological Drift", FALSE,
  "gini_coef",      "Gini Coefficient",     FALSE,
  "tau_mean",       "Mean Trust (τ̄)",        FALSE
)
```

```{r trust-run, cache=TRUE}
plan(multisession, workers = parallel::detectCores() - 1)

full_grid <- expand_grid(
  lambda = c(0.1, 0.5, 0.9),
  gamma  = c(0, 1, 2, 3),
  seed   = seq_len(100)
)

.t0_trust <- Sys.time()

sim_raw <- future_pmap_dfr(
  full_grid,
  function(lambda, gamma, seed) {
    simulate_with_trust(
      seed       = seed,
      n_agents   = 250,
      node_degree = 6,
      T          = T_SIM,
      r_op       = R_OP,
      r_pw       = R_PW,
      lambda     = lambda,
      gamma      = gamma
    ) |>
      mutate(lambda = lambda, gamma = gamma, seed = seed)
  },
  .options = furrr_options(seed = NULL)
)

plan(sequential)
.t1_trust  <- Sys.time()
.el_trust  <- as.numeric(difftime(.t1_trust, .t0_trust, units = "mins"))
.n_trust   <- nrow(full_grid)
cat(sprintf(
  "\n── trust-run ─────────────────────────────────────────\n  Combinations : 3 lambda × 4 gamma = 12\n  Seeds        : 100\n  Total runs   : %s\n  Workers      : %d\n  Elapsed      : %.1f min  (%.2f s / run)\n──────────────────────────────────────────────────────\n",
  formatC(.n_trust, format = "d", big.mark = ","),
  parallel::detectCores() - 1L,
  .el_trust, .el_trust * 60 / .n_trust
))

# Time-series summary: mean ± SD per (lambda, gamma, round) across seeds
ts_summary <- sim_raw |>
  group_by(lambda, gamma, round) |>
  summarise(
    across(
      delegation_rate:tau_mean,
      list(mean = \(x) mean(x, na.rm = TRUE),
           sd   = \(x) sd(x,   na.rm = TRUE)),
      .names = "{.col}__{.fn}"
    ),
    .groups = "drop"
  )

# Steady-state summary: mean ± SD per (lambda, gamma) over SS_WIN rounds
ss_summary <- sim_raw |>
  filter(round %in% SS_WIN) |>
  group_by(lambda, gamma, seed) |>
  summarise(
    across(delegation_rate:tau_mean, \(x) mean(x, na.rm = TRUE)),
    .groups = "drop"
  ) |>
  group_by(lambda, gamma) |>
  summarise(
    across(
      delegation_rate:tau_mean,
      list(mean = \(x) mean(x, na.rm = TRUE),
           sd   = \(x) sd(x,   na.rm = TRUE)),
      .names = "{.col}__{.fn}"
    ),
    .groups = "drop"
  ) |>
  mutate(
    delegation_rate__cv = if_else(delegation_rate__mean > 1e-9,
      delegation_rate__sd / delegation_rate__mean, NA_real_),
    lost_vote_rate__cv  = if_else(lost_vote_rate__mean  > 1e-9,
      lost_vote_rate__sd  / lost_vote_rate__mean,  NA_real_),
    avg_ideo_drift__cv  = if_else(avg_ideo_drift__mean  > 1e-9,
      avg_ideo_drift__sd  / avg_ideo_drift__mean,  NA_real_),
    gini_coef__cv       = if_else(gini_coef__mean       > 1e-9,
      gini_coef__sd       / gini_coef__mean,       NA_real_),
    tau_mean__cv        = if_else(abs(tau_mean__mean)   > 1e-9,
      tau_mean__sd  / abs(tau_mean__mean),          NA_real_)
  )
```

---

## 6. Trust Dynamics Over Time

Each panel shows one metric over 20 rounds for $\gamma \in \{0,1,2,3\}$
(colour) and $\lambda \in \{0.1, 0.5, 0.9\}$ (facet rows).
Ribbons: mean ± 1 SD across 100 seeds.
$\gamma = 0$ is the baseline (no trust modulation).

```{r time-series, fig.width=11, fig.height=9}
for (i in seq_len(nrow(metric_info))) {
  col      <- metric_info$col[i]
  label    <- metric_info$label[i]
  pct      <- metric_info$pct[i]
  mean_col <- paste0(col, "__mean")
  sd_col   <- paste0(col, "__sd")

  p <- ts_summary |>
    mutate(
      lambda_lab = paste0("λ == ", lambda),
      gamma_f    = factor(gamma)
    ) |>
    ggplot(aes(
      x    = round,
      y    = .data[[mean_col]],
      ymin = .data[[mean_col]] - .data[[sd_col]],
      ymax = .data[[mean_col]] + .data[[sd_col]],
      colour = gamma_f, fill = gamma_f
    )) +
    geom_ribbon(alpha = 0.12, colour = NA) +
    geom_line(linewidth = 0.85) +
    geom_vline(xintercept = min(SS_WIN), linetype = "dashed",
               colour = "grey50", linewidth = 0.4) +
    facet_wrap(~ lambda_lab, ncol = 1, labeller = label_parsed) +
    scale_colour_brewer(palette = "Dark2", name = expression(gamma)) +
    scale_fill_brewer(  palette = "Dark2", guide = "none") +
    labs(
      x        = "Round",
      y        = label,
      title    = label,
      subtitle = expression(r[op] == 2 ~~ r[pw] == 2 ~~ "(dashed line = steady-state window)")
    ) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "right")

  if (pct) p <- p + scale_y_continuous(
    labels = scales::percent_format(accuracy = 1)
  )

  print(p)
}
```

---

## 7. Steady-State Trust Parameter Profiles

Mean over rounds 15–20 (steady-state window), averaged across 100 seeds.
x-axis: $\gamma$ (trust sensitivity); colour: $\lambda$ (memory length).
Ribbons: mean ± 1 SD.

```{r ss-line-plots, fig.width=8, fig.height=4.5}
for (i in seq_len(nrow(metric_info))) {
  col      <- metric_info$col[i]
  label    <- metric_info$label[i]
  pct      <- metric_info$pct[i]
  mean_col <- paste0(col, "__mean")
  sd_col   <- paste0(col, "__sd")

  p <- ss_summary |>
    mutate(lambda_f = factor(paste0("λ == ", lambda))) |>
    ggplot(aes(
      x    = gamma,
      y    = .data[[mean_col]],
      ymin = .data[[mean_col]] - .data[[sd_col]],
      ymax = .data[[mean_col]] + .data[[sd_col]],
      colour = lambda_f, fill = lambda_f, group = lambda_f
    )) +
    geom_ribbon(alpha = 0.15, colour = NA) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2) +
    scale_x_continuous(breaks = c(0, 1, 2, 3)) +
    scale_colour_brewer(palette = "Set1", name = expression(lambda),
                        labels = scales::label_parse()) +
    scale_fill_brewer(  palette = "Set1", guide = "none") +
    labs(x = expression(gamma), y = label, title = label) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "right")

  if (pct) p <- p + scale_y_continuous(
    labels = scales::percent_format(accuracy = 1)
  )

  print(p)
}
```
---

## 8. Steady-State Comparison: Trust Sensitivity × Parameter Combinations

Steady-state means (rounds 15–20) for all $3 \times 3$ combinations of
$r_{op}, r_{pw} \in \{0,1,2\}$ as $\gamma$ increases from 0 to 3, with $\lambda = 0.5$
fixed throughout. Each line is one $(r_{op}, r_{pw})$ pair — colour encodes $r_{op}$,
line type encodes $r_{pw}$. $\gamma = 0$ is the no-trust baseline.
Averaged over 100 seeds per cell.

```{r ts-run, cache=TRUE}
plan(multisession, workers = parallel::detectCores() - 1)

final_grid <- expand_grid(
  r_op  = 0:2,
  r_pw  = 0:2,
  gamma = 0:3,
  seed  = seq_len(100)
)

.t0_ts <- Sys.time()

final_raw <- future_pmap_dfr(
  final_grid,
  function(r_op, r_pw, gamma, seed) {
    simulate_with_trust(
      seed        = seed,
      n_agents    = 250,
      node_degree = 6,
      T           = T_SIM,
      r_op        = r_op,
      r_pw        = r_pw,
      lambda      = 0.5,
      gamma       = gamma
    ) |>
      filter(round %in% SS_WIN) |>
      summarise(
        across(
          c(delegation_rate, lost_vote_rate, avg_ideo_drift,
            gini_coef, avg_chain_len, tau_mean),
          \(x) mean(x, na.rm = TRUE)
        )
      ) |>
      mutate(r_op = r_op, r_pw = r_pw, gamma = gamma, seed = seed)
  },
  .options = furrr_options(seed = NULL)
)

plan(sequential)
.t1_ts  <- Sys.time()
.el_ts  <- as.numeric(difftime(.t1_ts, .t0_ts, units = "mins"))
.n_ts   <- nrow(final_grid)
cat(sprintf(
  "\n── ts-run ────────────────────────────────────────────\n  Combinations : 3 r_op × 3 r_pw × 4 gamma = 36\n  Seeds        : 100\n  Total runs   : %s\n  Workers      : %d\n  Elapsed      : %.1f min  (%.2f s / run)\n──────────────────────────────────────────────────────\n",
  formatC(.n_ts, format = "d", big.mark = ","),
  parallel::detectCores() - 1L,
  .el_ts, .el_ts * 60 / .n_ts
))

final_ss <- final_raw |>
  group_by(r_op, r_pw, gamma) |>
  summarise(
    across(
      c(delegation_rate, lost_vote_rate, avg_ideo_drift,
        gini_coef, avg_chain_len, tau_mean),
      \(x) mean(x, na.rm = TRUE)
    ),
    .groups = "drop"
  )
```

```{r ts-plots, fig.width=11, fig.height=11}
final_metric_info <- tribble(
  ~col,             ~label,                   ~pct,
  "delegation_rate","Delegation Rate",          TRUE,
  "lost_vote_rate", "Lost Vote Rate",           TRUE,
  "avg_ideo_drift", "Avg. Ideological Drift",   FALSE,
  "gini_coef",      "Gini Coefficient",         FALSE,
  "avg_chain_len",  "Avg. Chain Length",        FALSE,
  "tau_mean",       "Mean Trust",               FALSE
)

final_long <- final_ss |>
  pivot_longer(
    cols      = c(delegation_rate, lost_vote_rate, avg_ideo_drift,
                  gini_coef, avg_chain_len, tau_mean),
    names_to  = "col",
    values_to = "mean"
  ) |>
  inner_join(final_metric_info, by = "col") |>
  mutate(
    label  = factor(label, levels = final_metric_info$label),
    r_op_f = factor(r_op),
    r_pw_f = factor(r_pw)
  )

ggplot(final_long,
       aes(x      = gamma,
           y      = mean,
           colour = r_op_f,
           linetype = r_pw_f,
           group  = interaction(r_op_f, r_pw_f))) +
  geom_line(linewidth = 0.85) +
  geom_point(size = 1.8) +
  facet_wrap(~ label, scales = "free_y", ncol = 2) +
  scale_x_continuous(breaks = 0:3) +
  scale_colour_brewer(
    palette = "Dark2",
    name    = expression(r[op])
  ) +
  scale_linetype_manual(
    values = c("solid", "dashed", "dotted"),
    name   = expression(r[pw])
  ) +
  labs(
    x        = expression(gamma ~ "(trust sensitivity)"),
    y        = NULL,
    title    = "Steady-state outcomes across parameter combinations",
    subtitle = expression(lambda == 0.5 ~ "(100 seeds, rounds 15-20)")
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "right",
    strip.text      = element_text(face = "bold")
  )
```

```{r timing-summary, echo=FALSE}
.total_runs <- .n_grid + .n_trust + .n_ts
.total_min  <- .el_grid + .el_trust + .el_ts
cat(sprintf(
  "\n══ SIMULATION SUMMARY ════════════════════════════════\n  grid-run   : %6s runs  —  %.1f min\n  trust-run  : %6s runs  —  %.1f min\n  ts-run     : %6s runs  —  %.1f min\n  ─────────────────────────────────────────────────────\n  TOTAL      : %6s runs  —  %.1f min  (%.1f h)\n══════════════════════════════════════════════════════\n",
  formatC(.n_grid,       format = "d", big.mark = ","), .el_grid,
  formatC(.n_trust,      format = "d", big.mark = ","), .el_trust,
  formatC(.n_ts,         format = "d", big.mark = ","), .el_ts,
  formatC(.total_runs,   format = "d", big.mark = ","), .total_min,
  .total_min / 60
))
```
