---
title: "Report 22, Part 2b -- T/n_voting_rounds calibration"
output: github_document
date: "2026-09-05"
---

> **Split note:** Runs directly after Part 2 (`Report_22_2_KDecision.Rmd`,
> which decides K) and before Part 3 (`Report_22_3_RSaturation.Rmd`, which
> sweeps r). Needs `CHOSEN_K_22` from Part 2, so the hidden `reused-k-decision`
> chunk below silently re-runs Part 2's full computation to obtain it, via the
> same `run_cached()` disk cache Part 2 populates -- see
> `Report_22_2_KDecision.Rmd` for the annotated version.
>
> **This file was reworked to fix an invalid earlier design**: it used to
> read multiple $n_v$ values off a *single* long simulation run (one run per
> $\Delta t$/r-level/seed, with `n_voting_rounds = NV_MAX`, indexing into its
> history arrays at each intermediate voting round). That is **not** valid:
> $n_v$ determines *which* rounds are voting rounds via
> `round(seq(T/n_v, T, length.out = n_v))`, and the model tracks
> `pct_T = t/T` per round (`Network.R`'s snapshot fields) -- so the state at
> round 50 of a $T=500$ run is not the same as the state at round 50 of a
> fresh $T=50$ run, even when both happen to share the same voting-round
> schedule up to that point. Every $(n_v, \Delta t)$ cell below is now its
> own independent, complete `simulate_liquid_democracy()` call.
>
> **This file currently implements Phase 1 only**: a 1-seed structural/
> sanity check that the sweep is built correctly and the plot shape looks
> plausible. Phase 2 (many seeds, averaged steady-state, saturation
> read-off, final $(n_v, \Delta t, T)$ decision) is deliberately **not**
> implemented yet, so this file no longer produces `CHOSEN_NV_22`/
> `CHOSEN_DT_22`/`CHOSEN_T_22`. **Known consequence, not fixed here (out of
> scope for this pass):** `Report_22_3_RSaturation.Rmd`'s own
> `reused-t-calibration` chunk still mirrors the old (invalid) single-long-run
> method and still produces those three values -- Part 3 needs rewiring to
> this file's real Phase 2 output once that exists.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 13, fig.height = 8)
library(tidyverse)
library(here)
library(knitr)
library(furrr)
library(digest)
library(igraph)
library(mgcv)

theme_set(theme_minimal(base_size = 12))
source(here::here("scripts/Network.R"))
dir.create(here::here("reports/Report_22_2b_TCalibration_files/figure-gfm"), recursive = TRUE, showWarnings = FALSE)
dir.create(here::here("data"), showWarnings = FALSE)

N_WORKERS <- max(1L, parallel::detectCores() - 1)

run_cached <- function(cache_name, fn, hash_obj) {
  hash       <- digest::digest(hash_obj)
  cache_file <- here::here(paste0("data/", cache_name, "_cache.rds"))
  if (file.exists(cache_file)) {
    cc <- readRDS(cache_file)
    if (identical(cc$hash, hash)) return(cc$data)
  }
  out <- fn()
  saveRDS(list(hash = hash, data = out), cache_file)
  out
}

metrics_core   <- c("flickerness", "delegation_rate", "lost_vote_rate", "chain_length")
metric_labels  <- c(flickerness = "flickerness", delegation_rate = "Delegation rate",
                     lost_vote_rate = "Lost-vote rate", chain_length = "Avg. chain length",
                     top5_power_share = "Top-5% voting-power share")

RATE_METRICS_01 <- c("flickerness", "Delegation rate", "Lost-vote rate", "Top-5% voting-power share")

y_anchor_01 <- function(metric_factor_levels, x_val) {
  present <- intersect(RATE_METRICS_01, metric_factor_levels)
  if (!length(present)) return(tibble(metric = factor(character(0), levels = metric_factor_levels), y = numeric(0), x = numeric(0)))
  tidyr::crossing(metric = factor(present, levels = metric_factor_levels), y = c(0, 1), x = x_val)
}

FACET_SPACING <- theme(panel.spacing.y = unit(1, "cm"))

steady_state_per_seed <- function(raw_df, round_max, window, id_cols, metrics = metrics_core) {
  raw_df |>
    filter(round > round_max - window) |>
    pivot_longer(all_of(metrics), names_to = "metric", values_to = "value") |>
    group_by(across(all_of(c(id_cols, "metric")))) |>
    summarise(steady_val = mean(value, na.rm = TRUE), .groups = "drop")
}
steady_state_summary <- function(per_seed_df, group_cols) {
  per_seed_df |>
    group_by(across(all_of(c(group_cols, "metric")))) |>
    summarise(mean_val = mean(steady_val, na.rm = TRUE),
              sd_val   = sd(steady_val, na.rm = TRUE),
              n_val    = sum(!is.na(steady_val)),
              se_val   = sd_val / sqrt(n_val), .groups = "drop")
}

# Context-sweep constants this part itself needs (same values as Part 3's
# copy). No T_CONTEXT/NV_CONTEXT here -- each (nv, dt) cell below derives its
# own T = nv * dt directly; this file's Phase 1 does not yet produce
# CHOSEN_T_22/CHOSEN_NV_22 (see the split note above).
N_BASE         <- 1000L
LAMBDA_CONTEXT <- 0.1
```

```{r reused-k-decision, include=FALSE}
R_SHARED_LEVELS <- c(0, 1, 4, 16, 32)

r_triples <- tibble(
  triple_id = seq_along(R_SHARED_LEVELS),
  r_op = R_SHARED_LEVELS,
  r_pw = R_SHARED_LEVELS,
  r_tr = R_SHARED_LEVELS
)
N_R_TRIPLES <- nrow(r_triples)

WARMUP_CONFIGS <- tribble(
  ~warmup_label, ~T_warmup, ~nv_warmup,
  "T=30, nv=10",  30L,       10L,
  "T=60, nv=20",  60L,       20L
) |> slice(1)  # FAST-RUN: must match Report_22_2's copy of this chunk
LAMBDA_WARMUP <- 0.1

K_SWEEP_THRESH <- seq(10, 150, by = 20)  # FAST-RUN: must match Report_22_2's copy of this chunk
A_THRESHOLD <- 1
TARGET_QUALIFYING <- 12
N_SEEDS_K <- 5L
N_THRESH  <- 1000L # FAST-RUN: lowest possible -- must match Report_22_2's copy of this chunk

count_qualifying_cell <- function(seed, K, triple_id, r_op, r_pw, r_tr, T_warmup, nv_warmup) {
  res <- simulate_liquid_democracy(
    seed = seed, n_per_community = N_THRESH, n_communities = 1, node_degree = K, T = T_warmup,
    n_voting_rounds = nv_warmup,
    r_op = r_op, r_pw = r_pw, gamma = r_tr, lambda = LAMBDA_WARMUP, k1 = 1, k2 = 1,
    self_reliance_c = 0.5, trust_mode = "relaxation",
    sigma_opinion = "auto", fast_path = TRUE, p_rewire = 0.05
  )
  gF  <- res$friendship_graph
  op  <- res$agents$opinion
  pow <- res$agents$power
  tau_list <- res$tau_final
  n_all <- nrow(res$agents)

  qualifying_counts <- vapply(seq_len(n_all), function(i) {
    nb <- as.integer(neighbors(gF, i, mode = "out"))
    if (!length(nb)) return(0L)
    a_op_pw   <- attractiveness_expose(op[i], op[nb], pow[i], pow[nb], r_op, r_pw)
    trust_mod <- if (!is.null(tau_list)) 2 * .sig(r_tr * (tau_list[[i]] - 1)) else 1
    sum(a_op_pw * trust_mod >= A_THRESHOLD)
  }, integer(1L))

  tibble(seed = seed, K = K, triple_id = triple_id, mean_qualifying = mean(qualifying_counts))
}

grid_thresh <- tidyr::crossing(K = K_SWEEP_THRESH, triple_id = r_triples$triple_id, seed = seq_len(N_SEEDS_K)) |>
  left_join(r_triples, by = "triple_id")

run_warmup_pass <- function(T_warmup, nv_warmup, cache_suffix) {
  run_cached(paste0("r22k_thresh_raw_", cache_suffix), function() {
    plan(multisession, workers = N_WORKERS)
    out <- future_pmap_dfr(grid_thresh, function(K, triple_id, seed, r_op, r_pw, r_tr)
      count_qualifying_cell(seed, K, triple_id, r_op, r_pw, r_tr, T_warmup, nv_warmup),
      .options = furrr_options(seed = NULL))
    plan(sequential)
    out
  }, list(grid_thresh, T_warmup, nv_warmup, A_THRESHOLD, N_THRESH, v = 1))
}

thresh_raw_all <- WARMUP_CONFIGS |>
  purrr::pmap(function(warmup_label, T_warmup, nv_warmup) {
    run_warmup_pass(T_warmup, nv_warmup, gsub("[^A-Za-z0-9]", "", warmup_label)) |>
      mutate(warmup_label = warmup_label)
  }) |>
  bind_rows() |>
  mutate(warmup_label = factor(warmup_label, levels = WARMUP_CONFIGS$warmup_label))

thresh_per_triple <- thresh_raw_all |>
  group_by(warmup_label, K, triple_id) |>
  summarise(mean_over_seeds = mean(mean_qualifying), .groups = "drop")

thresh_summary <- thresh_per_triple |>
  group_by(warmup_label, K) |>
  summarise(mean_val = mean(mean_over_seeds),
            min_val  = min(mean_over_seeds),
            max_val  = max(mean_over_seeds),
            sd_val   = sd(mean_over_seeds), .groups = "drop")

decision_by_warmup <- thresh_summary |>
  group_by(warmup_label) |>
  group_modify(~ {
    clearing <- .x |> filter(mean_val >= TARGET_QUALIFYING) |> pull(K)
    reached  <- length(clearing) > 0
    tibble(target_reached  = reached,
           chosen_k        = if (reached) min(clearing) else max(K_SWEEP_THRESH),
           max_qualifying  = max(.x$mean_val))
  }) |>
  ungroup()

WARMUPS_AGREE <- length(unique(decision_by_warmup$chosen_k)) == 1 &&
  all(decision_by_warmup$target_reached)
CHOSEN_K_22 <- if (WARMUPS_AGREE) decision_by_warmup$chosen_k[1] else max(decision_by_warmup$chosen_k)
```

# 2b. T/n_voting_rounds calibration (Phase 1: structural sweep)

T/n_voting_rounds has been a completely unaddressed placeholder throughout
this report (flagged explicitly in Sec 2.3 above as a source of circularity
in K's own calibration). This section calibrates it: `n_voting_rounds`
($n_v$) is swept for a handful of round-spacing values
$\Delta t \in \{1, 2, 3, 5, 10\}$; $T = n_v \times \Delta t$ is *derived*, not
swept independently. Robustness is checked against two background
responsiveness settings, $r_{op}=r_{pw}=r_{tr} \in \{0, 16\}$, at
K = `r CHOSEN_K_22` (this section's own `reused-k-decision` chunk). The same
`metrics_core` used throughout the rest of this report (no new metrics) are
used here too.

**Every $(n_v, \Delta t, r)$ cell below is its own independent, complete
`simulate_liquid_democracy()` call** at that cell's own $T = n_v \times
\Delta t$ -- not read off a single longer run at different checkpoints. An
earlier version of this section did exactly that (one run per $\Delta t$/r/
seed with `n_voting_rounds = NV_MAX`, indexing its history arrays at each
intermediate voting round to save compute) and it is invalid: $n_v$
determines *which* rounds are voting rounds
(`round(seq(T/n_v, T, length.out = n_v))`), and the model tracks
`pct_T = t/T` per round, so the state at round 50 of a much longer run is not
the same as the state at round 50 of a run whose $T$ actually is 50. By
construction below (a fresh `simulate_liquid_democracy()` call per grid row),
no two $n_v$ values on the same $\Delta t$/r line share a run object.

**This section currently implements Phase 1 only**: 1 seed per cell, purely
to confirm the sweep is built correctly and the plot shape looks plausible --
not a final, low-noise result. Phase 2 (Sec 2b.4) is not yet implemented.

## 2b.1 Grid and computation

```{r p2b-config}
DT_CANDIDATES <- c(1, 2, 3, 5, 10)   # Delta t candidates: spacing (rounds) between voting rounds
NV_SWEEP      <- c(1, 2, 5, 10, 20, 30, 40, 50)  # extend upward if no plateau is visible by
                                                 # nv=50 in Sec 2b.2's plot below -- don't fix
                                                 # this before actually seeing the plot
R_SETTINGS    <- c(0, 16)            # background r_op=r_pw=r_tr levels to cross-check against

N_SEEDS_T <- 1L  # Phase 1: structural check only. Phase 2 will raise this (e.g. to 10) and
                 # average steady-state across seeds -- not implemented yet, see Sec 2b.4.
```

```{r p2b-run}
run_nv_dt_cell <- function(seed, nv, dt, r_level) {
  T_val <- nv * dt
  res <- simulate_liquid_democracy(
    seed = seed, n_per_community = N_BASE, n_communities = 1, node_degree = CHOSEN_K_22,
    T = T_val, n_voting_rounds = nv,
    r_op = r_level, r_pw = r_level, gamma = r_level, lambda = LAMBDA_CONTEXT, k1 = 1, k2 = 1,
    self_reliance_c = 0.5, trust_mode = "relaxation",
    sigma_opinion = "auto", fast_path = TRUE, p_rewire = 0.05,
    snap_rounds = seq_len(T_val)  # needed for chain_length, which lives in res$snapshots
  )
  # Steady-state = mean of the last 5 rounds of THIS run only (or fewer, if
  # T_val itself is under 5 -- max(1, ...) guards the T_val<5 edge case).
  window   <- max(1, T_val - 4):T_val
  snap_idx <- match(window, res$snapshots$round)  # match on the round column, not
                                                   # positional indexing -- robust
  tibble(
    seed = seed, nv = nv, dt = dt, r_level = r_level,
    delegation_rate = mean(res$history_delegation[window], na.rm = TRUE),
    lost_vote_rate  = mean(res$history_lost[window], na.rm = TRUE),
    flickerness     = mean(res$history_flickerness[window], na.rm = TRUE),
    chain_length    = mean(res$snapshots$avg_chain_length_delegators[snap_idx], na.rm = TRUE)
  )
}

grid_nv_dt <- tidyr::crossing(nv = NV_SWEEP, dt = DT_CANDIDATES, r_level = R_SETTINGS,
                               seed = seq_len(N_SEEDS_T))

t_calib_raw <- run_cached(sprintf("r22t_calib_phase1_seeds%d", N_SEEDS_T), function() {
  plan(multisession, workers = N_WORKERS)
  out <- future_pmap_dfr(grid_nv_dt, function(nv, dt, r_level, seed)
    run_nv_dt_cell(seed, nv, dt, r_level),
    .options = furrr_options(seed = NULL))
  plan(sequential)
  out
}, list(grid_nv_dt, CHOSEN_K_22, N_BASE, LAMBDA_CONTEXT, N_SEEDS_T, v = 1))
# Cache key includes N_SEEDS_T (both in the cache_name string and this hash
# list) so Phase 2's later multi-seed run gets its own cache entry, not a
# false hit against this Phase 1 run.

t_summary <- t_calib_raw |>
  pivot_longer(all_of(metrics_core), names_to = "metric", values_to = "value") |>
  group_by(nv, dt, r_level, metric) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop") |>
  mutate(metric = factor(metric_labels[metric], levels = unname(metric_labels[metrics_core])))
```

## 2b.2 Plots

One plot per background r-setting, faceted by metric, $n_v$ on the x-axis,
one line per $\Delta t$ -- the direct analogue of `Report_20.Rmd`'s `1.2b`
plot (`facet_wrap(~metric)`, one coloured line per grouping variable), with
`x = nv` and `colour = factor(dt)` instead of `x = round`/`colour = factor(K)`.

```{r p2b-plot-fn}
plot_t_calib <- function(r_val) {
  t_summary |>
    filter(r_level == r_val) |>
    ggplot(aes(nv, mean_val, colour = factor(dt))) +
    geom_line(linewidth = 0.8) + geom_point(size = 1.8) +
    scale_colour_viridis_d(name = "Δt (round spacing)", end = 0.85) +
    facet_wrap(~metric, scales = "free_y", ncol = 2) +
    labs(title = sprintf("n_v / Δt sweep at r_op=r_pw=r_tr=%d", r_val),
         subtitle = sprintf("K=%d, N=%d, %d seed(s) per cell -- PHASE 1 (structural check only, not a final result)",
                             CHOSEN_K_22, N_BASE, N_SEEDS_T),
         x = "n_v (voting rounds)", y = NULL) +
    theme(legend.position = "top") + FACET_SPACING
}
```

```{r p2b-plot-r0, fig.height=9}
plot_t_calib(0)
```

```{r p2b-plot-r16, fig.height=9}
plot_t_calib(16)
```

## 2b.3 Phase 1 sanity check

Before doing anything else with this sweep: confirm above that the shape
looks plausible (no implausible sharp jump between $n_v=1$ and the rest of a
line, which is what the earlier, invalid single-long-run version produced).
With only `r N_SEEDS_T` seed per cell, the lines are expected to look noisy;
that alone is not a problem. If `NV_SWEEP`'s upper end
(`r max(NV_SWEEP)`) doesn't show a visible plateau for any metric/$\Delta t$/
r-setting combination above, extend `NV_SWEEP` further and re-knit before
moving on.

## 2b.4 Phase 2 -- not yet implemented

Once Sec 2b.2's plots have been reviewed and confirmed structurally sound,
Phase 2 will: raise `N_SEEDS_T` (e.g. to 10) and average steady-state values
across seeds per cell before plotting; read off, by inspection (annotated
directly in prose, no automated changepoint detector), the smallest $n_v$
beyond which further increases no longer visibly change any metric across
all $\Delta t$ lines, take the largest such $n_v$ across the 4 metrics, add a
small safety margin (e.g. +10%), and cross-check that the same $\Delta t$
also sits past its own saturation point; cross-check the two r-settings
against each other, taking the more conservative (larger) $n_v$/$\Delta t$
if they disagree; and finally state the chosen $(n_v, \Delta t, T)$
explicitly, in the same "plot-anchored modelling assumption" style used
elsewhere in this report for K, producing `CHOSEN_NV_22`/`CHOSEN_DT_22`/
`CHOSEN_T_22` for Part 3 to consume. **None of that is implemented in this
file yet** -- do not silently jump ahead to it.
