---
title: "RQ1-3 -- Effect of Responsiveness on Liquid Democracy Outcomes"
output: github_document
date: "2026-08-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 13, fig.height = 8)
library(tidyverse)
library(here)
library(knitr)
library(furrr)
library(digest)

theme_set(theme_minimal(base_size = 12))
source(here::here("scripts/Network.R"))
dir.create(here::here("reports/RQ1_files/figure-gfm"), recursive = TRUE, showWarnings = FALSE)
dir.create(here::here("data"), showWarnings = FALSE)

N_WORKERS <- max(1L, parallel::detectCores() - 1)

run_cached <- function(cache_name, fn, hash_obj) {
  hash       <- digest::digest(hash_obj)
  cache_file <- here::here(paste0("data/", cache_name, "_cache.rds"))
  if (file.exists(cache_file)) {
    cc <- readRDS(cache_file)
    if (identical(cc$hash, hash)) return(cc$data)
  }
  out <- fn()
  saveRDS(list(hash = hash, data = out), cache_file)
  out
}

# RQ1: delegation behaviour. RQ2: quality of representation (ideological
# drift). RQ3: concentration of voting power. All three are answered off the
# SAME simulation grid (see "Shared design" below) -- one run per (pair,
# fixed-level, seed) cell produces every metric needed for all three RQs at
# once, so the sweep only has to run once.
metrics_rq1 <- c("delegation_rate", "lost_vote_rate", "n_direct_voters", "chain_length", "flickerness")
metrics_rq2 <- c("avg_drift")
metrics_rq3 <- c("top5_power_share", "share_50pct")
metrics_all_rqs <- c(metrics_rq1, metrics_rq2, metrics_rq3)

metric_labels <- c(
  delegation_rate  = "Delegation rate",
  lost_vote_rate   = "Lost-vote rate",
  n_direct_voters  = "Direct voters (N x (1 - delegation rate))",
  chain_length     = "Avg. delegation-chain length",
  flickerness      = "Flickerness (delegation-target stability)",
  avg_drift        = "Avg. ideological drift",
  top5_power_share = "Top-5% voting-power share",
  share_50pct      = "Agents needed for 50% of power"
)
```

## Fixed inputs

```{r rq-fixed-inputs}
# ══════════════════════════════════════════════════════════════════════
# KEEP IN SYNC WITH REPORT_21 -- these are Report_21's Part 1-3 decisions,
# redefined here rather than sourced from that Rmd (more robust to
# Report_21 changing independently). Update whenever Report_21's N/K/T/c
# choices change. Values below match Report_21's current PROVISIONAL picks.
# ══════════════════════════════════════════════════════════════════════
CHOSEN_N          <- 500L
CHOSEN_K          <- 18L
CHOSEN_T          <- 60L    # T=60, 20 of them voting rounds -- see chat
CHOSEN_VOTING_RDS <- 20L
CHOSEN_C          <- 0.5    # Report_21 Part 3: value hitting the 50% baseline delegation target
# lambda: half-life = CHOSEN_VOTING_RDS/2 voting-round (trust-update) steps,
# i.e. (1-lambda)^h = 0.5 for h = n_voting_rounds/2 -- see Report_21.Rmd's
# global-config chunk for the full derivation and why this must stay a
# FIXED constant (not recomputed per experiment). ~0.0670 at
# CHOSEN_VOTING_RDS = 20, expressed relative to this report's own round
# design instead of a bare constant.
LAMBDA_TR         <- 1 - 0.5^(2 / CHOSEN_VOTING_RDS)
```

**N = `r CHOSEN_N`, K = `r CHOSEN_K`, T = `r CHOSEN_T`
(`r CHOSEN_VOTING_RDS` voting rounds), c = `r CHOSEN_C`.**

## Extreme behaviour: single-dial vs. isolated dimensions, r up to 512

Before the paired sweep below (bounded at $r\le32$), a wider look across
the full doubling range this thesis considers, up to $r=512$
($r \in \{0,1,2,4,8,16,32,64,128,256,512\}$) -- fixed at N=1000, K=16 --
showing every metric used in RQ1-3 below at once, to see where each one
saturates rather than assuming the $r\le32$ range used in the paired sweep
already captures the full picture. Four scenarios: **all**
($r_{op}=r_{pw}=r_{tr}=r$, one shared dial) and each dimension **isolated**
in turn ($r_{op}=r$ with $r_{pw}=r_{tr}=0$; $r_{pw}=r$ with
$r_{op}=r_{tr}=0$; $r_{tr}=r$ with $r_{op}=r_{pw}=0$) -- so the joint
effect and each mechanism's own isolated effect are visible side by side.

```{r rq-extreme-config}
N_EXTREME       <- 1000L
K_EXTREME       <- 16L
R_VALS_EXTREME  <- c(0, 2^(0:9))  # 0, 1, 2, 4, 8, 16, 32, 64, 128, 256, 512
N_SEEDS_EXTREME <- 20L

scenarios_extreme <- tribble(
  ~scenario,     ~use_op, ~use_pw, ~use_tr,
  "all (r=r=r)", TRUE,    TRUE,    TRUE,
  "op only",     TRUE,    FALSE,   FALSE,
  "pw only",     FALSE,   TRUE,    FALSE,
  "trust only",  FALSE,   FALSE,   TRUE
)
```

```{r rq-extreme-run}
run_extreme_cell <- function(seed, r_val, scenario, use_op, use_pw, use_tr) {
  r_op <- if (use_op) r_val else 0
  r_pw <- if (use_pw) r_val else 0
  r_tr <- if (use_tr) r_val else 0
  res <- simulate_liquid_democracy(
    seed = seed, n_per_community = N_EXTREME, n_communities = 1, node_degree = K_EXTREME, T = CHOSEN_T,
    n_voting_rounds = CHOSEN_VOTING_RDS,
    r_op = r_op, r_pw = r_pw, gamma = r_tr, lambda = LAMBDA_TR, k1 = 1, k2 = 1,
    decision_model = "expose", self_reliance_c = CHOSEN_C, trust_mode = "relaxation",
    sigma_opinion = "auto", fast_path = TRUE, p_rewire = 0.05
  )
  last5 <- max(1L, CHOSEN_T - 4L):CHOSEN_T
  snaps <- res$snapshots |> dplyr::filter(round >= max(1L, CHOSEN_T - 4L))
  delegation_rate <- mean(res$history_delegation[last5], na.rm = TRUE)
  tibble(
    seed = seed, r_val = r_val, scenario = scenario,
    delegation_rate  = delegation_rate,
    lost_vote_rate   = mean(res$history_lost[last5], na.rm = TRUE),
    n_direct_voters  = N_EXTREME * (1 - delegation_rate),
    chain_length     = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
    flickerness      = mean(res$history_flickerness[last5], na.rm = TRUE),
    avg_drift        = mean(snaps$avg_drift, na.rm = TRUE),
    top5_power_share = mean(snaps$top5_power_share, na.rm = TRUE),
    share_50pct      = mean(snaps$share_50pct, na.rm = TRUE)
  )
}

grid_extreme <- tidyr::crossing(scenarios_extreme, r_val = R_VALS_EXTREME, seed = seq_len(N_SEEDS_EXTREME))

extreme_raw <- run_cached("rq_extreme_raw", function() {
  plan(multisession, workers = N_WORKERS)
  out <- future_pmap_dfr(grid_extreme, run_extreme_cell, .options = furrr_options(seed = NULL))
  plan(sequential)
  out
}, list(N_EXTREME, K_EXTREME, CHOSEN_T, CHOSEN_VOTING_RDS, CHOSEN_C, N_SEEDS_EXTREME, R_VALS_EXTREME,
        scenarios_extreme, LAMBDA_TR, v = 6))  # v=6: seeds 2->20

extreme_summary <- extreme_raw |>
  pivot_longer(all_of(metrics_all_rqs), names_to = "metric", values_to = "value") |>
  group_by(scenario, r_val, metric) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop") |>
  mutate(metric   = factor(metric_labels[metric], levels = unname(metric_labels)),
         scenario = factor(scenario, levels = scenarios_extreme$scenario))
```

```{r rq-extreme-plot, fig.height=12}
extreme_summary |>
  ggplot(aes(factor(r_val, levels = R_VALS_EXTREME), mean_val, colour = scenario, group = scenario)) +
  geom_line(linewidth = 0.8) + geom_point(size = 1.5) +
  scale_colour_viridis_d(name = "responsiveness scenario", end = 0.85) +
  facet_wrap(~metric, scales = "free_y", ncol = 2) +
  labs(title = "Extreme behaviour -- N=1000, K=16, all vs. isolated dimensions (doubling to 512)",
       subtitle = sprintf("%d seeds per cell", N_SEEDS_EXTREME),
       x = "r (shared or isolated, doubling)", y = NULL) +
  theme(legend.position = "top")
```

---

## Shared design: paired sweep, third dimension held fixed

Same sweep grid for all three RQs. For each pair of responsiveness
dimensions, both are varied *together* (same shared value) across
$r \in \{1,2,4,8,16,32\}$, while the third dimension is held fixed at one of
4 levels ($0, 1, 8, 16$) -- shown as 4 coloured lines in the same plot.
Repeating this for each of the 3 possible pairs gives 3 plots per RQ (as many
metrics as that RQ needs, faceted), covering all three responsiveness
dimensions:

- $r_{op} = r_{pw} \in \{1,2,4,8,16,32\}$, $r_{tr} \in \{0,1,8,16\}$ (fixed)
- $r_{op} = r_{tr} \in \{1,2,4,8,16,32\}$, $r_{pw} \in \{0,1,8,16\}$ (fixed)
- $r_{pw} = r_{tr} \in \{1,2,4,8,16,32\}$, $r_{op} \in \{0,1,8,16\}$ (fixed)

```{r rq-config}
N_SEEDS_PAIR <- 20L

R_VALS_PAIR    <- c(1, 2, 4, 8, 16, 32)  # values for the two paired (jointly varied) dimensions
R_FIXED_LEVELS <- c(0, 1, 8, 16)          # levels for the third, held-fixed dimension
```

---

# Simulation grid (answers RQ1, RQ2 and RQ3 at once)


```{r rq-pair-run}
build_pair_grid <- function(pair_label, dim_fixed) {
  df <- tidyr::crossing(v = R_VALS_PAIR, f = R_FIXED_LEVELS, seed = seq_len(N_SEEDS_PAIR))
  df$pair_label <- pair_label
  df$r_op <- if (dim_fixed == "r_op") df$f else df$v
  df$r_pw <- if (dim_fixed == "r_pw") df$f else df$v
  df$r_tr <- if (dim_fixed == "r_tr") df$f else df$v
  df
}

grid_pair <- bind_rows(
  build_pair_grid("r_op x r_pw (r_tr fixed)", "r_tr"),
  build_pair_grid("r_op x r_tr (r_pw fixed)", "r_pw"),
  build_pair_grid("r_pw x r_tr (r_op fixed)", "r_op")
)

run_pair_cell <- function(seed, pair_label, v, f, r_op, r_pw, r_tr) {
  res <- simulate_liquid_democracy(
    seed = seed, n_per_community = CHOSEN_N, n_communities = 1, node_degree = CHOSEN_K, T = CHOSEN_T,
    n_voting_rounds = CHOSEN_VOTING_RDS,
    r_op = r_op, r_pw = r_pw, gamma = r_tr, lambda = LAMBDA_TR, k1 = 1, k2 = 1,
    decision_model = "expose", self_reliance_c = CHOSEN_C, trust_mode = "relaxation",
    sigma_opinion = "auto", fast_path = TRUE, p_rewire = 0.05
  )
  last5 <- max(1L, CHOSEN_T - 4L):CHOSEN_T
  snaps <- res$snapshots |> dplyr::filter(round >= max(1L, CHOSEN_T - 4L))
  delegation_rate <- mean(res$history_delegation[last5], na.rm = TRUE)
  tibble(
    seed = seed, pair_label = pair_label, v = v, f = f,
    r_op = r_op, r_pw = r_pw, r_tr = r_tr,
    # RQ1
    delegation_rate  = delegation_rate,
    lost_vote_rate   = mean(res$history_lost[last5], na.rm = TRUE),
    n_direct_voters  = CHOSEN_N * (1 - delegation_rate),
    chain_length     = mean(snaps$avg_chain_length_delegators, na.rm = TRUE),
    flickerness      = mean(res$history_flickerness[last5], na.rm = TRUE),
    # RQ2
    avg_drift        = mean(snaps$avg_drift, na.rm = TRUE),
    # RQ3
    top5_power_share = mean(snaps$top5_power_share, na.rm = TRUE),
    share_50pct      = mean(snaps$share_50pct, na.rm = TRUE)
  )
}

pair_raw <- run_cached("rq123_pair_raw", function() {
  plan(multisession, workers = N_WORKERS)
  out <- future_pmap_dfr(grid_pair, run_pair_cell, .options = furrr_options(seed = NULL))
  plan(sequential)
  out
}, list(CHOSEN_N, CHOSEN_K, CHOSEN_T, CHOSEN_VOTING_RDS, CHOSEN_C, N_SEEDS_PAIR, R_VALS_PAIR, R_FIXED_LEVELS, LAMBDA_TR, v = 5))  # v=5: seeds 2->20

pair_summary <- pair_raw |>
  pivot_longer(all_of(metrics_all_rqs), names_to = "metric", values_to = "value") |>
  group_by(pair_label, v, f, metric) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop") |>
  mutate(metric = factor(metric_labels[metric], levels = unname(metric_labels)))

# metrics_subset selects which of the 7 metrics (as their metric_labels) go
# into this particular plot -- each RQ section below passes its own subset.
plot_pair <- function(label, metrics_subset) {
  pair_summary |>
    filter(pair_label == label, metric %in% metric_labels[metrics_subset]) |>
    ggplot(aes(v, mean_val, colour = factor(f, levels = R_FIXED_LEVELS))) +
    geom_line(linewidth = 0.8) + geom_point(size = 2) +
    scale_x_log10(breaks = R_VALS_PAIR) +
    scale_colour_viridis_d(name = "fixed level", end = 0.85) +
    facet_wrap(~metric, scales = "free_y", ncol = 2) +
    labs(title = sprintf("Paired sweep -- %s", label),
         subtitle = sprintf("%d seeds per cell", N_SEEDS_PAIR),
         x = "shared r value (log scale)", y = NULL) +
    theme(legend.position = "top")
}
```

---

# RQ1: delegation behaviour

Metrics: delegation rate, lost-vote rate, number of direct voters, average
delegation-chain length, and flickerness (share of agents keeping the same
delegation target as the previous round -- `history_flickerness` from
`simulate_liquid_democracy()`, NA on round 1; see `Network.R`'s
`history_stability[t]` computation for the exact per-round definition).

```{r rq1-pair-plot-1, fig.height=14}
plot_pair("r_op x r_pw (r_tr fixed)", metrics_rq1)
```

```{r rq1-pair-plot-2, fig.height=14}
plot_pair("r_op x r_tr (r_pw fixed)", metrics_rq1)
```

```{r rq1-pair-plot-3, fig.height=14}
plot_pair("r_pw x r_tr (r_op fixed)", metrics_rq1)
```

## Answering RQ1

*(Fill in once the plots above reflect the real simulation run -- this is
a scaffold, not a pre-written conclusion.)*

- **Ideological responsiveness ($r_{op}$):** ...
- **Competence/power responsiveness ($r_{pw}$):** ...
- **Trust responsiveness ($r_{tr}$):** ...
- **Interactions between dimensions:** ...
- **Overall answer to RQ1:** ...

---

# RQ2: quality of political representation

Metric: average ideological drift (`avg_drift`) -- the mean distance between
an agent's own opinion and the opinion of the agent who ultimately casts a
vote on its behalf.

Expected: $r_{op}$ reduces drift (dissimilar neighbours become less
attractive delegates); $r_{pw}$ increases drift (powerful delegates stay
attractive even when ideologically distant); $r_{tr}$ reduces drift, same
mechanism as $r_{op}$. Because attractiveness is multiplicative, the joint
$r_{op} \times r_{pw}$ effect is expected to favour the opinion term (net
drift reduction), not a simple additive combination of the two individual
effects.

```{r rq2-pair-plot-1, fig.height=6}
plot_pair("r_op x r_pw (r_tr fixed)", metrics_rq2)
```

```{r rq2-pair-plot-2, fig.height=6}
plot_pair("r_op x r_tr (r_pw fixed)", metrics_rq2)
```

```{r rq2-pair-plot-3, fig.height=6}
plot_pair("r_pw x r_tr (r_op fixed)", metrics_rq2)
```

## Answering RQ2

*(Fill in once the plots above reflect the real simulation run -- this is
a scaffold, not a pre-written conclusion.)*

- **Ideological responsiveness ($r_{op}$):** ...
- **Competence/power responsiveness ($r_{pw}$):** ...
- **Trust responsiveness ($r_{tr}$):** ...
- **Joint $r_{op} \times r_{pw}$ effect (opinion term expected to dominate):** ...
- **Overall answer to RQ2:** ...

---

# RQ3: concentration of voting power

Metrics: voting-power share held by the top 5% of agents (`top5_power_share`)
and the fraction of agents needed to collectively hold 50% of total voting
power (`share_50pct`).

Expected: $r_{op}$ lowers concentration (more agents vote directly);
$r_{pw}$ raises concentration (rich-get-richer -- already-powerful agents
become more attractive delegates). For the joint $r_{op} \times r_{pw}$
effect: as in RQ2, low opinion-based attractiveness is expected to constrain
how much high voting power can raise a neighbour's overall attractiveness --
so increasing both together should yield *lower* concentration than high
$r_{pw}$ alone.

```{r rq3-pair-plot-1, fig.height=8}
plot_pair("r_op x r_pw (r_tr fixed)", metrics_rq3)
```

```{r rq3-pair-plot-2, fig.height=8}
plot_pair("r_op x r_tr (r_pw fixed)", metrics_rq3)
```

```{r rq3-pair-plot-3, fig.height=8}
plot_pair("r_pw x r_tr (r_op fixed)", metrics_rq3)
```

## Answering RQ3

*(Fill in once the plots above reflect the real simulation run -- this is
a scaffold, not a pre-written conclusion.)*

- **Ideological responsiveness ($r_{op}$):** ...
- **Competence/power responsiveness ($r_{pw}$):** ...
- **Joint $r_{op} \times r_{pw}$ effect (opinion term expected to constrain concentration):** ...
- **Overall answer to RQ3:** ...
