---
title: "Liquid Democracy — Experiment 2: Convergence Analysis"
output: 
  github_document:
date: "2026-03-12"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(
  fig.path = "output/figures/"
)

library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(tibble)
library(igraph)

source(here::here("scripts/Network.R"))
source(here::here("scripts/Summary.R"))
```

## Experimental Design

This experiment investigates *when* the Liquid Democracy system converges
rather than just *whether* it has converged by the final round.
We track three complementary metrics at every round:

- **Delegation stability**: fraction of lay agents not switching their
  delegation target between consecutive rounds
- **Edge similarity**: Jaccard similarity of the delegation graph between
  consecutive rounds — how similar are the two networks as a whole
- **Lost vote rate**: fraction of votes lost to cycles at each round

We run T = 500 rounds for a representative subset of parameter combinations
to observe the full convergence trajectory.
```{r conv-params}
responsiveness_vals <- c(0.1, 1, 5, 25)
p_rewire_vals       <- c(0, 0.05, 0.25)
seed_vals           <- 1:100

conv_params <- expand.grid(
  responsiveness = responsiveness_vals,
  p_rewire       = p_rewire_vals,
  seed           = seed_vals
)

cat("Convergence combinations:", nrow(conv_params), "\n")
```

---

## Helper Function
```{r stability-helper}
# Computes delegation stability for every consecutive round-pair.
# Returns a vector of length T-1.
compute_stability_over_time <- function(res) {
  n_graphs <- length(res$delegation_graphs)
  n_all    <- nrow(res$agents)
  lay_ids  <- which(res$agents$type == "lay")

  sapply(2:n_graphs, function(t) {
    g_prev <- res$delegation_graphs[[t - 1]]
    g_curr <- res$delegation_graphs[[t]]

    target_prev <- integer(n_all)
    target_curr <- integer(n_all)

    el_prev <- as_edgelist(g_prev, names = FALSE)
    el_curr <- as_edgelist(g_curr, names = FALSE)

    if (nrow(el_prev)) target_prev[el_prev[, 1]] <- el_prev[, 2]
    if (nrow(el_curr)) target_curr[el_curr[, 1]] <- el_curr[, 2]

    mean(target_prev[lay_ids] == target_curr[lay_ids])
  })
}
```

---

## Run Convergence Simulations
```{r conv-runs}
t_start <- proc.time()

library(furrr)

setting <- "T_500_seeds100" 
results_file <- here::here(paste0("simulations/output/results_sim3_", setting, ".rds"))

force_recompute <- FALSE

if (file.exists(results_file) && !force_recompute) {
  saved        <- readRDS(results_file)
  conv_raw     <- saved$conv_raw
  conv_history <- saved$conv_history
  cat("Loaded saved convergence results.\n")
  cat(sprintf("Raw rows: %d\n", nrow(conv_raw)))
  cat(sprintf("Averaged rows: %d\n", nrow(conv_history)))
  
} else {
  plan(multisession, workers = max(1, parallel::detectCores() - 1))

  T_val <- 500
  
  run_single_convergence <- function(responsiveness, p_rewire, seed) {
    
    res <- simulate_liquid_democracy(
      seed                    = seed,
      n_per_community         = 50,
      n_communities           = 5,
      node_degree             = 6,
      n_experts_per_community = 0,
      expert_connectedness    = 0,
      p_rewire                = p_rewire,
      responsiveness          = responsiveness,
      inertia                 = 0,
      T                       = T_val
    )
    
    stability <- compute_stability_over_time(res)
    
    edge_similarity <- c(NA_real_, sapply(2:T_val, function(t) {
      el_prev <- as_edgelist(res$delegation_graphs[[t - 1]], names = FALSE)
      el_curr <- as_edgelist(res$delegation_graphs[[t]],     names = FALSE)
      
      if (nrow(el_prev) == 0 && nrow(el_curr) == 0) return(1)
      if (nrow(el_prev) == 0 || nrow(el_curr) == 0) return(0)
      
      edges_prev <- paste(el_prev[, 1], el_prev[, 2], sep = "-")
      edges_curr <- paste(el_curr[, 1], el_curr[, 2], sep = "-")
      
      length(intersect(edges_prev, edges_curr)) / length(union(edges_prev, edges_curr))
    }))
    
    tibble(
      round           = 1:T_val,
      lost_vote_rate  = res$history_lost,
      stability       = c(NA_real_, stability),
      edge_similarity = edge_similarity
    )
  }
  
  conv_raw <- conv_params %>%
    mutate(sim = future_pmap(
      list(responsiveness, p_rewire, seed),
      run_single_convergence,
      .options = furrr_options(seed = TRUE)
    )) %>%
    unnest(sim)
  
  conv_history <- conv_raw %>%
    group_by(responsiveness, p_rewire, round) %>%
    summarise(
      lost_vote_rate  = mean(lost_vote_rate, na.rm = TRUE),
      stability       = mean(stability, na.rm = TRUE),
      edge_similarity = mean(edge_similarity, na.rm = TRUE),
      .groups = "drop"
    )
  
  plan(sequential)
  
  t_end   <- proc.time()
  elapsed <- (t_end - t_start)[["elapsed"]]
  cat(sprintf("Completed in %.1f seconds (%.1f min)\n", elapsed, elapsed / 60))
  
  saveRDS(
    list(
      conv_raw     = conv_raw,
      conv_history = conv_history
    ),
    results_file
  )
}

t_end   <- proc.time()
elapsed <- (t_end - t_start)[["elapsed"]]
cat(sprintf("Completed in %.1f seconds (%.1f min)\n", elapsed, elapsed / 60))
```
---

## Results

### Plot 1 — Delegation Stability over Time

A value of 1.0 means nobody switched their delegation target — the
structure is fully frozen. The red dashed line marks the 99% threshold
used as the convergence criterion.
```{r plot-conv-stability, fig.width = 10, fig.height = 6}
conv_history %>%
  filter(!is.na(stability)) %>%
  ggplot(aes(x = round, y = stability, color = factor(p_rewire))) +
  geom_line(alpha = 0.8) +
  geom_hline(yintercept = 0.99, linetype = "dashed", color = "red") +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 2) +
  labs(title    = "Delegation Stability over Time",
       subtitle = "Averaged over 100 seeds; red line = 99% stability threshold",
       x        = "Round",
       y        = "Share of Agents Not Switching Target",
       color    = "Rewiring (p)") +
  theme_minimal()
```

---

### Plot 2 — Network Structural Stability (Edge Similarity)

Jaccard similarity between the delegation graphs of consecutive rounds.
A value of 1.0 means the full edge set is identical between rounds.
```{r plot-conv-edge, fig.width = 10, fig.height = 6}
conv_history %>%
  filter(!is.na(edge_similarity)) %>%
  ggplot(aes(x = round, y = edge_similarity, color = factor(p_rewire))) +
  geom_line(alpha = 0.8) +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 2) +
  labs(title = "Network Structural Stability (Edge Similarity)",
       subtitle = "Averaged over 100 seeds",
       x     = "Round",
       y     = "Jaccard Edge Similarity",
       color = "Rewiring (p)") +
  theme_minimal()
```

---

### Plot 3 — Lost Vote Rate over Time

Tracks whether cycles self-resolve over time or persist indefinitely.
```{r plot-conv-lost, fig.width = 10, fig.height = 6}
conv_history %>%
  ggplot(aes(x = round, y = lost_vote_rate, color = factor(p_rewire))) +
  geom_line(alpha = 0.8) +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 2) +
  labs(title = "Lost Vote Rate over Time",
       subtitle = "Averaged over 100 seeds",
       x     = "Round",
       y     = "Lost Vote Rate",
       color = "Rewiring (p)") +
  theme_minimal()
```

---

### Table — Convergence Summary

At which round does each combination first reach 99% delegation stability?
Combinations that never reach the threshold within 500 rounds are marked >500.
```{r conv-table}
convergence_on_mean <- conv_history %>%
  filter(!is.na(stability)) %>%
  group_by(responsiveness, p_rewire) %>%
  summarise(
    converged_at = {
      idx <- which(stability >= 0.99)
      if (length(idx)) min(round[idx]) else NA_integer_
    },
    .groups = "drop"
  )

final_by_seed <- conv_raw %>%
  group_by(responsiveness, p_rewire, seed) %>%
  summarise(
    final_stability       = last(stability[!is.na(stability)]),
    final_edge_similarity = last(edge_similarity[!is.na(edge_similarity)]),
    final_lost_rate       = last(lost_vote_rate),
    lost_vote_sd_time     = sd(lost_vote_rate, na.rm = TRUE),
    .groups = "drop"
  )

final_by_seed %>%
  group_by(responsiveness, p_rewire) %>%
  summarise(
    final_stability_mean       = round(mean(final_stability, na.rm = TRUE), 3),
    final_stability_sd         = round(sd(final_stability, na.rm = TRUE), 3),
    final_edge_similarity_mean = round(mean(final_edge_similarity, na.rm = TRUE), 3),
    final_lost_rate_mean       = round(mean(final_lost_rate, na.rm = TRUE), 3),
    final_lost_rate_sd         = round(sd(final_lost_rate, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  left_join(convergence_on_mean, by = c("responsiveness", "p_rewire")) %>%
  arrange(responsiveness, p_rewire) %>%
  knitr::kable(
    col.names = c("Responsiveness", "Rewiring (p)",
                  "Final Stability (Mean)", "Final Stability (SD)",
                  "Final Edge Similarity (Mean)",
                  "Final Lost Rate (Mean)", "Final Lost Rate (SD)",
                  "Converged at Round"),
    caption = "Convergence diagnostics by parameter combination (averaged over 100 seeds)"
  )
```