---
title: "Simulation — Homophily: Effect on Liquid Democracy Dynamics"
output: github_document
date: "2026-04-16"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())
knitr::opts_chunk$set(fig.path = "output/figures/")

library(here)
library(tibble)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(furrr)

source(here::here("scripts/Network.R"))
source(here::here("scripts/Summary.R"))
```

## Experimental Design

This experiment adds the Metropolis opinion-shuffle step to the standard
Watts-Strogatz baseline. Rather than varying a process parameter,
`target_homophily` directly sets the **target opinion assortativity**
the shuffle must reach before the simulation loop starts. Three levels
plus a no-homophily baseline are compared. Responsiveness remains the
primary x-axis variable, as in Simulation 1.

- **Community**: 1 community, 250 lay agents, no experts
- **Seeds**: 1–100
- **Responsiveness**: 0.1, 0.5, 1, 2, 5, 10, 25
- **target_homophily**: NA (baseline), 0.2 (weak), 0.4 (moderate), 0.6 (strong)
- **homophily_t**: 5 (fixed — greedy convergence)
- **homophily_steps**: 100 000 (safety cap)
- **T**: 200 (snapshots at 50, 100, 150, 200)

```{r param-grid}
responsiveness_vals    <- c(0.1, 0.5, 1, 2, 5, 10, 25)
target_homophily_vals  <- c(NA, 0.2, 0.4, 0.6)
seed_vals              <- 1:100

param_grid <- expand.grid(
  responsiveness   = responsiveness_vals,
  target_homophily = target_homophily_vals,
  seed             = seed_vals
)

cat("Total combinations:", nrow(param_grid), "\n")
```

---

## Simulation Function

```{r sim-function}
run_single_simulation <- function(responsiveness, target_homophily, seed) {
  sim <- simulate_liquid_democracy(
    seed                    = seed,
    n_per_community         = 250,
    n_communities           = 1,
    node_degree             = 6,
    n_experts_per_community = 0,
    expert_connectedness    = 0,
    p_rewire                = 0.10,
    responsiveness          = responsiveness,
    inertia                 = 0,
    T                       = 200,
    selfweight_fn           = selfweight_argmax_log,
    attractiveness_fn       = attractiveness_log,
    sigma_opinion           = 0,
    sigma_pow               = 0,
    target_homophily        = if (is.na(target_homophily)) NULL else target_homophily,
    homophily_t             = 5,
    homophily_steps         = 100000
  )

  list(
    history = tibble(
      round                = seq_len(200),
      lost_vote_rate       = sim$history_lost,
      avg_drift            = sim$history_drift,
      delegation_rate      = sim$history_delegation,
      delegation_stability = sim$history_stability
    ),
    snapshots = sim$snapshots
  )
}
```

---

## Run Simulations

```{r run-sims}
results_file <- here::here("simulations/output/results_sim_homophily.rds")

if (file.exists(results_file)) {
  saved                  <- readRDS(results_file)
  results_time           <- saved$results_time
  results_snapshots      <- saved$results_snapshots
  results_snapshots_seed <- saved$results_snapshots_seed
  results                <- saved$results
  cat("Loaded saved results.\n")
  cat(sprintf("Snapshots: %d rows\n",   nrow(results_snapshots)))
  cat(sprintf("Time series: %d rows\n", nrow(results_time)))

} else {
  plan(multisession, workers = parallel::detectCores() - 1)
  t_start <- proc.time()

  raw <- param_grid %>%
    mutate(sim = furrr::future_pmap(
      list(responsiveness, target_homophily, seed),
      run_single_simulation,
      .options = furrr_options(seed = TRUE)
    ))

  results_time <- raw %>%
    mutate(h = map(sim, "history")) %>%
    select(responsiveness, target_homophily, h) %>%
    unnest(h) %>%
    group_by(responsiveness, target_homophily, round) %>%
    summarise(across(where(is.numeric), mean, .names = "{.col}"), .groups = "drop")

  results_snapshots_seed <- raw %>%
    mutate(s = map(sim, "snapshots")) %>%
    select(responsiveness, target_homophily, seed, s) %>%
    unnest(s)

  results_snapshots <- results_snapshots_seed %>%
    group_by(responsiveness, target_homophily, round, pct_T) %>%
    summarise(across(where(is.numeric), mean, .names = "{.col}"), .groups = "drop")

  results <- results_snapshots %>%
    filter(pct_T == 1.00) %>%
    select(-round, -pct_T)

  plan(sequential)
  t_end   <- proc.time()
  elapsed <- (t_end - t_start)[["elapsed"]]
  cat(sprintf("Completed: %d simulations in %.1f seconds (%.1f min)\n",
              nrow(param_grid), elapsed, elapsed / 60))

  saveRDS(list(
    results_time           = results_time,
    results_snapshots      = results_snapshots,
    results_snapshots_seed = results_snapshots_seed,
    results                = results
  ), results_file)

  cat(sprintf("Results saved to %s\n", results_file))
}
```

---

## Results

```{r plot-setup, include=FALSE}
x_scale <- scale_x_log10(breaks = c(0.1, 0.5, 1, 2, 5, 10, 25),
                          labels = c("0.1","0.5","1","2","5","10","25"))

hom_colors <- c("baseline" = "grey60",
                "0.2"      = "#4DBBD5",
                "0.4"      = "#E64B35",
                "0.6"      = "#3C5488")

# Replace NA with a readable label for plots
results_snapshots <- results_snapshots %>%
  mutate(homophily_label = ifelse(is.na(target_homophily), "baseline",
                                  as.character(target_homophily)))
```

### Ideological Drift

```{r plot-drift-snapshots}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = avg_drift,
             color = homophily_label)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  scale_color_manual(values = hom_colors) +
  labs(title  = "Average Ideological Drift at T = 50 / 100 / 150 / 200",
       x      = "Responsiveness to Power",
       y      = "Avg. Ideological Drift",
       color  = "Target Homophily") +
  theme_minimal()
```

### Lost Vote Rate

```{r plot-lost-snapshots}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = lost_vote_rate,
             color = homophily_label)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  scale_color_manual(values = hom_colors) +
  labs(title  = "Lost Vote Rate at T = 50 / 100 / 150 / 200",
       x      = "Responsiveness to Power",
       y      = "Lost Vote Rate",
       color  = "Target Homophily") +
  theme_minimal()
```

### Delegation Rate

```{r plot-delegation-snapshots}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = delegation_rate,
             color = homophily_label)) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  scale_color_manual(values = hom_colors) +
  labs(title  = "Delegation Rate at T = 50 / 100 / 150 / 200",
       x      = "Responsiveness to Power",
       y      = "Share of Lay Agents Delegating",
       color  = "Target Homophily") +
  theme_minimal()
```
