## Experimental Design

This experiment runs a (Watts-Strogatz) Network of 250 agents (directed) over 100 random seeds
to obtain stable average estimates of key outcomes, removing seed-specific noise.

- **Community**: 1 community, 250 lay agents
- **Seeds**: 1–100
- **Responsiveness**: 0.1, 0.5, 1, 2, 5, 10, 25
- **Rewiring (p)**: 0, 0.05, 0.10, 0.25
- **T**: 200
```{r param-grid}
responsiveness_vals <- c(0.1, 0.5, 1, 2, 5, 10, 25)
p_rewire_vals       <- c(0, 0.05, 0.10, 0.25)
seed_vals           <- 1:100

param_grid <- expand.grid(
  responsiveness = responsiveness_vals,
  p_rewire       = p_rewire_vals,
  seed           = seed_vals
)

cat("Total combinations:", nrow(param_grid), "\n")
```

---

## Simulation Function
```{r sim-function}
run_single_simulation <- function(responsiveness, p_rewire, seed) {
  sim <- simulate_liquid_democracy(
    seed           = seed,
    n_per_community = 250,
    n_communities  = 1,
    node_degree    = 6,
    n_experts_per_community = 0,
    expert_connectedness    = 0,
    p_rewire       = p_rewire,
    responsiveness = responsiveness,
    inertia        = 0,
    T              = 200
  )

  list(
    # Full time series for cheap metrics (200 rows)
    history = tibble(
      round           = seq_len(200),
      lost_vote_rate  = sim$history_lost,
      avg_drift       = sim$history_drift,
      delegation_rate = sim$history_delegation,
      delegation_stability = sim$history_stability
    ),
    # Snapshots with all metrics (4 rows)
    snapshots = sim$snapshots
  )
}
```

---

## Run Simulations
```{r run-simulations}
library(furrr)
library(tidyverse)

plan(multisession, workers = parallel::detectCores() - 1)

t_start <- proc.time()

# Run all simulations in parallel
raw <- param_grid %>%
  mutate(sim = furrr::future_pmap(
    list(responsiveness, p_rewire, seed),
    run_single_simulation,
    .options = furrr_options(seed = TRUE)
  ))

# Full time series averaged over seeds
results_time <- raw %>%
  mutate(h = map(sim, "history")) %>%
  select(responsiveness, p_rewire, seed, h) %>%
  unnest(h) %>%
  group_by(responsiveness, p_rewire, round) %>%
  summarise(across(where(is.numeric), mean, .names = "{.col}"),
            .groups = "drop")

# Snapshots averaged over seeds
results_snapshots <- raw %>%
  mutate(s = map(sim, "snapshots")) %>%
  select(responsiveness, p_rewire, seed, s) %>%
  unnest(s) %>%
  group_by(responsiveness, p_rewire, round, pct_T) %>%
  summarise(across(where(is.numeric), mean, .names = "{.col}"),
            .groups = "drop")

# Final round only
results <- results_snapshots %>%
  filter(pct_T == 1.00) %>%
  select(-round, -pct_T)

t_end   <- proc.time()
elapsed <- (t_end - t_start)[["elapsed"]]
cat(sprintf("Completed: %d simulations in %.1f seconds (%.1f min)\n",
            nrow(param_grid), elapsed, elapsed / 60))

# Reset to sequential after run
plan(sequential)
```


```{r x-axis-setup, include=FALSE}
x_scale <- scale_x_log10(
  breaks = c(0.1, 0.5, 1, 2, 5, 10, 25),
  labels = c("0.1", "0.5", "1", "2", "5", "10", "25")
)
```

## Results

### Plot 1 — Average Ideological Drift (Snapshots)
```{r plot-drift-snapshots, fig.width = 10, fig.height = 5}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = avg_drift,
             color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  labs(title   = "Average Ideological Drift at T = 50 / 100 / 150 / 200",
       x       = "Responsiveness to Power",
       y       = "Avg. Ideological Drift",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Plot 2 — Lost Vote Rate (Snapshots)
```{r plot-lost-snapshots, fig.width = 10, fig.height = 5}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = lost_vote_rate,
             color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  labs(title   = "Lost Vote Rate at T = 50 / 100 / 150 / 200",
       x       = "Responsiveness to Power",
       y       = "Lost Vote Rate",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Plot 3 — Delegation Rate (Snapshots)
```{r plot-delegation-snapshots, fig.width = 10, fig.height = 5}
results_snapshots %>%
  ggplot(aes(x = responsiveness, y = delegation_rate,
             color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  labs(title   = "Delegation Rate at T = 50 / 100 / 150 / 200",
       x       = "Responsiveness to Power",
       y       = "Share of Lay Agents Delegating",
       color   = "Rewiring (p)") +
  theme_minimal()
```
---

---
### Plot 4
### Table — Yes / No / Lost Votes by Snapshot
```{r table-yes-no, fig.width = 10, fig.height = 5}
results_snapshots %>%
  mutate(
    lost_votes = direct_yes + direct_no - liquid_yes - liquid_no
  ) %>%
  select(responsiveness, p_rewire, round,
         direct_yes, direct_no,
         liquid_yes, liquid_no, lost_votes) %>%
  mutate(across(c(direct_yes, direct_no,
                  liquid_yes, liquid_no, lost_votes),
                ~ round(.x, 1))) %>%  # <- explizit als lambda
  arrange(round, responsiveness, p_rewire) %>%
  knitr::kable(
    col.names = c("Responsiveness", "Rewiring (p)", "T",
                  "Direct Yes", "Direct No",
                  "Liquid Yes", "Liquid No", "Lost Votes"),
    caption   = "Yes / No Vote Counts at T = 50 / 100 / 150 / 200 (averaged over seeds)"
  )
```

### Plot — Majority Margins: Liquid vs. Direct Democracy (Snapshots)
```{r plot-margins-snapshots, fig.width = 10, fig.height = 5}
results_snapshots %>%
  tidyr::pivot_longer(
    cols      = c(liquid_margin, direct_margin),
    names_to  = "system",
    values_to = "margin"
  ) %>%
  mutate(system = recode(system,
    "liquid_margin" = "Liquid Democracy",
    "direct_margin" = "Direct Democracy")) %>%
  ggplot(aes(x = responsiveness, y = margin, color = system)) +
  geom_line() + geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  facet_wrap(~round, labeller = label_both, ncol = 4) +
  x_scale +
  labs(title   = "Majority Margins: Liquid vs. Direct Democracy",
       x       = "Responsiveness to Power",
       y       = "Majority Margin (Yes − No)",
       color   = "System") +
  theme_minimal()
```

---

### Plot 5 — Average Ideological Drift over Time (Time Series)
```{r plot-drift-time, fig.width = 10, fig.height = 5}
results_time %>%
  ggplot(aes(x = round, y = avg_drift,
             color = factor(p_rewire))) +
  geom_line(alpha = 0.8) +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 4) +
  labs(title   = "Average Ideological Drift over Time",
       x       = "Round",
       y       = "Avg. Ideological Drift",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Plot 6 — Lost Vote Rate over Time (Time Series)
```{r plot-lost-time, fig.width = 10, fig.height = 5}
results_time %>%
  ggplot(aes(x = round, y = lost_vote_rate,
             color = factor(p_rewire))) +
  geom_line(alpha = 0.8) +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 4) +
  labs(title   = "Lost Vote Rate over Time",
       x       = "Round",
       y       = "Lost Vote Rate",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Plot 7 — Delegation Rate over Time (Time Series)
```{r plot-delegation-time, fig.width = 10, fig.height = 5}
results_time %>%
  ggplot(aes(x = round, y = delegation_rate,
             color = factor(p_rewire))) +
  geom_line(alpha = 0.8) +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 4) +
  labs(title   = "Delegation Rate over Time",
       x       = "Round",
       y       = "Share of Lay Agents Delegating",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Plot 8 — Delegation Stability over Time (Time Series)
```{r plot-stability-time, fig.width = 10, fig.height = 5}
results_time %>%
  filter(!is.na(delegation_stability)) %>%
  ggplot(aes(x = round, y = delegation_stability,
             color = factor(p_rewire))) +
  geom_line(alpha = 0.8) +
  geom_hline(yintercept = 0.99, linetype = "dashed", color = "red") +
  facet_wrap(~responsiveness, labeller = label_both, ncol = 4) +
  labs(title    = "Delegation Stability over Time",
       subtitle = "Red line = 99% stability threshold",
       x        = "Round",
       y        = "Share of Agents Not Switching Target",
       color    = "Rewiring (p)") +
  theme_minimal()
```
