---
title: "Liquid Democracy — Experiment 1: Responsiveness to Power"
output: 
  github_document:
date: "2026-03-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = here::here())

# Output-Ordner für Plots direkt im Simulations-Ordner setzen
knitr::opts_chunk$set(
  fig.path = "Simulations/Simulations_files/figure-gfm/"
)

library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(tibble)

source("scripts/Network.R")
source("scripts/Summary.R")
```
```

## Experimental Design

This experiment explores how responsiveness to power affects the behaviour of
the Liquid Democracy system (without experts only lays)

We systematically vary three parameters:

- **Responsiveness to power**: 0.1 → 25
- **Network rewiring probability**: 0 → 0.25
- **Delegation rounds T**: 50 → 500

```{r param-grid}
responsiveness_vals <- c(0.1, 0.5, 1, 2, 5, 10, 25)
p_rewire_vals       <- c(0, 0.01, 0.05, 0.1, 0.25)
T_vals              <- c(50, 100, 200, 500)

param_grid <- expand.grid(
  responsiveness = responsiveness_vals,
  p_rewire       = p_rewire_vals,
  T              = T_vals
)

cat("Total parameter combinations:", nrow(param_grid), "\n")
```

---

## Simulation Function

Each row of the parameter grid is passed to `run_single_simulation()`, which
runs the full model and extracts all outcome metrics.

```{r sim-function}
run_single_simulation <- function(responsiveness, p_rewire, T) {

  sim <- simulate_liquid_democracy(
    seed                    = 123,
    n_per_community         = 50,
    n_communities           = 5,
    node_degree             = 6,
    n_experts_per_community = 0,
    expert_connectedness    = 0,
    p_rewire                = p_rewire,
    responsiveness          = responsiveness,
    T                       = T
  )

  stats <- summary_metrics(sim)

  # only return outcome metrics — parameter columns already exist in param_grid
  tibble(
    # Representation quality
    avg_ideological_drift       = stats$dynamic_evaluation$avg_ideological_drift,
    med_ideological_drift       = stats$dynamic_evaluation$med_ideological_drift,
    max_ideological_drift       = stats$dynamic_evaluation$max_ideological_drift,
    pct_high_drift              = stats$dynamic_evaluation$pct_high_drift,

    # Vote loss and delegation behaviour
    lost_vote_rate              = stats$network_description$lost_vote_rate,
    delegation_rate             = stats$network_description$delegation_rate,

    # Power concentration
    gini_power                  = stats$network_description$gini_power_inequality,
    top5_power_share            = stats$network_description$top5_power_share,

    # Chain structure
    avg_chain_length_all        = stats$network_description$avg_chain_length_all,
    avg_chain_length_delegators = stats$network_description$avg_chain_length_delegators,

    # Network structure
    largest_voting_bloc_share   = stats$network_description$largest_voting_bloc_share,

    # Outcome margins
    direct_yes                  = stats$dynamic_evaluation$direct_yes,
    direct_no                   = stats$dynamic_evaluation$direct_no,
    liquid_yes                  = stats$dynamic_evaluation$liquid_yes,
    liquid_no                   = stats$dynamic_evaluation$liquid_no,
    liquid_margin               = stats$dynamic_evaluation$liquid_margin,
    direct_margin               = stats$dynamic_evaluation$direct_margin,

    # Convergence
    convergence_sd              = stats$convergence$convergence_sd,
    delegation_stability        = stats$convergence$delegation_stability
  )
}
```

---

## Run All Simulations

```{r run-simulations}
t_start <- proc.time()

results <- param_grid %>%
  mutate(sim = purrr::pmap(
    list(responsiveness, p_rewire, T),
    run_single_simulation
  )) %>%
  unnest(sim)

t_end   <- proc.time()
elapsed <- (t_end - t_start)[["elapsed"]]

cat(sprintf("Grid completed: %d simulations in %.1f seconds (%.1f min)\n",
            nrow(param_grid), elapsed, elapsed / 60))
```

```{r x-axis-setup, include=FALSE}
# Shared x-axis scale applied to all plots — shows every simulated value
x_scale <- scale_x_log10(
  breaks = c(0.1, 0.5, 1, 2, 5, 10, 25),
  labels = c("0.1", "0.5", "1", "2", "5", "10", "25")
)
```

---

## Results

### Representation Quality

#### Plot 1 — Average Ideological Drift

Higher responsiveness makes agents delegate to ideologically proximate and
powerful agents, reducing the distance between their own preference and the
vote cast on their behalf.

```{r plot-drift-avg, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = avg_ideological_drift,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Average Ideological Drift",
       x       = "Responsiveness to Power",
       y       = "Avg. Ideological Drift",
       color   = "Rewiring (p)") +
  theme_minimal()
```

#### Plot 2 — Share of High-Drift Agents (drift > 0.3)

Average drift can mask heavy-tailed distributions. This plot shows the share
of agents who are severely misrepresented (drift > 0.3).

```{r plot-drift-pct, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = pct_high_drift,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Share of Severely Misrepresented Agents (Drift > 0.3)",
       x       = "Responsiveness to Power",
       y       = "Fraction of Agents",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Vote Loss and Delegation Behaviour

#### Plot 3 — Lost Vote Rate

Cycles cause votes to be lost entirely. High responsiveness reduces random
delegation and drives agents toward stable hubs, lowering cycle frequency.

```{r plot-lost, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = lost_vote_rate,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Lost Vote Rate",
       x       = "Responsiveness to Power",
       y       = "Lost Vote Rate",
       color   = "Rewiring (p)") +
  theme_minimal()
```

#### Plot 4 — Delegation Rate

The share of lay agents who delegated in the final round. 

```{r plot-delegation-rate, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = delegation_rate,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Delegation Rate (Final Round)",
       x       = "Responsiveness to Power",
       y       = "Share of Lay Agents Delegating",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Power Concentration

#### Plot 5 — Gini Coefficient of Voting Power

Overall inequality in the power distribution among represented agents.
A Gini of 0 means all agents hold equal power; 1 means a single agent
holds all votes.

```{r plot-gini, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = gini_power,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Power Inequality (Gini Coefficient)",
       x       = "Responsiveness to Power",
       y       = "Gini Coefficient of Voting Power",
       color   = "Rewiring (p)") +
  theme_minimal()
```

#### Plot 6 — Top-5% Power Share

Captures the "guru effect" (Kahng et al. 2021): the fraction of total votes
held by the most powerful 5% of agents among represented agents. More
sensitive to extreme concentration than the Gini coefficient.

```{r plot-top5, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = top5_power_share,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Top-5% Power Share (Guru Effect)",
       x       = "Responsiveness to Power",
       y       = "Share of Total Votes (Top 5%)",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Chain Structure

#### Plot 7 — Average Chain Length (Delegating Agents Only)

Chain length among agents who actually delegated. Longer chains indicate
more transitive delegation steps before reaching a final voter. Direct
voters (chain = 0) are excluded to allow fair comparison across conditions
with different delegation rates.

```{r plot-chain, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = avg_chain_length_delegators,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Average Delegation Chain Length (Delegating Agents Only)",
       x       = "Responsiveness to Power",
       y       = "Avg. Chain Length",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Network Structure

#### Plot 8 — Largest Voting Bloc Share

The largest connected delegation cluster as a share of all agents. A large
bloc indicates that a single representative controls a disproportionate
fraction of the electorate.

```{r plot-bloc, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = largest_voting_bloc_share,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Largest Voting Bloc Share",
       x       = "Responsiveness to Power",
       y       = "Share of All Agents",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Convergence

#### Plot 9a — Delegation Stability (Final 20 Rounds)

Fraction of lay agents who kept the same delegation target between
consecutive rounds, averaged over the last 20 round-pairs. A value of 1.0
means nobody changed their target — the delegation structure has fully
converged.

```{r plot-stability, fig.width = 9, fig.height = 4}
ggplot(results,
       aes(x = responsiveness, y = delegation_stability,
           color = factor(p_rewire))) +
  geom_line() + geom_point() +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Delegation Stability (Final 20 Rounds)",
       x       = "Responsiveness to Power",
       y       = "Share of Agents Not Switching Target",
       color   = "Rewiring (p)") +
  theme_minimal()
```

---

### Outcome Comparison

#### Plot 10 — Majority Margins: Liquid vs. Direct Democracy

`liquid_margin` and `direct_margin` show the majority margin (yes votes
minus no votes) under each system. Both are shown side by side to make
visible whether and how strongly delegation shifts the collective outcome.
A positive value indicates a Yes majority; negative indicates a No majority.

Note: `liquid_margin` is computed over represented agents only (cycles
excluded), while `direct_margin` covers all agents — so they are not
directly commensurable in absolute terms, but the direction and sign
of each margin is meaningful on its own.

```{r plot-margins, fig.width = 9, fig.height = 5}
results %>%
  group_by(responsiveness, T) %>%
  summarise(
    liquid_margin = mean(liquid_margin),
    direct_margin = mean(direct_margin),
    .groups = "drop"
  ) %>%
  tidyr::pivot_longer(cols = c(liquid_margin, direct_margin),
                      names_to  = "system",
                      values_to = "margin") %>%
  mutate(system = recode(system,
                         "liquid_margin" = "Liquid Democracy",
                         "direct_margin" = "Direct Democracy")) %>%
  ggplot(aes(x = responsiveness, y = margin, color = system)) +
  geom_line() + geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
  facet_wrap(~T, labeller = label_both) +
  x_scale +
  labs(title   = "Majority Margins: Liquid vs. Direct Democracy",
       x       = "Responsiveness to Power",
       y       = "Majority Margin (Yes − No)",
       color   = "System") +
  theme_minimal()
```
## Yes/No Table
```{r Table-Yes/No, fig.width = 9, fig.height = 5}
results %>%
  group_by(responsiveness, T) %>%
  summarise(
    direct_yes  = mean(direct_yes),
    direct_no   = mean(direct_no),
    liquid_yes  = mean(liquid_yes),
    liquid_no   = mean(liquid_no),
    lost_votes  = mean((direct_yes + direct_no) - (liquid_yes + liquid_no)),
    .groups = "drop"
  ) %>%
  mutate(across(c(direct_yes, direct_no, liquid_yes, liquid_no, lost_votes), round, 1)) %>%
  arrange(T, responsiveness) %>%
  knitr::kable(
    col.names = c("Responsiveness", "T",
                  "Direct Yes", "Direct No",
                  "Liquid Yes", "Liquid No",
                  "Lost Votes"),
    caption = "Yes/No Vote Counts: Direct vs. Liquid Democracy (averaged over rewiring conditions)"
  )
```



---

## Full Results Table

```{r results-table}
results %>%
  arrange(p_rewire, T, responsiveness) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4)))
```
